### 2017-08-15 paper.tex
First LaTeX version. Currently missing figures

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%

\documentclass{nature}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{natbib}

\usepackage{hyperref}
\usepackage{url}
\usepackage{microtype}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{tabularx}

\title{How and Why to run a Hack Week}

%% Notice placement of commas and superscripts and use of &
%% in the author list

\author{Jake VanderPlas, Karthik Ram, Daniela Huppenkothen, David W. Hogg, Ariel Rokem, Anthony Arendt \& Tal Yarkoni}

\begin{document}

\maketitle

\begin{affiliations}
 \item UW Seattle
 \item NYU
 \item UC Berkeley
\end{affiliations}

\begin{abstract}
Across almost all scientific disciplines, the instruments that record our experimental data and the methods required for storage and data analysis are rapidly increasing in complexity. This gives rise to the need for scientific communities to adapt on shorter time scales than traditional university curricula allow for, and therefore requires new modes of knowledge transfer. At the same time, as disciplines become more specialized, data analysis problems often tend to be fairly universal. It is thus desirable to foster exchange of ideas and computational workflows across disciplines. Among the different approaches put forward toward addressing these issues, hack weeks have in recent years emerged as both an effective tool for training researchers in modern data analysis workflows, and in fostering exchange between subdisciplines within and between scientific fields. While interpreted differently in the various disciplines where they have been implemented, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and free-form project work. In this paper, we present the concept of a hack week in the larger context of scientific meetings and point out similarities and differences to traditional conferences. We motivate the need for such an event and present in detail its strengths and challenges. We find that hack weeks are successful at cultivating collaboration and the exchange of knowledge. Participants self-report that these events help them both in their day-to-day research as well as their careers. Based on our results, we conclude that hack weeks present an effective, easy-to-implement, fairly low-cost tool to positively impact data analysis literacy in academic disciplines, foster collaboration and cultivate best practices.
\end{abstract}

\section*{Introduction}
\label{sec:introduction}

As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
At the same time, formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998\footnote{\url{https://f1000research.com/articles/3-62/v2}}.
At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.; in the Astronomy community, one of the longest-running has been at the Penn State Center for Astrostatistics *Summer School in Statistics for Astronomers*, consisting of ''intensive week covering basic statistical inference, several fields of applied statistics, and the R computing environment``\footnote{\url{http://astrostatistics.psu.edu/su16/}}. In the geospatial community, some software providers are taking the lead on educating the community, for example through the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}.
Another model of extra-curricular meeting is less pedagogical and more project-focused, perhaps typified by the dotAstronomy meetings, a week-long conference series which focuses specifically on leveraging web technologies for novel applications in Astronomy\footnote{\url{http://dotastronomy.com}}.
Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, where Astronomers gather to work intensively on collaborative projects, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\footnote{\url{https://gigascience.biomedcentral.com/articles/10.1186/s13742-016-0121-x}}, and a hackathon associate with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
In broad-brush, pedagogically-focused workshops and summer schools tend to follow a classic academic model where novices learn a skillset from experts, while project-focused workshops focus on people collaboratively exercising the skillset they already have.
A disadvantage of the summer school model is that it can tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
A disadvantage of the hackathon model is the common perception (whether accurate or not) that the week is designed for experts in technical tools, which may discourage others from attending.
In 2014, we started Astro Hack Week to try to fill the gaps between these models. The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for collaborative hacks or creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.

\begin{figure}
\begin{center}
\includegraphics[width=9cm]{fig/HackSpectrum.png}
\caption{Comparison of Extracurricular Workshop Models}
\label{fig:hackspectrum}
\end{center}
\end{figure}

As of the publication of this paper, we have run five such hack week events: three focused on Astronomy, one focused on Neuroscience, and one focused on Geoscience.
Below we will share some of the philosophy behind the hack week model, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.

\section*{What is a hackathon?}

Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \citep{Decker2015}. Hackathons have historically focused on software development and technology design as a way to motivate innovation within industry. In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries. In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to 'learn by doing' within a constructivist educational framework \citep{Bransford2000-lu,Papert1980-fh}. With this in mind, hackathons around scientific topics, designed to foster collaboration \citep{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \citep{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.

The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon. Following the typology of \citep{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines. Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \citep{Drouhard2017}. ''Catalytic`` hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge. ''Contributive`` hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project. Finally, ''Communal`` hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.

Our past hack week events most closely follow the communal hackathon model as it applies to scientific communities of practice. Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure 1). For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials. In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.

We note that the terminology for these events is constantly evolving, and that the ''hackathon`` concept may have implicit connotations that are disfavored in some communities. One criticism of hackathons is that they propel the ''geek`` stereotype and may present a barrier to creating an inclusive working environment, especially for women \citep{Decker2015}. Another problem is the competitive atmosphere of many industry hackathons, where teams actively compete for prizes. Because certain groups are more risk-averse than others, and risk-aversion traces demographics like gender, we urge caution when contemplating making an event competitive. Other related nomenclature includes the ''unconference`` which is also a time-bounded event gathering together a specific community, but these often have minimal structure and do not necessarily have a focus on software. The ''sprint`` or ''scrum`` label typically refers to an event focused on rapid software development on a specific set of code. This may be a component of our hack week model but it ignores the important pedagogical components.

\section*{Why run a Hack Week?}

There are several reasons to run a hack week of the sort described here.

\begin{itemize}
\item{\textit{Education and training}: Some hack weeks are more focused on education than others (see Figure 1). However, even those that are not explicitly focused on skills-development will include extensive discussions on reproducible and open science practices that are valuable to participants. Discussing these topics in a diverse group setting helps to ingrain within participants the basic conventions of the broader open source community. Participants can then serve as ''ambassadors`` of these practices and foster their adoption by other researchers in the field. Furthermore, the kind of co-working that is inherent to the hack week format facilitates lateral knolwedge transfer, where participants learn from each other about practices and tools of the trade. Often these tools are not described in research products such as papers, or even in the software implementations.}
\item{\textit{Tool development}: hack weeks present an opportunity for computational tool developers to get together with users of these tools, and to critically evaluate the use of the tools in addressing particular scientific questions.}

\item{\textit{Community building}: a hack week is an opportunity to galvanize a community within a field that cares about computational issues, and relates to specific open-source software projects. For example, because of its focus on free and open-source tools, Neurohack week helps galvanize the Neuroimaging in Python (http://nipy.org) community. It also allows researchers who may otherwise be isolated within their departments and subfield on account of their computational specialization to come together and exchange experiences. As one participant at Astro Hack Week expressed in the survey, \textit{''The main thing I took from AstroHackWeek is the fact that I'm not alone``}.}

\item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g., between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines. These experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}. This is despite the fact that interdisciplinary research can also be very impactful \citep{Hall2012-hi}. A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}

\item{\textit{Recruitment and networking}: hackweeks often bring together research, government and industry communities, providing an opportunity for participants to learn about each other's interests and their abilities in close quarters. This kind of extended opportunity for interaction and ''on-the-job`` testing is considered an excellent form of evaluation of prospective job candidates [citation needed]. Conversely, participants have an opportunity to gain exposure to potential career opportunities in communities other than their own. As an example, one of the participants in Neurohack week 2016 will be joining the eScience Institute as a post-doc in Fall of 2017 (and as an instructor in Neurohack week 2017).}

\item{\textit{It's fun}: the opportunity to focus for a week on a slightly more whimsical and adventurous sort of scientific activities can be both fruitful, and enjoyable. Even aside from the prospect of interdisciplinary research, it can provide a laboratory for more experimental approaches and high-risk projects than scientist will take on in their day-to-day research.}

\end{itemize}

It is worth noting that the reasons for participants to attend a hack week are as diverse as the goals that the organizing committee explicitly states. This is largely a function of the diversity of the participants (see section below), and should be something that organizers design for. Beginner participants may attend primarily to learn a new technique or method, while more experienced participants attend to gain more experience mentoring and teaching. Some researchers may come with particular projects and collaborators in mind, while others come with a focus on learning and with no explicit plan for the project-based work. While it is very hard to design a workshop that is universally useful to all participants, in practice many researchers attend with the understanding that some components will be more congruent with their own goals than others, but also bring an open mind and a willingness to learn. In addition, the diversity of goals can effectively be a strength of the workshop, if the organizers can facilitate matching up participants with complementary goals.

\section*{Audience and Participant Selection}

Because hack weeks are at their core non-traditional workshops, it is worth thinking in detail about the target audience for a hack week, along with methods to select an appropriate group of participants. Here, appropriate essentially refers to 'most likely to achieve the stated goal of the hack week', though measuring that achievement remains a difficult issue itself. This is especially important because hack weeks, particularly in their more project-heavy iterations, are very participant-driven workshops. At a traditional conference or summer school, the program will be largely set in advance by the organizing committee using the careful selection of speakers and lecturers. At a hack week much of the program is in flux both before and during the workshop. In this context, a successful hack week requires participants who are both interested in participation and feel comfortable participating in discussions and taking risks. Therefore, a robust, effective participant selection procedure is of crucial importance to the outcomes of the hack week and may require both careful engineering and an in-depth discussion among the organizers.

Perhaps unsurprisingly, the target audience for a hack week depends directly on the stated goal of the workshop. At conferences, researchers typically report recent results to the community and network with other scientists largely from their discipline, leading to a fairly narrow target distribution in academic seniority and represented fields. At a summer school, the aim is generally for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows. Hack weeks differ in that knowledge transfer occurs across many levels of senority, due to the relative newness of the content and the distributed knowledge base. Because the potential for knowledge transfer and community building is highest when the group is heterogeneous, participant diverse in both academic background and geographic location should be selected. This increases the chances that participants will take their newfound knowledge into their respective communities and spread it further.

Aside from these broad considerations, the tuning of the audience and participant selection will depend directly on several of the axes described in detail above. In particular, education-focused workshops might be more homogeneous than project-focused workshops, and similarly single-discipline hack weeks may be more homogeneous than interdisciplinary ones. Additional goals may be not directly related to the workshop itself, but on higher-order goals that a community might find desirable and that can be directly addressed by engineering the participant selection. For example, women are still a small minority in computational astrophysics. One goal of a hack week could be to increase the participation of gender minorities in computation-focused research. Similarly, if a particular sub-field within a discipline is ahead of others in their computational approaches, it might be worthwhile to engineer the mix of participants such that a number of experts from that sub-field are present. One interesting axis to consider is the mix of academics versus participants from industry. The latter may often be more focused on practically useful applications of methods and data sets, and can therefore provide an important counterpoint to more theory-focused academics. In particular in fields where the career path from academia to industry is less immediately obvious (as this the case e.g. in astronomy, where little directly relevant industry exists), inviting back participants who transitioned to industry successfully can provide important role models and networking opportunities for junior participants. Finally, with much of data science currently performed in the private sector, industry participants can bring directly relevant knowledge and new scientific approaches to the table.

The need to carefully select participants dovetails with another requirement: transparency. Up to date, all hack weeks we have organized have been oversubscribed by at least a factor of two, and thus necessarily require turning applicants away. The selection procedure should be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance. Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the workshop's success as well as the larger community.

In order to maximize transparency in our selection process we are working to remove the human element from the decision making as far as possible and transfer some of these tasks to a well-designed algorithm. Here, well-designed refers to several properties: (1) it is interpretable, i.e. the algorithm, its parameters and outputs can be understood by humans, (2) it is openly available, and thus the decision process can be inspected, (3) the algorithm itself does not perpetuate intrinsic biases in the data (though we will show below that an algorithm itself does not entirely remove the requirement of humans to be aware of their biases).

In practice, most selection procedures follow two separate steps, though in the case of human committees, these steps are often conflated. In the first step, the merit of a candidate must be assessed: does the candidate fulfill the requirements posed by the goals of the workshop? For example, at Astro Hack Week, it may be unrealistic to admit an undergraduate student in biochemistry, unless they had a compelling reason for attending. This first step is exceedingly difficult to automate, because it requires judgment calls, often based on long-form answers, that are as of yet difficult for computers to parse and classify. In the past, we have performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop. The second step in the selection procedure thus requires tie-breaking between equally acceptable candidates. It is here where one may impose outside constraints on the selection based on the goals of the workshop. If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which provably optimal algorithms exist that will far outperform any human selection procedure.

In particular, the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}} employs one such algorithm and is capable of efficient selection of an optimal candidate set based on human target fractions. It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest. If these targets reproduce the distribution of the overall sample of candidates, the selection will become essentially random. Any human biases involved in setting these target fractions will be perpetuated in the selection procedure. Secondly, perhaps more obviously, the algorithm can only act on information that has been collected. Biased participant sets may still result from selection procedures that fail to include crucial categories (for example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority).

Within our hack weeks the goal of the selection procedure has always been diversity among several axes. This includes academic subdiscipline, to allow for maximally possible knowledge exchange between different domain fields, as well as with researchers from adjacent fields including statistics, computer science and data science. Other important categories revolve around the participants' previous knowledge in terms of machine learning, statistics and programming. Unlike for a summer school, we have generally selected for a diverse set of participants in order to be able to efficiently pair up novices and experts in projects. Finally, we also include demographic information with two main goals: firstly, representation has been proven to be an effective way to foster an inclusive environment where minority participants feel comfortable to participate and take risks. Secondly, we would like to boost the participation of minority researchers in computational astrophysics, and our hack week model provides a useful framework for doing so.

\section*{Themes}

To date, all organized hack weeks have been subject-specific. That is, they aimed to bring together a community with a shared scientific interest, e.g.\ neuroscience, astronomy or geoscience. This has several clear advantages. In particular, communities organized by subject generally share the same language, which drastically cuts down on the time participants spend simply agreeing on scientific terms. Similarly, the scientific objectives and questions are broadly clear to all participants: there is little need to explain in detail why a particular question may be of interest to the field. Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data. This leaves more time for active collaboration on cutting-edge science.

However, there are also drawbacks to limiting the theme of a hack week to a single field. Homogeneity may be a disadvantage if it leads to \textit{group think} and inhibits new, creative solutions. Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing. These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead. In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration. On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.

An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/](http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of 2017. These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research. Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others. In addition, participants included computer vision researchers, who develop algorithms that can be applied to answer scientific questions. Similarly, a series of events at UC Berkeley (Text XD) focused on the analysis of text data across different domains of research.

\section*{Design considerations}

Design considerations depend very strongly on goals. In general, longer events allow for a larger taught component, more ambitious projects and especially for cross-disciplinary events are more likely to provide enough time for different groups to effectively communicate across barriers of professional terminology. On the other hand, because of their large participatory component, hack weeks tend to be exhausting, and long events may lead to fatigue among attendees, where active participation and measurable outcomes may drop sharply in the latter days of the workshops. This is particularly important if---as is often the case---participants continue participating in workshop-relevant activities even after the official component ends for the day.

Ways to mitigate these effects include capping taught components during the day, providing a reasonable clear schedule, and limiting parallel components to avoid decision fatigue. At Astro Hack Week, we have limited the scheduled taught components to no more than three out of nine hours for each of the five workshop days, and unscheduled (shorter) tutorials to a maximum of two per day. Even so, in our surveys, participants occasionally report concerns about choosing to attend tutorials versus working on projects.

Aside from time, space plays a particularly important role in facilitating a successful hack week. Universities in general provide a convenient venue with existing structures to facilitate hack weeks (e.g. access to scientific publications, institutional support including staff and funding). On the other hand, most spaces in universities are designed for lectures, which are diametrically opposed to unconference-style events. Thus, finding an appropriate physical space within a university may be a challenge. However, with the rise of active learning as a preferred teaching methods, traditional lecture spaces are transforming into more flexible spaces that are generally appropriate for hack weeks. As a general rule, the smaller a hack week's emphasis on a taught component, the more flexible the space has to be, with ample opportunity for re-configuration. At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine. It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not (examples are the Lorentz Center in Leiden, the Aspen Center something or other, and **others**?).

Another important design consideration is group size. With all previous hack weeks severely oversubscribed, it seems natural to simply admit more participants. However, this can counter the ideals of the workshop, in particular when building a community is one of its stated goals. If the group is too large, participants are unlikely to even meet each other, and workshop cohesion may be lost as the workshop fractures into smaller groups, often among participants who already know each other. This may inhibit knowledge transfer by clustering participants into small in-groups. Additionally, it is likely that the number participant-led components in the schedule may increase with workshop size. While generally desirable, a programme that is too crowded may lead to fatigue On the other hand, if the size of the workshop is too small, it is unlikely to achieve the desired diversity among participants to foster new collaborations across sub-fields and disciplines. In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.

As mentioned above, the balance between pedagogy and working depends both on the goals of the workshop and the topics around which the workshop is organized. If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other. In communities where a shared understanding exists, tutorials can focus on more advanced or innovative topics, and less time may be allocated for them, leaving more time for active participation. In astronomy, a relatively small field, even students generally share a common knowledge base and have rudimentary knowledge about the types of data used and the challenges associated with each. Thus, we have focused the taught component of Astro Hack Week less on domain-specific knowledge, and instead offered tutorials in topics from domain-adjacent fields like statistics and computer science that attendees are unlikely to have encountered in their regular education.

Hack week outcomes, in turn, depend strongly on participants and are often a function of their interests and seniority. A number of attendees arrive with the stated goal of writing a specific scientific article, though these tend to be more advanced participants with significant pre-knowledge of both hackathons and their topic of interest. Conversely, many attendees arrive with the stated goal of learning about a specific topic (such as machine learning) and bring a specific data set they believe the new knowledge may be applicable to. This leads to a wide variety of project types from sandbox-style explorations to focused work efforts. It is worth noting that while a scientific paper need not be the stated goal of a hack (and is unlikely to be completed in the short time allocated in any case), results may still be published as short reports or unconference proceedings. For example, Neurohack week provides a venue for participants to publish a short (two-page) ''project report`` summarizing the hack that participants did during the week of NHW. Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.

\subsection*{Box 1: Impostor Syndrome}

The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear. Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \citep{Clance1978-ef}.
Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{http://astrohackweek.org/blog/ethnographic-notes.html} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.  This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee. This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}. Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations. Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
We are working within our hack weeks to mitigate IS using various techniques. With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS. More generally, being open about the presence and prevalence of IS can help participants feel more at ease (e.g. one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.  I think it helped ease the feeling!!``}) . Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.

\section*{Measures of success}

Measuring the success of a hack week objectively is complicated by the diversity of goals of both organizing committee and attendees. A participant who attends primarily in order to hack on a specific project may find the tutorials less overall useful than a graduate student exposed to the tutorial topics for the first time. Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success. Additionally, it is difficult to define in the first place what success means to organizers and success. For more project-driven events, one key metric is the number of publications that result from hack week projects. However, this implies a very narrow definition of success in line with standard academic performance indicators, and may be hard to collect in practice, if participants do not explicitly include a reference to the event in their publication. Another, somewhat less traditional and complementary indicator is the activity of participants in terms of code written and committed to a public code repository (e.g. GitHub). This could be defined in lines of code committed, the number of new open-source projects spawned, and the number of new developers contributing to open-source projects. This assumes that participants work largely in the open during a hack week, and that most projects have a strong programming component. Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes. In particular, it ignores learning, community building as well as networking outcomes, which are often equally important objectives. One possible approach here is to ask participants to self-report their success via a post-workshop survey.
If participants are selected as described above, it is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not. This would then provide a somewhat objective measure of the impact the hack week has had. In practice, this has not been done for any of the hack weeks conducted thus far.

For practical purposes, we recommend an approach that combines all of these metrics: publications directly resulting out of hack week collaborations, new code generated and new contributors to open-source projects, self-reported learning outcomes in the topics covered in the tutorials, and survey results probing the usefulness of the workshop in terms of career and networking goals. Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.

If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week. Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent. In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating. It is important to establish procedures to conduct such studies and to obtain approval from an IRB.

\section*{Results}

In the following we report outcomes of previous events held in 2016 in three different fields (neuroscience, geoscience and astrophysics) and participants attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations. Because all three events are less than a year in the past, evaluating long-term outcomes (number of publications, citations, career outcomes) will require a longer baseline. There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \citep{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}. [NOTE: HERE IS WHERE WE SHOULD REFER TO THE BOXES ONCE THEY ARE MERGED.]
We asked participants of all three hack weeks to complete surveys to validate the hack week's success. All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.

We find that at all three hack weeks, participant self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 2, left panel). The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (see Figure 2, middle and right panels). While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
Another important goal of the hack weeks was centred on community building and fostering collaborations. We find that overall, the majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel). This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 3, right panel). 
Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants. We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories. This is expected: in particular participants new to the field may participate to learn rather than to teach.

Reproducibility and open science have recently drawn much attention from the scientific community, and all three hack weeks have been working to promote positive attitudes towards reproducibility and open science in their respective fields. We find that hack weeks have been largely successful at encouraging open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 4, left panel) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 4, middle panel). When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields. In particular, astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences. Neuroscience seems to fall somewhere in between. This also implies that in this respect, hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop. This is also reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%). Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible. 
\begin{figure*}[htbp]
\begin{center}
\includegraphics[\textwidth]{fig/eval_techskills.png}
\caption{Self-reported technical learning outcomes at all three hack weeks}
\label{fig:techskills}
\end{center}
\end{figure*}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[\textwidth]{fig/eval_collab.png}
\caption{Hack week outcomes with respect to collaboration and peer-learning}
\label{fig:techskills}
\end{center}
\end{figure*}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[\textwidth]{fig/eval_openscience.png}
\caption{Changing attitudes towards open science at hack weeks}
\label{fig:techskills}
\end{center}
\end{figure*}

\subsection*{Box 2: Examples of Hack Week Outcomes}
\subsubsection*{Example 1: Astro Hack Week}
In 2015, a small team used the opportunity of Astro Hack Week to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of algorithms for time series analysis often used in X-ray astronomy.  
The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects, role models of successful projects, and an environment encouraging scientific risk taking.
Since its beginnings at Astro Hack Week, Stingray has matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.
\subsubsection*{Example 2: Geo Hack Week} 
In 2016, a Geo Hack Week project team used Google Earth Engine to explore spatial patterns in climate, topography and population data with the goal of mapping the most suitable locations for renewable energy sites in the United States. The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine. George Richardson, one of the project leads, now works for a renewable resource company in Seattle. More information can be found [here](http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/).
\subsubsection*{Example 3: Neuro Hack Week} 
Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies.
This is a particular concern in studies in which children or patients are studied, as they are more likely to move.
In the 2016 Neuro Hack Week one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off were varied.
They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the Neuro Hack Week was over, and they eventually wrote and published a paper describing these results in the open access journal Research Ideas and Outcomes \citep{leonard2017}.

\section*{Conclusions}

The fast-paced changes of the computational and methodological landscape require traditional fields of science to rapidly adapt to new data analysis challenges. Traditional modes of learning, including university curricula, are often too slow to incorporate new developments on short enough time scales to meet their acute need in scientific advancement. To address this imbalance, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences. 
Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields.

As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility. Initial results from three events held in 2016 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community. Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods. 
We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.

%Spelling must be British English (Oxford English Dictionary)

%In addition, a cover letter needs to be written with the
%following:
%\begin{enumerate}
% \item A 100 word or less summary indicating on scientific grounds
%why the paper should be considered for a wide-ranging journal like
%\textsl{Nature} instead of a more narrowly focussed journal.
% \item A 100 word or less summary aimed at a non-scientific audience,
%written at the level of a national newspaper.  It may be used for
%\textsl{Nature}'s press release or other general publicity.
% \item The cover letter should state clearly what is included as the
%submission, including number of figures, supporting manuscripts
%and any Supplementary Information (specifying number of items and
%format).
% \item The cover letter should also state the number of
%words of text in the paper; the number of figures and parts of
%figures (for example, 4 figures, comprising 16 separate panels in
%total); a rough estimate of the desired final size of figures in
%terms of number of pages; and a full current postal address,
%telephone and fax numbers, and current e-mail address.
%\end{enumerate}

%See \textsl{Nature}'s website
%(\texttt{http://www.nature.com/nature/submit/gta/index.html}) for
%complete submission guidelines.

%\begin{methods}
%Put methods in here.  If you are going to subsection it, use
%\verb|\subsection| commands.  Methods section should be less than
%800 words and if it is less than 200 words, it can be incorporated
%into the main text.

%\end{methods}

%% Put the bibliography here, most people will use BiBTeX in
%% which case the environment below should be replaced with
%% the \bibliography{} command.

% \begin{thebibliography}{1}
% \bibitem{dummy} Articles are restricted to 50 references, Letters
% to 30.
% \bibitem{dummyb} No compound references -- only one source per
% reference.
% \end{thebibliography}

\bibliographystyle{naturemag}
\bibliography{paper}

%% Here is the endmatter stuff: Supplementary Info, etc.
%% Use \item's to separate, default label is "Acknowledgements"

\begin{addendum}
 \item Anyone thanked here: *Feel free to move up to the Author list!* It is a pleasure to thank Laura Norén (NYU) for help on ethics and IRB. This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington. Neurohackweek is supported through a grant from the National Institute for Mental Health (1R25MH112480). Daniela Huppenkothen is partially supported by the James Arthur Postdoctoral Fellowship at NYU.
 \item[Competing Interests] The authors declare that they have no
competing financial interests.
 \item[Correspondence] Correspondence and requests for materials
should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
\end{addendum}

%%
%% TABLES
%%
%% If there are any tables, put them here.
%%

%\begin{table}
%\centering
%\caption{This is a table with scientific results.}
%\medskip
%\begin{tabular}{ccccc}
%\hline
%1 & 2 & 3 & 4 & 5\\
%\hline
%aaa & bbb & ccc & ddd & eee\\
%aaaa & bbbb & cccc & dddd & eeee\\
%aaaaa & bbbbb & ccccc & ddddd & eeeee\\
%aaaaaa & bbbbbb & cccccc & dddddd & eeeeee\\
%1.000 & 2.000 & 3.000 & 4.000 & 5.000\\
%\hline
%\end{tabular}
%\end{table}

\end{document}

### 2017-08-15 paper.tex
Line break after every full stop

@@
 A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998\footnote{\url{https://f1000research.com/articles/3-62/v2}}.
-At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.; in the Astronomy community, one of the longest-running has been at the Penn State Center for Astrostatistics *Summer School in Statistics for Astronomers*, consisting of ''intensive week covering basic statistical inference, several fields of applied statistics, and the R computing environment``\footnote{\url{http://astrostatistics.psu.edu/su16/}}.
-In the geospatial community, some software providers are taking the lead on educating the community, for example through the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}.
-Another model of extra-curricular meeting is less pedagogical and more project-focused, perhaps typified by the dotAstronomy meetings, a week-long conference series which focuses specifically on leveraging web technologies for novel applications in Astronomy\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, where Astronomers gather to work intensively on collaborative projects, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\footnote{\url{https://gigascience.biomedcentral.com/articles/10.1186/s13742-016-0121-x}}, and a hackathon associate with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field. 
+Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, and more project-focused than pedagogical meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
+Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\footnote{\url{https://gigascience.biomedcentral.com/articles/10.1186/s13742-016-0121-x}}, and a hackathon associate with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
 In broad-brush, pedagogically-focused workshops and summer schools tend to follow a classic academic model where novices learn a skillset from experts, while project-focused workshops focus on people collaboratively exercising the skillset they already have.
@@
 
-The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon.
+%The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon. 
 Following the typology of \citep{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
@@
 Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing. 
-These disadvantages may be mitigated by organizing a hack week around a technique (e.g.
-Gaussian Processes) or modality (e.g.
-imaging) instead.
+These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead. 
 In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
@@
 
-An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/](http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of 2017.
+An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of 2017.
 These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
@@
 Aside from time, space plays a particularly important role in facilitating a successful hack week. 
-Universities in general provide a convenient venue with existing structures to facilitate hack weeks (e.g.
-access to scientific publications, institutional support including staff and funding).
+Universities in general provide a convenient venue with existing structures to facilitate hack weeks (e.g. access to scientific publications, institutional support including staff and funding). 
 On the other hand, most spaces in universities are designed for lectures, which are diametrically opposed to unconference-style events. 
@@
 As a general rule, the smaller a hack week's emphasis on a taught component, the more flexible the space has to be, with ample opportunity for re-configuration. 
-At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g.
-a company headquarter) to engage participants and break the routine.
-It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not (examples are the Lorentz Center in Leiden, the Aspen Center something or other, and **others**?).
+At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine. 
+It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not.
 
@@
 Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \citep{Clance1978-ef}.
-Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{http://astrohackweek.org/blog/ethnographic-notes.html} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
+Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohackweek.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
 This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
@@
 one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
-I think it helped ease the feeling!!``}) .
+I think it helped ease the feeling!!``}). 
 Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
@@
 However, this implies a very narrow definition of success in line with standard academic performance indicators, and may be hard to collect in practice, if participants do not explicitly include a reference to the event in their publication.
-Another, somewhat less traditional and complementary indicator is the activity of participants in terms of code written and committed to a public code repository (e.g.
-GitHub).
+Another, somewhat less traditional and complementary indicator is the activity of participants in terms of code written and committed to a public code repository (e.g. GitHub). 
 This could be defined in lines of code committed, the number of new open-source projects spawned, and the number of new developers contributing to open-source projects. 
 This assumes that participants work largely in the open during a hack week, and that most projects have a strong programming component. 
-Again, this may not be true for some hacks (e.g.
-tutorials, other written text, visualizations), and focuses entirely on material outcomes.
+Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes. 
 In particular, it ignores learning, community building as well as networking outcomes, which are often equally important objectives. 
@@
 In 2016, a Geo Hack Week project team used Google Earth Engine to explore spatial patterns in climate, topography and population data with the goal of mapping the most suitable locations for renewable energy sites in the United States.
-The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine.
+The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine\footnote{\url{http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/}}.
 George Richardson, one of the project leads, now works for a renewable resource company in Seattle.
-More information can be found [here](http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/).
 \subsubsection*{Example 3: Neuro Hack Week} 
@@
 To address this imbalance, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
-Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields.
+Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields
 
@@
 \begin{addendum}
- \item Anyone thanked here: *Feel free to move up to the Author list!* It is a pleasure to thank Laura Norén (NYU) for help on ethics and IRB.
+ \item{Anyone thanked here: *Feel free to move up to the Author list!* It is a pleasure to thank Laura Norén (NYU) for help on ethics and IRB.
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.
 Neurohackweek is supported through a grant from the National Institute for Mental Health (1R25MH112480).
-Daniela Huppenkothen is partially supported by the James Arthur Postdoctoral Fellowship at NYU.
- \item[Competing Interests] The authors declare that they have no
-competing financial interests.
- \item[Correspondence] Correspondence and requests for materials
-should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
+Daniela Huppenkothen is partially supported by the James Arthur Postdoctoral Fellowship at NYU.}
+ \item[Competing Interests]{The authors declare that they have no competing financial interests.}
+ \item[Correspondence]{Correspondence and requests for materials should be addressed to D.Huppenkothen.~(email: daniela.huppenkothen@nyu.edu).}
 \end{addendum}

### 2017-08-17 paper.tex

@@
 The ''sprint`` or ''scrum`` label typically refers to an event focused on rapid software development on a specific set of code. 
-This may be a component of our hack week model but it ignores the important pedagogical components.
+%This may be a component of our hack week model but it ignores the important pedagogical components.
 
@@
 \begin{itemize}
-\item{\textit{Education and training}: Some hack weeks are more focused on education than others (see Figure 1). 
-However, even those that are not explicitly focused on skills-development will include extensive discussions on reproducible and open science practices that are valuable to participants. 
-Discussing these topics in a diverse group setting helps to ingrain within participants the basic conventions of the broader open source community. 
+\item{\textit{Education and training}: %Some hack weeks are more focused on education than others (see Figure 1). 
+Even those hackathons not explicitly focused on skills-development will include extensive discussions on reproducible and open science practices that are valuable to participants. 
+%Discussing these topics in a diverse group setting helps to ingrain within participants the basic conventions of the broader open source community. 
 Participants can then serve as ''ambassadors`` of these practices and foster their adoption by other researchers in the field. 
-Furthermore, the kind of co-working that is inherent to the hack week format facilitates lateral knolwedge transfer, where participants learn from each other about practices and tools of the trade. 
-Often these tools are not described in research products such as papers, or even in the software implementations.}
+Furthermore, the kind of co-working that is inherent to the hack week format facilitates lateral knowledge transfer, where participants learn from each other about practices and tools of the trade.}
+%Often these tools are not described in research products such as papers, or even in the software implementations.}
 \item{\textit{Tool development}: hack weeks present an opportunity for computational tool developers to get together with users of these tools, and to critically evaluate the use of the tools in addressing particular scientific questions.}
+\item{\textit{Community building}: a hack week is an opportunity to galvanize a community within a field.% that cares about computational issues, and relates to specific open-source software projects.
+For example, because of its focus on free and open-source tools, Neurohack week helps galvanize the Neuroimaging in Python (http://nipy.org) community. 
+It also allows researchers who may otherwise be isolated within their departments and subfield on account of their computational specialization to come together and exchange experiences.}
+%As one participant at Astro Hack Week expressed in the survey, \textit{''The main thing I took from Astrohack week is the fact that I'm not alone``}.}
+\item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g.\ between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines.
+Despite the impact of interdisciplinary research \citep{Hall2012-hi}, these experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.}
+%This is despite the fact that interdisciplinary research can also be very impactful \citep{Hall2012-hi}. 
+%A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}
 
-\item{\textit{Community building}: a hack week is an opportunity to galvanize a community within a field that cares about computational issues, and relates to specific open-source software projects. 
-For example, because of its focus on free and open-source tools, Neurohack week helps galvanize the Neuroimaging in Python (http://nipy.org) community. 
-It also allows researchers who may otherwise be isolated within their departments and subfield on account of their computational specialization to come together and exchange experiences. 
-As one participant at Astro Hack Week expressed in the survey, \textit{''The main thing I took from AstroHackWeek is the fact that I'm not alone``}.}
-
-\item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g., between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines.
-These experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.
-This is despite the fact that interdisciplinary research can also be very impactful \citep{Hall2012-hi}. 
-A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}
-
-\item{\textit{Recruitment and networking}: hackweeks often bring together research, government and industry communities, providing an opportunity for participants to learn about each other's interests and their abilities in close quarters.
-This kind of extended opportunity for interaction and ''on-the-job`` testing is considered an excellent form of evaluation of prospective job candidates [citation needed]. 
-Conversely, participants have an opportunity to gain exposure to potential career opportunities in communities other than their own. 
-As an example, one of the participants in Neurohack week 2016 will be joining the eScience Institute as a post-doc in Fall of 2017 (and as an instructor in Neurohack week 2017).}
-
-\item{\textit{It's fun}: the opportunity to focus for a week on a slightly more whimsical and adventurous sort of scientific activities can be both fruitful, and enjoyable. 
-Even aside from the prospect of interdisciplinary research, it can provide a laboratory for more experimental approaches and high-risk projects than scientist will take on in their day-to-day research.}
-
+\item{\textit{Recruitment and networking}: hack weeks often bring together research, government and industry communities, providing an opportunity for participants to learn about each other's interests and their abilities in close quarters.
+This kind of extended opportunity for interaction and ''on-the-job`` testing is considered an excellent form of evaluation of prospective job candidates [CITATION NEEDED]. 
+Conversely, participants have an opportunity to gain exposure to potential career opportunities in communities other than their own.}
+%As an example, one of the participants in Neurohack week 2016 will be joining the eScience Institute as a post-doc in Fall of 2017 (and as an instructor in Neurohack week 2017).}
+%\item{\textit{It's fun}: the opportunity to focus for a week on a slightly more whimsical and adventurous sort of scientific activities can be both fruitful, and enjoyable. 
+%Even aside from the prospect of interdisciplinary research, it can provide a laboratory for more experimental approaches and high-risk projects than scientist will take on in their day-to-day research.}
 \end{itemize}
@@
 Some researchers may come with particular projects and collaborators in mind, while others come with a focus on learning and with no explicit plan for the project-based work. 
-While it is very hard to design a workshop that is universally useful to all participants, in practice many researchers attend with the understanding that some components will be more congruent with their own goals than others, but also bring an open mind and a willingness to learn.
-In addition, the diversity of goals can effectively be a strength of the workshop, if the organizers can facilitate matching up participants with complementary goals.
+%While it is very hard to design a workshop that is universally useful to all participants, in practice many researchers attend with the understanding that some components will be more congruent with their own goals than others, but also bring an open mind and a willingness to learn.
+%In addition, the diversity of goals can effectively be a strength of the workshop, if the organizers can facilitate matching up participants with complementary goals.
 
@@
 Here, appropriate essentially refers to 'most likely to achieve the stated goal of the hack week', though measuring that achievement remains a difficult issue itself.
-This is especially important because hack weeks, particularly in their more project-heavy iterations, are very participant-driven workshops. 
-At a traditional conference or summer school, the program will be largely set in advance by the organizing committee using the careful selection of speakers and lecturers. 
-At a hack week much of the program is in flux both before and during the workshop. 
+This is especially important because hack weeks are very participant-driven workshops. 
+%At a traditional conference or summer school, the program will be largely set in advance by the organizing committee using the careful selection of speakers and lecturers. 
+%At a hack week much of the program is in flux both before and during the workshop. 
 In this context, a successful hack week requires participants who are both interested in participation and feel comfortable participating in discussions and taking risks. 
@@
 At a summer school, the aim is generally for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows.
-Hack weeks differ in that knowledge transfer occurs across many levels of senority, due to the relative newness of the content and the distributed knowledge base. 
-Because the potential for knowledge transfer and community building is highest when the group is heterogeneous, participant diverse in both academic background and geographic location should be selected.
-This increases the chances that participants will take their newfound knowledge into their respective communities and spread it further.
+Hack weeks differ in that knowledge transfer occurs across many levels of seniority, due to the relative newness of the content and the distributed knowledge base. 
+%Because the potential for knowledge transfer and community building is highest when the group is heterogeneous, participant diverse in both academic background and geographic location should be selected.
+%This increases the chances that participants will take their newfound knowledge into their respective communities and spread it further.
 
@@
 In particular, education-focused workshops might be more homogeneous than project-focused workshops, and similarly single-discipline hack weeks may be more homogeneous than interdisciplinary ones. 
-Additional goals may be not directly related to the workshop itself, but on higher-order goals that a community might find desirable and that can be directly addressed by engineering the participant selection.
- For example, women are still a small minority in computational astrophysics. 
- One goal of a hack week could be to increase the participation of gender minorities in computation-focused research. 
- Similarly, if a particular sub-field within a discipline is ahead of others in their computational approaches, it might be worthwhile to engineer the mix of participants such that a number of experts from that sub-field are present.
- One interesting axis to consider is the mix of academics versus participants from industry. 
- The latter may often be more focused on practically useful applications of methods and data sets, and can therefore provide an important counterpoint to more theory-focused academics. 
- In particular in fields where the career path from academia to industry is less immediately obvious (as this the case e.g.
+Additional goals may be not directly related to the workshop itself, but on higher-order goals (e.g.\ increasing minority participation, knowledge transfer between academia and industry) that a community might find desirable and that can be directly addressed by engineering the participant selection.
+% For example, women are still a small minority in computational astrophysics. 
+%One goal of a hack week could be to increase the participation of gender minorities in computation-focused research. 
+%Similarly, if a particular sub-field within a discipline is ahead of others in their computational approaches, it might be worthwhile to engineer the mix of participants such that a number of experts from that sub-field are present.
+%One interesting axis to consider is the mix of academics versus participants from industry. 
+%The latter may often be more focused on practically useful applications of methods and data sets, and can therefore provide an important counterpoint to more theory-focused academics. 
+%In particular in fields where the career path from academia to industry is less immediately obvious (as this the case e.g.
 in astronomy, where little directly relevant industry exists), inviting back participants who transitioned to industry successfully can provide important role models and networking opportunities for junior participants.
- Finally, with much of data science currently performed in the private sector, industry participants can bring directly relevant knowledge and new scientific approaches to the table.
+%Finally, with much of data science currently performed in the private sector, industry participants can bring directly relevant knowledge and new scientific approaches to the table.
 
 The need to carefully select participants dovetails with another requirement: transparency. 
-Up to date, all hack weeks we have organized have been oversubscribed by at least a factor of two, and thus necessarily require turning applicants away. 
+%Up to date, all hack weeks we have organized have been oversubscribed by at least a factor of two, and thus necessarily require turning applicants away. 
 The selection procedure should be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance. 
@@
 In the first step, the merit of a candidate must be assessed: does the candidate fulfill the requirements posed by the goals of the workshop?
-For example, at Astro Hack Week, it may be unrealistic to admit an undergraduate student in biochemistry, unless they had a compelling reason for attending.
-This first step is exceedingly difficult to automate, because it requires judgment calls, often based on long-form answers, that are as of yet difficult for computers to parse and classify. 
+%For example, at Astro Hack Week, it may be unrealistic to admit an undergraduate student in biochemistry, unless they had a compelling reason for attending.
+This first step is difficult to automate, because it requires judgment calls, often based on long-form answers. 
 In the past, we have performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
@@
 It is here where one may impose outside constraints on the selection based on the goals of the workshop. 
-If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which provably optimal algorithms exist that will far outperform any human selection procedure.
+If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
 
@@
 
-Within our hack weeks the goal of the selection procedure has always been diversity among several axes. 
-This includes academic subdiscipline, to allow for maximally possible knowledge exchange between different domain fields, as well as with researchers from adjacent fields including statistics, computer science and data science.
-Other important categories revolve around the participants' previous knowledge in terms of machine learning, statistics and programming. 
-Unlike for a summer school, we have generally selected for a diverse set of participants in order to be able to efficiently pair up novices and experts in projects. 
-Finally, we also include demographic information with two main goals: firstly, representation has been proven to be an effective way to foster an inclusive environment where minority participants feel comfortable to participate and take risks.
-Secondly, we would like to boost the participation of minority researchers in computational astrophysics, and our hack week model provides a useful framework for doing so.
+%Within our hack weeks the goal of the selection procedure has always been diversity among several axes. 
+%This includes academic subdiscipline, to allow for maximally possible knowledge exchange between different domain fields, as well as with researchers from adjacent fields including statistics, computer science and data science.
+%Other important categories revolve around the participants' previous knowledge in terms of machine learning, statistics and programming. 
+%Unlike for a summer school, we have generally selected for a diverse set of participants in order to be able to efficiently pair up novices and experts in projects. 
+%Finally, we also include demographic information with two main goals: firstly, representation has been proven to be an effective way to foster an inclusive environment where minority participants feel comfortable to participate and take risks.
+%Secondly, we would like to boost the participation of minority researchers in computational astrophysics, and our hack week model provides a useful framework for doing so.
 
@@
 This has several clear advantages. 
-In particular, communities organized by subject generally share the same language, which drastically cuts down on the time participants spend simply agreeing on scientific terms. 
-Similarly, the scientific objectives and questions are broadly clear to all participants: there is little need to explain in detail why a particular question may be of interest to the field. 
+In particular, communities organized by subject generally share the same language.%, which drastically cuts down on the time participants spend simply agreeing on scientific terms. 
+Similarly, the scientific objectives and questions are broadly clear to all participants.%: there is little need to explain in detail why a particular question may be of interest to the field. 
 Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data.
@@
 These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead. 
-In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
+In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task.%, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
 On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.
@@
 These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
-Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others. 
-In addition, participants included computer vision researchers, who develop algorithms that can be applied to answer scientific questions. 
-Similarly, a series of events at UC Berkeley (Text XD) focused on the analysis of text data across different domains of research.
+%Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others.
+%In addition, participants included computer vision researchers, who develop algorithms that can be applied to answer scientific questions. 
+%Similarly, a series of events at UC Berkeley (Text XD) focused on the analysis of text data across different domains of research.
 
@@
 On the other hand, because of their large participatory component, hack weeks tend to be exhausting, and long events may lead to fatigue among attendees, where active participation and measurable outcomes may drop sharply in the latter days of the workshops.
-This is particularly important if---as is often the case---participants continue participating in workshop-relevant activities even after the official component ends for the day.
-
+%This is particularly important if---as is often the case---participants continue participating in workshop-relevant activities even after the official component ends for the day.
 Ways to mitigate these effects include capping taught components during the day, providing a reasonable clear schedule, and limiting parallel components to avoid decision fatigue. 
-At Astro Hack Week, we have limited the scheduled taught components to no more than three out of nine hours for each of the five workshop days, and unscheduled (shorter) tutorials to a maximum of two per day.
-Even so, in our surveys, participants occasionally report concerns about choosing to attend tutorials versus working on projects.
+%At Astro Hack Week, we have limited the scheduled taught components to no more than three out of nine hours for each of the five workshop days, and unscheduled (shorter) tutorials to a maximum of two per day.
+%Even so, in our surveys, participants occasionally report concerns about choosing to attend tutorials versus working on projects.
 
@@
 As a general rule, the smaller a hack week's emphasis on a taught component, the more flexible the space has to be, with ample opportunity for re-configuration. 
-At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine. 
-It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not.
+%At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine. 
+%It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not.
 
 Another important design consideration is group size. 
-With all previous hack weeks severely oversubscribed, it seems natural to simply admit more participants. 
-However, this can counter the ideals of the workshop, in particular when building a community is one of its stated goals. 
+%With all previous hack weeks severely oversubscribed, it seems natural to simply admit more participants. 
+%However, this can counter the ideals of the workshop, in particular when building a community is one of its stated goals. 
 If the group is too large, participants are unlikely to even meet each other, and workshop cohesion may be lost as the workshop fractures into smaller groups, often among participants who already know each other.
 This may inhibit knowledge transfer by clustering participants into small in-groups. 
-Additionally, it is likely that the number participant-led components in the schedule may increase with workshop size. 
-While generally desirable, a programme that is too crowded may lead to fatigue On the other hand, if the size of the workshop is too small, it is unlikely to achieve the desired diversity among participants to foster new collaborations across sub-fields and disciplines.
+%Additionally, it is likely that the number participant-led components in the schedule may increase with workshop size. 
+%While generally desirable, a programme that is too crowded may lead to fatigue 
+On the other hand, if the size of the workshop is too small, it is unlikely to achieve the desired diversity among participants to foster new collaborations across sub-fields and disciplines. 
 In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.
@@
 In communities where a shared understanding exists, tutorials can focus on more advanced or innovative topics, and less time may be allocated for them, leaving more time for active participation. 
-In astronomy, a relatively small field, even students generally share a common knowledge base and have rudimentary knowledge about the types of data used and the challenges associated with each. 
-Thus, we have focused the taught component of Astro Hack Week less on domain-specific knowledge, and instead offered tutorials in topics from domain-adjacent fields like statistics and computer science that attendees are unlikely to have encountered in their regular education.
+%In astronomy, a relatively small field, even students generally share a common knowledge base and have rudimentary knowledge about the types of data used and the challenges associated with each. 
+%Thus, we have focused the taught component of Astro Hack Week less on domain-specific knowledge, and instead offered tutorials in topics from domain-adjacent fields like statistics and computer science that attendees are unlikely to have encountered in their regular education.
 
 Hack week outcomes, in turn, depend strongly on participants and are often a function of their interests and seniority. 
-A number of attendees arrive with the stated goal of writing a specific scientific article, though these tend to be more advanced participants with significant pre-knowledge of both hackathons and their topic of interest.
-Conversely, many attendees arrive with the stated goal of learning about a specific topic (such as machine learning) and bring a specific data set they believe the new knowledge may be applicable to. 
+some attendees arrive with the stated goal of writing a specific scientific article, usually more advanced participants with significant pre-knowledge of both hackathons and their topic of interest. 
+Many attendees arrive with the plan to learn a specific topic (such as machine learning) or bring a specific data set they believe the new knowledge may be applicable to. 
 This leads to a wide variety of project types from sandbox-style explorations to focused work efforts. 
-It is worth noting that while a scientific paper need not be the stated goal of a hack (and is unlikely to be completed in the short time allocated in any case), results may still be published as short reports or unconference proceedings.
-For example, Neurohack week provides a venue for participants to publish a short (two-page) ''project report`` summarizing the hack that participants did during the week of NHW. 
-Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.
+%It is worth noting that while a scientific paper need not be the stated goal of a hack (and is unlikely to be completed in the short time allocated in any case), results may still be published as short reports or unconference proceedings.
+%For example, Neurohack week provides a venue for participants to publish a short (two-page) ''project report`` summarizing the hack that participants did during the week of NHW. 
+%Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.
 
@@
 Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \citep{Clance1978-ef}.
-Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohackweek.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
+Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
 This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
@@
 With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS. 
-More generally, being open about the presence and prevalence of IS can help participants feel more at ease (e.g.
+More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g.
 one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
-I think it helped ease the feeling!!``}). 
+%I think it helped ease the feeling!!``}). 
 Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
@@
 Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success. 
-Additionally, it is difficult to define in the first place what success means to organizers and success. 
 For more project-driven events, one key metric is the number of publications that result from hack week projects. 
@@
 Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes. 
-In particular, it ignores learning, community building as well as networking outcomes, which are often equally important objectives. 
+These measures ignore learning, community building as well as networking outcomes, which are often equally important objectives. 
 One possible approach here is to ask participants to self-report their success via a post-workshop survey.
-If participants are selected as described above, it is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
-This would then provide a somewhat objective measure of the impact the hack week has had. 
-In practice, this has not been done for any of the hack weeks conducted thus far.
-
-For practical purposes, we recommend an approach that combines all of these metrics: publications directly resulting out of hack week collaborations, new code generated and new contributors to open-source projects, self-reported learning outcomes in the topics covered in the tutorials, and survey results probing the usefulness of the workshop in terms of career and networking goals.
-Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
+It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not. 
+%This would then provide a somewhat objective measure of the impact the hack week has had. 
+%In practice, this has not been done for any of the hack weeks conducted thus far.
+For practical purposes, we recommend an approach that combines all of these metrics: publications, new code and projects generated, self-reported learning outcomes, and survey results. 
+%Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
 
@@
 In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating. 
-It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
+%It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
 
@@
 We asked participants of all three hack weeks to complete surveys to validate the hack week's success. 
-All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.
+%All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.
 
@@
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (see Figure 2, middle and right panels).
-While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
-Another important goal of the hack weeks was centred on community building and fostering collaborations. 
+%While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
+%Another important goal of the hack weeks was centred on community building and fostering collaborations. 
 We find that overall, the majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel).
@@
 We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories.
-This is expected: in particular participants new to the field may participate to learn rather than to teach.
+This is expected: participants new to the field may participate to learn rather than to teach.
 
@@
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields. 
-In particular, astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences. 
-Neuroscience seems to fall somewhere in between. 
-This also implies that in this respect, hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
-This is also reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%).
+In particular, astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
+%Neuroscience seems to fall somewhere in between. 
+This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
+Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%).
 Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible. 
@@
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.
-Neurohackweek is supported through a grant from the National Institute for Mental Health (1R25MH112480).
+Neurohack week is supported through a grant from the National Institute for Mental Health (1R25MH112480).
 Daniela Huppenkothen is partially supported by the James Arthur Postdoctoral Fellowship at NYU.}

### 2017-08-17 supplementary_material.tex

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%

\documentclass{nature}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{natbib}

\usepackage{hyperref}
\usepackage{url}
\usepackage{microtype}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{tabularx}

%\title{How and Why to run a Hack Week}

%% Notice placement of commas and superscripts and use of &
%% in the author list

%\author{Jake VanderPlas, Karthik Ram, Daniela Huppenkothen, David W. Hogg, Ariel Rokem, Anthony Arendt \& Tal Yarkoni}

\begin{document}

%\maketitle

%\begin{affiliations}
% \item UW Seattle
% \item NYU
% \item UC Berkeley
%\end{affiliations}

\section{Supplementary Material}

While the main text of this paper remains fairly theoretical, we aim to lay out more practical advice regarding the organization of a hack week in the appendix of this document.

\subsection{What to do before}

\subsubsection{Obtaining permission for tracking and studying participants}
A goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes. To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities. These scientists then reviewed their observations and and gave valuable feedback about the hack week outcomes. In order to be able to track and study participants, approval from Internal Review Boards is required for workshops organized in the United States. This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally don't involve human subject research. Ideally, participants should consent to be contacted at the application stage, which opens up the potential of tracking both participants and non-participants and allows for evaluation of the hack week.

\subsubsection{Goal Planning}
Unlike traditional conferences, where the programme is primarily driven by speaker selection, careful goal planning is essential for a successful hack week. The stated goals will continue to determine almost all of the remaining organizational questions, including space, participant selection and workshop evaluation. Is it aimed more at students or mature researchers, or a mix of both? Should the taught component dominate over project work or should it be only a minor programme point? Is the community to be invited more software-focused or should outcomes skew towards publishable results? Does the workshop aim to maximize its output by carrying best practices into universities and institutions that may be particularly underrepresented? These questions, among others, should have a very clear answer before any further workshop organization takes place.

\subsubsection{Hack Week Location}
When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration. The non-traditional nature of hack weeks means that some students report challenges in receiving their superior's approval to attend. Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be the highest priority when deciding on a location. Conference centers like the Lorentz Center at the University of Leiden provide funding along with the collaborative space, but have long lead times when applying for their spaces (typically 12-18 months). The same can be true for collaborative spaces within universities and active learning environments, which tend to be popular. Therefore, the search for a location should commence no later than twelve months before the workshop, ideally much earlier.
The space itself should be large and ideally configurable, with moveable tables and chairs, several screens or projectors, and an ample number of white-/chalkboards. A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.

\subsubsection{Funding}

We have worked to minimize the cost of registration for our events in order to enable broad participation, especially for participants from institutions, countries and demographics that are generally underrepresented at other meetings. Locations that offer some funding or provide free space are preferable to those that do not.
At Astro Hack Week, we have found that travel grants are of crucial importance especially to undergraduates, students from minority backgrounds and participants from underrepresented countries. During the organization of Python in Astronomy 2017, we found that minority participants disproportionally decline to attend, and often cite lack of sufficient travel funding as the primary reason. In particular if increasing diversity in the more technical aspects of a field is one of the workshop's goals, providing enough travel funding for between five to ten participants can make a significant difference in allowing minority researchers to attend.

\subsubsection{Code of Conduct}

A hack week should have a code of conduct. Organizers should be aware that disparities in backgrounds, knowledge and experience, together with the close collaboration much of the hacking requires, creates the potential for conduct that others might find offensive, especially when it comes to language. The goal of every hack week should be to provide a comfortable, inclusive environment for participants to learn and work. It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in advance.
A part of this could and should be the recognition that participants (and, indeed, members of the organizing committee) may occasionally violate the code of conduct in minor ways, and allow for tools and mechanisms to resolve these infractions. Major infractions, conversely, should be taken seriously and in severe cases lead to exclusion of participants.
Aside from important provisions banning discrimination on the basis of demographic characteristics, there may be domain- or workshop-specific clauses that may require recognition. For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome, defensive and inhibit learning on a larger scale.

\subsubsection{Communicating with the Outside World}

Because hack weeks have enjoyed considerable popularity, they have often been oversubscribed. This leaves organizers with an important decision: given the necessity of turning a large fraction of applicants away, what methods does the group wish to employ in order to communicate during the hack week with the wider community not in attendance? Tutorials may be live-streamed via popular services such as youtube, but aside from the technical complexity of setting up a reliable live stream, their interactive format may not easily lend themselves to the one-directional communication of a video feed. Even so, live streams have proven to be both useful and popular in the past, but require considerable effort to set up. At the very least, there should be some method of real-time communication with the in-group. Twitter has proven itself to be an effective medium of general communication with the wider community, including hack pitches, questions during tutorials and collaboration with researchers outside of Astro Hack Week. However, the closed nature of Twitter to those with an account may not make it the ideal platform for participation. Other open-source formats that do not require sign-up, such as Matrix, may be more compatible with the hack week format\footnote{\url{https://dave.cheney.net/2017/04/11/why-slack-is-inappropriate-for-open-source-communications}}. Important documents such as programme (and especially programme changes) as well as hack descriptions should be open to the public and communicated clearly to facilitate collaborations with outside researchers.

\subsection{During a Hack Week}

Unlike traditional conference, where the programme is generally set by pre-selected talks, hack weeks require additional involvement from the organization committee during the hack week itself.

\subsubsection{Tutorials}
Tutorials provide a mechanism for rapid knowledge sharing and are often one of the most useful components of a hack week. Unlike traditional lectures, in which more time is available to explore numerous topics in depth, tutorials aim to deliver only the most essential information while ensuring students remain engaged and interactive. Moreover, the need for an instructor to balance prepared material with spontaneous, interactive learning, while attending to community building and inclusive practices, can be difficult for even the most seasoned educator. Therefore careful planning and preparation is necessary for effective delivery of a hack week tutorial.

We have experimented with numerous designs in delivering our hack week tutorials in a way that is student-centered and focused on meeting each student at their particular skill level. We have observed that, especially during long tutorials and with a particularly diverse group of students, there is a natural separation between participants who are focused on gaining new knowledge as it is being presented, versus those who already possess this knowledge and hence turn their attention to other subjects as the tutorial is ongoing. In practice, this concern can be mitigated by explicitly involving experienced participants in the tutorials. We encourage instructors to call upon experts in the audience to identify themselves and effectively act as teaching assistants. Expert participants often find this gratifying because it allows them to test their own understand and improve their skills by learning through teaching. At the same time, beginners benefit by having increased interactions with other hack week participants, thereby increasing team cohesion and building community. Another technique to engage expert participants during the tutorial is to invite them to work through more challenging content located in the online lesson, which will not be covered in the session, but for which they can obtain help during the tutorial.

Some of our hack week tutorials follow the structure of the Software Carpentry model. The content is divided in to a series of lessons, each having a well-defined set of learning objectives, key questions, and expected outcomes. Interspersed throughout each lesson are a series of exercises which participants work through at their own pace. We frequently use a method drawn from Software Carpentry whereby students indicate their need for help on an exercise via posting a colored sticky-note on their computer. This provides a simple method to deliver help to those students who need it without interrupting the flow of the tutorial. We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks or following the markdown templates provided by Software Carpentry. This way the hack week "curriculum" can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics. Depending on the content, some instructors convey concepts during the tutorial via "live-coding" while others display static code and their results. Either way, we find divding code into short blocks, each with a specific task, helps participants from becoming saturated with informaiton.

While the short duration of the tutorial format helps us minimize the problem of information overload, it can be challenging to decide on the scope and breadth of material to be covered. In-depth collaboration among instructors in advance of the event is particularly valuable in this regard, to assess which methods or subjects will be most useful for the specific audience. We suggest that the primary goals of the tutorials should be to provide an entry point into an exploration of participants' datasets, opening the door to more thorough study outside of the hack week. Given the broad span of available tools and topics, the organizing committee should seek to teach content that represents the state-of-the-art and is deemed to be of greatest use to the broadest span of participants. A common theme across all of our hack weeks is to teach initial tutorials in version control, command line interfaces, data science platforms and practices of reproducible research. These tutorials endeavor to provide a common baseline of techniques that participants can build on in other tutorials and project work. Subsequent tutorials are then delivered in more domain-specific fields, but can be arranged in a way that builds knowledge constructively through the week. In some cases, it may be advisable to recommend some theoretical texts as reading before attending the hack week, though it is likely that speakers should not count on every participant arriving with the same knowledge baseline. Because of their practical nature, however, software requirements should be announced as early and clearly as possible, with the expectation that participants will have installed necessary software in advance. It may be helpful to designate some time early on in the meeting to trouble-shoot installations before tutorials begin in order to avoid loosing significant time during the actual tutorials.

It is noteworthy that in practice, the mixed audience and interactive setting can often lead to impostor syndrome among the speakers, who lose their explicit status as expert among the participants. It is therefore imperative that speakers are made aware very early about the audience so that they can plan their tutorial accordingly. It also helps to make them aware that participants may be taking the role of teaching assistants during exercises and include them in the teaching, rather than see them as adversaries who may seek to correct them. Conversely, experts may be used to ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and encourage questions especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome. At Astro Hack Week, we encourage the organizing committee and experienced participants to ask questions, particularly when they know the answer but think the concept may not yet be clear to parts of the audience. This has in the past led to exchanges between experts that have greatly contributed to the tutorials in ways that were both unplanned and unexpected for all involved, beginners and experts alike. Including experts in the teaching also takes pressure off teachers during exercises, when demands on the speaker are generally high.

To summarize, a good tutorial will:
\begin{itemize}
\item be very clearly tailored to the audience and narrowly scoped,
\item strictly limit the amount of lecture-style teaching to less than 50\% ,
\item use experts in the audience to ask key questions and act as teaching assistants during exercises
\item communicate technical requirements at least a week before the hack week
\end{itemize}

\subsubsection{Break-Out Sessions}

Because instructors and organizing committees are unlikely to know in advance what projects and data sets participants will bring, often there will be topics and methods not covered in the pre-planned tutorials that are are of interest to the audience. Here, break-out sessions offer an alternative: short (30-45 minute) tutorials that are fairly spontaneously organized (with as little a lead time of a few hours) and often taught by expert participants in the audience. These tutorials can be a more in-depth treatment of one specific method of interest (e.g. Gaussian Processes, K-Means Clustering, Deep Learning), or cover a practical skill that may be useful to the audience, but is not formally part of the hack week (this is especially the case for skills related to software-development, such as code testing, documentation, and profiling). They are generally more informal than tutorials and can be taught to a subset of the group rather than requiring all participants to attend.
Break-out sessions should be limited to one or two a day, in order to allow participants to attend break-out sessions without having to sacrifice a significant fraction of their time reserved for project work. It is also possible to intentionally leave a tutorial slot free, to be filled with one or two self-organized break-out sessions instead of a pre-planned tutorial. In practical terms, it is advisable to keep a physical board with requests and potential teachers in a prominent location during the week. Decisions on which option to choose from the list can be done via an informal voting process, under the condition that a volunteer to hold the tutorial is found. Giving a break-out session can be a daunting task: while some experts may have relevant tutorials already prepared from other workshops, often it requires holding a thirty-minute talk with little to no preparation. Organizers can and should take steps to provide a positive, encouraging environment for participants to volunteer their knowledge. They should particularly encourage junior participants, who may often be the most knowledgable about the topics usually requested, to volunteer for break-out sessions. They provide a valuable teaching experience in a friendly environment and offer the opportunity of networking with the larger community.

To summarize, break-out sessions:
\begin{itemize}
\item consist of short tutorials about a specific topic or practical skill not covered in the tutorials
\item are more informal and may be taught to a subset of interested participants
\item should be limited to one to two per day and no longer than 45 minutes to avoid conflicts with project-work
\item can be daunting to teach, especially for junior participants, and requires the organizing committee to be pro-active about encouraging volunteer teachers
\end{itemize}

\subsubsection{Hack Sessions}

For hack weeks that place less emphasis on teaching, hack sessions are at the core of the workshop. They provide the opportunity for participants to actively collaborate and in many cases use the knowledge and skills learned in the tutorials on their own data sets. Even though hack sessions are strongly driven by the participants themselves, they require careful planning and vigilant supervision by the organizing committee. In general, hack sessions follow a standard pattern. In the initial stage, participants pitch projects to the group and in some cases request specific expertise (of a certain method, programming language or other skill). Ideally, the organizing committee will encourage participants to post potential projects as early as possible online in a central document, which will allow some organization even before the begin of the hack week. If the hack week admits very junior participants (e.g. undergraduate students or junior graduate students) it may be advisable for the organizing committee to contact experienced participants in advance and ask them to suggest hacks that are appropriate for those junior participants.
At the hack week itself, the time designated for project work will generally be prefaced by a pitch session, where participants with projects give a short description of their project to the group and request help. For junior participants and those new to the format, this can be daunting, even if (or especially when) they do not have a project themselves. In particular junior participants tend to believe that they do not yet have the skills to contribute meaningfully to any projects. Here, it can be useful to require those participants who are not pitching projects to state their expertise and affirm that every participant has expertise in something. This broad definition of hacking can help the organizers to facilitate inclusion by emphasizing the value in skills such as writing (e.g. for tutorials and documentation), testing code to be released, or other tasks. After pitching, participants self-organize into teams to work on specific projects. Here, organizers should remain aware of participants who may not have found a group, or are having trouble finding a task to work on. This remains true during the week, as some projects are abandoned and others appear. At least one member of the organizing committee should be familiar with the current roster of projects and the experience in the room to help match participants as needed. Tracking ongoing projects can be accomplished online, via real-time collaborative documents, as well as in the physical space, via a diagram of where in the space teams are located.
Because hack weeks are generally longer than typical hackathons (often limited to just one or two days), this will lead to a variety of shorter projects completed within a day or two, and longer projects going on throughout the week. Regular check-in sessions are helpful for pitching new projects, requesting help on projects stuck at a certain point, and re-matching participants after hacks have been completed or abandoned. These sessions can also be used for completed hacks to be showcased, or intermediate achievements to be presented. At the end of the week we provide ample time for participants to present their projects to the community. We endeavor to design these presentations in a way that builds confidence, especially for junior researchers, and that welcomes all projects regardless of their level of sophistication. The open-ended format of a hack week encourages experimental approaches and new directions, which necessarily will not all be successful. Too often, showcases focus on the most successful and impressive projects, which distorts reality and produces both an unreasonable expectation of what the result of a hack week ought to be as well as a pressure to be successful that can in practice inhibit participants. Therefore, it should be encouraged that participants show failed ideas, not least because the knowledge of why a project failed can be a valuable learning experience beyond the team working on that particular project.

\subsection{What to do after}

Feedback is a crucial component of a hack week. Because of its experimental design, and the differences in group from year to year, it is important to routinely check whether the workshop still matches up with what participants expect and find useful. Additionally, computational fields tend to be fast-paced; a tutorial given one year might be outdated a few years later. It is thus useful to put some detailed thoughts into the survey. Following from discussions the committee has likely had during the initial stage of preparing the announcement and selecting participants, organizers will want to confront their own expectations before the workshop, and design a survey that allows for both measurements of specific outcomes and goals the organizers are interested in, as well as leave enough free space for participants to self-report outcomes that arose spontaneously or were out of scope of the initial design. It is helpful to design the survey early on and allow for some time near the end of the workshop for participants to fill it out there; in our experience, this leads to much higher response rates than if the survey is distributed after the end of the workshop. While some improvements and changes in the survey design from year to year are inevitable, it is advisable to set a number of core questions that do *not* change from year to year. These questions will allow organizers to assess the workshop's function within their community, and assess how changes to the format between one year and another may affect outcomes of interest.

For tracking tangible outcomes, a central real-time document used during the workshop to keep track of projects can function as a record of productivity. This document might be rewritten as conference proceedings. Additionally, organizers should encourage participants to identify written outcomes with the hack week: for publications, acknowledgments provide a convenient venue. Code repositories can allow this kind of acknowledgment in a \textit{readme} document; GitHub also allows attaching tags to a repository, thus all code repositories that experienced major contributions during the workshop may include a tag specific to the hack week.

On a larger scale, it might be useful to track measures of success more long-term. Those could for example be publication networks for scientists who attended the workshop, their publications as well as their career development. For this purpose, it is useful to keep a way to contact previous participants, via a mailing list or real-time messaging services like \textit{Matrix} or \textit{Slack}.

%Spelling must be British English (Oxford English Dictionary)

%In addition, a cover letter needs to be written with the
%following:
%\begin{enumerate}
% \item A 100 word or less summary indicating on scientific grounds
%why the paper should be considered for a wide-ranging journal like
%\textsl{Nature} instead of a more narrowly focussed journal.
% \item A 100 word or less summary aimed at a non-scientific audience,
%written at the level of a national newspaper.  It may be used for
%\textsl{Nature}'s press release or other general publicity.
% \item The cover letter should state clearly what is included as the
%submission, including number of figures, supporting manuscripts
%and any Supplementary Information (specifying number of items and
%format).
% \item The cover letter should also state the number of
%words of text in the paper; the number of figures and parts of
%figures (for example, 4 figures, comprising 16 separate panels in
%total); a rough estimate of the desired final size of figures in
%terms of number of pages; and a full current postal address,
%telephone and fax numbers, and current e-mail address.
%\end{enumerate}

%See \textsl{Nature}'s website
%(\texttt{http://www.nature.com/nature/submit/gta/index.html}) for
%complete submission guidelines.

%\begin{methods}
%Put methods in here.  If you are going to subsection it, use
%\verb|\subsection| commands.  Methods section should be less than
%800 words and if it is less than 200 words, it can be incorporated
%into the main text.

%\end{methods}

%% Put the bibliography here, most people will use BiBTeX in
%% which case the environment below should be replaced with
%% the \bibliography{} command.

% \begin{thebibliography}{1}
% \bibitem{dummy} Articles are restricted to 50 references, Letters
% to 30.
% \bibitem{dummyb} No compound references -- only one source per
% reference.
% \end{thebibliography}

%%
%% TABLES
%%
%% If there are any tables, put them here.
%%

%\begin{table}
%\centering
%\caption{This is a table with scientific results.}
%\medskip
%\begin{tabular}{ccccc}
%\hline
%1 & 2 & 3 & 4 & 5\\
%\hline
%aaa & bbb & ccc & ddd & eee\\
%aaaa & bbbb & cccc & dddd & eeee\\
%aaaaa & bbbbb & ccccc & ddddd & eeeee\\
%aaaaaa & bbbbbb & cccccc & dddddd & eeeeee\\
%1.000 & 2.000 & 3.000 & 4.000 & 5.000\\
%\hline
%\end{tabular}
%\end{table}

\end{document}

### 2017-08-17 paper.tex
Divide into sections, for easier work moving forward.

@@
 
-\begin{abstract}
-Across almost all scientific disciplines, the instruments that record our experimental data and the methods required for storage and data analysis are rapidly increasing in complexity. 
-This gives rise to the need for scientific communities to adapt on shorter time scales than traditional university curricula allow for, and therefore requires new modes of knowledge transfer. 
-At the same time, as disciplines become more specialized, data analysis problems often tend to be fairly universal. 
-It is thus desirable to foster exchange of ideas and computational workflows across disciplines. 
-Among the different approaches put forward toward addressing these issues, hack weeks have in recent years emerged as both an effective tool for training researchers in modern data analysis workflows, and in fostering exchange between subdisciplines within and between scientific fields.
-While interpreted differently in the various disciplines where they have been implemented, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and free-form project work.
-In this paper, we present the concept of a hack week in the larger context of scientific meetings and point out similarities and differences to traditional conferences. 
-We motivate the need for such an event and present in detail its strengths and challenges. 
-We find that hack weeks are successful at cultivating collaboration and the exchange of knowledge. 
-Participants self-report that these events help them both in their day-to-day research as well as their careers. 
-Based on our results, we conclude that hack weeks present an effective, easy-to-implement, fairly low-cost tool to positively impact data analysis literacy in academic disciplines, foster collaboration and cultivate best practices.
-\end{abstract}
-
-\section*{Introduction}
-\label{sec:introduction}
-
-As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
-At the same time, formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
-A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998\footnote{\url{https://f1000research.com/articles/3-62/v2}}.
-At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field. 
-Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, and more project-focused than pedagogical meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\footnote{\url{https://gigascience.biomedcentral.com/articles/10.1186/s13742-016-0121-x}}, and a hackathon associate with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-In broad-brush, pedagogically-focused workshops and summer schools tend to follow a classic academic model where novices learn a skillset from experts, while project-focused workshops focus on people collaboratively exercising the skillset they already have.
-A disadvantage of the summer school model is that it can tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
-A disadvantage of the hackathon model is the common perception (whether accurate or not) that the week is designed for experts in technical tools, which may discourage others from attending.
-In 2014, we started Astro Hack Week to try to fill the gaps between these models. 
-The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for collaborative hacks or creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=9cm]{fig/HackSpectrum.png}
-\caption{Comparison of Extracurricular Workshop Models}
-\label{fig:hackspectrum}
-\end{center}
-\end{figure}
-
-As of the publication of this paper, we have run five such hack week events: three focused on Astronomy, one focused on Neuroscience, and one focused on Geoscience.
-Below we will share some of the philosophy behind the hack week model, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.
-
-\section*{What is a hackathon?}
-
-Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \citep{Decker2015}. 
-Hackathons have historically focused on software development and technology design as a way to motivate innovation within industry. 
-In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries. 
-In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to 'learn by doing' within a constructivist educational framework \citep{Bransford2000-lu,Papert1980-fh}.
-With this in mind, hackathons around scientific topics, designed to foster collaboration \citep{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \citep{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.
-
-%The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon. 
-Following the typology of \citep{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
-Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \citep{Drouhard2017}.
-''Catalytic`` hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge. 
-''Contributive`` hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project. 
-Finally, ''Communal`` hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.
-
-Our past hack week events most closely follow the communal hackathon model as it applies to scientific communities of practice. 
-Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. 
-Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure 1). 
-For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials. 
-In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.
-
-We note that the terminology for these events is constantly evolving, and that the ''hackathon`` concept may have implicit connotations that are disfavored in some communities. 
-One criticism of hackathons is that they propel the ''geek`` stereotype and may present a barrier to creating an inclusive working environment, especially for women \citep{Decker2015}. 
-Another problem is the competitive atmosphere of many industry hackathons, where teams actively compete for prizes. 
-Because certain groups are more risk-averse than others, and risk-aversion traces demographics like gender, we urge caution when contemplating making an event competitive. 
-Other related nomenclature includes the ''unconference`` which is also a time-bounded event gathering together a specific community, but these often have minimal structure and do not necessarily have a focus on software.
-The ''sprint`` or ''scrum`` label typically refers to an event focused on rapid software development on a specific set of code. 
-%This may be a component of our hack week model but it ignores the important pedagogical components.
-
-\section*{Why run a Hack Week?}
-
-There are several reasons to run a hack week of the sort described here.
-
-\begin{itemize}
-\item{\textit{Education and training}: %Some hack weeks are more focused on education than others (see Figure 1). 
-Even those hackathons not explicitly focused on skills-development will include extensive discussions on reproducible and open science practices that are valuable to participants. 
-%Discussing these topics in a diverse group setting helps to ingrain within participants the basic conventions of the broader open source community. 
-Participants can then serve as ''ambassadors`` of these practices and foster their adoption by other researchers in the field. 
-Furthermore, the kind of co-working that is inherent to the hack week format facilitates lateral knowledge transfer, where participants learn from each other about practices and tools of the trade.}
-%Often these tools are not described in research products such as papers, or even in the software implementations.}
-\item{\textit{Tool development}: hack weeks present an opportunity for computational tool developers to get together with users of these tools, and to critically evaluate the use of the tools in addressing particular scientific questions.}
-\item{\textit{Community building}: a hack week is an opportunity to galvanize a community within a field.% that cares about computational issues, and relates to specific open-source software projects.
-For example, because of its focus on free and open-source tools, Neurohack week helps galvanize the Neuroimaging in Python (http://nipy.org) community. 
-It also allows researchers who may otherwise be isolated within their departments and subfield on account of their computational specialization to come together and exchange experiences.}
-%As one participant at Astro Hack Week expressed in the survey, \textit{''The main thing I took from Astrohack week is the fact that I'm not alone``}.}
-\item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g.\ between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines.
-Despite the impact of interdisciplinary research \citep{Hall2012-hi}, these experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.}
-%This is despite the fact that interdisciplinary research can also be very impactful \citep{Hall2012-hi}. 
-%A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}
-
-\item{\textit{Recruitment and networking}: hack weeks often bring together research, government and industry communities, providing an opportunity for participants to learn about each other's interests and their abilities in close quarters.
-This kind of extended opportunity for interaction and ''on-the-job`` testing is considered an excellent form of evaluation of prospective job candidates [CITATION NEEDED]. 
-Conversely, participants have an opportunity to gain exposure to potential career opportunities in communities other than their own.}
-%As an example, one of the participants in Neurohack week 2016 will be joining the eScience Institute as a post-doc in Fall of 2017 (and as an instructor in Neurohack week 2017).}
-%\item{\textit{It's fun}: the opportunity to focus for a week on a slightly more whimsical and adventurous sort of scientific activities can be both fruitful, and enjoyable. 
-%Even aside from the prospect of interdisciplinary research, it can provide a laboratory for more experimental approaches and high-risk projects than scientist will take on in their day-to-day research.}
-\end{itemize}
-
-It is worth noting that the reasons for participants to attend a hack week are as diverse as the goals that the organizing committee explicitly states. 
-This is largely a function of the diversity of the participants (see section below), and should be something that organizers design for. 
-Beginner participants may attend primarily to learn a new technique or method, while more experienced participants attend to gain more experience mentoring and teaching. 
-Some researchers may come with particular projects and collaborators in mind, while others come with a focus on learning and with no explicit plan for the project-based work. 
-%While it is very hard to design a workshop that is universally useful to all participants, in practice many researchers attend with the understanding that some components will be more congruent with their own goals than others, but also bring an open mind and a willingness to learn.
-%In addition, the diversity of goals can effectively be a strength of the workshop, if the organizers can facilitate matching up participants with complementary goals.
-
-\section*{Audience and Participant Selection}
-
-Because hack weeks are at their core non-traditional workshops, it is worth thinking in detail about the target audience for a hack week, along with methods to select an appropriate group of participants.
-Here, appropriate essentially refers to 'most likely to achieve the stated goal of the hack week', though measuring that achievement remains a difficult issue itself.
-This is especially important because hack weeks are very participant-driven workshops. 
-%At a traditional conference or summer school, the program will be largely set in advance by the organizing committee using the careful selection of speakers and lecturers. 
-%At a hack week much of the program is in flux both before and during the workshop. 
-In this context, a successful hack week requires participants who are both interested in participation and feel comfortable participating in discussions and taking risks. 
-Therefore, a robust, effective participant selection procedure is of crucial importance to the outcomes of the hack week and may require both careful engineering and an in-depth discussion among the organizers.
-
-Perhaps unsurprisingly, the target audience for a hack week depends directly on the stated goal of the workshop. 
-At conferences, researchers typically report recent results to the community and network with other scientists largely from their discipline, leading to a fairly narrow target distribution in academic seniority and represented fields.
-At a summer school, the aim is generally for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows.
-Hack weeks differ in that knowledge transfer occurs across many levels of seniority, due to the relative newness of the content and the distributed knowledge base. 
-%Because the potential for knowledge transfer and community building is highest when the group is heterogeneous, participant diverse in both academic background and geographic location should be selected.
-%This increases the chances that participants will take their newfound knowledge into their respective communities and spread it further.
-
-Aside from these broad considerations, the tuning of the audience and participant selection will depend directly on several of the axes described in detail above. 
-In particular, education-focused workshops might be more homogeneous than project-focused workshops, and similarly single-discipline hack weeks may be more homogeneous than interdisciplinary ones. 
-Additional goals may be not directly related to the workshop itself, but on higher-order goals (e.g.\ increasing minority participation, knowledge transfer between academia and industry) that a community might find desirable and that can be directly addressed by engineering the participant selection.
-% For example, women are still a small minority in computational astrophysics. 
-%One goal of a hack week could be to increase the participation of gender minorities in computation-focused research. 
-%Similarly, if a particular sub-field within a discipline is ahead of others in their computational approaches, it might be worthwhile to engineer the mix of participants such that a number of experts from that sub-field are present.
-%One interesting axis to consider is the mix of academics versus participants from industry. 
-%The latter may often be more focused on practically useful applications of methods and data sets, and can therefore provide an important counterpoint to more theory-focused academics. 
-%In particular in fields where the career path from academia to industry is less immediately obvious (as this the case e.g.
-in astronomy, where little directly relevant industry exists), inviting back participants who transitioned to industry successfully can provide important role models and networking opportunities for junior participants.
-%Finally, with much of data science currently performed in the private sector, industry participants can bring directly relevant knowledge and new scientific approaches to the table.
-
-The need to carefully select participants dovetails with another requirement: transparency. 
-%Up to date, all hack weeks we have organized have been oversubscribed by at least a factor of two, and thus necessarily require turning applicants away. 
-The selection procedure should be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance. 
-Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the workshop's success as well as the larger community.
-
-In order to maximize transparency in our selection process we are working to remove the human element from the decision making as far as possible and transfer some of these tasks to a well-designed algorithm.
-Here, well-designed refers to several properties: (1) it is interpretable, i.e.
-the algorithm, its parameters and outputs can be understood by humans, (2) it is openly available, and thus the decision process can be inspected, (3) the algorithm itself does not perpetuate intrinsic biases in the data (though we will show below that an algorithm itself does not entirely remove the requirement of humans to be aware of their biases).
-
-In practice, most selection procedures follow two separate steps, though in the case of human committees, these steps are often conflated. 
-In the first step, the merit of a candidate must be assessed: does the candidate fulfill the requirements posed by the goals of the workshop?
-%For example, at Astro Hack Week, it may be unrealistic to admit an undergraduate student in biochemistry, unless they had a compelling reason for attending.
-This first step is difficult to automate, because it requires judgment calls, often based on long-form answers. 
-In the past, we have performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
-When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop. 
-The second step in the selection procedure thus requires tie-breaking between equally acceptable candidates. 
-It is here where one may impose outside constraints on the selection based on the goals of the workshop. 
-If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
-
-In particular, the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}} employs one such algorithm and is capable of efficient selection of an optimal candidate set based on human target fractions.
-It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest. 
-If these targets reproduce the distribution of the overall sample of candidates, the selection will become essentially random. 
-Any human biases involved in setting these target fractions will be perpetuated in the selection procedure. 
-Secondly, perhaps more obviously, the algorithm can only act on information that has been collected. 
-Biased participant sets may still result from selection procedures that fail to include crucial categories (for example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority).
-
-%Within our hack weeks the goal of the selection procedure has always been diversity among several axes. 
-%This includes academic subdiscipline, to allow for maximally possible knowledge exchange between different domain fields, as well as with researchers from adjacent fields including statistics, computer science and data science.
-%Other important categories revolve around the participants' previous knowledge in terms of machine learning, statistics and programming. 
-%Unlike for a summer school, we have generally selected for a diverse set of participants in order to be able to efficiently pair up novices and experts in projects. 
-%Finally, we also include demographic information with two main goals: firstly, representation has been proven to be an effective way to foster an inclusive environment where minority participants feel comfortable to participate and take risks.
-%Secondly, we would like to boost the participation of minority researchers in computational astrophysics, and our hack week model provides a useful framework for doing so.
-
-\section*{Themes}
-
-To date, all organized hack weeks have been subject-specific. 
-That is, they aimed to bring together a community with a shared scientific interest, e.g.\ neuroscience, astronomy or geoscience. 
-This has several clear advantages. 
-In particular, communities organized by subject generally share the same language.%, which drastically cuts down on the time participants spend simply agreeing on scientific terms. 
-Similarly, the scientific objectives and questions are broadly clear to all participants.%: there is little need to explain in detail why a particular question may be of interest to the field. 
-Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data.
-This leaves more time for active collaboration on cutting-edge science.
-
-However, there are also drawbacks to limiting the theme of a hack week to a single field. 
-Homogeneity may be a disadvantage if it leads to \textit{group think} and inhibits new, creative solutions. 
-Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing. 
-These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead. 
-In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task.%, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
-On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.
-
-An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of 2017.
-These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
-%Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others.
-%In addition, participants included computer vision researchers, who develop algorithms that can be applied to answer scientific questions. 
-%Similarly, a series of events at UC Berkeley (Text XD) focused on the analysis of text data across different domains of research.
-
-\section*{Design considerations}
-
-Design considerations depend very strongly on goals. 
-In general, longer events allow for a larger taught component, more ambitious projects and especially for cross-disciplinary events are more likely to provide enough time for different groups to effectively communicate across barriers of professional terminology.
-On the other hand, because of their large participatory component, hack weeks tend to be exhausting, and long events may lead to fatigue among attendees, where active participation and measurable outcomes may drop sharply in the latter days of the workshops.
-%This is particularly important if---as is often the case---participants continue participating in workshop-relevant activities even after the official component ends for the day.
-Ways to mitigate these effects include capping taught components during the day, providing a reasonable clear schedule, and limiting parallel components to avoid decision fatigue. 
-%At Astro Hack Week, we have limited the scheduled taught components to no more than three out of nine hours for each of the five workshop days, and unscheduled (shorter) tutorials to a maximum of two per day.
-%Even so, in our surveys, participants occasionally report concerns about choosing to attend tutorials versus working on projects.
-
-Aside from time, space plays a particularly important role in facilitating a successful hack week. 
-Universities in general provide a convenient venue with existing structures to facilitate hack weeks (e.g. access to scientific publications, institutional support including staff and funding). 
-On the other hand, most spaces in universities are designed for lectures, which are diametrically opposed to unconference-style events. 
-Thus, finding an appropriate physical space within a university may be a challenge. 
-However, with the rise of active learning as a preferred teaching methods, traditional lecture spaces are transforming into more flexible spaces that are generally appropriate for hack weeks. 
-As a general rule, the smaller a hack week's emphasis on a taught component, the more flexible the space has to be, with ample opportunity for re-configuration. 
-%At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine. 
-%It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not.
-
-Another important design consideration is group size. 
-%With all previous hack weeks severely oversubscribed, it seems natural to simply admit more participants. 
-%However, this can counter the ideals of the workshop, in particular when building a community is one of its stated goals. 
-If the group is too large, participants are unlikely to even meet each other, and workshop cohesion may be lost as the workshop fractures into smaller groups, often among participants who already know each other.
-This may inhibit knowledge transfer by clustering participants into small in-groups. 
-%Additionally, it is likely that the number participant-led components in the schedule may increase with workshop size. 
-%While generally desirable, a programme that is too crowded may lead to fatigue 
-On the other hand, if the size of the workshop is too small, it is unlikely to achieve the desired diversity among participants to foster new collaborations across sub-fields and disciplines. 
-In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.
-
-As mentioned above, the balance between pedagogy and working depends both on the goals of the workshop and the topics around which the workshop is organized. 
-If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other. 
-In communities where a shared understanding exists, tutorials can focus on more advanced or innovative topics, and less time may be allocated for them, leaving more time for active participation. 
-%In astronomy, a relatively small field, even students generally share a common knowledge base and have rudimentary knowledge about the types of data used and the challenges associated with each. 
-%Thus, we have focused the taught component of Astro Hack Week less on domain-specific knowledge, and instead offered tutorials in topics from domain-adjacent fields like statistics and computer science that attendees are unlikely to have encountered in their regular education.
-
-Hack week outcomes, in turn, depend strongly on participants and are often a function of their interests and seniority. 
-some attendees arrive with the stated goal of writing a specific scientific article, usually more advanced participants with significant pre-knowledge of both hackathons and their topic of interest. 
-Many attendees arrive with the plan to learn a specific topic (such as machine learning) or bring a specific data set they believe the new knowledge may be applicable to. 
-This leads to a wide variety of project types from sandbox-style explorations to focused work efforts. 
-%It is worth noting that while a scientific paper need not be the stated goal of a hack (and is unlikely to be completed in the short time allocated in any case), results may still be published as short reports or unconference proceedings.
-%For example, Neurohack week provides a venue for participants to publish a short (two-page) ''project report`` summarizing the hack that participants did during the week of NHW. 
-%Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.
-
-\subsection*{Box 1: Impostor Syndrome}
-
-The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \citep{Clance1978-ef}.
-Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
-This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
-The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee. 
-This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
-Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
- Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
-We are working within our hack weeks to mitigate IS using various techniques. 
-With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS. 
-More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g.
-one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
-%I think it helped ease the feeling!!``}). 
-Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
-
-\section*{Measures of success}
-
-Measuring the success of a hack week objectively is complicated by the diversity of goals of both organizing committee and attendees. 
-A participant who attends primarily in order to hack on a specific project may find the tutorials less overall useful than a graduate student exposed to the tutorial topics for the first time. 
-Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success. 
-For more project-driven events, one key metric is the number of publications that result from hack week projects. 
-However, this implies a very narrow definition of success in line with standard academic performance indicators, and may be hard to collect in practice, if participants do not explicitly include a reference to the event in their publication.
-Another, somewhat less traditional and complementary indicator is the activity of participants in terms of code written and committed to a public code repository (e.g. GitHub). 
-This could be defined in lines of code committed, the number of new open-source projects spawned, and the number of new developers contributing to open-source projects. 
-This assumes that participants work largely in the open during a hack week, and that most projects have a strong programming component. 
-Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes. 
-These measures ignore learning, community building as well as networking outcomes, which are often equally important objectives. 
-One possible approach here is to ask participants to self-report their success via a post-workshop survey.
-It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not. 
-%This would then provide a somewhat objective measure of the impact the hack week has had. 
-%In practice, this has not been done for any of the hack weeks conducted thus far.
-For practical purposes, we recommend an approach that combines all of these metrics: publications, new code and projects generated, self-reported learning outcomes, and survey results. 
-%Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
-
-If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
-Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent. 
-In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating. 
-%It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
-
-\section*{Results}
-
-In the following we report outcomes of previous events held in 2016 in three different fields (neuroscience, geoscience and astrophysics) and participants attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
-Because all three events are less than a year in the past, evaluating long-term outcomes (number of publications, citations, career outcomes) will require a longer baseline. 
-There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \citep{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
-[NOTE: HERE IS WHERE WE SHOULD REFER TO THE BOXES ONCE THEY ARE MERGED.]
-We asked participants of all three hack weeks to complete surveys to validate the hack week's success. 
-%All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.
-
-We find that at all three hack weeks, participant self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 2, left panel).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (see Figure 2, middle and right panels).
-%While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
-%Another important goal of the hack weeks was centred on community building and fostering collaborations. 
-We find that overall, the majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel).
-This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 3, right panel). 
-Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants. 
-We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories.
-This is expected: participants new to the field may participate to learn rather than to teach.
-
-Reproducibility and open science have recently drawn much attention from the scientific community, and all three hack weeks have been working to promote positive attitudes towards reproducibility and open science in their respective fields.
-We find that hack weeks have been largely successful at encouraging open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 4, left panel) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 4, middle panel).
-When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields. 
-In particular, astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
-%Neuroscience seems to fall somewhere in between. 
-This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
-Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%).
-Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible. 
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[\textwidth]{fig/eval_techskills.png}
-\caption{Self-reported technical learning outcomes at all three hack weeks}
-\label{fig:techskills}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[\textwidth]{fig/eval_collab.png}
-\caption{Hack week outcomes with respect to collaboration and peer-learning}
-\label{fig:techskills}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[\textwidth]{fig/eval_openscience.png}
-\caption{Changing attitudes towards open science at hack weeks}
-\label{fig:techskills}
-\end{center}
-\end{figure*}
-
-\subsection*{Box 2: Examples of Hack Week Outcomes}
-\subsubsection*{Example 1: Astro Hack Week}
-In 2015, a small team used the opportunity of Astro Hack Week to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of algorithms for time series analysis often used in X-ray astronomy.
-The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects, role models of successful projects, and an environment encouraging scientific risk taking.
-Since its beginnings at Astro Hack Week, Stingray has matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.
-\subsubsection*{Example 2: Geo Hack Week} 
-In 2016, a Geo Hack Week project team used Google Earth Engine to explore spatial patterns in climate, topography and population data with the goal of mapping the most suitable locations for renewable energy sites in the United States.
-The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine\footnote{\url{http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/}}.
-George Richardson, one of the project leads, now works for a renewable resource company in Seattle.
-\subsubsection*{Example 3: Neuro Hack Week} 
-Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies.
-This is a particular concern in studies in which children or patients are studied, as they are more likely to move.
-In the 2016 Neuro Hack Week one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
-To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off were varied.
-They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the Neuro Hack Week was over, and they eventually wrote and published a paper describing these results in the open access journal Research Ideas and Outcomes \citep{leonard2017}.
-
-\section*{Conclusions}
-
-The fast-paced changes of the computational and methodological landscape require traditional fields of science to rapidly adapt to new data analysis challenges. 
-Traditional modes of learning, including university curricula, are often too slow to incorporate new developments on short enough time scales to meet their acute need in scientific advancement. 
-To address this imbalance, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
-Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields
-
-As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
-Initial results from three events held in 2016 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
-Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.
-We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.
+\input{sections/abstract}
+\input{sections/intro}
+\input{sections/what}
+\input{sections/why}
+\input{sections/participants}
+\input{sections/themes}
+\input{sections/design}
+\input{sections/results}
+\input{sections/conclusions}
 
@@
 \begin{addendum}
- \item{Anyone thanked here: *Feel free to move up to the Author list!* It is a pleasure to thank Laura Norén (NYU) for help on ethics and IRB.
+ \item{Anyone thanked here: *Feel free to move up to the Author list!* It is a pleasure to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB.
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.

### 2017-08-17 sections/conclusions.tex
Divide into sections, for easier work moving forward.

\section*{Conclusions}

The fast-paced changes of the computational and methodological landscape require traditional fields of science to rapidly adapt to new data analysis challenges.
Traditional modes of learning, including university curricula, are often too slow to incorporate new developments on short enough time scales to meet their acute need in scientific advancement.
To address this imbalance, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields

As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
Initial results from three events held in 2016 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.
We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.

### 2017-08-17 sections/design.tex
Divide into sections, for easier work moving forward.

\section*{Design considerations}

Design considerations depend very strongly on goals.
In general, longer events allow for a larger taught component, more ambitious projects and especially for cross-disciplinary events are more likely to provide enough time for different groups to effectively communicate across barriers of professional terminology.
On the other hand, because of their large participatory component, hack weeks tend to be exhausting, and long events may lead to fatigue among attendees, where active participation and measurable outcomes may drop sharply in the latter days of the workshops.
%This is particularly important if---as is often the case---participants continue participating in workshop-relevant activities even after the official component ends for the day.
Ways to mitigate these effects include capping taught components during the day, providing a reasonable clear schedule, and limiting parallel components to avoid decision fatigue.
%At Astro Hack Week, we have limited the scheduled taught components to no more than three out of nine hours for each of the five workshop days, and unscheduled (shorter) tutorials to a maximum of two per day.
%Even so, in our surveys, participants occasionally report concerns about choosing to attend tutorials versus working on projects.

Aside from time, space plays a particularly important role in facilitating a successful hack week.
Universities in general provide a convenient venue with existing structures to facilitate hack weeks (e.g. access to scientific publications, institutional support including staff and funding).
On the other hand, most spaces in universities are designed for lectures, which are diametrically opposed to unconference-style events.
Thus, finding an appropriate physical space within a university may be a challenge.
However, with the rise of active learning as a preferred teaching methods, traditional lecture spaces are transforming into more flexible spaces that are generally appropriate for hack weeks.
As a general rule, the smaller a hack week's emphasis on a taught component, the more flexible the space has to be, with ample opportunity for re-configuration.
%At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine.
%It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not.

Another important design consideration is group size.
%With all previous hack weeks severely oversubscribed, it seems natural to simply admit more participants.
%However, this can counter the ideals of the workshop, in particular when building a community is one of its stated goals.
If the group is too large, participants are unlikely to even meet each other, and workshop cohesion may be lost as the workshop fractures into smaller groups, often among participants who already know each other.
This may inhibit knowledge transfer by clustering participants into small in-groups.
%Additionally, it is likely that the number participant-led components in the schedule may increase with workshop size.
%While generally desirable, a programme that is too crowded may lead to fatigue
On the other hand, if the size of the workshop is too small, it is unlikely to achieve the desired diversity among participants to foster new collaborations across sub-fields and disciplines.
In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.

As mentioned above, the balance between pedagogy and working depends both on the goals of the workshop and the topics around which the workshop is organized.
If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
In communities where a shared understanding exists, tutorials can focus on more advanced or innovative topics, and less time may be allocated for them, leaving more time for active participation.
%In astronomy, a relatively small field, even students generally share a common knowledge base and have rudimentary knowledge about the types of data used and the challenges associated with each.
%Thus, we have focused the taught component of Astro Hack Week less on domain-specific knowledge, and instead offered tutorials in topics from domain-adjacent fields like statistics and computer science that attendees are unlikely to have encountered in their regular education.

Hack week outcomes, in turn, depend strongly on participants and are often a function of their interests and seniority.
some attendees arrive with the stated goal of writing a specific scientific article, usually more advanced participants with significant pre-knowledge of both hackathons and their topic of interest.
Many attendees arrive with the plan to learn a specific topic (such as machine learning) or bring a specific data set they believe the new knowledge may be applicable to.
This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
%It is worth noting that while a scientific paper need not be the stated goal of a hack (and is unlikely to be completed in the short time allocated in any case), results may still be published as short reports or unconference proceedings.
%For example, Neurohack week provides a venue for participants to publish a short (two-page) ''project report`` summarizing the hack that participants did during the week of NHW.
%Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.

\subsection*{Box 1: Impostor Syndrome}

The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \citep{Clance1978-ef}.
Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
 Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
We are working within our hack weeks to mitigate IS using various techniques.
With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g. one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
%I think it helped ease the feeling!!``}).
Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.

### 2017-08-17 sections/results.tex
Divide into sections, for easier work moving forward.

\section*{Measures of success}

Measuring the success of a hack week objectively is complicated by the diversity of goals of both organizing committee and attendees.
A participant who attends primarily in order to hack on a specific project may find the tutorials less overall useful than a graduate student exposed to the tutorial topics for the first time.
Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
For more project-driven events, one key metric is the number of publications that result from hack week projects.
However, this implies a very narrow definition of success in line with standard academic performance indicators, and may be hard to collect in practice, if participants do not explicitly include a reference to the event in their publication.
Another, somewhat less traditional and complementary indicator is the activity of participants in terms of code written and committed to a public code repository (e.g. GitHub).
This could be defined in lines of code committed, the number of new open-source projects spawned, and the number of new developers contributing to open-source projects.
This assumes that participants work largely in the open during a hack week, and that most projects have a strong programming component.
Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes.
These measures ignore learning, community building as well as networking outcomes, which are often equally important objectives.
One possible approach here is to ask participants to self-report their success via a post-workshop survey.
It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
%This would then provide a somewhat objective measure of the impact the hack week has had.
%In practice, this has not been done for any of the hack weeks conducted thus far.
For practical purposes, we recommend an approach that combines all of these metrics: publications, new code and projects generated, self-reported learning outcomes, and survey results.
%Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.

If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent.
In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating.
%It is important to establish procedures to conduct such studies and to obtain approval from an IRB.

\section*{Results}

In the following we report outcomes of previous events held in 2016 in three different fields (neuroscience, geoscience and astrophysics) and participants attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
Because all three events are less than a year in the past, evaluating long-term outcomes (number of publications, citations, career outcomes) will require a longer baseline.
There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \citep{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
[NOTE: HERE IS WHERE WE SHOULD REFER TO THE BOXES ONCE THEY ARE MERGED.]
We asked participants of all three hack weeks to complete surveys to validate the hack week's success.
%All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.

We find that at all three hack weeks, participant self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 2, left panel).
The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (see Figure 2, middle and right panels).
%While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
%Another important goal of the hack weeks was centred on community building and fostering collaborations.
We find that overall, the majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel).
This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 3, right panel).
Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories.
This is expected: participants new to the field may participate to learn rather than to teach.

Reproducibility and open science have recently drawn much attention from the scientific community, and all three hack weeks have been working to promote positive attitudes towards reproducibility and open science in their respective fields.
We find that hack weeks have been largely successful at encouraging open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 4, left panel) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 4, middle panel).
When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
In particular, astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
%Neuroscience seems to fall somewhere in between.
This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%).
Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
\begin{figure*}[htbp]
\begin{center}
\includegraphics[\textwidth]{fig/eval_techskills.png}
\caption{Self-reported technical learning outcomes at all three hack weeks}
\label{fig:techskills}
\end{center}
\end{figure*}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[\textwidth]{fig/eval_collab.png}
\caption{Hack week outcomes with respect to collaboration and peer-learning}
\label{fig:techskills}
\end{center}
\end{figure*}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[\textwidth]{fig/eval_openscience.png}
\caption{Changing attitudes towards open science at hack weeks}
\label{fig:techskills}
\end{center}
\end{figure*}

\subsection*{Box 2: Examples of Hack Week Outcomes}
\subsubsection*{Example 1: Astro Hack Week}
In 2015, a small team used the opportunity of Astro Hack Week to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of algorithms for time series analysis often used in X-ray astronomy.
The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects, role models of successful projects, and an environment encouraging scientific risk taking.
Since its beginnings at Astro Hack Week, Stingray has matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.
\subsubsection*{Example 2: Geo Hack Week}
In 2016, a Geo Hack Week project team used Google Earth Engine to explore spatial patterns in climate, topography and population data with the goal of mapping the most suitable locations for renewable energy sites in the United States.
The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine\footnote{\url{http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/}}.
George Richardson, one of the project leads, now works for a renewable resource company in Seattle.
\subsubsection*{Example 3: Neuro Hack Week}
Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies.
This is a particular concern in studies in which children or patients are studied, as they are more likely to move.
In the 2016 Neuro Hack Week one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off were varied.
They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the Neuro Hack Week was over, and they eventually wrote and published a paper describing these results in the open access journal Research Ideas and Outcomes \citep{leonard2017}.

### 2017-08-17 sections/themes.tex
Divide into sections, for easier work moving forward.

\section*{Themes}

To date, all organized hack weeks have been subject-specific.
That is, they aimed to bring together a community with a shared scientific interest, e.g.\ neuroscience, astronomy or geoscience.
This has several clear advantages.
In particular, communities organized by subject generally share the same language.%, which drastically cuts down on the time participants spend simply agreeing on scientific terms.
Similarly, the scientific objectives and questions are broadly clear to all participants.%: there is little need to explain in detail why a particular question may be of interest to the field.
Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data.
This leaves more time for active collaboration on cutting-edge science.

However, there are also drawbacks to limiting the theme of a hack week to a single field.
Homogeneity may be a disadvantage if it leads to \textit{group think} and inhibits new, creative solutions.
Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing.
These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead.
In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task.%, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.

An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of 2017.
These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
%Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others.
%In addition, participants included computer vision researchers, who develop algorithms that can be applied to answer scientific questions.
%Similarly, a series of events at UC Berkeley (Text XD) focused on the analysis of text data across different domains of research.

### 2017-08-17 sections/what.tex
Divide into sections, for easier work moving forward.

\section*{What is a hackathon?}

Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \citep{Decker2015}.
Hackathons have historically focused on software development and technology design as a way to motivate innovation within industry.
In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to 'learn by doing' within a constructivist educational framework \citep{Bransford2000-lu,Papert1980-fh}.
With this in mind, hackathons around scientific topics, designed to foster collaboration \citep{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \citep{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.

%The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon.
Following the typology of \citep{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \citep{Drouhard2017}.
''Catalytic`` hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
''Contributive`` hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
Finally, ''Communal`` hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.

Our past hack week events most closely follow the communal hackathon model as it applies to scientific communities of practice.
Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure 1).
For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.

We note that the terminology for these events is constantly evolving, and that the ''hackathon`` concept may have implicit connotations that are disfavored in some communities.
One criticism of hackathons is that they propel the ''geek`` stereotype and may present a barrier to creating an inclusive working environment, especially for women \citep{Decker2015}.
Another problem is the competitive atmosphere of many industry hackathons, where teams actively compete for prizes.
Because certain groups are more risk-averse than others, and risk-aversion traces demographics like gender, we urge caution when contemplating making an event competitive.
Other related nomenclature includes the ''unconference`` which is also a time-bounded event gathering together a specific community, but these often have minimal structure and do not necessarily have a focus on software.
The ''sprint`` or ''scrum`` label typically refers to an event focused on rapid software development on a specific set of code.
%This may be a component of our hack week model but it ignores the important pedagogical components.

### 2017-08-17 sections/why.tex
Divide into sections, for easier work moving forward.

\section*{Why run a Hack Week?}

There are several reasons to run a hack week of the sort described here.

\begin{itemize}
\item{\textit{Education and training}: %Some hack weeks are more focused on education than others (see Figure 1).
Even those hackathons not explicitly focused on skills-development will include extensive discussions on reproducible and open science practices that are valuable to participants.
%Discussing these topics in a diverse group setting helps to ingrain within participants the basic conventions of the broader open source community.
Participants can then serve as ''ambassadors`` of these practices and foster their adoption by other researchers in the field.
Furthermore, the kind of co-working that is inherent to the hack week format facilitates lateral knowledge transfer, where participants learn from each other about practices and tools of the trade.}
%Often these tools are not described in research products such as papers, or even in the software implementations.}
\item{\textit{Tool development}: hack weeks present an opportunity for computational tool developers to get together with users of these tools, and to critically evaluate the use of the tools in addressing particular scientific questions.}
\item{\textit{Community building}: a hack week is an opportunity to galvanize a community within a field.% that cares about computational issues, and relates to specific open-source software projects.
For example, because of its focus on free and open-source tools, Neurohack week helps galvanize the Neuroimaging in Python (http://nipy.org) community.
It also allows researchers who may otherwise be isolated within their departments and subfield on account of their computational specialization to come together and exchange experiences.}
%As one participant at Astro Hack Week expressed in the survey, \textit{''The main thing I took from Astrohack week is the fact that I'm not alone``}.}
\item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g.\ between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines.
Despite the impact of interdisciplinary research \citep{Hall2012-hi}, these experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.}
%This is despite the fact that interdisciplinary research can also be very impactful \citep{Hall2012-hi}.
%A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}

\item{\textit{Recruitment and networking}: hack weeks often bring together research, government and industry communities, providing an opportunity for participants to learn about each other's interests and their abilities in close quarters.
This kind of extended opportunity for interaction and ''on-the-job`` testing is considered an excellent form of evaluation of prospective job candidates [CITATION NEEDED].
Conversely, participants have an opportunity to gain exposure to potential career opportunities in communities other than their own.}
%As an example, one of the participants in Neurohack week 2016 will be joining the eScience Institute as a post-doc in Fall of 2017 (and as an instructor in Neurohack week 2017).}
%\item{\textit{It's fun}: the opportunity to focus for a week on a slightly more whimsical and adventurous sort of scientific activities can be both fruitful, and enjoyable.
%Even aside from the prospect of interdisciplinary research, it can provide a laboratory for more experimental approaches and high-risk projects than scientist will take on in their day-to-day research.}
\end{itemize}

It is worth noting that the reasons for participants to attend a hack week are as diverse as the goals that the organizing committee explicitly states.
This is largely a function of the diversity of the participants (see section below), and should be something that organizers design for.
Beginner participants may attend primarily to learn a new technique or method, while more experienced participants attend to gain more experience mentoring and teaching.
Some researchers may come with particular projects and collaborators in mind, while others come with a focus on learning and with no explicit plan for the project-based work.
%While it is very hard to design a workshop that is universally useful to all participants, in practice many researchers attend with the understanding that some components will be more congruent with their own goals than others, but also bring an open mind and a willingness to learn.
%In addition, the diversity of goals can effectively be a strength of the workshop, if the organizers can facilitate matching up participants with complementary goals.

### 2017-08-17 sections/design.tex
Use the correct citation format. Replace all "citep" => "cite"
Bingo!

@@
 The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \citep{Clance1978-ef}.
+Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \cite{Clance1978-ef}.
 Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
@@
 The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
-This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
+This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
 Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.

### 2017-08-17 sections/why.tex
Use the correct citation format. Replace all "citep" => "cite"
Bingo!

@@
 \item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g.\ between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines.
-Despite the impact of interdisciplinary research \citep{Hall2012-hi}, these experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.}
-%This is despite the fact that interdisciplinary research can also be very impactful \citep{Hall2012-hi}.
+Despite the impact of interdisciplinary research \cite{Hall2012-hi}, these experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.}
+%This is despite the fact that interdisciplinary research can also be very impactful \cite{Hall2012-hi}.
 %A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}

### 2017-08-17 sections/abstract.tex

\begin{abstract}
Across almost all scientific disciplines, the instruments that record our experimental data and the methods required for storage and data analysis are rapidly increasing in complexity.
This gives rise to the need for scientific communities to adapt on shorter time scales than traditional university curricula allow for, and therefore requires new modes of knowledge transfer.
At the same time, as disciplines become more specialized, data analysis problems often tend to be fairly universal.
It is thus desirable to foster exchange of ideas and computational workflows across disciplines.
Among the different approaches put forward toward addressing these issues, hack weeks have in recent years emerged as both an effective tool for training researchers in modern data analysis workflows, and in fostering exchange between subdisciplines within and between scientific fields.
While interpreted differently in the various disciplines where they have been implemented, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and free-form project work.
In this paper, we present the concept of a hack week in the larger context of scientific meetings and point out similarities and differences to traditional conferences.
We motivate the need for such an event and present in detail its strengths and challenges.
We find that hack weeks are successful at cultivating collaboration and the exchange of knowledge.
Participants self-report that these events help them both in their day-to-day research as well as their careers.
Based on our results, we conclude that hack weeks present an effective, easy-to-implement, fairly low-cost tool to positively impact data analysis literacy in academic disciplines, foster collaboration and cultivate best practices.
\end{abstract}

### 2017-08-17 sections/intro.tex

\section*{Introduction}
\label{sec:introduction}

As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
At the same time, formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998\footnote{\url{https://f1000research.com/articles/3-62/v2}}.
At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.
Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, and more project-focused than pedagogical meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\footnote{\url{https://gigascience.biomedcentral.com/articles/10.1186/s13742-016-0121-x}}, and a hackathon associate with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
In broad-brush, pedagogically-focused workshops and summer schools tend to follow a classic academic model where novices learn a skillset from experts, while project-focused workshops focus on people collaboratively exercising the skillset they already have.
A disadvantage of the summer school model is that it can tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
A disadvantage of the hackathon model is the common perception (whether accurate or not) that the week is designed for experts in technical tools, which may discourage others from attending.
In 2014, we started Astro Hack Week to try to fill the gaps between these models.
The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for collaborative hacks or creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.

\begin{figure}
\begin{center}
\includegraphics[width=9cm]{fig/HackSpectrum.png}
\caption{Comparison of Extracurricular Workshop Models}
\label{fig:hackspectrum}
\end{center}
\end{figure}

As of the publication of this paper, we have run five such hack week events: three focused on Astronomy, one focused on Neuroscience, and one focused on Geoscience.
Below we will share some of the philosophy behind the hack week model, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.

### 2017-08-17 sections/participants.tex

\section*{Audience and Participant Selection}

Because hack weeks are at their core non-traditional workshops, it is worth thinking in detail about the target audience for a hack week, along with methods to select an appropriate group of participants. Here, appropriate essentially refers to 'most likely to achieve the stated goal of the hack week', though measuring that achievement remains a difficult issue itself.
This is especially important because hack weeks are very participant-driven workshops.
%At a traditional conference or summer school, the program will be largely set in advance by the organizing committee using the careful selection of speakers and lecturers.
%At a hack week much of the program is in flux both before and during the workshop.
In this context, a successful hack week requires participants who are both interested in participation and feel comfortable participating in discussions and taking risks.
Therefore, a robust, effective participant selection procedure is of crucial importance to the outcomes of the hack week and may require both careful engineering and an in-depth discussion among the organizers.

Perhaps unsurprisingly, the target audience for a hack week depends directly on the stated goal of the workshop.
At conferences, researchers typically report recent results to the community and network with other scientists largely from their discipline, leading to a fairly narrow target distribution in academic seniority and represented fields.
At a summer school, the aim is generally for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows.
Hack weeks differ in that knowledge transfer occurs across many levels of seniority, due to the relative newness of the content and the distributed knowledge base.
%Because the potential for knowledge transfer and community building is highest when the group is heterogeneous, participant diverse in both academic background and geographic location should be selected.
%This increases the chances that participants will take their newfound knowledge into their respective communities and spread it further.

Aside from these broad considerations, the tuning of the audience and participant selection will depend directly on several of the axes described in detail above.
In particular, education-focused workshops might be more homogeneous than project-focused workshops, and similarly single-discipline hack weeks may be more homogeneous than interdisciplinary ones.
Additional goals may be not directly related to the workshop itself, but on higher-order goals (e.g.\ increasing minority participation, knowledge transfer between academia and industry) that a community might find desirable and that can be directly addressed by engineering the participant selection.
% For example, women are still a small minority in computational astrophysics.
%One goal of a hack week could be to increase the participation of gender minorities in computation-focused research.
%Similarly, if a particular sub-field within a discipline is ahead of others in their computational approaches, it might be worthwhile to engineer the mix of participants such that a number of experts from that sub-field are present.
%One interesting axis to consider is the mix of academics versus participants from industry.
%The latter may often be more focused on practically useful applications of methods and data sets, and can therefore provide an important counterpoint to more theory-focused academics.
%In particular in fields where the career path from academia to industry is less immediately obvious (as this the case e.g. in astronomy, where little directly relevant industry exists), inviting back participants who transitioned to industry successfully can provide important role models and networking opportunities for junior participants.
%Finally, with much of data science currently performed in the private sector, industry participants can bring directly relevant knowledge and new scientific approaches to the table.

The need to carefully select participants dovetails with another requirement: transparency.
%Up to date, all hack weeks we have organized have been oversubscribed by at least a factor of two, and thus necessarily require turning applicants away.
The selection procedure should be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance.
Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the workshop's success as well as the larger community.

In order to maximize transparency in our selection process we are working to remove the human element from the decision making as far as possible and transfer some of these tasks to a well-designed algorithm.
Here, well-designed refers to several properties: (1) it is interpretable, i.e. the algorithm, its parameters and outputs can be understood by humans, (2) it is openly available, and thus the decision process can be inspected, (3) the algorithm itself does not perpetuate intrinsic biases in the data (though we will show below that an algorithm itself does not entirely remove the requirement of humans to be aware of their biases).

In practice, most selection procedures follow two separate steps, though in the case of human committees, these steps are often conflated.
In the first step, the merit of a candidate must be assessed: does the candidate fulfill the requirements posed by the goals of the workshop? %For example, at Astro Hack Week, it may be unrealistic to admit an undergraduate student in biochemistry, unless they had a compelling reason for attending.
This first step is difficult to automate, because it requires judgment calls, often based on long-form answers.
In the past, we have performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
The second step in the selection procedure thus requires tie-breaking between equally acceptable candidates.
It is here where one may impose outside constraints on the selection based on the goals of the workshop.
If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.

In particular, the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}} employs one such algorithm and is capable of efficient selection of an optimal candidate set based on human target fractions.
It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.
If these targets reproduce the distribution of the overall sample of candidates, the selection will become essentially random.
Any human biases involved in setting these target fractions will be perpetuated in the selection procedure.
Secondly, perhaps more obviously, the algorithm can only act on information that has been collected.
Biased participant sets may still result from selection procedures that fail to include crucial categories (for example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority).

%Within our hack weeks the goal of the selection procedure has always been diversity among several axes.
%This includes academic subdiscipline, to allow for maximally possible knowledge exchange between different domain fields, as well as with researchers from adjacent fields including statistics, computer science and data science.
%Other important categories revolve around the participants' previous knowledge in terms of machine learning, statistics and programming.
%Unlike for a summer school, we have generally selected for a diverse set of participants in order to be able to efficiently pair up novices and experts in projects.
%Finally, we also include demographic information with two main goals: firstly, representation has been proven to be an effective way to foster an inclusive environment where minority participants feel comfortable to participate and take risks.
%Secondly, we would like to boost the participation of minority researchers in computational astrophysics, and our hack week model provides a useful framework for doing so.

### 2017-08-17 sections/results.tex

@@
 Because all three events are less than a year in the past, evaluating long-term outcomes (number of publications, citations, career outcomes) will require a longer baseline.
-There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \citep{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
+There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
 [NOTE: HERE IS WHERE WE SHOULD REFER TO THE BOXES ONCE THEY ARE MERGED.]
@@
 They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the Neuro Hack Week was over, and they eventually wrote and published a paper describing these results in the open access journal Research Ideas and Outcomes \citep{leonard2017}.
+The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the Neuro Hack Week was over, and they eventually wrote and published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.

### 2017-08-17 sections/what.tex

@@
 
-Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \citep{Decker2015}.
+Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
 Hackathons have historically focused on software development and technology design as a way to motivate innovation within industry.
 In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
-In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to 'learn by doing' within a constructivist educational framework \citep{Bransford2000-lu,Papert1980-fh}.
-With this in mind, hackathons around scientific topics, designed to foster collaboration \citep{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \citep{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.
+In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to 'learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
+With this in mind, hackathons around scientific topics, designed to foster collaboration \cite{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \cite{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.
 
 %The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon.
-Following the typology of \citep{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
-Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \citep{Drouhard2017}.
+Following the typology of \cite{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
+Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \cite{Drouhard2017}.
 ''Catalytic`` hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
@@
 We note that the terminology for these events is constantly evolving, and that the ''hackathon`` concept may have implicit connotations that are disfavored in some communities.
-One criticism of hackathons is that they propel the ''geek`` stereotype and may present a barrier to creating an inclusive working environment, especially for women \citep{Decker2015}.
+One criticism of hackathons is that they propel the ''geek`` stereotype and may present a barrier to creating an inclusive working environment, especially for women \cite{Decker2015}.
 Another problem is the competitive atmosphere of many industry hackathons, where teams actively compete for prizes.

### 2017-08-17 sections/design.tex
Moved impostor syndrome box to supplementary materials

@@
 %Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.
-
-\subsection*{Box 1: Impostor Syndrome}
-
-The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \cite{Clance1978-ef}.
-Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
-This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
-The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
-This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
-Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
- Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
-We are working within our hack weeks to mitigate IS using various techniques.
-With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
-More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g.
-one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
-%I think it helped ease the feeling!!``}).
-Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.

### 2017-08-17 supplementary_material.tex
Moved impostor syndrome box to supplementary materials

@@
 
+\subsection{Box 1: Impostor Syndrome}
+
+The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
+Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \cite{Clance1978-ef}.
+Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
+This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
+The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
+This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
+Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
+ Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
+We are working within our hack weeks to mitigate IS using various techniques.
+With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
+More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g.
+one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
+%I think it helped ease the feeling!!``}).
+Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
+
 \subsection{What to do after}

### 2017-08-17 sections/themes.tex
Shortened themes to single paragraph

@@
 
-To date, all organized hack weeks have been subject-specific.
-That is, they aimed to bring together a community with a shared scientific interest, e.g.\ neuroscience, astronomy or geoscience.
-This has several clear advantages.
-In particular, communities organized by subject generally share the same language.%, which drastically cuts down on the time participants spend simply agreeing on scientific terms.
-Similarly, the scientific objectives and questions are broadly clear to all participants.%: there is little need to explain in detail why a particular question may be of interest to the field.
-Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data.
-This leaves more time for active collaboration on cutting-edge science.
+To date, all organized hack weeks have been subject-specific, i.e.\ aimed to bring together a community with a shared scientific interest, such as neuroscience.
+Advantages to this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.
+On the other hand, homogeneity may lead to \textit{group think} and inhibit new, creative solutions. 
+In this case, it may be advantageous to design a hack week around a technique (e.g.\ Gaussian Processes) or modality (e.g.\ imaging), such as the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings.
+For these events, building a shared vocabulary and understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.
 
-However, there are also drawbacks to limiting the theme of a hack week to a single field.
-Homogeneity may be a disadvantage if it leads to \textit{group think} and inhibits new, creative solutions.
-Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing.
-These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead.
-In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task.%, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
-On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.
+%This has several clear advantages.
+%In particular, communities organized by subject generally share the same language.%, which drastically cuts down on the time participants spend simply agreeing on scientific terms.
+%Similarly, the scientific objectives and questions are broadly clear to all participants.%: there is little need to explain in detail why a particular question may be of interest to the field.
+%Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data.
+%This leaves more time for active collaboration on cutting-edge science.
 
-An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of 2017.
-These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
+%However, there are also drawbacks to limiting the theme of a hack week to a single field.
+%Homogeneity may be a disadvantage if it leads to \textit{group think} and inhibits new, creative solutions.
+%Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing.
+%These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead.
+%In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task.%, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
+%On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.
+
+%An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of %2017.
+%These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
 %Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others.

### 2017-08-17 sections/themes.tex
Minor typo

@@
 
-To date, all organized hack weeks have been subject-specific, i.e.\ aimed to bring together a community with a shared scientific interest, such as neuroscience.
+To date, all organized hack weeks have been subject-specific, i.e.\ aimed at bringing together a community with a shared scientific interest, such as neuroscience.
 Advantages to this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.

### 2017-08-18 sections/results.tex
Consolidate and shorten metrics/results sections.

@@
-\section*{Measures of success}
+\section*{Results}
 
-Measuring the success of a hack week objectively is complicated by the diversity of goals of both organizing committee and attendees.
-A participant who attends primarily in order to hack on a specific project may find the tutorials less overall useful than a graduate student exposed to the tutorial topics for the first time.
+Measuring the success of a hack week objectively is complicated by the variety of objectives that a hack week might have (see \ref{sec:why}).
 Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
-For more project-driven events, one key metric is the number of publications that result from hack week projects.
-However, this implies a very narrow definition of success in line with standard academic performance indicators, and may be hard to collect in practice, if participants do not explicitly include a reference to the event in their publication.
-Another, somewhat less traditional and complementary indicator is the activity of participants in terms of code written and committed to a public code repository (e.g. GitHub).
-This could be defined in lines of code committed, the number of new open-source projects spawned, and the number of new developers contributing to open-source projects.
-This assumes that participants work largely in the open during a hack week, and that most projects have a strong programming component.
-Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes.
-These measures ignore learning, community building as well as networking outcomes, which are often equally important objectives.
-One possible approach here is to ask participants to self-report their success via a post-workshop survey.
-It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
+One key metric is the number of publications that result from hack week projects, but this is a fairly narrow definition of success, in line with standard academic performance indicators.
+Assuming that participants work largely in the open during a hack week, and that most projects have a strong programming component another indicator of success is the activity of participants in terms of code written and committed to a public code repository.
+Still, these measures ignore learning, community-building as well as networking outcomes.
+These outcomes can be assessed through post-workshop surveys.
+%It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
 %This would then provide a somewhat objective measure of the impact the hack week has had.
 %In practice, this has not been done for any of the hack weeks conducted thus far.
-For practical purposes, we recommend an approach that combines all of these metrics: publications, new code and projects generated, self-reported learning outcomes, and survey results.
+Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see \ref{box:outcomes})
 %Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
 
-If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
-Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent.
-In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating.
+%If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
+%Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent.
+%In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating.
 %It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
 
-\section*{Results}
-
-In the following we report outcomes of previous events held in 2016 in three different fields (neuroscience, geoscience and astrophysics) and participants attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
-Because all three events are less than a year in the past, evaluating long-term outcomes (number of publications, citations, career outcomes) will require a longer baseline.
-There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
-[NOTE: HERE IS WHERE WE SHOULD REFER TO THE BOXES ONCE THEY ARE MERGED.]
-We asked participants of all three hack weeks to complete surveys to validate the hack week's success.
-%All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.
+We report the outcomes of astro-, geo- and neuro- hack weeks from 2016.
+In the results from post-workshop surveys, we focus on attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
 
@@
 Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
+
 \begin{figure*}[htbp]
@@
 
+Because all three events are relatively recent, it is still early to evaluate long-term outcomes, such as publications and collaborations resulting from these events.
+There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
+For specific examples, see also \ref{box:outcomes}.
+
 \subsection*{Box 2: Examples of Hack Week Outcomes}
+\label{box:outcomes}
 \subsubsection*{Example 1: Astro Hack Week}

### 2017-08-18 sections/results.tex
Consolidate the three into one figure. Caption.

@@
 \includegraphics[\textwidth]{fig/eval_techskills.png}
-\caption{Self-reported technical learning outcomes at all three hack weeks}
-\label{fig:techskills}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[htbp]
-\begin{center}
 \includegraphics[\textwidth]{fig/eval_collab.png}
-\caption{Hack week outcomes with respect to collaboration and peer-learning}
-\label{fig:techskills}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[htbp]
-\begin{center}
 \includegraphics[\textwidth]{fig/eval_openscience.png}
-\caption{Changing attitudes towards open science at hack weeks}
-\label{fig:techskills}
+\caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
+We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
+\label{fig:survey}
 \end{center}

### 2017-08-20 sections/participants.tex
Rewrote participant selection section, to shorten it down.

@@
 
-Because hack weeks are at their core non-traditional workshops, it is worth thinking in detail about the target audience for a hack week, along with methods to select an appropriate group of participants.
-Here, appropriate essentially refers to 'most likely to achieve the stated goal of the hack week', though measuring that achievement remains a difficult issue itself.
-This is especially important because hack weeks are very participant-driven workshops.
-%At a traditional conference or summer school, the program will be largely set in advance by the organizing committee using the careful selection of speakers and lecturers.
-%At a hack week much of the program is in flux both before and during the workshop.
-In this context, a successful hack week requires participants who are both interested in participation and feel comfortable participating in discussions and taking risks.
-Therefore, a robust, effective participant selection procedure is of crucial importance to the outcomes of the hack week and may require both careful engineering and an in-depth discussion among the organizers.
-
-Perhaps unsurprisingly, the target audience for a hack week depends directly on the stated goal of the workshop.
-At conferences, researchers typically report recent results to the community and network with other scientists largely from their discipline, leading to a fairly narrow target distribution in academic seniority and represented fields.
-At a summer school, the aim is generally for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows.
-Hack weeks differ in that knowledge transfer occurs across many levels of seniority, due to the relative newness of the content and the distributed knowledge base.
-%Because the potential for knowledge transfer and community building is highest when the group is heterogeneous, participant diverse in both academic background and geographic location should be selected.
-%This increases the chances that participants will take their newfound knowledge into their respective communities and spread it further.
-
-Aside from these broad considerations, the tuning of the audience and participant selection will depend directly on several of the axes described in detail above.
-In particular, education-focused workshops might be more homogeneous than project-focused workshops, and similarly single-discipline hack weeks may be more homogeneous than interdisciplinary ones.
-Additional goals may be not directly related to the workshop itself, but on higher-order goals (e.g.\ increasing minority participation, knowledge transfer between academia and industry) that a community might find desirable and that can be directly addressed by engineering the participant selection.
-% For example, women are still a small minority in computational astrophysics.
-%One goal of a hack week could be to increase the participation of gender minorities in computation-focused research.
-%Similarly, if a particular sub-field within a discipline is ahead of others in their computational approaches, it might be worthwhile to engineer the mix of participants such that a number of experts from that sub-field are present.
-%One interesting axis to consider is the mix of academics versus participants from industry.
-%The latter may often be more focused on practically useful applications of methods and data sets, and can therefore provide an important counterpoint to more theory-focused academics.
-%In particular in fields where the career path from academia to industry is less immediately obvious (as this the case e.g.
-in astronomy, where little directly relevant industry exists), inviting back participants who transitioned to industry successfully can provide important role models and networking opportunities for junior participants.
-%Finally, with much of data science currently performed in the private sector, industry participants can bring directly relevant knowledge and new scientific approaches to the table.
-
-The need to carefully select participants dovetails with another requirement: transparency.
-%Up to date, all hack weeks we have organized have been oversubscribed by at least a factor of two, and thus necessarily require turning applicants away.
-The selection procedure should be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance.
-Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the workshop's success as well as the larger community.
-
-In order to maximize transparency in our selection process we are working to remove the human element from the decision making as far as possible and transfer some of these tasks to a well-designed algorithm.
-Here, well-designed refers to several properties: (1) it is interpretable, i.e.
-the algorithm, its parameters and outputs can be understood by humans, (2) it is openly available, and thus the decision process can be inspected, (3) the algorithm itself does not perpetuate intrinsic biases in the data (though we will show below that an algorithm itself does not entirely remove the requirement of humans to be aware of their biases).
-
-In practice, most selection procedures follow two separate steps, though in the case of human committees, these steps are often conflated.
+A thoughtful robust, and effective participant selection procedure is of crucial importance to the success of a hack week and may require both careful engineering and an in-depth discussion among the organizers.
+At conferences, researchers typically report recent results to the community and network with other scientists, largely from their discipline.
+In contrast, traditional summer schools aim for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows.
+Hack weeks differ from both of these in that knowledge transfer occurs across many levels of seniority and across disciplinary boundaries, as well as novelty of the topics discussed,ranging from established theory to new findings.
+In addition, the content is primarily participant-driven, and much of it is generated during the event itself.
+This mean that a successful hack week requires participants who are both interested in participation, and also feel comfortable participating in discussions and taking risks.
+Participant selection also addresses other objectives of a hack week such as increasing minority participation and gender diversity, knowledge transfer between academia and industry, etc.
+These factors can also be addressed by thoughtful participant selection.
+To meet these goals, a thoughtful participant selection process needs to be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance in selecting participants.
+Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
+One way to maximize transparency in the selection process is to remove the human element from the decision making as far as possible and transfer some of these tasks to a well-designed algorithm.
+In this context, well-designed means \cite{o2017weapons}: (1) interpretable, i.e.
+the algorithm, its parameters and outputs can be understood by humans, (2) openly available, and thus the decision process can be inspected, (3) the algorithm itself does not perpetuate intrinsic biases in the data (though we will show below that an algorithm itself does not entirely remove the requirement of humans to be aware of their biases).
+In practice, most selection procedures follow two separate steps.
 In the first step, the merit of a candidate must be assessed: does the candidate fulfill the requirements posed by the goals of the workshop?
-%For example, at Astro Hack Week, it may be unrealistic to admit an undergraduate student in biochemistry, unless they had a compelling reason for attending.
 This first step is difficult to automate, because it requires judgment calls, often based on long-form answers.
-In the past, we have performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
+To avoid implicit/unconcsious bias, in the past, we have sometimes performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
 When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
-The second step in the selection procedure thus requires tie-breaking between equally acceptable candidates.
+The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
 It is here where one may impose outside constraints on the selection based on the goals of the workshop.
@@
 
-In particular, the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}} employs one such algorithm and is capable of efficient selection of an optimal candidate set based on human target fractions.
+On solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.
+The algorithm maximizes the variance of participants on specified dimensions (e.g., career stage, geographic location, etc.) to meet pre-set fractions set by the organizers.
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.
@@
 Secondly, perhaps more obviously, the algorithm can only act on information that has been collected.
-Biased participant sets may still result from selection procedures that fail to include crucial categories (for example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority).
-
-%Within our hack weeks the goal of the selection procedure has always been diversity among several axes.
-%This includes academic subdiscipline, to allow for maximally possible knowledge exchange between different domain fields, as well as with researchers from adjacent fields including statistics, computer science and data science.
-%Other important categories revolve around the participants' previous knowledge in terms of machine learning, statistics and programming.
-%Unlike for a summer school, we have generally selected for a diverse set of participants in order to be able to efficiently pair up novices and experts in projects.
-%Finally, we also include demographic information with two main goals: firstly, representation has been proven to be an effective way to foster an inclusive environment where minority participants feel comfortable to participate and take risks.
-%Secondly, we would like to boost the participation of minority researchers in computational astrophysics, and our hack week model provides a useful framework for doing so.
+Biased participant sets may still result from selection procedures that fail to include crucial categories.
+For example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority, and impossible to correct gender bias in the pool of applicants, if no information is available about the gender of participants.

### 2017-08-21 sections/participants.tex
More accurate description of what Entrofy does.
Addresses https://github.com/uwescience/HackWeek-Writeup/pull/61#pullrequestreview-57370470

@@
 
-On solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.
-The algorithm maximizes the variance of participants on specified dimensions (e.g., career stage, geographic location, etc.) to meet pre-set fractions set by the organizers.
+One solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.
+The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
+For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.

### 2017-08-21 sections/participants.tex
Add a comma.

@@
 
-A thoughtful robust, and effective participant selection procedure is of crucial importance to the success of a hack week and may require both careful engineering and an in-depth discussion among the organizers.
+A thoughtful, robust, and effective participant selection procedure is of crucial importance to the success of a hack week and may require both careful engineering and an in-depth discussion among the organizers.
 At conferences, researchers typically report recent results to the community and network with other scientists, largely from their discipline.

### 2017-08-21 sections/results.tex
Add a comment that clarifies that we have IRB approval.

@@
 We report the outcomes of astro-, geo- and neuro- hack weeks from 2016.
-In the results from post-workshop surveys, we focus on attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
+In the results from post-workshop surveys (collected with the approval of the Institutional Review Boards at UW, NYU and UC Berkeley, and with participants' informed consent), we focus on attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
 

### 2017-08-21 sections/design.tex
shorten design section

@@
 
-Design considerations depend very strongly on goals.
-In general, longer events allow for a larger taught component, more ambitious projects and especially for cross-disciplinary events are more likely to provide enough time for different groups to effectively communicate across barriers of professional terminology.
-On the other hand, because of their large participatory component, hack weeks tend to be exhausting, and long events may lead to fatigue among attendees, where active participation and measurable outcomes may drop sharply in the latter days of the workshops.
-%This is particularly important if---as is often the case---participants continue participating in workshop-relevant activities even after the official component ends for the day.
-Ways to mitigate these effects include capping taught components during the day, providing a reasonable clear schedule, and limiting parallel components to avoid decision fatigue.
-%At Astro Hack Week, we have limited the scheduled taught components to no more than three out of nine hours for each of the five workshop days, and unscheduled (shorter) tutorials to a maximum of two per day.
-%Even so, in our surveys, participants occasionally report concerns about choosing to attend tutorials versus working on projects.
+Scheduling, group size and venue are important design considerations contributing to the success of a hack week.   
+Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges. 
+By spending more time together, participants are more likely to overcome barriers of professional terminology.
+On the other hand, events that are too long may lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
+A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and open project work.
 
-Aside from time, space plays a particularly important role in facilitating a successful hack week.
-Universities in general provide a convenient venue with existing structures to facilitate hack weeks (e.g. access to scientific publications, institutional support including staff and funding).
-On the other hand, most spaces in universities are designed for lectures, which are diametrically opposed to unconference-style events.
-Thus, finding an appropriate physical space within a university may be a challenge.
-However, with the rise of active learning as a preferred teaching methods, traditional lecture spaces are transforming into more flexible spaces that are generally appropriate for hack weeks.
-As a general rule, the smaller a hack week's emphasis on a taught component, the more flexible the space has to be, with ample opportunity for re-configuration.
-%At Astro Hack Week, we have found it beneficial to spend one day mid-week at a different location (e.g. a company headquarter) to engage participants and break the routine.
-%It is worth noting that collaborative spaces for non-traditional workshops are becoming more prevalent in academia (some attached to universities, some not) and may provide support and infrastructure that a traditional university location may not.
+A hackweek requires a flexible workspace environment with ample opportunity for re-configuration. 
+Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
+Workspaces must also accommodate interactive project work where small teams can gather and work together uninterrupted by other groups.    
+Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
+However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
+Fortunately active learning teaching methods are creating new environments that foster interactivity through modified seating arrangements and integration of computer monitors at group workstations.
 
 Another important design consideration is group size.
-%With all previous hack weeks severely oversubscribed, it seems natural to simply admit more participants.
-%However, this can counter the ideals of the workshop, in particular when building a community is one of its stated goals.
-If the group is too large, participants are unlikely to even meet each other, and workshop cohesion may be lost as the workshop fractures into smaller groups, often among participants who already know each other.
-This may inhibit knowledge transfer by clustering participants into small in-groups.
-%Additionally, it is likely that the number participant-led components in the schedule may increase with workshop size.
-%While generally desirable, a programme that is too crowded may lead to fatigue
-On the other hand, if the size of the workshop is too small, it is unlikely to achieve the desired diversity among participants to foster new collaborations across sub-fields and disciplines.
+If the group is too large, chances for random participant exchanges are reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
+If the group is too small, it is unlikely to achieve the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines.
 In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.
 
-As mentioned above, the balance between pedagogy and working depends both on the goals of the workshop and the topics around which the workshop is organized.
+As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
 If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
-In communities where a shared understanding exists, tutorials can focus on more advanced or innovative topics, and less time may be allocated for them, leaving more time for active participation.
-%In astronomy, a relatively small field, even students generally share a common knowledge base and have rudimentary knowledge about the types of data used and the challenges associated with each.
-%Thus, we have focused the taught component of Astro Hack Week less on domain-specific knowledge, and instead offered tutorials in topics from domain-adjacent fields like statistics and computer science that attendees are unlikely to have encountered in their regular education.
+In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.
 
-Hack week outcomes, in turn, depend strongly on participants and are often a function of their interests and seniority.
-some attendees arrive with the stated goal of writing a specific scientific article, usually more advanced participants with significant pre-knowledge of both hackathons and their topic of interest.
-Many attendees arrive with the plan to learn a specific topic (such as machine learning) or bring a specific data set they believe the new knowledge may be applicable to.
+Hack week outcomes depend strongly on the interest and engagement of participants.
+Some attendees, usually those with a strong background in hackathons and their specific topic of study, arrive with the goal of writing a scientific article.
+Other attendees plan to learn a specific topic, such as machine learning, or to analyze a specific data set that relates to the tools covered in the tutorials.
 This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
-%It is worth noting that while a scientific paper need not be the stated goal of a hack (and is unlikely to be completed in the short time allocated in any case), results may still be published as short reports or unconference proceedings.
-%For example, Neurohack week provides a venue for participants to publish a short (two-page) ''project report`` summarizing the hack that participants did during the week of NHW.
-%Similarly, Python in Astronomy gathers all documents produced during the workshop (unconference transcripts, talk summaries, descriptions of sprint and hack projects) into citeable unconference proceedings.
+This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.

### 2017-08-21 supplementary_material.tex
Removed box from impostor syndrome section
move imposter syndrome box here for now

@@
 
-\subsection{Box 1: Impostor Syndrome}
-
-The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \cite{Clance1978-ef}.
-Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
-This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
-The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
-This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
-Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
- Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
-We are working within our hack weeks to mitigate IS using various techniques.
-With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
-More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g.
-one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
-%I think it helped ease the feeling!!``}).
-Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
-
 \subsection{What to do after}
@@
 For this purpose, it is useful to keep a way to contact previous participants, via a mailing list or real-time messaging services like \textit{Matrix} or \textit{Slack}.
+
+\subsection*{Box 1: Impostor Syndrome}
+
+The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
+Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \cite{Clance1978-ef}.
+Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
+This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
+The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
+This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
+Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
+ Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
+We are working within our hack weeks to mitigate IS using various techniques.
+With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
+More generally, being open about the presence and prevalence of IS can help participants feel more at ease.
+Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
 

### 2017-08-23 sections/results.tex
some changes

--- 2017-08-18 sections/results.tex
@@
+\section*{Measures of success}
+
+Objectively measuring the success of a hack week is complicated by the diversity of goals of both organizing committee and attendees.
+The participant-driven, open-ended format of a hack week facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
+In addition, hack week goals such as networking and community building are difficult to measure and may evolve after the event has ended. 
+For more project-driven events, one key metric is the number of publications that result from hack week projects.
+However, this implies a very narrow definition of success in line with standard academic performance indicators, and may be hard to collect in practice, if participants do not explicitly include a reference to the event in their publication.
+Another complementary indicator is the quantity of code written and committed to a public code repository (e.g. GitHub).
+This could be defined in lines of code committed, the number of new open-source projects spawned, and the number of new developers contributing to open-source projects.
+This assumes that participants work largely in the open during a hack week, and that most projects have a strong programming component.
+Again, this may not be true for some hacks (e.g. tutorials, other written text, visualizations), and focuses entirely on material outcomes.
+These measures ignore learning, community building as well as networking outcomes, which are often equally important objectives.
+One possible approach here is to ask participants to self-report their success via a post-workshop survey.
+It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
+%This would then provide a somewhat objective measure of the impact the hack week has had.
+%In practice, this has not been done for any of the hack weeks conducted thus far.
+For practical purposes, we recommend an approach that combines all of these metrics: publications, new code and projects generated, self-reported learning outcomes, and survey results.
+%Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
+
+If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
+Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent.
+In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating.
+%It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
+
 \section*{Results}
 
-Measuring the success of a hack week objectively is complicated by the variety of objectives that a hack week might have (see \ref{sec:why}).
-Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
-One key metric is the number of publications that result from hack week projects, but this is a fairly narrow definition of success, in line with standard academic performance indicators.
-Assuming that participants work largely in the open during a hack week, and that most projects have a strong programming component another indicator of success is the activity of participants in terms of code written and committed to a public code repository.
-Still, these measures ignore learning, community-building as well as networking outcomes.
-These outcomes can be assessed through post-workshop surveys.
-%It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
-%This would then provide a somewhat objective measure of the impact the hack week has had.
-%In practice, this has not been done for any of the hack weeks conducted thus far.
-Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see \ref{box:outcomes})
-%Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
-
-%If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
-%Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent.
-%In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating.
-%It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
-
-We report the outcomes of astro-, geo- and neuro- hack weeks from 2016.
-In the results from post-workshop surveys, we focus on attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
+In the following we report outcomes of previous events held in 2016 in three different fields (neuroscience, geoscience and astrophysics) and participants attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
+Because all three events are less than a year in the past, evaluating long-term outcomes (number of publications, citations, career outcomes) will require a longer baseline.
+There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
+[NOTE: HERE IS WHERE WE SHOULD REFER TO THE BOXES ONCE THEY ARE MERGED.]
+We asked participants of all three hack weeks to complete surveys to validate the hack week's success.
+%All surveys included a number of questions allowing participants to self-report their attitudes about the hack week in general as well as questions specifically designed to probe outcomes directly related to the hack week goals.
 
@@
 Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
-
 \begin{figure*}[htbp]
@@
 
-Because all three events are relatively recent, it is still early to evaluate long-term outcomes, such as publications and collaborations resulting from these events.
-There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
-For specific examples, see also \ref{box:outcomes}.
-
 \subsection*{Box 2: Examples of Hack Week Outcomes}
-\label{box:outcomes}
 \subsubsection*{Example 1: Astro Hack Week}

### 2017-08-26 sections/intro.tex
some changes
Added citations for SWC + DC #27

@@
 At the same time, formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
-A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998\footnote{\url{https://f1000research.com/articles/3-62/v2}}.
+A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998 \cite{b:wilson-swc-lessons-2016,teal2015data}.
 At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.

### 2017-08-29 sections/why.tex

@@
 \begin{itemize}
-\item{\textit{Education and training}: %Some hack weeks are more focused on education than others (see Figure 1).
-Even those hackathons not explicitly focused on skills-development will include extensive discussions on reproducible and open science practices that are valuable to participants.
-%Discussing these topics in a diverse group setting helps to ingrain within participants the basic conventions of the broader open source community.
-Participants can then serve as ''ambassadors`` of these practices and foster their adoption by other researchers in the field.
-Furthermore, the kind of co-working that is inherent to the hack week format facilitates lateral knowledge transfer, where participants learn from each other about practices and tools of the trade.}
-%Often these tools are not described in research products such as papers, or even in the software implementations.}
-\item{\textit{Tool development}: hack weeks present an opportunity for computational tool developers to get together with users of these tools, and to critically evaluate the use of the tools in addressing particular scientific questions.}
-\item{\textit{Community building}: a hack week is an opportunity to galvanize a community within a field.% that cares about computational issues, and relates to specific open-source software projects.
-For example, because of its focus on free and open-source tools, Neurohack week helps galvanize the Neuroimaging in Python (http://nipy.org) community.
-It also allows researchers who may otherwise be isolated within their departments and subfield on account of their computational specialization to come together and exchange experiences.}
-%As one participant at Astro Hack Week expressed in the survey, \textit{''The main thing I took from Astrohack week is the fact that I'm not alone``}.}
-\item{\textit{Interdisciplinary research}: a short time-bounded event is an excellent opportunity to experiment with concepts, questions and methods that traverse boundaries within a discipline (e.g.\ between researchers using different methods, or studying different systems, or experimental models), as well as across disciplines.
-Despite the impact of interdisciplinary research \cite{Hall2012-hi}, these experiments are often discouraged in traditional disciplinary research, because of the difficulty to achieve results in both disciplines, and because of the risks inherent in taking a less trodden path\footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.}
-%This is despite the fact that interdisciplinary research can also be very impactful \cite{Hall2012-hi}.
-%A diversity of research backgrounds, and interactions with scientists from other disciplines, further facilitates these experiments.}
+\item{\textit{Education and Training}: %Some hack weeks are more focused on education than others (see Figure 1).
+While some hack weeks are focused more on education than others (Figure 1), there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
+Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices in their respective fields.
+This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
 
-\item{\textit{Recruitment and networking}: hack weeks often bring together research, government and industry communities, providing an opportunity for participants to learn about each other's interests and their abilities in close quarters.
-This kind of extended opportunity for interaction and ''on-the-job`` testing is considered an excellent form of evaluation of prospective job candidates [CITATION NEEDED].
-Conversely, participants have an opportunity to gain exposure to potential career opportunities in communities other than their own.}
-%As an example, one of the participants in Neurohack week 2016 will be joining the eScience Institute as a post-doc in Fall of 2017 (and as an instructor in Neurohack week 2017).}
-%\item{\textit{It's fun}: the opportunity to focus for a week on a slightly more whimsical and adventurous sort of scientific activities can be both fruitful, and enjoyable.
-%Even aside from the prospect of interdisciplinary research, it can provide a laboratory for more experimental approaches and high-risk projects than scientist will take on in their day-to-day research.}
+\item{\textit{Tool Development}: Hack weeks present an opportunity for scientific software developers to meaningfully engage with users and critically evaluate applications to particular scientific issues.}
+
+\item{\textit{Community Building}: Hack weeks provide a tremendous opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
+These events allow computationally minded researchers to break from the isolation of their academic departments and build connections and spark new collaborations.}
+
+\item{\textit{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
+Despite the fact that such interdisciplinary experiments are highly impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.
+
+\item{\textit{Recruitment and Networking}: Hack weeks are often a melting pot of participants from academia, government, and industry and provide numerous opportunities for networking.
+Close collaboration in diverse groups exposes skills that might be suitable for careers outside of one's narrow domain.}
+
+\item{\textit{It's fun}: Hack weeks provide a respite from day-to-day research activities and provide a low-stress venue to learn new skills and attempt high-risk projects.}
 \end{itemize}

### 2017-08-29 sections/why.tex

@@
 \item{\textit{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
-Despite the fact that such interdisciplinary experiments are highly impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.
+Despite the fact that such interdisciplinary experiments are highly impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}}.
 
@@
 \item{\textit{It's fun}: Hack weeks provide a respite from day-to-day research activities and provide a low-stress venue to learn new skills and attempt high-risk projects.}
+
 \end{itemize}

### 2017-08-29 sections/why.tex
Various cleanups: figures embedded, references tidied. Authorship order.

--- 2017-08-29 sections/why.tex
@@
 \item{\textit{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
-Despite the fact that such interdisciplinary experiments are highly impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \footnote{\url{https://www.ncbi.nlm.nih.gov/labs/articles/12970550} and \url{https://www.researchgate.net/publication/8126355_EDUCATION_Risks_and_Rewards_of_an_Interdisciplinary_Research_Path}}.
+Despite the fact that such interdisciplinary experiments are highly impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \cite{Sung2003-go, Rhoten2004-fk}}.
 

### 2017-08-30 sections/results.tex

--- 2017-08-21 sections/results.tex
@@
 
-Measuring the success of a hack week objectively is complicated by the variety of objectives that a hack week might have (see \ref{sec:why}).
+Measuring the success of a hack week objectively is complicated by the variety of objectives that a hack week might have (see above).
 Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
 \begin{center}
-\includegraphics[\textwidth]{fig/eval_techskills.png}
-\includegraphics[\textwidth]{fig/eval_collab.png}
-\includegraphics[\textwidth]{fig/eval_openscience.png}
+\includegraphics[width=12cm]{fig/eval_techskills}
+\includegraphics[width=12cm]{fig/eval_collab}
+\includegraphics[width=12cm]{fig/eval_openscience}
 \caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
-For specific examples, see also \ref{box:outcomes}.
+For specific examples, see also \emph{Box 1}.
 
-\subsection*{Box 2: Examples of Hack Week Outcomes}
+\subsection{Box 1: Examples of Hack Week Outcomes}
 \label{box:outcomes}

### 2017-08-30 sections/what.tex

@@
 %The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon.
-Following the typology of \cite{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
+Following the typology of Drouhard et al.
+\cite{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
 Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \cite{Drouhard2017}.

### 2017-08-30 paper.tex
Various cleanups: figures embedded, references tidied. Authorship order.

@@
 \usepackage{amssymb}	% Extra maths symbols
-\usepackage{natbib}
+%\usepackage{natbib}
 
@@
 
-\author{Jake VanderPlas, Karthik Ram, Daniela Huppenkothen, David W. Hogg, Ariel Rokem, Anthony Arendt \& Tal Yarkoni}
+\author{Daniela Huppenkothen, Anthony Arendt, David W. Hogg, Karthik Ram, Jake VanderPlas, Tal Yarkoni, \& Ariel Rokem}
+
+\makeatletter
+\let\saved@includegraphics\includegraphics
+\AtBeginDocument{\let\includegraphics\saved@includegraphics}
+\renewenvironment*{figure}{\@float{figure}}{\end@float}
+\makeatother
 
@@
  \item UC Berkeley
+ \item UT Austin
 \end{affiliations}
@@
 
+\section{References}
 \bibliographystyle{naturemag}
@@
 \begin{addendum}
- \item{Anyone thanked here: *Feel free to move up to the Author list!* It is a pleasure to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB.
+ \item{The authors would like to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB, Stuart Geiger for helping to formulate the questionnaires that served as the basis for the results presented here.
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.

### 2017-08-30 sections/intro.tex
Various cleanups: figures embedded, references tidied. Authorship order.

@@
 Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, and more project-focused than pedagogical meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\footnote{\url{https://gigascience.biomedcentral.com/articles/10.1186/s13742-016-0121-x}}, and a hackathon associate with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon associated with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
 In broad-brush, pedagogically-focused workshops and summer schools tend to follow a classic academic model where novices learn a skillset from experts, while project-focused workshops focus on people collaboratively exercising the skillset they already have.
@@
 \begin{center}
-\includegraphics[width=9cm]{fig/HackSpectrum.png}
+\includegraphics[width=9cm]{fig/HackSpectrum}
 \caption{Comparison of Extracurricular Workshop Models}
@@
 
-As of the publication of this paper, we have run five such hack week events: three focused on Astronomy, one focused on Neuroscience, and one focused on Geoscience.
+As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
 Below we will share some of the philosophy behind the hack week model, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.

### 2017-08-30 sections/abstract.tex
wordsmithing / shortening

@@
 This gives rise to the need for scientific communities to adapt on shorter time scales than traditional university curricula allow for, and therefore requires new modes of knowledge transfer.
-At the same time, as disciplines become more specialized, data analysis problems often tend to be fairly universal.
-It is thus desirable to foster exchange of ideas and computational workflows across disciplines.
-Among the different approaches put forward toward addressing these issues, hack weeks have in recent years emerged as both an effective tool for training researchers in modern data analysis workflows, and in fostering exchange between subdisciplines within and between scientific fields.
-While interpreted differently in the various disciplines where they have been implemented, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and free-form project work.
+The universal applicability of data science tools to a broad range of problems has generated new opportunities to foster exchange of ideas and computational workflows across disciplines.
+In recent years, hack weeks have emerged as an effective tool for fostering these exchanges by providing training in modern data analysis workflows.
+While there are variations in hack week implementation, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and project work in a collaborative environment.
 In this paper, we present the concept of a hack week in the larger context of scientific meetings and point out similarities and differences to traditional conferences.

### 2017-08-30 sections/intro.tex
wordsmithing / shortening

@@
 As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
-At the same time, formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
-A well-known example is the Software Carpentry and Data Carpentry workshops, whose interdisciplinary workshops in research computing skills have reached more than 16,000 participants since its inception in 1998 \cite{b:wilson-swc-lessons-2016,teal2015data}.
+Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
+Well-known examples are the Software add Data Carpentry workshops providing training in research computing skills through a volunteer instructor program  \cite{b:wilson-swc-lessons-2016,teal2015data}.
 At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.
 Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, and more project-focused than pedagogical meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon associated with the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-In broad-brush, pedagogically-focused workshops and summer schools tend to follow a classic academic model where novices learn a skillset from experts, while project-focused workshops focus on people collaboratively exercising the skillset they already have.
-A disadvantage of the summer school model is that it can tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
-A disadvantage of the hackathon model is the common perception (whether accurate or not) that the week is designed for experts in technical tools, which may discourage others from attending.
+Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+Generally, pedagogically-focused events follow a classic academic model where novices learn new skills from experts, while project-focused workshops emphasize collaborative activities using existing skills.
+A disadvantage of the pedagogical model is that it can tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
+A disadvantage of the project model is the common perception that the week is designed for technical experts, which may discourage others from attending.
 In 2014, we started Astro Hack Week to try to fill the gaps between these models.
-The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for collaborative hacks or creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
 

### 2017-08-30 sections/what.tex
wordsmith / shorten

@@
 In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
-In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to 'learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
+In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
 With this in mind, hackathons around scientific topics, designed to foster collaboration \cite{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \cite{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.
 
-%The recent surge in popularity of these events has resulted in a broad spectrum of ways to define the hackathon.
 Following the typology of Drouhard et al.
@@
 Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \cite{Drouhard2017}.
-''Catalytic`` hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
-''Contributive`` hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
-Finally, ''Communal`` hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.
+``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
+``Contributive'' hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
+Finally, ``Communal'' hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.
 
@@
 
-We note that the terminology for these events is constantly evolving, and that the ''hackathon`` concept may have implicit connotations that are disfavored in some communities.
-One criticism of hackathons is that they propel the ''geek`` stereotype and may present a barrier to creating an inclusive working environment, especially for women \cite{Decker2015}.
+We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
+One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
 Another problem is the competitive atmosphere of many industry hackathons, where teams actively compete for prizes.
 Because certain groups are more risk-averse than others, and risk-aversion traces demographics like gender, we urge caution when contemplating making an event competitive.
-Other related nomenclature includes the ''unconference`` which is also a time-bounded event gathering together a specific community, but these often have minimal structure and do not necessarily have a focus on software.
-The ''sprint`` or ''scrum`` label typically refers to an event focused on rapid software development on a specific set of code.
-%This may be a component of our hack week model but it ignores the important pedagogical components.

### 2017-08-30 sections/participants.tex
shorten / restructure

--- 2017-08-21 sections/participants.tex
@@
 
-A thoughtful robust, and effective participant selection procedure is of crucial importance to the success of a hack week and may require both careful engineering and an in-depth discussion among the organizers.
-At conferences, researchers typically report recent results to the community and network with other scientists, largely from their discipline.
-In contrast, traditional summer schools aim for senior researchers to transfer knowledge about a particular subject or field to novice researchers, primarily graduate students and early postdoctoral fellows.
-Hack weeks differ from both of these in that knowledge transfer occurs across many levels of seniority and across disciplinary boundaries, as well as novelty of the topics discussed,ranging from established theory to new findings.
-In addition, the content is primarily participant-driven, and much of it is generated during the event itself.
-This mean that a successful hack week requires participants who are both interested in participation, and also feel comfortable participating in discussions and taking risks.
-Participant selection also addresses other objectives of a hack week such as increasing minority participation and gender diversity, knowledge transfer between academia and industry, etc.
-These factors can also be addressed by thoughtful participant selection.
-To meet these goals, a thoughtful participant selection process needs to be as transparent to the applicants as possible, and the organizing committee should be held accountable for their performance in selecting participants.
+Hack weeks differ from traditional conferences or summer schools in that knowledge transfer occurs across many levels of seniority, disciplinary boundaries, and novelty of the topics discussed.
+In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from participants.
+Therefore in order to maximize learning outcomes and the liklihood for collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
+In our experience, a participant group that is diverse across categories of diversity, gender, discipline and career track helps to ensure we meet these objectives. 
+To achieve this diversity, we advocate for a selection process that is as transparent as possible, enabling particiants to hold organizers accountable for their selection decisions.
 Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
-One way to maximize transparency in the selection process is to remove the human element from the decision making as far as possible and transfer some of these tasks to a well-designed algorithm.
-In this context, well-designed means \cite{o2017weapons}: (1) interpretable, i.e.
-the algorithm, its parameters and outputs can be understood by humans, (2) openly available, and thus the decision process can be inspected, (3) the algorithm itself does not perpetuate intrinsic biases in the data (though we will show below that an algorithm itself does not entirely remove the requirement of humans to be aware of their biases).
-In practice, most selection procedures follow two separate steps.
-In the first step, the merit of a candidate must be assessed: does the candidate fulfill the requirements posed by the goals of the workshop?
-This first step is difficult to automate, because it requires judgment calls, often based on long-form answers.
-To avoid implicit/unconcsious bias, in the past, we have sometimes performed this step by blinding ourselves to a candidate's other attributes (including name and other personal information), and assess their candidacy based solely upon key questions asked specifically for this purpose.
+One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to transfer some steps to an algorithm that is easily interpreted, openly available, and immune to the perpetuation of intrinsic biases.
+We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.
+We try to minimize bias in this step by blinding ourselves to a candidate's other attributes, including name and other personal information, and assess their candidacy based soley on questions asked specifically for this purpose.
 When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.

### 2017-08-30 sections/results.tex
wordsmith / shorten

@@
 
-We report the outcomes of astro-, geo- and neuro- hack weeks from 2016.
-In the results from post-workshop surveys (collected with the approval of the Institutional Review Boards at UW, NYU and UC Berkeley, and with participants' informed consent), we focus on attitudes towards open science and reproducibility, their own learning outcomes, as well as community building and collaborations.
-
-We find that at all three hack weeks, participant self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 2, left panel).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (see Figure 2, middle and right panels).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks from 2016 (collected with the approval of the Institutional Review Boards at UW, NYU and UC Berkeley, and with participants' informed consent), we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 2, left panel).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure 2, middle and right panels).
 %While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
 %Another important goal of the hack weeks was centred on community building and fostering collaborations.
-We find that overall, the majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel).
+The majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel).
 This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 3, right panel).
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ''somewhat agree``, ''agree`` and ''strongly agree``; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories.
+We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories.
 This is expected: participants new to the field may participate to learn rather than to teach.
 
-Reproducibility and open science have recently drawn much attention from the scientific community, and all three hack weeks have been working to promote positive attitudes towards reproducibility and open science in their respective fields.
-We find that hack weeks have been largely successful at encouraging open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 4, left panel) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 4, middle panel).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 4, left panel) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 4, middle panel).
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.

### 2017-09-01 sections/why.tex

@@
 \begin{itemize}
-\item{\textit{Education and Training}: %Some hack weeks are more focused on education than others (see Figure 1).
+\item{\textit{Education and Training}: 
 While some hack weeks are focused more on education than others (Figure 1), there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
@@
 
-It is worth noting that the reasons for participants to attend a hack week are as diverse as the goals that the organizing committee explicitly states.
-This is largely a function of the diversity of the participants (see section below), and should be something that organizers design for.
-Beginner participants may attend primarily to learn a new technique or method, while more experienced participants attend to gain more experience mentoring and teaching.
-Some researchers may come with particular projects and collaborators in mind, while others come with a focus on learning and with no explicit plan for the project-based work.
-%While it is very hard to design a workshop that is universally useful to all participants, in practice many researchers attend with the understanding that some components will be more congruent with their own goals than others, but also bring an open mind and a willingness to learn.
-%In addition, the diversity of goals can effectively be a strength of the workshop, if the organizers can facilitate matching up participants with complementary goals.
+It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event. 
+Beginner participants may attend primarily to learn a new technique and may not have an explicit plan for project based work, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress.

### 2017-09-04 paper.tex
Move Tal from authors to acknoweldgements.

@@
 
-\author{Daniela Huppenkothen, Anthony Arendt, David W. Hogg, Karthik Ram, Jake VanderPlas, Tal Yarkoni, \& Ariel Rokem}
+\author{Daniela Huppenkothen, Anthony Arendt, David W. Hogg, Karthik Ram, Jake VanderPlas, \& Ariel Rokem}
 
@@
  \item{The authors would like to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB, Stuart Geiger for helping to formulate the questionnaires that served as the basis for the results presented here.
+Thanks to Tal Yarkoni for advice regarding automated selection procedures.
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.
-Neurohack week is supported through a grant from the National Institute for Mental Health (1R25MH112480).
+Neuro hack week is supported through a grant from the National Institute for Mental Health (1R25MH112480).
 Daniela Huppenkothen is partially supported by the James Arthur Postdoctoral Fellowship at NYU.}

### 2017-09-19 sections/intro.tex
Some cleanups and small things.

@@
 Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
-Well-known examples are the Software add Data Carpentry workshops providing training in research computing skills through a volunteer instructor program  \cite{b:wilson-swc-lessons-2016,teal2015data}.
+Well-known examples are the Software and Data Carpentry workshops providing training in research computing skills through a volunteer instructor program  \cite{b:wilson-swc-lessons-2016,teal2015data}.
 At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.
-Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, and more project-focused than pedagogical meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have started in conjunction with conferences, such Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, as well as a variety of project-focused (rather than pedagogical) meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
+Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
 Generally, pedagogically-focused events follow a classic academic model where novices learn new skills from experts, while project-focused workshops emphasize collaborative activities using existing skills.
@@
 A disadvantage of the project model is the common perception that the week is designed for technical experts, which may discourage others from attending.
-In 2014, we started Astro Hack Week to try to fill the gaps between these models.
+In 2014, we initiated an alternative model of ``Hack Weeks'' that try to fill the gaps between these models.
 The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.

### 2017-09-19 sections/what.tex
Some cleanups and small things.

@@
 
-Following the typology of Drouhard et al.
-\cite{Drouhard2017}, we assert that core elements of all hackathons include opportunities for networking, the strengthening of social ties and the building of community connections, both within and across disciplines.
-Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \cite{Drouhard2017}.
+However, in addition to these goals, core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
+Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \cite{Drouhard2017}:
 ``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
@@
 
-Our past hack week events most closely follow the communal hackathon model as it applies to scientific communities of practice.
+Our past hack week events follow most closely the communal hackathon model, as it applies to scientific communities of practice.
 Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.

### 2017-09-19 sections/intro.tex
One more small thing.

@@
 In 2014, we initiated an alternative model of ``Hack Weeks'' that try to fill the gaps between these models.
-The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+These are week-long events that combine pedagogy (often focused on statistical and computational techniques) with room for hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
 

### 2017-09-19 sections/what.tex
Correct the historical record (comment from BFG).

@@
 Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
-Hackathons have historically focused on software development and technology design as a way to motivate innovation within industry.
+Hackathons originated from the open-source software movement, and have historically focused on software development and technology design as a way to motivate innovation, eventually being adopted also within the technology industry.
 In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.

### 2017-09-19 sections/what.tex
Attempts to address comment made by BFG regarding hackathon criticsms.

@@
 One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
-Another problem is the competitive atmosphere of many industry hackathons, where teams actively compete for prizes.
-Because certain groups are more risk-averse than others, and risk-aversion traces demographics like gender, we urge caution when contemplating making an event competitive.
+This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter}.
+Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes that are awarded prizes, academic hackathons, and specifically the hack weeks that we have organized events are not explicitely competitive

### 2017-09-19 sections/intro.tex
Another small wording thing.

--- 2017-09-19 sections/intro.tex
@@
 In 2014, we initiated an alternative model of ``Hack Weeks'' that try to fill the gaps between these models.
-The hack week model combines pedagogy (often focused on statistical and computational techniques) with room for hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+These are week-long events that combine pedagogy (often focused on statistical and computational techniques) together with time for hacks and creative projects, and with the goal of encouraging collaboration and learning among people at various stages of their career.
 

### 2017-09-20 paper.tex
Remove cruft.
Some cleanups and small things.
Fix typo.

@@
 
-\author{Daniela Huppenkothen, Anthony Arendt, David W. Hogg, Karthik Ram, Jake VanderPlas, \& Ariel Rokem}
+\author{Daniela Huppenkothen$^1$, Anthony Arendt$^2$, David W. Hogg$^1$, Karthik Ram$^3$, Jake VanderPlas$^2$, and Ariel Rokem$^2$}
 
@@
 \begin{affiliations}
- \item UW Seattle
- \item NYU
- \item UC Berkeley
- \item UT Austin
+ \item Center for Data Science, New York University
+ \item The University of Washington eScience Institute
+ \item The Berkeley Institute for Data Science, UC Berkeley
 \end{affiliations}
@@
 \begin{addendum}
- \item{The authors would like to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB, Stuart Geiger for helping to formulate the questionnaires that served as the basis for the results presented here.
-Thanks to Tal Yarkoni for advice regarding automated selection procedures.
+ \item{The authors would like to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB, Stuart Geiger for helping to formulate the questionnaires that served as the basis for the results presented here, Brittany Fiore-Gartland for comments on the manuscript and Tal Yarkoni for advice regarding automated selection procedures.
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.

### 2017-09-20 sections/participants.tex

@@
 Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
-One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to transfer some steps to an algorithm that is easily interpreted, openly available, and immune to the perpetuation of intrinsic biases.
+One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to transfer some steps to an algorithm that is easily interpreted, openly available, and can be designed to counter the perpetuation of intrinsic biases.
 We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.

### 2017-09-20 sections/what.tex

@@
 This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter}.
-Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes that are awarded prizes, academic hackathons, and specifically the hack weeks that we have organized events are not explicitely competitive
+Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitely competitive

### 2017-09-20 sections/intro.tex
Changes in response to comments by Ariel's neuroscience colleague

@@
 As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
-Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, third-party workshops.
+In addition, the recent reproducibility crises in several scientific fields [REF] has led to the growing realization that improving awareness of open science and reproducibility as well as practical skills in making research reproducible is essential to scientific progress \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
+Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts such as workshops (an overview and typography of such efforts in the data science context can be found in \cite{demasi2017}).
 Well-known examples are the Software and Data Carpentry workshops providing training in research computing skills through a volunteer instructor program  \cite{b:wilson-swc-lessons-2016,teal2015data}.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=9cm]{fig/HackSpectrum}
-\caption{Comparison of Extracurricular Workshop Models}
-\label{fig:hackspectrum}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=9cm]{fig/HackSpectrum}
+%\caption{Comparison of Extracurricular Workshop Models}
+%\label{fig:hackspectrum}
+%\end{center}
+%\end{figure}
 

### 2017-09-20 sections/results.tex
Changes in response to comments by Ariel's neuroscience colleague

@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks from 2016 (collected with the approval of the Institutional Review Boards at UW, NYU and UC Berkeley, and with participants' informed consent), we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 2, left panel).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure 2, middle and right panels).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks from 2016 (collected with the approval of the Institutional Review Boards at UW, NYU and UC Berkeley, and with participants' informed consent), we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 1, top panel, left).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure 1, top panel, right and middle).
 %While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
 %Another important goal of the hack weeks was centred on community building and fostering collaborations.
-The majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 3, middle panel).
-This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 3, right panel).
+The majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 1, middle panel, middle).
+This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 1, middle panel, right).
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 3, left panel), though responses are not as unequivocal as they are in some of the other categories.
+We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 1, middle panel, left), though responses are not as unequivocal as they are in some of the other categories.
 This is expected: participants new to the field may participate to learn rather than to teach.
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 4, left panel) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 4, middle panel).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 1, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 1, bottom panel, middle).
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
@@
 The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects, role models of successful projects, and an environment encouraging scientific risk taking.
+Astro Hack Week enabled participants to seed a new collaboration around a software project needed by the larger community.
 Since its beginnings at Astro Hack Week, Stingray has matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.

### 2017-09-20 sections/results.tex
added acknowledgment, fixed references and added sentence about long-term outcomes

@@
 
-Because all three events are relatively recent, it is still early to evaluate long-term outcomes, such as publications and collaborations resulting from these events.
+From the very nature of the activities that we encourage in hack weeks, participants in these events produce digital records of their research online.
+This means that it will be fairly straightforward to evaluate the long-term impact of these activities on participants' productivity (e.g., through contributions to open-source software) in the future.
+Because all three events are relatively recent, it is still early to evaluate such long-term outcomes, as well as others including publications and collaborations resulting from these events.
 There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.

### 2017-09-20 sections/what.tex

@@
 Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
-Hackathons originated from the open-source software movement, and have historically focused on software development and technology design as a way to motivate innovation, eventually being adopted also within the technology industry.
+Hackathons originated from the open-source software movement, and have historically focused on software development, particularly in the form of coding sprints, and technology design as a way to motivate innovation, eventually being adopted also within the technology industry.
 In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
@@
 Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
-Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure 1).
+Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
 For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.

### 2017-09-20 sections/why.tex

@@
 \item{\textit{Education and Training}: 
-While some hack weeks are focused more on education than others (Figure 1), there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
+While some hack weeks are focused more on education than others, there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
 Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices in their respective fields.

### 2017-09-25 sections/intro.tex
Made figure two-column, fixed typo

@@
-\section*{Introduction}
+%\section*{}
 \label{sec:introduction}
@@
 As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
-In addition, the recent reproducibility crises in several scientific fields [REF] has led to the growing realization that improving awareness of open science and reproducibility as well as practical skills in making research reproducible is essential to scientific progress \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
+In addition, the recent reproducibility crises in several scientific fields has led to the growing realization that improving awareness of open science and reproducibility as well as practical skills in making research reproducible is essential to scientific progress \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
 Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts such as workshops (an overview and typography of such efforts in the data science context can be found in \cite{demasi2017}).

### 2017-09-25 sections/results.tex
Made figure two-column, fixed typo
Not sure we can have a box in the new format
Added Figure 1 back, fixed links to figures

@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks from 2016 (collected with the approval of the Institutional Review Boards at UW, NYU and UC Berkeley, and with participants' informed consent), we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 1, top panel, left).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure 1, top panel, right and middle).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks from 2016, we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, top panel, left).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure \ref{fig:survey}, top panel, right and middle).
 %While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
 %Another important goal of the hack weeks was centred on community building and fostering collaborations.
-The majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure 1, middle panel, middle).
-This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure 1, middle panel, right).
+The majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure \ref{fig:survey}, middle panel, middle).
+This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure  \ref{fig:survey}, middle panel, right).
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure 1, middle panel, left), though responses are not as unequivocal as they are in some of the other categories.
+We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, middle panel, left), though responses are not as unequivocal as they are in some of the other categories.
 This is expected: participants new to the field may participate to learn rather than to teach.
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure 1, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure 1, bottom panel, middle).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure \ref{fig:survey}, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure \ref{fig:survey}, bottom panel, middle).
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[h!]
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
@@
 There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
-For specific examples, see also \emph{Box 1}.
+For specific examples, see also below.
 
-\subsection{Box 1: Examples of Hack Week Outcomes}
-\label{box:outcomes}
+\subsection*{Examples of Hack Week Outcomes}
+\label{sec:outcomes}
 \subsubsection*{Example 1: Astro Hack Week}

### 2017-09-25 sections/what.tex
Added Figure 1 back, fixed links to figures

@@
 Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
-Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
+Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure \ref{fig:hackspectrum}).
 For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
@@
 This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter}.
-Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitely competitive
+Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive

### 2017-09-25 sections/why.tex
Added Figure 1 back, fixed links to figures

@@
 \item{\textit{Education and Training}: 
-While some hack weeks are focused more on education than others, there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
+While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
 Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices in their respective fields.

### 2017-09-25 sections/design.tex
Added sentence about impostor syndrome to design section
added a couple of additional sentences about space in response to Laura's comments

@@
 Fortunately active learning teaching methods are creating new environments that foster interactivity through modified seating arrangements and integration of computer monitors at group workstations.
+There is a natural tension between keeping the group together and providing physically separated break-out spaces.
+On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas.
+On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical for hack week projects.
+It may hence be advisable to allow for diffusion of the group into adjacent rooms, while providing ample venue for the entire group to congregate for tutorials, breaks and reports.
 
@@
 This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
+The large variety in participant backgrounds and experiences can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it 
+is important to take this into account during workshop design.

### 2017-09-28 sections/design.tex
Added reference and Ariel's suggestion in design section, moved figure 2 to a better position

@@
 However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
-Fortunately active learning teaching methods are creating new environments that foster interactivity through modified seating arrangements and integration of computer monitors at group workstations.
+Fortunately, many universities are experimenting with other types of spaces: for example, the adoption of active learning teaching methods \cite{prince2004} have led to the development of modular classrooms, where seating arrangements can be flexibly modified and group activities can be more readily undertaken.
 There is a natural tension between keeping the group together and providing physically separated break-out spaces.

### 2017-09-28 sections/results.tex
Added reference and Ariel's suggestion in design section, moved figure 2 to a better position
Added numbers from NHW

@@
 \section*{Results}
+
+\begin{figure*}[h!]
+\begin{center}
+\includegraphics[width=12cm]{fig/eval_techskills}
+\includegraphics[width=12cm]{fig/eval_collab}
+\includegraphics[width=12cm]{fig/eval_openscience}
+\caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
+We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
+\label{fig:survey}
+\end{center}
+\end{figure*}
 
@@
 %In practice, this has not been done for any of the hack weeks conducted thus far.
-Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see \ref{box:outcomes})
+Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see \ref{box:outcomes}).
 %Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
@@
 
-\begin{figure*}[h!]
-\begin{center}
-\includegraphics[width=12cm]{fig/eval_techskills}
-\includegraphics[width=12cm]{fig/eval_collab}
-\includegraphics[width=12cm]{fig/eval_openscience}
-\caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
-\label{fig:survey}
-\end{center}
-\end{figure*}
-
 From the very nature of the activities that we encourage in hack weeks, participants in these events produce digital records of their research online.

### 2017-09-30 sections/participants.tex
Minor changes; this is the version submitted to PNAS

@@
 In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from participants.
-Therefore in order to maximize learning outcomes and the liklihood for collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
+Therefore in order to maximize learning outcomes and the likelihood for collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
+
 In our experience, a participant group that is diverse across categories of diversity, gender, discipline and career track helps to ensure we meet these objectives.
-To achieve this diversity, we advocate for a selection process that is as transparent as possible, enabling particiants to hold organizers accountable for their selection decisions.
+To achieve this diversity, we advocate for a selection process that is as transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
 Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
+
 One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to transfer some steps to an algorithm that is easily interpreted, openly available, and can be designed to counter the perpetuation of intrinsic biases.
@@
 When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
+
 The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
@@
 For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
+
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.

### 2017-09-30 sections/what.tex
Minor changes; this is the version submitted to PNAS

@@
 This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter}.
-Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive
+Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.

### 2017-09-30 sections/why.tex
Minor changes; this is the version submitted to PNAS

@@
 \item{\textit{Community Building}: Hack weeks provide a tremendous opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
-These events allow computationally minded researchers to break from the isolation of their academic departments and build connections and spark new collaborations.}
+These events allow computationally minded researchers to break from the isolation of their academic departments, build connections and spark new collaborations.}
 

### 2017-09-30 sections/design.tex

--- 2017-09-25 sections/design.tex
@@
 However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
-Fortunately active learning teaching methods are creating new environments that foster interactivity through modified seating arrangements and integration of computer monitors at group workstations.
+Fortunately, many universities are experimenting with other types of spaces: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, where seating arrangements can be flexibly modified and group activities can be more readily undertaken.
 There is a natural tension between keeping the group together and providing physically separated break-out spaces.

### 2017-09-30 sections/intro.tex
Added Figure 1 back, fixed links to figures

@@
 \label{sec:introduction}
-
-As data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
+\dropcap{A}s data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
 In addition, the recent reproducibility crises in several scientific fields has led to the growing realization that improving awareness of open science and reproducibility as well as practical skills in making research reproducible is essential to scientific progress \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
@@
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=9cm]{fig/HackSpectrum}
-%\caption{Comparison of Extracurricular Workshop Models}
-%\label{fig:hackspectrum}
-%\end{center}
-%\end{figure}
+\begin{figure}
+\begin{center}
+\includegraphics[width=9cm]{fig/HackSpectrum}
+\caption{Comparison of Extracurricular Workshop Models}
+\label{fig:hackspectrum}
+\end{center}
+\end{figure}
 

### 2017-09-30 sections/results.tex

--- 2017-09-25 sections/results.tex
@@
 \section*{Results}
+
+\begin{figure*}[h!]
+\begin{center}
+\includegraphics[width=11cm]{fig/eval_techskills}
+\includegraphics[width=11cm]{fig/eval_collab}
+\includegraphics[width=11cm]{fig/eval_openscience}
+\caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
+We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
+\label{fig:survey}
+\end{center}
+\end{figure*}
 
@@
 Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
+
 One key metric is the number of publications that result from hack week projects, but this is a fairly narrow definition of success, in line with standard academic performance indicators.
 Assuming that participants work largely in the open during a hack week, and that most projects have a strong programming component another indicator of success is the activity of participants in terms of code written and committed to a public code repository.
-Still, these measures ignore learning, community-building as well as networking outcomes.
-These outcomes can be assessed through post-workshop surveys.
+Still, these measures ignore learning, community-building as well as networking outcomes, which can be assessed through post-workshop surveys.
 %It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
@@
 %In practice, this has not been done for any of the hack weeks conducted thus far.
-Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see \ref{box:outcomes})
+Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see the following section).
 %Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks from 2016, we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, top panel, left).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016, we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, top panel, left).
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure \ref{fig:survey}, top panel, right and middle).
@@
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
-In particular, astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
-%Neuroscience seems to fall somewhere in between.
+Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
 This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
@@
 Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
-
-\begin{figure*}[h!]
-\begin{center}
-\includegraphics[width=12cm]{fig/eval_techskills}
-\includegraphics[width=12cm]{fig/eval_collab}
-\includegraphics[width=12cm]{fig/eval_openscience}
-\caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
-\label{fig:survey}
-\end{center}
-\end{figure*}
 
@@
 \subsubsection*{Example 1: Astro Hack Week}
-In 2015, a small team used the opportunity of Astro Hack Week to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of algorithms for time series analysis often used in X-ray astronomy.
+In 2015, a small team used the opportunity of AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of time series analysis algorithms often used in X-ray astronomy.
 The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects, role models of successful projects, and an environment encouraging scientific risk taking.
@@
 \subsubsection*{Example 2: Geo Hack Week}
-In 2016, a Geo Hack Week project team used Google Earth Engine to explore spatial patterns in climate, topography and population data with the goal of mapping the most suitable locations for renewable energy sites in the United States.
+In 2016, a GHW project team used Google Earth Engine to explore spatial patterns in climate, topography and population data with the goal of mapping the most suitable locations for renewable energy sites in the United States.
 The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine\footnote{\url{http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/}}.
@@
 \subsubsection*{Example 3: Neuro Hack Week}
-Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies.
-This is a particular concern in studies in which children or patients are studied, as they are more likely to move.
-In the 2016 Neuro Hack Week one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
+Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies, particularly in studies of children or patients, as they are more likely to move.
+During NHW 2016 one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
 To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off were varied.
-They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the Neuro Hack Week was over, and they eventually wrote and published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.
+%They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
+The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the end of Neuro Hack Week, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.

### 2017-09-30 old_nature_draft/paper.tex

--- 2017-09-20 paper.tex
@@
 \begin{addendum}
- \item{The authors would like to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB, Stuart Geiger for helping to formulate the questionnaires that served as the basis for the results presented here, Brittany Fiore-Gartland for comments on the manuscript and Tal Yarkoni for advice regarding automated selection procedures.
+ \item{The authors would like to thank Laura Nor\'{e}n (NYU) for help on ethics and IRB, Stuart Geiger for helping to formulate the questionnaires that served as the basis for the results presented here, Brittany Fiore-Gartland and Jason Yeatman for comments on the manuscript, and Tal Yarkoni for advice regarding automated selection procedures.
 This work was partially supported by the Moore-Sloan Data Science Environments at UC Berkeley, New York University, and the University of Washington.

### 2017-12-05 sections/what.tex
a few sentences about summer school versus hack week

@@
 
+Three design considerations distinguish the hack week from the summer school model: (1) hack week transfer knowledge across many levels of senority, with participants routinely leading tutorials, in part due to the relative newness of the content and the distributed knowledge base; (2) hack weeks contain a significant amount of unstructured time, creating the conditions for improvised exchanges and project development; and (3) hack weeks intentionally priortize participant diversity because this is known to maximize the liklihood for the co-creation of learning during the event.
+
 We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.

### 2017-12-06 sections/results.tex
Updated text and figures with new results

@@
 \begin{center}
-\includegraphics[width=11cm]{fig/eval_techskills}
-\includegraphics[width=11cm]{fig/eval_collab}
-\includegraphics[width=11cm]{fig/eval_openscience}
+\includegraphics[width=11cm]{fig/eval_techskills.eps}
+\includegraphics[width=11cm]{fig/eval_collab.eps}
+\includegraphics[width=11cm]{fig/eval_openscience.eps}
 \caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
+\textbf{For GHW and NHW, response rates were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
+At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.} We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
 \label{fig:survey}
@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016, we find that all participants self-report successful learning outcomes in new topics, tools or methods (AHW: 86\%, GHW: 94\%, NHW: 76\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, top panel, left).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016, we find that \textbf{most} participants self-report successful learning outcomes in new topics, tools or methods (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, top panel, left).
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure \ref{fig:survey}, top panel, right and middle).
@@
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 83\%, GHW: 69\%, NHW: 71\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, middle panel, left), though responses are not as unequivocal as they are in some of the other categories.
+We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, middle panel, left), though responses are not as unequivocal as they are in some of the other categories.
 This is expected: participants new to the field may participate to learn rather than to teach.
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 70\% of all participants (AHW: 79\%, GHW: 72\%, NHW:85\%; Figure \ref{fig:survey}, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure \ref{fig:survey}, bottom panel, middle).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure \ref{fig:survey}, bottom panel, middle).
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
 Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
-This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
+%This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
 Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%).
 Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
+\textbf{While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, including open-source software, data sharing and reproducibility.
+In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.}
 

### 2017-12-06 sections/conclusions.tex
Added some correlation searches to paper

@@
 As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
-Initial results from three events held in 2016 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
+Initial results from three events held in 2016 \textbf{and 2017} in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
 Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.

### 2017-12-06 sections/results.tex
Added some correlation searches to paper

@@
 This is expected: participants new to the field may participate to learn rather than to teach.
+\textbf{It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
+In particular, participants from all career stages between graduate students and tenured faculty report similar engagement in teaching.
+This supports our hypothesis that compared to e.g.\ a traditional summer school, a hack week is less hierarchical and encourages lateral knowledge transfer between participants at different career stages.
+We similar find no correlations with gender identity or race/ethnicity.
+Responses to the other two questions regarding whether participants' contributions to hack teams were valued and whether they made valuable connections at the hack week were similarly evenly distributed across career stages, gender and racial/ethnic identities.}
 

### 2017-12-06 sections/participants.tex
Updated participant selection section, part one

@@
 
-One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to transfer some steps to an algorithm that is easily interpreted, openly available, and can be designed to counter the perpetuation of intrinsic biases.
-We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.
-We try to minimize bias in this step by blinding ourselves to a candidate's other attributes, including name and other personal information, and assess their candidacy based soley on questions asked specifically for this purpose.
-When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
+Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible [REFs]. 
+In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 
+Because human decision makers tend to be swayed by unrelated characteristics including name [REF] or gender [REF] of an applicant, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
+A merit selection could then perform via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate [REF], it may be beneficial to subsequently set a fairly tolerant threshold for acceptance, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 
-The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
-It is here where one may impose outside constraints on the selection based on the goals of the workshop.
-If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
-
-One solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.
+One solution to the latter problem is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}). 
 The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
 For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
-
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
@@
 For example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority, and impossible to correct gender bias in the pool of applicants, if no information is available about the gender of participants.
+
+Blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a very broad population. 
+For more targeted workshops, organizers should be mindful that blinding to demographics might be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
+In this case, it may be beneficial to include demographic characteristics in the first stage.
+
+No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate the different steps in their procedure.
+For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded.
+
+%One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to . 
+%We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.
+%We try to minimize bias in this step by blinding ourselves to a candidate's other attributes, including name and other personal information, and assess their candidacy based soley on questions asked specifically for this purpose.
+%When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
+
+%The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
+%It is here where one may impose outside constraints on the selection based on the goals of the workshop.
+%If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
+
+%One solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.

### 2017-12-07 sections/participants.tex
Minor fiddles
added references

@@
 In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 
-Because human decision makers tend to be swayed by unrelated characteristics including name [REF] or gender [REF] of an applicant, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
+Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012} of an applicant, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
 A merit selection could then perform via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate [REF], it may be beneficial to subsequently set a fairly tolerant threshold for acceptance, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate, ~\cite{highhouse2008} it may be beneficial to subsequently set a fairly tolerant threshold for acceptance, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 
@@
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
-If these targets reproduce the distribution of the overall sample of candidates, the selection will become essentially random. 
 Any human biases involved in setting these target fractions will be perpetuated in the selection procedure. 
@@
 Blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a very broad population. 
-For more targeted workshops, organizers should be mindful that blinding to demographics might be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
+For more targeted workshops, organizers should be mindful that blinding to demographics might be counterproductive~\cite{behaghel2015unintended} if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
 In this case, it may be beneficial to include demographic characteristics in the first stage.

### 2017-12-07 sections/participants.tex
Added more references

@@
 
-Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible [REFs]. 
+Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible~\cite{sunstein2015wiser,bohnet2016}.
 In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 

### 2017-12-07 supplementary_material.tex
some notes

--- 2017-08-17 supplementary_material.tex
@@
 
-\documentclass{nature}
+%\documentclass{nature}
 
@@
 %% LaTeX can find them
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
-\usepackage{natbib}
+%\usepackage{graphicx}	% Including figure files
+%\usepackage{amsmath}	% Advanced maths commands
+%\usepackage{amssymb}	% Extra maths symbols
+%\usepackage{natbib}
 
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage{microtype}
-\usepackage{rotating}
-\usepackage{booktabs}
-\usepackage{threeparttable}
-\usepackage{tabularx}
+%\usepackage{hyperref}
+%\usepackage{url}
+%\usepackage{microtype}
+%\usepackage{rotating}
+%\usepackage{booktabs}
+%\usepackage{threeparttable}
+%\usepackage{tabularx}
 
@@
 
-\begin{document}
+%\begin{document}
 
@@
 
+\subsection{Event Size and Duration}
+
+- add a few sentences about size and interaction with goals
+- add a few sentences about duration and interaction with goals
+- add that we're still figuring this out
+- we're experimenting with length and duration
+
+\subsubsection{Participant selection}
+
+Add more on participant selection, especially practical details. 
+- add problem of asking good questions: especially questions that require candidates to calibrate should be written in such a way that they allow for objective calibration (e.g.
+ask for specific actions people have taken, rather than "beginner", "intermediate", "expert"
+- add other approaches taken by the three different hack weeks
+
+- add different criteria for selection
+
 \subsubsection{Hack Week Location}
@@
 At Astro Hack Week, we have found that travel grants are of crucial importance especially to undergraduates, students from minority backgrounds and participants from underrepresented countries.
-During the organization of Python in Astronomy 2017, we found that minority participants disproportionally decline to attend, and often cite lack of sufficient travel funding as the primary reason.
+We found that minority participants disproportionally decline to attend, and often cite lack of sufficient travel funding as the primary reason.
 In particular if increasing diversity in the more technical aspects of a field is one of the workshop's goals, providing enough travel funding for between five to ten participants can make a significant difference in allowing minority researchers to attend.
@@
 
-\subsection{Box 1: Impostor Syndrome}
+\subsection{Ensuring Participation}
+to add here:
+- things about inclusion and how to make people participate
+- add sentence about that to main paper
+- ANthony needs to add stuff here
+- add e.g. that we planted designated question askers
+- have organizers as dedicated facilitators
+
+- want to have some good
+
+\subsubsection{Impostor Syndrome}
 
@@
 
-\end{document}
+%\end{document}

### 2017-12-08 sections/design.tex
Added lots of text to supplementary materials, a few changes here and there

@@
 This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
-The large variety in participant backgrounds and experiences can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it 
-is important to take this into account during workshop design.
+\textbf{The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week.
+Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
+In particular the presence of impostor syndrome experienced by many participants is important to take this into account during workshop design (see also supplementary materials for concrete suggestions).}
+%can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it

### 2017-12-08 sections/conclusions.tex

--- 2017-08-17 sections/conclusions.tex
@@
 We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.
+\textbf{While we focus here on hack weeks in scientific fields, the concept could be extended to other areas, and is more generally useful in any area (1) where useful tools can be learned in short tutorials, (2) where results and outcomes can be produced on the timescale of a few days, and (3) that would benefit from collaborative approaches that cross traditional boundaries.
+Such areas could include for example the social sciences, music and art.}

### 2017-12-08 sections/participants.tex

--- 2017-12-06 sections/participants.tex
@@
 
-Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible [REFs]. 
-In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 
-Because human decision makers tend to be swayed by unrelated characteristics including name [REF] or gender [REF] of an applicant, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
-A merit selection could then perform via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate [REF], it may be beneficial to subsequently set a fairly tolerant threshold for acceptance, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to transfer some steps to an algorithm that is easily interpreted, openly available, and can be designed to counter the perpetuation of intrinsic biases.
+We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.
+We try to minimize bias in this step by blinding ourselves to a candidate's other attributes, including name and other personal information, and assess their candidacy based soley on questions asked specifically for this purpose.
+When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
 
-One solution to the latter problem is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}). 
+The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
+It is here where one may impose outside constraints on the selection based on the goals of the workshop.
+If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
+
+One solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.
 The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
 For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
+
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.
@@
 For example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority, and impossible to correct gender bias in the pool of applicants, if no information is available about the gender of participants.
-
-Blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a very broad population. 
-For more targeted workshops, organizers should be mindful that blinding to demographics might be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
-In this case, it may be beneficial to include demographic characteristics in the first stage.
-
-No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate the different steps in their procedure.
-For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded.
-
-%One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to . 
-%We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.
-%We try to minimize bias in this step by blinding ourselves to a candidate's other attributes, including name and other personal information, and assess their candidacy based soley on questions asked specifically for this purpose.
-%When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
-
-%The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
-%It is here where one may impose outside constraints on the selection based on the goals of the workshop.
-%If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
-
-%One solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.

### 2017-12-08 supplementary_material.tex

--- 2017-08-17 supplementary_material.tex
@@
-%% Template for a preprint Letter or Article for submission
-%% to the journal Nature.
-%% Written by Peter Czoschke, 26 February 2004
-%%
-
-\documentclass{nature}
-
-%% make sure you have the nature.cls and naturemag.bst files where
-%% LaTeX can find them
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
-\usepackage{natbib}
-
-\usepackage{hyperref}
-\usepackage{url}
-\usepackage{microtype}
-\usepackage{rotating}
-\usepackage{booktabs}
-\usepackage{threeparttable}
-\usepackage{tabularx}
-
-%\title{How and Why to run a Hack Week}
-
-%% Notice placement of commas and superscripts and use of &
-%% in the author list
-
-%\author{Jake VanderPlas, Karthik Ram, Daniela Huppenkothen, David W. Hogg, Ariel Rokem, Anthony Arendt \& Tal Yarkoni}
-
-\begin{document}
-
-%\maketitle
-
-%\begin{affiliations}
-% \item UW Seattle
-% \item NYU
-% \item UC Berkeley
-%\end{affiliations}
-
 \section{Supplementary Material}
@@
 \subsubsection{Goal Planning}
+
 Unlike traditional conferences, where the programme is primarily driven by speaker selection, careful goal planning is essential for a successful hack week.
@@
 
+\subsection{Event Size and Duration}
+
+Event size (i.e.
+the number of attendees) and event durations are important to consider during the design and planning phase of a hack week.
+All hack weeks held so far have been five days, with group sizes between 40-60 participants.
+The number of attendees resulted from a number of different design considerations.
+One practical consideration was the size of the available location: especially the project-focused parts of a hack week generally require a reconfigurable space with more seats than attendees, such that spontaneous organization in sub-groups is easy.
+However, group size is also important in the context of community building, group dynamics and programme decisions.
+We limited group size to a maximum of 70 participants, because previous anecdotal evidence has shown us that this is about the maximum group size where interactions between all participants is still possible.
+We wanted to ensure that all participants had at least met all other participants in order to facilitate community building and keep the group from fracturing into cliques.
+Another important consideration is representation along the many axes of diversity we considered.
+For example, a single undergraduate participant might feel like an outsider; admitting three or four can mitigate alienation effects.
+Similarly, in order to foster active collaboration among many different subfields, we ensured admitting participants with many different types of expertise, some of which overlapped.
+Having a set of participants that is too small would interfere both with these goals of representation and of fostering interdisciplinary collaborations and networking, thus we would advise admitting at least $\sim$30 or so participants if these objectives are an important part of the hack week.
+
+All hack weeks organized up to now have been five days.
+This duration was chosen with several different design objectives in mind.
+In particular, it was chosen to allow both learning and project work at the same time.
+Participants often come with ambitious project ideas, or with the goal of taking newly learned methods and applying them to their data sets right away.
+Thus many projects tended to extend over multiple days.
+Attendees who participated in multiple projects often used the hack week to try out several different ideas.
+A five-day hack week allows for ample time for these ambitious projects or experimentation, where a shorter hack week would necessarily have to be more constrained and focused.
+At the same time, hack weeks tend to be exhausting.
+Project work often extends into the evenings, and requires prolonged phases of high concentration of participants, thus we limited our hack weeks to five days in order to avoid excessive fatigue among our attendees.
+
+It is worth noting, however, that these guidelines were developed out of design considerations rather than experimentation and testing.
+In the future, we plan to experiment with the format, duration and size of these events and continue to evaluate their interplay with our objectives.
+
+\subsubsection{Participant selection}
+
+As we lay out in the main text, participant selection is one of the most important design decisions as part of a hack week.
+Conversations about participant selection, and planning of selection strategies should commence early, ideally at the start of the organizing process.
+Because selection procedures are subject to a whole range of human biases, it is worth thinking critically about established procedures and how to improve them.
+At the same time, hack weeks are comparatively lower-stakes events than other situations where cohort selection is relevant, like e.g.\ graduate school admissions, thus hack weeks may also serve as test beds for experimentation and improvements that can inform selection in other contexts.
+
+The hack weeks presented here used different approaches, that nevertheless several common themes and ideas.
+In particular, they focus on making the selection procedure as quantitative as possible and use the data collected during each stage of the process to critically evaluate the efficacy of the process and its subject to human bias.
+Secondly, much energy is expended early on in the design of the application form, which often use a combination of open-answer and multiple-choice questions to assess candidates' suitability, skill levels and backgrounds.
+Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' or ``expert'' are have no absolute scaling and may have different meanings to different applicants.
+They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates.
+It is thus advisable to calibrate skill levels on concrete milestones.
+For example, instead of asking participants whether they are ``beginners'' or ``experts'' at machine learning, we instead asked them whether they had done machine learning for class work, had used techniques from machine learning in their own research, or had developed and implemented their own methods.
+Responding to questions about demographic characteristics should generally be optional, and organizers should consider how these questions might interact adversely with workshop goals, e.g.\ via stereotype threat.
+For example, Astro Hack Week only asked whether an applicant considered themselves a minority with respect to gender identity within astronomy, rather than asking them to explicitly self-identify within a specific category.
+
+The overall selection procedure for Astro Hack Week comprised a two-step process.
+In the initial step, we performed a blind assessment for suitability using open-ended questions querying candidates' motivations for attending the workshop, and their goals for attending.
+We only excluded a small fraction ($<10\%$) of applicants at this step, because our goals aligned with admitting a wide range of different skill levels and backgrounds.
+Applicants were generally either excluded for being an obvious mismatch with the workshop goals (as in the case of an Android developer applying to work on an unrelated game) or for being duplicate entries.
+We subsequently used a range of categories queried with multiple-choice questions  on the application to maximize diversity across these categories, which included career stage, gender identity, racial/ethnic identity, geographic origin, programming background, machine learning skills, statistics knowledge, and previous attendance at Astro Hack Week.
+Targets within each categories were set according to the different workshop goals.
+For example, we set ambitious goals of admitting 50\% attendees from underrepresented minorities with respect to gender, race and ethnicity with the goal of increasing representation of these groups with astronomy and data science.
+Similarly, we set targets for other categories (e.g.
+skill levels) in order to maximize goals around collaboration, networking and fostering interdisciplinary exchange of knowledge.
+Subsequently, we used a greedy optimization algorithm to jointly optimize over all categories simultaneously.
+We evaluated our results by directly comparing the characteristics of the selected set of participants with both our targets and the distribution of the set of all candidates.
+For demographic characteristics specifically, 42\% of participants in both 2016 and 2017 considered themselves a minority with respect to gender identity in astronomy, and 32\% of participants in 2016 (2017: 33\%) considered themselves a minority with respect to racial or ethnic identity.
+About half of the attendees were students ($\sim$40\% graduate students, $\sim$10\% undergraduate students), another $25\5$ postdoctoral fellow, and the remainder about equally split between research staff, tenured and tenure-track faculty, and participants in non-academic careers.
+
 \subsubsection{Funding}
@@
 Important documents such as programme (and especially programme changes) as well as hack descriptions should be open to the public and communicated clearly to facilitate collaborations with outside researchers.
+
+It is, in principle, possible to organize an entire hack week virtually.
+However, the format rests firmly on a fundament of community building and creating opportunities for participants to exchange ideas and knowledge.
+While not impossible to achieve in a virtual setting, both would be significantly harder than in a single physical location, to a point where the lack of immediate exchange may severely limit to generate successful new collaborations and ideas.
+We thus recommend against running a purely virtual hack week.
 
@@
 
-It is noteworthy that in practice, the mixed audience and interactive setting can often lead to impostor syndrome among the speakers, who lose their explicit status as expert among the participants.
+It is noteworthy that in practice, the mixed audience and interactive setting can often lead to impostor syndrome (see below) among the speakers, who lose their explicit status as expert among the participants.
 It is therefore imperative that speakers are made aware very early about the audience so that they can plan their tutorial accordingly.
@@
 
-\subsection{Box 1: Impostor Syndrome}
+\subsection{Facilitating Participation}
+
+Admitting a diverse group of participants is only the first step in encouraging collaboration and knowledge exchange.
+Community building and careful facilitation of both tutorials and project works are arguably the cornerstones of a successful hack week.
+This is especially important because the group will inevitably include a wide range of personalities, some of which may find it easier to form new bonds and interact with the group than others.
+At all hack weeks, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
+Clearly communicating values and expectations at the beginning for the week can help setting the stage for an environment that a large range of different personalities can thrive in.
+This includes both communication of the code of conduct, putting an emphasis on collaborative rather than competitive behaviour, as well as a discussion of potential issues participants might experience, such as impostor syndrome (see also section below).
+
+At Astro Hack Week, we aim to influence group dynamics in specific ways to allow timid participants to speak up.
+As mentioned above, we ask a few participants known to us to occasionally ask questions during tutorials, in order to encourage other participants to do the same.
+We also designate at least two members of the organizing committee whose role it is to facilitate both tutorials and hack sessions.
+It is their task to for example manage dominant personalities, which can be both a boon and a hindrance to the success of a hack week.
+Often, these attendees can be effective at generating excitement and engagement, but can also intimidate and dominate discussions.
+It is up to the organizers to balance the participation of these attendees with the needs of the rest of the group.
+Similarly, facilitators at Astro Hack Week spend the majority of hacking time circulating the room, connecting participants without a group to a project, help connect groups who run into a problem with experts who can help out, and make sure group work proceeds in such a way that all members feel valued.
+
+\subsubsection{Impostor Syndrome}
 
@@
 For this purpose, it is useful to keep a way to contact previous participants, via a mailing list or real-time messaging services like \textit{Matrix} or \textit{Slack}.
-
-%Spelling must be British English (Oxford English Dictionary)
-
-%In addition, a cover letter needs to be written with the
-%following:
-%\begin{enumerate}
-% \item A 100 word or less summary indicating on scientific grounds
-%why the paper should be considered for a wide-ranging journal like
-%\textsl{Nature} instead of a more narrowly focussed journal.
-% \item A 100 word or less summary aimed at a non-scientific audience,
-%written at the level of a national newspaper.  It may be used for
-%\textsl{Nature}'s press release or other general publicity.
-% \item The cover letter should state clearly what is included as the
-%submission, including number of figures, supporting manuscripts
-%and any Supplementary Information (specifying number of items and
-%format).
-% \item The cover letter should also state the number of
-%words of text in the paper; the number of figures and parts of
-%figures (for example, 4 figures, comprising 16 separate panels in
-%total); a rough estimate of the desired final size of figures in
-%terms of number of pages; and a full current postal address,
-%telephone and fax numbers, and current e-mail address.
-%\end{enumerate}
-
-%See \textsl{Nature}'s website
-%(\texttt{http://www.nature.com/nature/submit/gta/index.html}) for
-%complete submission guidelines.
-
-%\begin{methods}
-%Put methods in here.  If you are going to subsection it, use
-%\verb|\subsection| commands.  Methods section should be less than
-%800 words and if it is less than 200 words, it can be incorporated
-%into the main text.
-
-%\end{methods}
-
-%% Put the bibliography here, most people will use BiBTeX in
-%% which case the environment below should be replaced with
-%% the \bibliography{} command.
-
-% \begin{thebibliography}{1}
-% \bibitem{dummy} Articles are restricted to 50 references, Letters
-% to 30.
-% \bibitem{dummyb} No compound references -- only one source per
-% reference.
-% \end{thebibliography}
-
-%%
-%% TABLES
-%%
-%% If there are any tables, put them here.
-%%
-
-%\begin{table}
-%\centering
-%\caption{This is a table with scientific results.}
-%\medskip
-%\begin{tabular}{ccccc}
-%\hline
-%1 & 2 & 3 & 4 & 5\\
-%\hline
-%aaa & bbb & ccc & ddd & eee\\
-%aaaa & bbbb & cccc & dddd & eeee\\
-%aaaaa & bbbbb & ccccc & ddddd & eeeee\\
-%aaaaaa & bbbbbb & cccccc & dddddd & eeeeee\\
-%1.000 & 2.000 & 3.000 & 4.000 & 5.000\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-\end{document}

### 2017-12-08 supplementary_material.tex
Pass through some parts of the supplementary materials.

@@
 \subsubsection{Obtaining permission for tracking and studying participants}
-A goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes.
+A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes.
 To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities.
 These scientists then reviewed their observations and and gave valuable feedback about the hack week outcomes.
-In order to be able to track and study participants, approval from Internal Review Boards is required for workshops organized in the United States.
+In order to be able to track and study participants, approval from an Internal Review Board is required.
 This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally don't involve human subject research.
@@
 The stated goals will continue to determine almost all of the remaining organizational questions, including space, participant selection and workshop evaluation.
-Is it aimed more at students or mature researchers, or a mix of both?
+Is the target audience primarily students, or mature researchers, or a mix of both?
 Should the taught component dominate over project work or should it be only a minor programme point?
@@
 When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration.
-The non-traditional nature of hack weeks means that some students report challenges in receiving their superior's approval to attend.
-Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be the highest priority when deciding on a location.
+The non-traditional nature of hack weeks means that some students report challenges in receiving support from their supervisor or department to attend.
+Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be a high priority when deciding on a location.
 Conference centers like the Lorentz Center at the University of Leiden provide funding along with the collaborative space, but have long lead times when applying for their spaces (typically 12-18 months).
@@
 A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.
+There should be plenty of power outlets for participants to plug in their computers.
+The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access to visitors from other institutions.
+Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available.
 
@@
 We limited group size to a maximum of 70 participants, because previous anecdotal evidence has shown us that this is about the maximum group size where interactions between all participants is still possible.
-We wanted to ensure that all participants had at least met all other participants in order to facilitate community building and keep the group from fracturing into cliques.
+We wanted to ensure that all participants had an opportunity to at least meet all other participants, in order to facilitate community building and keep the group from fracturing into cliques.
 Another important consideration is representation along the many axes of diversity we considered.
@@
 Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' or ``expert'' are have no absolute scaling and may have different meanings to different applicants.
-They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates.
+They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates (the Dunning Kruger effect).
 It is thus advisable to calibrate skill levels on concrete milestones.
 For example, instead of asking participants whether they are ``beginners'' or ``experts'' at machine learning, we instead asked them whether they had done machine learning for class work, had used techniques from machine learning in their own research, or had developed and implemented their own methods.
+Similarly, questions about programming skills may ask whether participants can perform a particular task in one of a list of languages  (e.g., "write a module or a library").
 Responding to questions about demographic characteristics should generally be optional, and organizers should consider how these questions might interact adversely with workshop goals, e.g.\ via stereotype threat.
 For example, Astro Hack Week only asked whether an applicant considered themselves a minority with respect to gender identity within astronomy, rather than asking them to explicitly self-identify within a specific category.
+Meanwhile, neuro hack week opted to ask about gender ethnicity and race, using questions that are standardly used for example, when collecting statistics about study participants.
 
@@
 For demographic characteristics specifically, 42\% of participants in both 2016 and 2017 considered themselves a minority with respect to gender identity in astronomy, and 32\% of participants in 2016 (2017: 33\%) considered themselves a minority with respect to racial or ethnic identity.
-About half of the attendees were students ($\sim$40\% graduate students, $\sim$10\% undergraduate students), another $25\5$ postdoctoral fellow, and the remainder about equally split between research staff, tenured and tenure-track faculty, and participants in non-academic careers.
+About half of the attendees were students ($\sim$40\% graduate students, $\sim$10\% undergraduate students), another $25\%$ postdoctoral fellow, and the remainder about equally split between research staff, tenured and tenure-track faculty, and participants in non-academic careers.
+
+Neuro hack week took a different approach: participants provided CV as part of their application, in addition to answering questions about their technical and research background, and writing a short statement of the purpose of their participation.
+Each application was reviewed by several instructors using 1-10 scales on several criteria: technical and computing experience, research experience, the quality of the statement, an assessment of the contribution of participation to the applicant's career, and an assessment of the degree to which participation would increase the representation of groups that are usually under-represented in neuroscience and computing.
+The scores were combined, applicants were ranked according to the combined score, and the 40 highest-scoring applicants were then selected to participate.
+As an example,  this procedure resulted in gender parity in 2016 and near-parity (17 women and 23 men) in 2017.
+In neuro hack week as well about half of the participants were students (though no undergraduate students participated), about $25\%$ were post-docs, with the remainder including research staff and faculty members.
 
@@
 Locations that offer some funding or provide free space are preferable to those that do not.
-At Astro Hack Week, we have found that travel grants are of crucial importance especially to undergraduates, students from minority backgrounds and participants from underrepresented countries.
+We have found that travel grants are of crucial importance especially to undergraduates, students from minority backgrounds and participants from underrepresented countries.
 We found that minority participants disproportionally decline to attend, and often cite lack of sufficient travel funding as the primary reason.
 In particular if increasing diversity in the more technical aspects of a field is one of the workshop's goals, providing enough travel funding for between five to ten participants can make a significant difference in allowing minority researchers to attend.
+Funding for some activities can be obtained through commercial vendors.
+For example, cloud computing providers have educational programs that provide cloud computing credits for students.
+Other corporate sponsorships of certain activities or of the event as a whole can be used to lower participation costs, and make participation more feasible for many participants.
+Federal funding agencies are also sometimes able to fund hack weeks and similar activities under traditional funding mechanisms that support training.
+For example, Neuro hack week 2017 was supported through a grant from the National Institute of Mental Health through their standard R25 training grant mechanism, and this grant continues to support it in extended format in coming years.
 
@@
 
-At Astro Hack Week, we aim to influence group dynamics in specific ways to allow timid participants to speak up.
+From our experience, there are ways in which organizers can influence group dynamics in specific ways to allow timid participants to speak up.
 As mentioned above, we ask a few participants known to us to occasionally ask questions during tutorials, in order to encourage other participants to do the same.

### 2017-12-09 supplementary_material.tex
Format supplement.

@@
-\section{Supplementary Material}
+%% Template for a preprint Letter or Article for submission
+%% to the journal Nature.
+%% Written by Peter Czoschke, 26 February 2004
+%%
 
-While the main text of this paper remains fairly theoretical, we aim to lay out more practical advice regarding the organization of a hack week in the appendix of this document.
+\documentclass{nature}
 
-\subsection{What to do before}
+%% make sure you have the nature.cls and naturemag.bst files where
+%% LaTeX can find them
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+\usepackage{amssymb}	% Extra maths symbols
+%\usepackage{natbib}
 
-\subsubsection{Obtaining permission for tracking and studying participants}
+\usepackage{hyperref}
+\usepackage{url}
+\usepackage{microtype}
+\usepackage{rotating}
+\usepackage{booktabs}
+\usepackage{threeparttable}
+\usepackage{tabularx}
+\linespread{1.2}
+
+\title{Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"}
+
+%% Notice placement of commas and superscripts and use of &
+%% in the author list
+
+\author{Daniela Huppenkothen$^1$, Anthony Arendt$^2$, David W. Hogg$^1$, Karthik Ram$^3$, Jake VanderPlas$^2$, and Ariel Rokem$^2$}
+
+\makeatletter
+\let\saved@includegraphics\includegraphics
+\AtBeginDocument{\let\includegraphics\saved@includegraphics}
+\renewenvironment*{figure}{\@float{figure}}{\end@float}
+\makeatother
+
+\begin{document}
+
+\maketitle
+
+\begin{affiliations}
+ \item Center for Data Science, New York University
+ \item The University of Washington eScience Institute
+ \item The Berkeley Institute for Data Science, UC Berkeley
+\end{affiliations}
+
+\noindent While the main text of this paper remains fairly theoretical, we aim to lay out more practical advice regarding the organization of a hack week in this supplement.
+
+\section{What to do before}
+
+\subsection{Obtaining permission for tracking and studying participants}
 A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes.
@@
 
-\subsubsection{Goal Planning}
+\subsection{Goal Planning}
 
@@
 
-\subsubsection{Hack Week Location}
+\subsection{Hack Week Location}
 When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration.
@@
 
-\subsection{Event Size and Duration}
+\section{Event Size and Duration}
 
@@
 
-\subsubsection{Participant selection}
+\subsection{Participant selection}
 
@@
 
-\subsubsection{Funding}
+\subsection{Funding}
 
@@
 
-\subsubsection{Code of Conduct}
+\subsection{Code of Conduct}
 
@@
 
-\subsubsection{Communicating with the Outside World}
+\subsection{Communicating with the Outside World}
 
@@
 
-\subsection{During a Hack Week}
+\section{During a Hack Week}
 
@@
 
-\subsubsection{Tutorials}
+\subsection{Tutorials}
 Tutorials provide a mechanism for rapid knowledge sharing and are often one of the most useful components of a hack week.
@@
 
-\subsubsection{Break-Out Sessions}
+\subsection{Break-Out Sessions}
 
@@
 
-\subsubsection{Hack Sessions}
+\subsection{Hack Sessions}
 
@@
 
-\subsection{Facilitating Participation}
+\section{Facilitating Participation}
 
@@
 
-\subsubsection{Impostor Syndrome}
+\subsection{Impostor Syndrome}
 
@@
 
-\subsection{What to do after}
+\section{What to do after}
 
@@
 For this purpose, it is useful to keep a way to contact previous participants, via a mailing list or real-time messaging services like \textit{Matrix} or \textit{Slack}.
+
+\section{References}
+\bibliographystyle{naturemag}
+\bibliography{paper}
+
+\end{document}

### 2017-12-09 supplementary_material.tex
I prefer "setting objectives" to "goal planning".
But that's at the level of bike-shedding.
Edits to the supplementary materials.

@@
 
-\subsection{Goal Planning}
+\subsection{Determining your objectives}
 
@@
 
-A hack week should have a code of conduct.
-Organizers should be aware that disparities in backgrounds, knowledge and experience, together with the close collaboration much of the hacking requires, creates the potential for conduct that others might find offensive, especially when it comes to language.
-The goal of every hack week should be to provide a comfortable, inclusive environment for participants to learn and work.
+An overarching goal of every hack week is to provide a welcoming, comfortable, and inclusive environment for participants to learn and work.
+However, disparities in backgrounds, knowledge and experience, together with the close collaboration much of the hacking requires, creates the potential for conduct that might offend participants, and preclude them from participating fully.
+For this reason, a hack week should have a clear and enforcable code of conduct, that can guide the behavior of participants, and help organizers maintain a professional and welcoming environment.
 It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in advance.
@@
 For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome, defensive and inhibit learning on a larger scale.
+Even if no code of conduct infarctions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are thoughtful of creating a welcoming environment.
+This by itself helps create a welcoming environment.
 
@@
 However, the closed nature of Twitter to those with an account may not make it the ideal platform for participation.
-Other open-source formats that do not require sign-up, such as Matrix, may be more compatible with the hack week format\footnote{\url{https://dave.cheney.net/2017/04/11/why-slack-is-inappropriate-for-open-source-communications}}.
+Other open-source formats that do not require sign-up, such as Matrix, may be more compatible with the hack week format \footnote{\url{https://dave.cheney.net/2017/04/11/why-slack-is-inappropriate-for-open-source-communications}}.
 Important documents such as programme (and especially programme changes) as well as hack descriptions should be open to the public and communicated clearly to facilitate collaborations with outside researchers.
@@
 It is, in principle, possible to organize an entire hack week virtually.
-However, the format rests firmly on a fundament of community building and creating opportunities for participants to exchange ideas and knowledge.
+However, the format rests firmly on a foundation of community-building and creating opportunities for participants to exchange ideas and knowledge.
 While not impossible to achieve in a virtual setting, both would be significantly harder than in a single physical location, to a point where the lack of immediate exchange may severely limit to generate successful new collaborations and ideas.
@@
 Therefore careful planning and preparation is necessary for effective delivery of a hack week tutorial.
+Tutorials may also be video-taped and shared with a broader community.
+We have found that video recordings of tutorials are rather popular, with some of them drawing more than 1,000 views.
 
@@
 
-Some of our hack week tutorials follow the structure of the Software Carpentry model.
+Some of our hack week tutorials follow the structure of the Software Carpentry model \cite{b:wilson-swc-lessons-2016}.
 The content is divided in to a series of lessons, each having a well-defined set of learning objectives, key questions, and expected outcomes.
@@
 This provides a simple method to deliver help to those students who need it without interrupting the flow of the tutorial.
-We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks or following the markdown templates provided by Software Carpentry.
+We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \cite{kluyver2016jupyter} or following the markdown templates provided by Software Carpentry.
 This way the hack week "curriculum" can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
@@
 Either way, we find divding code into short blocks, each with a specific task, helps participants from becoming saturated with informaiton.
+In some cases, tutorials require substantial software installation and data.
+Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometime so complicated as to slow down teaching, and hold back.
+In some cases, we have experimented using \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \cite{holdgraf2017portable}.
+Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \cite{Holdgraf2017-pd, Titus_Brown_undated-pc}
 
@@
 It also helps to make them aware that participants may be taking the role of teaching assistants during exercises and include them in the teaching, rather than see them as adversaries who may seek to correct them.
-Conversely, experts may be used to ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and encourage questions especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome.
+Conversely, experts may be used to ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and to encourage questions, especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome.
 At Astro Hack Week, we encourage the organizing committee and experienced participants to ask questions, particularly when they know the answer but think the concept may not yet be clear to parts of the audience.

### 2017-12-09 sections/what.tex
A few more pointers to the supplement.

--- 2017-09-30 sections/what.tex
@@
 One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
-This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter}.
+This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter} (see supplementary material).
 Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.

### 2017-12-12 supplementary_material.tex
geohack facilitation

@@
 For example, instead of asking participants whether they are ``beginners'' or ``experts'' at machine learning, we instead asked them whether they had done machine learning for class work, had used techniques from machine learning in their own research, or had developed and implemented their own methods.
-Similarly, questions about programming skills may ask whether participants can perform a particular task in one of a list of languages  (e.g., "write a module or a library").
+Similarly, questions about programming skills may ask whether participants can perform a particular task in one of a list of languages  (e.g., ``write a module or a library'').
 Responding to questions about demographic characteristics should generally be optional, and organizers should consider how these questions might interact adversely with workshop goals, e.g.\ via stereotype threat.
@@
 We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \cite{kluyver2016jupyter} or following the markdown templates provided by Software Carpentry.
-This way the hack week "curriculum" can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
-Depending on the content, some instructors convey concepts during the tutorial via "live-coding" while others display static code and their results.
+This way the hack week ``curriculum'' can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
+Depending on the content, some instructors convey concepts during the tutorial via ``live-coding'' while others display static code and their results.
 Either way, we find divding code into short blocks, each with a specific task, helps participants from becoming saturated with informaiton.
@@
 
+At geohack week we employed ``Liberating Structures'', a facilitation approach in which we introduce small shifts in the way groups engage in order to increase relational coordination and trust.
+We invited a professional facilitator to assist us in leading a series of conversations and activities aimed at fostering group engagement.
+In one activity, participants faced one another and were given a brief statement inviting a two minute conversation, after which the rows rotated to maximize the mixing of ideas.
+Invitations included statements such as ``I belong at geohack because...'' or ``A challenge I hope we can make progress on is...''.
+Importantly, event organizers joined in the activity, which we found helped participants feel as if they were not alone in facing some of the challenges inherent in data science studies.
+Another activity was designed to address the challenges of forming project teams.
+Individuals were invited to briefly pitch project ideas to the full group, and then form stations around the room.
+The remaining participants were then invited to visit any station that drew their interest, and to rotate between up to two other stations if they desired.
+We found this approach minimized the pressure on participants, giving them time to get a feel for project ideas and group dynamics before joining a team.
+
 \subsection{Impostor Syndrome}
@@
 The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ''found out``, shamed and expelled from their environment \cite{Clance1978-ef}.
+Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \cite{Clance1978-ef}.
 Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
@@
 Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
+We also invited our organizing committees to model open communication about their own handling of IS through the course of their career.
+We found that transparency about one's own uncertainty, espcially from community leaders, helped create a hospitable atmosphere and break down barriers.
 

### 2017-12-19 sections/intro.tex

@@
 As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
-Below we will share some of the philosophy behind the hack week model, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.
+Below we will share some of the philosophy behind the hack week model, results from surveys of participants, and practical lessons we have learned in organizing these events, as well as recommendations for future hack weeks in other disciplines (supplementary materials provide additional details on the practical aspects of organizing these events).

### 2017-12-19 sections/participants.tex

--- 2017-12-07 sections/participants.tex
@@
 No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate the different steps in their procedure.
-For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded.
+For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded (see also the Supplementary materials for more details on how the previous hack weeks approached this problem).
 

### 2017-12-19 sections/why.tex

@@
 It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
-Beginner participants may attend primarily to learn a new technique and may not have an explicit plan for project based work, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress.
+Beginner participants may attend primarily to learn a new technique and may not have an explicit plan for project based work, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives for a hack week, see supplementary materials)

### 2017-12-21 sections/results.tex

@@
 \begin{figure*}[h!]
-\begin{center}
+%\begin{center}
+\begin{subfigure}[t]{\textwidth}
+\centering
+\caption{}
 \includegraphics[width=11cm]{fig/eval_techskills.eps}
+\end{subfigure}
+
+\begin{subfigure}[t]{\textwidth}
+\centering
+\caption{}
+\includegraphics[width=11cm]{fig/longitudinal_visualization.eps}
+\end{subfigure}
+
+\begin{subfigure}[t]{\textwidth}
+\centering
+\caption{}
 \includegraphics[width=11cm]{fig/eval_collab.eps}
+\end{subfigure}
+
+\begin{subfigure}[t]{\textwidth}
+\centering
+\caption{}
 \includegraphics[width=11cm]{fig/eval_openscience.eps}
+\end{subfigure}
+
 \caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
 \textbf{For GHW and NHW, response rates were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
-At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.} We report here about results in three different domains: the development of technical skills (top), collaboration and learning (middle), and shifts in attitudes towards reproducibility and open science (bottom)}
+At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
+We report here about results in three different domains: the development of technical skills (a and b), collaboration and learning (c), and shifts in attitudes towards reproducibility and open science (d)}}
 \label{fig:survey}
-\end{center}
+%\end{center}
 \end{figure*}
@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016, we find that \textbf{most} participants self-report successful learning outcomes in new topics, tools or methods (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, top panel, left).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist ( Figure \ref{fig:survey}, top panel, right and middle).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016, we find that \textbf{most} participants self-report successful learning outcomes.
+in new topics, tools or methods (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a), left).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (a), right and middle).
+\textbf{More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (b), left).
+We asked participants to subjectively rate their knowledge before the hack week, and find skill levels to be broadly distributed.
+This is expected, given the goal of diversity during the selection stage, but we also note that there might be systematic effects, since the way the question was posed may be subject to individual biases.
+We then asked participants to rate their learning outcomes at the hack weeks in this category, and find that most participants have positive learning outcomes at all three hack weeks for data visualization, but that outcomes vary strongly, as is expected, too, with a group that is this diverse \textit{a priori}.}
 %While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.

### 2017-12-21 sections/what.tex

--- 2017-12-05 sections/what.tex
@@
 One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
-This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter}.
+
+This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter} (see supplementary material).
 Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.

### 2017-12-21 sections/what.tex
Added a bit of text to the hackathon section

--- 2017-12-09 sections/what.tex
@@
 
+The hack week model sets itself apart from other events notably in the goals that are pursued (see also Figure \ref{fig:hackspectrum}).
+Many industry hackathons are focused on delivering solutions to a specific problem or challenge on a very short timescale.
+Hack weeks are generally more open-ended than these hackathons and allow for a wider variety of outcomes and goals among participants.
+Within academia, the majority of venues for scientific collaboration, learning and communication fall into one of three categories: traditional conferences, summer schools, and, more recently, unconferences.
+All three types share elements with hack weeks, but have goals that are more narrowly defined.
+Traditional conferences are focused on the exchange of recent results, as well as networking.
+However, collaborative components are usually absent, or restricted to short time intervals (e.g.\ coffee breaks and ``hack days'').
+Unconferences are less structured and allow for greater input from participants, but are generally less focused on knowledge transfer and learning, but on achieving measurable outcomes.
+However, these outcomes are often in the form of policies, planning and white papers, whereas a hack week encourages participants to directly produce outcomes such as a paper, a visualization, or code.
+
+Three design considerations distinguish the hack week from the summer school model: (1) hack week transfer knowledge across many levels of seniority, with participants routinely leading tutorials, in part due to the relative newness of the content and the distributed knowledge base; (2) hack weeks contain a significant amount of unstructured time, creating the conditions for improvised exchanges and project development; and (3) hack weeks intentionally priortize participant diversity because this is known to maximize the likelihood for the co-creation of learning during the event.
+
 We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
 One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
+
 This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter} (see supplementary material).
-Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.
+%Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.

### 2017-12-21 sections/participants.tex
Added sentence to participants section referring to supplementary materials

@@
 
+We note that participant selection is only the first step in creating a successful hack week.
+An equal amount of thought and effort should be put into how to facilitate tutorials, break-out sessions and project work such that all participants can successfully and comfortably participate.
+In the supplementary materials, we give additional suggestions for successful facilitation.
 %One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to . 

### 2017-12-28 sections/design.tex
Minor wording changes

@@
 Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
-Workspaces must also accommodate interactive project work where small teams can gather and work together uninterrupted by other groups.    
-Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
-However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
-Fortunately, many universities are experimenting with other types of spaces: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, where seating arrangements can be flexibly modified and group activities can be more readily undertaken.
-There is a natural tension between keeping the group together and providing physically separated break-out spaces.
-On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas.
-On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical for hack week projects.
-It may hence be advisable to allow for diffusion of the group into adjacent rooms, while providing ample venue for the entire group to congregate for tutorials, breaks and reports.
+Workspaces must also accommodate interactive project work where small teams can gather and work together.    
+%Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
+%However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
+Fortunately, many universities are experimenting with other types of spaces: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.
 
@@
 Hack week outcomes depend strongly on the interest and engagement of participants.
-Some attendees, usually those with a strong background in hackathons and their specific topic of study, arrive with the goal of writing a scientific article.
-Other attendees plan to learn a specific topic, such as machine learning, or to analyze a specific data set that relates to the tools covered in the tutorials.
+Some attendees arrive with the goal of writing a scientific article.
+Other attendees plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
 This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
@@
 Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
-In particular the presence of impostor syndrome experienced by many participants is important to take this into account during workshop design (see also supplementary materials for concrete suggestions).}
+In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also supplementary materials for concrete suggestions).}
 %can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it

### 2017-12-28 sections/participants.tex
Minor wording changes

--- 2017-12-19 sections/participants.tex
@@
 In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from participants.
-Therefore in order to maximize learning outcomes and the likelihood for collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
+Therefore in order to maximize learning outcomes and collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
 
 In our experience, a participant group that is diverse across categories of diversity, gender, discipline and career track helps to ensure we meet these objectives.
-To achieve this diversity, we advocate for a selection process that is as transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
-Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
+To achieve this diversity, we advocate for a selection process that is as quantitative and transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
+%Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
 
@@
 In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 
-Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012} of an applicant, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
+Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
 A merit selection could then perform via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
@@
 The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
-For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
+%For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
@@
 Biased participant sets may still result from selection procedures that fail to include crucial categories.
-For example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority, and impossible to correct gender bias in the pool of applicants, if no information is available about the gender of participants.
+%For example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority, and impossible to correct gender bias in the pool of applicants, if no information is available about the gender of participants.
 
@@
 In this case, it may be beneficial to include demographic characteristics in the first stage. 
+Because systemic biases likely also enter at the application stage (where underrepresented groups may be less likely to apply) organizers should consider oversampling traditionally disenfranchised groups compared to the population of applicants.
 
-No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate the different steps in their procedure.
-For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded (see also the Supplementary materials for more details on how the previous hack weeks approached this problem).
+No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate the different steps in their procedure%.
+For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded
+(see also the Supplementary materials for more details on how the previous hack weeks approached this problem).
 

### 2017-12-28 sections/results.tex
Minor wording changes

@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016, we find that \textbf{most} participants self-report successful learning outcomes.
-in new topics, tools or methods (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a), left).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that \textbf{most} participants self-report successful learning outcomes in new topics, tools or methods (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a), left).
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (a), right and middle).
@@
 We asked participants to subjectively rate their knowledge before the hack week, and find skill levels to be broadly distributed.
-This is expected, given the goal of diversity during the selection stage, but we also note that there might be systematic effects, since the way the question was posed may be subject to individual biases.
-We then asked participants to rate their learning outcomes at the hack weeks in this category, and find that most participants have positive learning outcomes at all three hack weeks for data visualization, but that outcomes vary strongly, as is expected, too, with a group that is this diverse \textit{a priori}.}
+This is expected, given the goal of diversity during the selection stage.
+We find that most participants have positive learning outcomes at all three hack weeks for data visualization, but that outcomes vary strongly, as is expected, too, with a group that is this diverse.
+We also note that there might be systematic effects and biases due to the subjectivity of rating knowledge and learning outcomes.}
 %While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
 %Another important goal of the hack weeks was centred on community building and fostering collaborations.
-The majority of participants felt that their contributions to their hack teams was valued, and that they built valuable connections to other researchers (Figure \ref{fig:survey}, middle panel, middle).
+The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, middle panel, middle).
 This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure  \ref{fig:survey}, middle panel, right).
@@
 \textbf{It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
-In particular, participants from all career stages between graduate students and tenured faculty report similar engagement in teaching.
-This supports our hypothesis that compared to e.g.\ a traditional summer school, a hack week is less hierarchical and encourages lateral knowledge transfer between participants at different career stages.
-We similar find no correlations with gender identity or race/ethnicity.
-Responses to the other two questions regarding whether participants' contributions to hack teams were valued and whether they made valuable connections at the hack week were similarly evenly distributed across career stages, gender and racial/ethnic identities.}
+Participants at all career stages report similar engagement in teaching.
+This suggests there is some evidence for our hypothesis that compared to e.g.\ a traditional summer school, a hack week is less hierarchical and encourages lateral knowledge transfer between participants at different career stages.
+We similarly find no correlations with gender identity or race/ethnicity.}
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed this practice before the event (Figure \ref{fig:survey}, bottom panel, middle).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed \textbf{a regular practice of publishing code} before the event (Figure \ref{fig:survey}, bottom panel, middle).
 When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
@@
 %This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
-Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, closely followed by neuroscience (95\%), while there was somewhat less of an impact on participants' attitudes in astronomy (72\%).
-Overall, our results show that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
+Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
+Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
 \textbf{While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, including open-source software, data sharing and reproducibility.
@@
 
-From the very nature of the activities that we encourage in hack weeks, participants in these events produce digital records of their research online.
+%From the very nature of the activities that we encourage in hack weeks, participants in these events produce digital records of their research online.
 This means that it will be fairly straightforward to evaluate the long-term impact of these activities on participants' productivity (e.g., through contributions to open-source software) in the future.
-Because all three events are relatively recent, it is still early to evaluate such long-term outcomes, as well as others including publications and collaborations resulting from these events.
+Because all three events are relatively recent, it is still early to evaluate long-term outcomes, as well as others including publications and collaborations resulting from these events.
 There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
@@
 \subsubsection*{Example 1: Astro Hack Week}
-In 2015, a small team used the opportunity of AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of time series analysis algorithms often used in X-ray astronomy.
-The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects, role models of successful projects, and an environment encouraging scientific risk taking.
+In 2015, a small team used AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of time series analysis algorithms often used in X-ray astronomy.
+The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects and role models of successful projects.
 Astro Hack Week enabled participants to seed a new collaboration around a software project needed by the larger community.
-Since its beginnings at Astro Hack Week, Stingray has matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.
+Stingray has since matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.
 \subsubsection*{Example 2: Geo Hack Week}

### 2017-12-28 sections/what.tex
Minor wording changes

--- 2017-12-09 sections/what.tex
@@
 Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
-Hackathons originated from the open-source software movement, and have historically focused on software development, particularly in the form of coding sprints, and technology design as a way to motivate innovation, eventually being adopted also within the technology industry.
+Hackathons originated from the open-source software movement, and have historically focused on software development, eventually being adopted also within the technology industry.
 In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
@@
 
-Our past hack week events follow most closely the communal hackathon model, as it applies to scientific communities of practice.
+Our past hack week events follow most closely the communal hackathon model.
 Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
@@
 
+The hack week model sets itself apart from other events notably in the goals that are pursued (see also Figure \ref{fig:hackspectrum}). 
+Many industry hackathons are focused on delivering solutions to a specific problem or challenge on a very short timescale. 
+Hack weeks are  more open-ended than these hackathons and allow for a wider variety of outcomes and goals among participants.
+Within academia, the majority of venues for scientific collaboration, learning and communication fall into one of three categories: traditional conferences, summer schools, and, more recently, unconferences.
+All three types share elements with hack weeks, but have goals that are more narrowly defined.
+Traditional conferences are focused on the exchange of recent results, as well as networking.
+However, collaborative components are usually absent, or restricted to short time intervals (e.g.\ coffee breaks).
+Unconferences are less structured and allow for greater input from participants, but are generally less focused on knowledge transfer and learning, but on achieving measurable outcomes, often in the form of policies, proposals and white papers.
+%However, these outcomes are often in the form of policies, planning and white papers, whereas a hack week encourages participants to directly produce outcomes such as a paper, a visualization, or code.
+
+Three design considerations distinguish the hack week from the summer school model: (1) hack week transfer knowledge across many levels of seniority, with participants routinely leading tutorials; (2) hack weeks contain a significant amount of unstructured time, creating the conditions for improvised exchanges and project development; and (3) hack weeks intentionally priortize participant diversity because this is known to maximize the likelihood for the co-creation of learning during the event.
+Additionally, hack weeks tend to focus on \textit{established} knowledge, whereas tutorials at hack weeks favour cutting-edge methodology and experimental approaches.
+
 We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
-One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
-This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter} (see supplementary material).
-Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.
+%One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
+
+%This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter} (see supplementary material).
+%Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.

### 2017-12-28 sections/results.tex
Updated Figure 2, fiddled some with text

@@
 %\begin{center}
-\begin{subfigure}[t]{\textwidth}
+%\begin{subfigure}[t]{\textwidth}
 \centering
-\caption{}
-\includegraphics[width=11cm]{fig/eval_techskills.eps}
-\end{subfigure}
-
-\begin{subfigure}[t]{\textwidth}
-\centering
-\caption{}
-\includegraphics[width=11cm]{fig/longitudinal_visualization.eps}
-\end{subfigure}
-
-\begin{subfigure}[t]{\textwidth}
-\centering
-\caption{}
-\includegraphics[width=11cm]{fig/eval_collab.eps}
-\end{subfigure}
-
-\begin{subfigure}[t]{\textwidth}
-\centering
-\caption{}
-\includegraphics[width=11cm]{fig/eval_openscience.eps}
-\end{subfigure}
+%\caption{}
+\includegraphics[width=\textwidth]{fig/f2.eps}
+%\end{subfigure}
 
@@
 At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
-We report here about results in three different domains: the development of technical skills (a and b), collaboration and learning (c), and shifts in attitudes towards reproducibility and open science (d)}}
+We report here about results in three different domains: the development of technical skills (a -- e), collaboration and learning (f,g), and shifts in attitudes towards reproducibility and open science (h -- j)}}
 \label{fig:survey}
@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that \textbf{most} participants self-report successful learning outcomes in new topics, tools or methods (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a), left).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (a), right and middle).
-\textbf{More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (b), left).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that \textbf{most} participants self-report successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (b, c) ).
+\textbf{More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
 We asked participants to subjectively rate their knowledge before the hack week, and find skill levels to be broadly distributed.
@@
 We also note that there might be systematic effects and biases due to the subjectivity of rating knowledge and learning outcomes.}
-%While each hack week also probed participants' attitudes in more detail toward specific topics, methods, and modes of learning, the hack weeks differed substantially on that level, making the results difficult to compare.
-%Another important goal of the hack weeks was centred on community building and fostering collaborations.
-The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, middle panel, middle).
-This is especially true for Neuro Hack Week, where more than 64\% of participants strongly agreed that they formed valuable connections at NHW (Figure  \ref{fig:survey}, middle panel, right).
+
+The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, (g)), especially at Neuro Hack Week, where more than 64\% of participants strongly agreed with this statement.
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\% for combined responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, middle panel, left), though responses are not as unequivocal as they are in some of the other categories.
-This is expected: participants new to the field may participate to learn rather than to teach.
+We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\%; Figure \ref{fig:survey}, (f)), though responses are not as unequivocal as they are in some of the other categories.
+Similarly to learning outcomes, this is expected given the range of skill levels at the workshops.
 \textbf{It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
@@
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, bottom panel, left) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed \textbf{a regular practice of publishing code} before the event (Figure \ref{fig:survey}, bottom panel, middle).
-When asking participants whether they had made their code or data openly available in the past, the overall behaviour reflects how general conventions and attitudes differ in different fields.
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed \textbf{a regular practice of publishing code} before the event (Figure \ref{fig:survey}, (i)).
+The overall behaviour reflects how general conventions and attitudes differ in different fields.
 Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
 %This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
-Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science: again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
+Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science (Figure \ref{fig:survey}, (j)): again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
 Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
-\textbf{While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, including open-source software, data sharing and reproducibility.
+\textbf{While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks like open-source software and data sharing.
 In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.}
@@
 \subsubsection*{Example 1: Astro Hack Week}
-In 2015, a small team used AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested, well-documented implementations of time series analysis algorithms often used in X-ray astronomy.
+In 2015, a small team used AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested implementations of time series analysis algorithms often used in X-ray astronomy.
 The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects and role models of successful projects.
 Astro Hack Week enabled participants to seed a new collaboration around a software project needed by the larger community.
-Stingray has since matured into an enduring collaboration within the community with five active maintainers, a number of contributors and four Google Summer of Code projects.
+Stingray has since matured into an enduring collaboration within the community with five active maintainers and four Google Summer of Code projects.
 \subsubsection*{Example 2: Geo Hack Week}
@@
 %They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four different researchers from four different institutions in two different countries) continued to work on this project remotely after the end of Neuro Hack Week, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.
+The team (composed of four different researchers from four institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.

### 2017-12-28 sections/participants.tex

@@
 \section*{Audience and Participant Selection}
-
+\begin{bf}
 Hack weeks differ from traditional conferences or summer schools in that knowledge transfer occurs across many levels of seniority, disciplinary boundaries, and novelty of the topics discussed.
@@
 (see also the Supplementary materials for more details on how the previous hack weeks approached this problem). 
-
+\end{bf}
 %One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to . 

### 2017-12-28 sections/why.tex

--- 2017-09-30 sections/why.tex
@@
 \item{\textit{Education and Training}:
-While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component that entails extensive discussion on reproducible research and open science practices.
-Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices in their respective fields.
+While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component that entails discussions about reproducible research.
+Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices.
 This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
@@
 \item{\textit{Community Building}: Hack weeks provide a tremendous opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
-These events allow computationally minded researchers to break from the isolation of their academic departments, build connections and spark new collaborations.}
+These events allow computationally minded researchers to break from the isolation of their academic departments and spark new collaborations.}
 
@@
 It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
-Beginner participants may attend primarily to learn a new technique and may not have an explicit plan for project based work, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress.
+Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see supplementary materials).

### 2017-12-28 supplementary_material.tex

@@
 A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.
+There is a natural tension between keeping the group together and providing physically separated break-out spaces.
+On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas.
+On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical for hack week projects.
+It may hence be advisable to allow for diffusion of the group into adjacent rooms, while providing ample venue for the entire group to congregate for tutorials, breaks and reports.
 There should be plenty of power outlets for participants to plug in their computers.
@@
 Event size (i.e.
-the number of attendees) and event durations are important to consider during the design and planning phase of a hack week.
+the number of attendees) and event duration are important to consider during the design and planning phase of a hack week.
 All hack weeks held so far have been five days, with group sizes between 40-60 participants.
@@
 It is helpful to design the survey early on and allow for some time near the end of the workshop for participants to fill it out there; in our experience, this leads to much higher response rates than if the survey is distributed after the end of the workshop.
-While some improvements and changes in the survey design from year to year are inevitable, it is advisable to set a number of core questions that do *not* change from year to year.
+While some improvements and changes in the survey design from year to year are inevitable, it is advisable to set a number of core questions that do \textit{not} change from year to year.
 These questions will allow organizers to assess the workshop's function within their community, and assess how changes to the format between one year and another may affect outcomes of interest.

### 2018-01-02 supplementary_material.tex
minor wordsmithing

@@
 To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities.
-These scientists then reviewed their observations and and gave valuable feedback about the hack week outcomes.
+These scientists then reviewed their observations and gave valuable feedback about the hack week outcomes.
 In order to be able to track and study participants, approval from an Internal Review Board is required.
@@
 Is the community to be invited more software-focused or should outcomes skew towards publishable results?
-Does the workshop aim to maximize its output by carrying best practices into universities and institutions that may be particularly underrepresented?
-These questions, among others, should have a very clear answer before any further workshop organization takes place.
+Does the hack week aim to maximize its output by carrying best practices into universities and institutions that may be particularly underrepresented?
+These questions, among others, should have a very clear answer before any further organization takes place.
 
@@
 the number of attendees) and event duration are important to consider during the design and planning phase of a hack week.
-All hack weeks held so far have been five days, with group sizes between 40-60 participants.
+All hack weeks held so far have had group sizes between 40-60 participants.
 The number of attendees resulted from a number of different design considerations.
@@
 Another important consideration is representation along the many axes of diversity we considered.
-For example, a single undergraduate participant might feel like an outsider; admitting three or four can mitigate alienation effects.
-Similarly, in order to foster active collaboration among many different subfields, we ensured admitting participants with many different types of expertise, some of which overlapped.
+For example, whereas a single undergraduate participant might feel like an outsider, admitting three or four can mitigate alienation.
+Similarly, in order to foster active collaboration among many different subfields, we endeavored to admit participants with many different types of overlapping expertise.
 Having a set of participants that is too small would interfere both with these goals of representation and of fostering interdisciplinary collaborations and networking, thus we would advise admitting at least $\sim$30 or so participants if these objectives are an important part of the hack week.
@@
 All hack weeks organized up to now have been five days.
-This duration was chosen with several different design objectives in mind.
-In particular, it was chosen to allow both learning and project work at the same time.
-Participants often come with ambitious project ideas, or with the goal of taking newly learned methods and applying them to their data sets right away.
+This duration was chosen to allow enough time for both learning and project work at the same time.
+Participants often arrived with ambitious project ideas, or with the goal of taking newly learned methods and applying them to their data sets right away.
 Thus many projects tended to extend over multiple days.
@@
 Secondly, much energy is expended early on in the design of the application form, which often use a combination of open-answer and multiple-choice questions to assess candidates' suitability, skill levels and backgrounds.
-Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' or ``expert'' are have no absolute scaling and may have different meanings to different applicants.
+Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' or ``expert'' have no absolute scaling and may have different meanings to different applicants.
 They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates (the Dunning Kruger effect).
@@
 For this reason, a hack week should have a clear and enforcable code of conduct, that can guide the behavior of participants, and help organizers maintain a professional and welcoming environment.
-It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in advance.
+It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in wadvance.
 A part of this could and should be the recognition that participants (and, indeed, members of the organizing committee) may occasionally violate the code of conduct in minor ways, and allow for tools and mechanisms to resolve these infractions.
@@
 For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome, defensive and inhibit learning on a larger scale.
-Even if no code of conduct infarctions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are thoughtful of creating a welcoming environment.
-This by itself helps create a welcoming environment.
+Even if no code of conduct infarctions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are thoughtful of creating a welcoming environment, which in turn feeds back to the creation of such a space.
 
@@
 In some cases, tutorials require substantial software installation and data.
-Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometime so complicated as to slow down teaching, and hold back.
-In some cases, we have experimented using \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \cite{holdgraf2017portable}.
-Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \cite{Holdgraf2017-pd, Titus_Brown_undated-pc}
+Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometime so complicated that it slows down teaching.
+In some cases, we have experimented with \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \cite{holdgraf2017portable}.
+Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \cite{Holdgraf2017-pd, Titus_Brown_undated-pc}.
 
@@
 Because of their practical nature, however, software requirements should be announced as early and clearly as possible, with the expectation that participants will have installed necessary software in advance.
-It may be helpful to designate some time early on in the meeting to trouble-shoot installations before tutorials begin in order to avoid loosing significant time during the actual tutorials.
+It may be helpful to designate some time early on in the meeting to trouble-shoot installations before tutorials begin in order to avoid losing significant time during the actual tutorials.
 
@@
 It also helps to make them aware that participants may be taking the role of teaching assistants during exercises and include them in the teaching, rather than see them as adversaries who may seek to correct them.
-Conversely, experts may be used to ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and to encourage questions, especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome.
+Conversely, experts may ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and to encourage questions, especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome.
 At Astro Hack Week, we encourage the organizing committee and experienced participants to ask questions, particularly when they know the answer but think the concept may not yet be clear to parts of the audience.
@@
 At all hack weeks, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
-Clearly communicating values and expectations at the beginning for the week can help setting the stage for an environment that a large range of different personalities can thrive in.
+Clearly communicating values and expectations at the beginning for the week can help setting the stage for an environment in which a large range of different personalities can thrive.
 This includes both communication of the code of conduct, putting an emphasis on collaborative rather than competitive behaviour, as well as a discussion of potential issues participants might experience, such as impostor syndrome (see also section below).
@@
 In one activity, participants faced one another and were given a brief statement inviting a two minute conversation, after which the rows rotated to maximize the mixing of ideas.
-Invitations included statements such as ``I belong at geohack because...'' or ``A challenge I hope we can make progress on is...''.
+Invitations included statements such as ``I belong at geohack week because...'' or ``A challenge I hope we can make progress on is...''.
 Importantly, event organizers joined in the activity, which we found helped participants feel as if they were not alone in facing some of the challenges inherent in data science studies.
@@
 Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \cite{Clance1978-ef}.
-Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS: participants might feel the need to be experts in multiple aspects of the activities pursued during a hack week: expertise in a scientific domain, as well as expertise in a the variety of technical tools used.
+Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS, namely that participants feel the need to be experts in multiple scientific and technical aspects of the activities pursued during a hack week.
 This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
@@
 We also invited our organizing committees to model open communication about their own handling of IS through the course of their career.
-We found that transparency about one's own uncertainty, espcially from community leaders, helped create a hospitable atmosphere and break down barriers.
+We found that transparency about one's own uncertainty, espcially from community leaders, helped create a hospitable atmosphere and removed barriers.
 

### 2018-01-05 sections/intro.tex

--- 2017-09-30 sections/intro.tex
@@
 \label{sec:introduction}
-\dropcap{A}s data becomes cheaper to gather and store, research across a wide range of disciplines has become increasingly reliant on computational workflows involving a familiarity with aspects of statistical modeling, machine learning, scalable computation, and related skills.
-In addition, the recent reproducibility crises in several scientific fields has led to the growing realization that improving awareness of open science and reproducibility as well as practical skills in making research reproducible is essential to scientific progress \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
-Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts such as workshops (an overview and typography of such efforts in the data science context can be found in \cite{demasi2017}).
+\dropcap{A}s data become cheaper to gather and store, researchers have become increasingly reliant on computational workflows requiring skills in statistical modeling, machine learning and scalable computation.
+In addition, the recent reproducibility crises in several scientific fields has motivated the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
+Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts such as workshops \cite{demasi2017}.
 Well-known examples are the Software and Data Carpentry workshops providing training in research computing skills through a volunteer instructor program \cite{b:wilson-swc-lessons-2016,teal2015data}.
-At the same time, there has been a rise in the number of domain-specific courses focusing on statistics and computation within their field.
+At the same time, there has been a rise in the number of statistical and computational courses designed for specific scientific disciplines.
 Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, as well as a variety of project-focused (rather than pedagogical) meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons that take place in conjunction with meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-Generally, pedagogically-focused events follow a classic academic model where novices learn new skills from experts, while project-focused workshops emphasize collaborative activities using existing skills.
-A disadvantage of the pedagogical model is that it can tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
+Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+In general, events to date have been designed withing either a pedagogical or project-based framework.
+Pedagogically-focused events follow a classic academic model where novices learn new skills from experts, while project-focused workshops emphasize collaborative activities using existing skills.
+A disadvantage of the pedagogical model is that it tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
 A disadvantage of the project model is the common perception that the week is designed for technical experts, which may discourage others from attending.
-In 2014, we initiated an alternative model of ``Hack Weeks'' that try to fill the gaps between these models.
-These are week-long events that combine pedagogy (often focused on statistical and computational techniques) together with time for hacks and creative projects, and with the goal of encouraging collaboration and learning among people at various stages of their career.
+In 2014, we initiated an alternative model of ``Hack Weeks'' that aim to combine the best of each of these models.
+These week-long events combine pedagogy (often focused on statistical and computational techniques) with hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
 
@@
 As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
-Below we will share some of the philosophy behind the hack week model, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.
+Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines (supplementary materials provide additional details on the practical aspects of organizing these events).

### 2018-01-05 sections/what.tex

@@
-\section*{What is a hackathon?}
+\section*{What is a hack week?}
 
-Hackathons are time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}.
-Hackathons originated from the open-source software movement, and have historically focused on software development, eventually being adopted also within the technology industry.
-In recent years, hackathons have expanded into a model for intensive short-term collaboration across disciplinary and topical boundaries.
-In addition, because of their focus on participatory engagement, hackathons provide numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
-With this in mind, hackathons around scientific topics, designed to foster collaboration \cite{Groen2015-cj,Moller2013-ah}, or provide an opportunity to learn \cite{Kienzler2015-zu,Lamers2014-xf}, are becoming more common.
-
-However, in addition to these goals, core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
-Building on these core elements, there are various implementations of the hackathon concept with respect to the overall purpose, mode of participation, style of work environment and motivation \cite{Drouhard2017}:
+Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. 
+Our focus on both pedagogy and open project work aims to fill an educational space we feel is not currently addressed by existing models (Figure \ref{fig:hackspectrum}). 
+The hackathon, a time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events. 
+Hackathons originated from the open-source software movement and have historically focused on software and technology development. 
+In recent years hackathons have evolved into a model providing opportunities for intensive, interdisciplinary collaboration \cite{Groen2015-cj,Moller2013-ah} and education \cite{Kienzler2015-zu,Lamers2014-xf} in the sciences.
+Core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
+Building on these core elements hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}.
 ``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
@@
 
-Our past hack week events follow most closely the communal hackathon model.
-Our approach aims to combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
-Within the communal model we see these tools being implemented across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice (Figure \ref{fig:hackspectrum}).
+Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning, forming the second axis of our events. 
+Teaching within our events draws from a spectrum of approaches such as those traditionally employed in summer schools and software workshops. 
+However, by embedding these approaches in a more open hackathon framework, we can explore unique ways of enhance the traditional summer school model.
+For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
+Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over the transfer established knowledge. 
+Finally, the social structure of hackathons enable us to priortize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.
+
+We have found that these multiple dimensions of the hack week model provide signifcant flexibilty across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
 For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
@@
 
-The hack week model sets itself apart from other events notably in the goals that are pursued (see also Figure \ref{fig:hackspectrum}). 
-Many industry hackathons are focused on delivering solutions to a specific problem or challenge on a very short timescale. 
-Hack weeks are  more open-ended than these hackathons and allow for a wider variety of outcomes and goals among participants.
+We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
+
+%\begin{bf}
+%Many industry hackathons are focused on delivering solutions to a specific problem or challenge on a very short timescale. 
+%Hack weeks are  more open-ended than these hackathons and allow for a wider variety of outcomes and goals among participants.
 Within academia, the majority of venues for scientific collaboration, learning and communication fall into one of three categories: traditional conferences, summer schools, and, more recently, unconferences.
 All three types share elements with hack weeks, but have goals that are more narrowly defined.
-Traditional conferences are focused on the exchange of recent results, as well as networking.
+%Traditional conferences are focused on the exchange of recent results, as well as networking.
 However, collaborative components are usually absent, or restricted to short time intervals (e.g.\ coffee breaks).
 Unconferences are less structured and allow for greater input from participants, but are generally less focused on knowledge transfer and learning, but on achieving measurable outcomes, often in the form of policies, proposals and white papers.
-%However, these outcomes are often in the form of policies, planning and white papers, whereas a hack week encourages participants to directly produce outcomes such as a paper, a visualization, or code.
+%\end{bf}
 
-Three design considerations distinguish the hack week from the summer school model: (1) hack week transfer knowledge across many levels of seniority, with participants routinely leading tutorials; (2) hack weeks contain a significant amount of unstructured time, creating the conditions for improvised exchanges and project development; and (3) hack weeks intentionally priortize participant diversity because this is known to maximize the likelihood for the co-creation of learning during the event.
-Additionally, hack weeks tend to focus on \textit{established} knowledge, whereas tutorials at hack weeks favour cutting-edge methodology and experimental approaches.
-
-We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
 %One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.

### 2018-01-07 sections/methods.tex
Remove boldface for revision highlighting, add methods.

We performed post-attendance surveys for AHW, GHW and NHW in 2016 and 2017. All surveys contained general questions about attitudes towards the workshop as well as open science and reproducibility, and their own skills in statistical and computational methods, shared among all three surveys. The NHW and GHW surveys and the AHW 2017 survey were administered on site on the last day of the workshop; AHW participants in 2016 were e-mailed immediately after the workshop, with reminders after several weeks. All responses were collected within four weeks of the end of the workshop. The experimental procedures were approved by the Institutional Review Boards at UW, NYU and UC Berkeley. All participants gave their informed consent.

Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100\% in both years; the response rate for AHW was 71\% (35 out of 49) in 2016 and 82\% (37 out of 45) in 2017. The lower response rates for AHW can be explained by the generally lower rates of response for time-delayed survey in 2016. In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered. For the AHW surveys, we checked whether the demographics of the respondents different significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and find no significant difference in any of the categories except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees). Because both undergraduate and graduate students are at the beginning of their careers for whom AHW is likely the first exposure of many of the concepts around data science and open science, we believe that this discrepancy has little impact on the overall results presented here.

Participants were asked to respond to statements regarding these topics using a six-point Likert-type scale. All questions were anonymously recorded. No responses were discarded, and no pre-processing was performed on the data. We test for correlations between categorical variables using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cramer's V \cite{bergsma2013}. Only $p$-values with $p < 0.0018$ (equivalent to $p < 0.05$ corrected for $N=27$ trials) were reported as potential candidates. The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.

### 2018-01-07 sections/conclusions.tex

--- 2017-12-06 sections/conclusions.tex
@@
 As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
-Initial results from three events held in 2016 \textbf{and 2017} in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
+Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
 Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.
 We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.
+While we focus here on hack weeks in scientific fields, the concept could be extended to other areas, and is more generally useful in any area (1) where useful tools can be learned in short tutorials, (2) where results and outcomes can be produced on the timescale of a few days, and (3) that would benefit from collaborative approaches that cross traditional boundaries.
+Such areas could include for example the social sciences, music and art.

### 2018-01-07 sections/design.tex

@@
 This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
-\textbf{The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week.
+The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week.
 Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
-In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also supplementary materials for concrete suggestions).}
+In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also supplementary materials for concrete suggestions.
 %can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it

### 2018-01-07 sections/intro.tex

@@
 \label{sec:introduction}
-\dropcap{A}s data become cheaper to gather and store, researchers have become increasingly reliant on computational workflows requiring skills in statistical modeling, machine learning and scalable computation.
+As data become cheaper to gather and store, researchers have become increasingly reliant on computational workflows requiring skills in statistical modeling, machine learning and scalable computation.
 In addition, the recent reproducibility crises in several scientific fields has motivated the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.

### 2018-01-07 sections/participants.tex

--- 2017-12-28 sections/participants.tex
@@
 \section*{Audience and Participant Selection}
-
 Hack weeks differ from traditional conferences or summer schools in that knowledge transfer occurs across many levels of seniority, disciplinary boundaries, and novelty of the topics discussed.
@@
 (see also the Supplementary materials for more details on how the previous hack weeks approached this problem).
-
 %One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to . 

### 2018-01-07 sections/results.tex

@@
 \caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-\textbf{For GHW and NHW, response rates were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
+For GHW and NHW, response rates were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
 At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
-We report here about results in three different domains: the development of technical skills (a -- e), collaboration and learning (f,g), and shifts in attitudes towards reproducibility and open science (h -- j)}}
+We report here about results in three different domains: the development of technical skills (a -- e), collaboration and learning (f,g), and shifts in attitudes towards reproducibility and open science (h -- j)}
 \label{fig:survey}
@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that \textbf{most} participants self-report successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that most participants self-report successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (b, c) ).
-\textbf{More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
+More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
 We asked participants to subjectively rate their knowledge before the hack week, and find skill levels to be broadly distributed.
@@
 We find that most participants have positive learning outcomes at all three hack weeks for data visualization, but that outcomes vary strongly, as is expected, too, with a group that is this diverse.
-We also note that there might be systematic effects and biases due to the subjectivity of rating knowledge and learning outcomes.}
+We also note that there might be systematic effects and biases due to the subjectivity of rating knowledge and learning outcomes.
 
@@
 Similarly to learning outcomes, this is expected given the range of skill levels at the workshops.
-\textbf{It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
+It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
 Participants at all career stages report similar engagement in teaching.
 This suggests there is some evidence for our hypothesis that compared to e.g.\ a traditional summer school, a hack week is less hierarchical and encourages lateral knowledge transfer between participants at different career stages.
-We similarly find no correlations with gender identity or race/ethnicity.}
+We similarly find no correlations with gender identity or race/ethnicity.
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed \textbf{a regular practice of publishing code} before the event (Figure \ref{fig:survey}, (i)).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed a regular practice of publishing code before the event (Figure \ref{fig:survey}, (i)).
 The overall behaviour reflects how general conventions and attitudes differ in different fields.
@@
 Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
-\textbf{While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks like open-source software and data sharing.
-In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.}
+While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks like open-source software and data sharing.
+In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.
 

### 2018-01-08 sections/conclusions.tex
Minor changes before submitting

@@
 To address this imbalance, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
-Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields
+Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields.
 

### 2018-01-08 sections/intro.tex
Minor changes before submitting

--- 2018-01-05 sections/intro.tex
@@
 \dropcap{A}s data become cheaper to gather and store, researchers have become increasingly reliant on computational workflows requiring skills in statistical modeling, machine learning and scalable computation.
-In addition, the recent reproducibility crises in several scientific fields has motivated the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,frye2015,gezelter2015,baker2016}.
+In addition, the recent reproducibility crises in several scientific fields has motivated the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,baker2016}.
 Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts such as workshops \cite{demasi2017}.
@@
 Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-In general, events to date have been designed withing either a pedagogical or project-based framework.
+In general, events to date have been designed within either a pedagogical or project-based framework.
 Pedagogically-focused events follow a classic academic model where novices learn new skills from experts, while project-focused workshops emphasize collaborative activities using existing skills.
@@
 \begin{center}
-\includegraphics[width=9cm]{fig/HackSpectrum}
+\includegraphics[width=8cm]{HackSpectrum.pdf}
 \caption{Comparison of Extracurricular Workshop Models}

### 2018-01-08 sections/results.tex
Minor changes before submitting

@@
 %\caption{}
-\includegraphics[width=\textwidth]{fig/f2.eps}
+\includegraphics[width=\textwidth]{f2.eps}
 %\end{subfigure}
-
-\caption{{\bf Post-workshop surveys from three hack weeks}: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
+\caption{Post-workshop surveys from three hack weeks: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
 For GHW and NHW, response rates were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
@@
 Measuring the success of a hack week objectively is complicated by the variety of objectives that a hack week might have (see above). 
-Additionally, the participant-driven, open-ended format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
+Additionally, the participant-driven format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
 
@@
 Still, these measures ignore learning, community-building as well as networking outcomes, which can be assessed through post-workshop surveys.
-%It is in principle possible to measure gains in knowledge, networks and productivity within the pool of acceptable candidates for both those that attended the hack week and those that did not.
-%This would then provide a somewhat objective measure of the impact the hack week has had.
-%In practice, this has not been done for any of the hack weeks conducted thus far.
-Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications, new code and projects generated (see the following section).
-%Open-ended questions allow participants to provide feedback about outcomes, problems and goals that organizers had not anticipated.
-
-%If hack week organizers plan to conduct research that involves hack week participants (for example, using assessments and evaluations in a paper such as this one) it is important to obtain approval from the Institutional Review Board (IRB), or equivalent body that approves research with human participants at the institution hosting the hack week.
-%Though this research would usually fall under the category of ''minimal risk``, it is still important to establish procedures to manage these data, and to obtain informed consent.
-%In particular, in some cases hack week organizers may be interested in studying not only the participants in the hack week, but also the applicants who did not end up participating.
-%It is important to establish procedures to conduct such studies and to obtain approval from an IRB.
+Here, we have taken an approach that combines these metrics: we start with survey results, and anecdotally report about publications and projects generated (see the following section).
 
@@
 This suggests there is some evidence for our hypothesis that compared to e.g.\ a traditional summer school, a hack week is less hierarchical and encourages lateral knowledge transfer between participants at different career stages.
+However, we note that the sample sizes per category are relatively small, and future work is needed to confirm this hypothesis.
 We similarly find no correlations with gender identity or race/ethnicity.
@@
 We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed a regular practice of publishing code before the event (Figure \ref{fig:survey}, (i)).
-The overall behaviour reflects how general conventions and attitudes differ in different fields.
+The overall behaviour hints at how general conventions and attitudes differ in different fields.
 Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
-%This implies that hack weeks can have the highest impacts in field where a priori engagement in reproducibility efforts is low and significant progress can be made towards changing researchers' attitudes during a collaborative workshop.
 Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science (Figure \ref{fig:survey}, (j)): again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
@@
 
-%From the very nature of the activities that we encourage in hack weeks, participants in these events produce digital records of their research online.
-This means that it will be fairly straightforward to evaluate the long-term impact of these activities on participants' productivity (e.g., through contributions to open-source software) in the future.
 Because all three events are relatively recent, it is still early to evaluate long-term outcomes, as well as others including publications and collaborations resulting from these events.
@@
 \subsubsection*{Example 3: Neuro Hack Week}
-Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies, particularly in studies of children or patients, as they are more likely to move.
+Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies.
 During NHW 2016 one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
@@
 %They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four different researchers from four institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.
+The team (composed of four researchers from four institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.

### 2018-01-08 supplementary_material.tex
Minor changes before submitting

@@
 
-\author{Daniela Huppenkothen$^1$, Anthony Arendt$^2$, David W. Hogg$^1$, Karthik Ram$^3$, Jake VanderPlas$^2$, and Ariel Rokem$^2$}
+\author{Daniela Huppenkothen$^{1,2,3}$, Anthony Arendt$^{4,5}$, David W. Hogg$^{2,1,6,7}$, Karthik Ram$^8$, Jake VanderPlas$^5$, and Ariel Rokem$^5$}
 
@@
 \begin{affiliations}
- \item Center for Data Science, New York University
- \item The University of Washington eScience Institute
- \item The Berkeley Institute for Data Science, UC Berkeley
+\item{Center for Data Science, New York University, 65 5th Avenue, 7th Floor, New York, NY 10003, USA}
+\item{Center for Cosmology and Particle Physics, New York University, 726 Broadway, 10th Floor, New York, NY 10003, USA}
+\item{DIRAC Institute, Department of Astronomy, University of Washington, 3910 15th Ave NE, Seattle, WA 98195, USA}
+\item{Polar Science Center/Applied Physics Laboratory, University of Washington, 1013 NE 40th Street, Box 355640, Seattle, WA 98105-6698}
+\item{The University of Washington eScience Institute, The WRF Data Science Studio, Physics/Astronomy Tower, 6th Floor, 3910 15th Ave NE, Campus Box 351570, University of Washington, Seattle, WA 98105, USA}
+\item{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}
+\item{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, USA}
+\item{Berkeley Institute for Data Science \& Berkeley Initiative in Global Change Biology, University of California, Berkeley,  Berkeley CA 94720}
 \end{affiliations}
@@
 In order to be able to track and study participants, approval from an Internal Review Board is required.
-This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally don't involve human subject research.
+This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally do not involve human subject research.
 Ideally, participants should consent to be contacted at the application stage, which opens up the potential of tracking both participants and non-participants and allows for evaluation of the hack week.
@@
 The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access to visitors from other institutions.
-Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available.
+Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available?
 
@@
 
-The hack weeks presented here used different approaches, that nevertheless several common themes and ideas.
+The hack weeks presented here used different approaches, that nevertheless share several common themes and ideas.
 In particular, they focus on making the selection procedure as quantitative as possible and use the data collected during each stage of the process to critically evaluate the efficacy of the process and its subject to human bias.
@@
 However, disparities in backgrounds, knowledge and experience, together with the close collaboration much of the hacking requires, creates the potential for conduct that might offend participants, and preclude them from participating fully.
-For this reason, a hack week should have a clear and enforcable code of conduct, that can guide the behavior of participants, and help organizers maintain a professional and welcoming environment.
-It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in wadvance.
+For this reason, a hack week should have a clear and enforceable code of conduct, that can guide the behavior of participants, and help organizers maintain a professional and welcoming environment.
+It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in advance.
 A part of this could and should be the recognition that participants (and, indeed, members of the organizing committee) may occasionally violate the code of conduct in minor ways, and allow for tools and mechanisms to resolve these infractions.
@@
 For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome, defensive and inhibit learning on a larger scale.
-Even if no code of conduct infarctions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are thoughtful of creating a welcoming environment, which in turn feeds back to the creation of such a space.
+Even if no code of conduct infractions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are thoughtful of creating a welcoming environment, which in turn feeds back to the creation of such a space.
 
@@
 Depending on the content, some instructors convey concepts during the tutorial via ``live-coding'' while others display static code and their results.
-Either way, we find divding code into short blocks, each with a specific task, helps participants from becoming saturated with informaiton.
+Either way, we find dividing code into short blocks, each with a specific task, helps participants from becoming saturated with information.
 In some cases, tutorials require substantial software installation and data.
@@
 With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
-More generally, being open about the presence and prevalence of IS can help participants feel more at ease.%(e.g.
+More generally, being open about the presence and prevalence of IS can help participants feel more at ease.
+%(e.g.
 one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.

### 2018-03-27 why.tex
Reorganize so it builds. New Figure 1

--- 2017-12-28 sections/why.tex
@@
 \item{\textit{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
-Despite the fact that such interdisciplinary experiments are highly impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \cite{Sung2003-go, Rhoten2004-fk}}.
+Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \cite{Sung2003-go}}.
 
@@
 
-It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
+\noindent It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
 Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see supplementary materials).

### 2018-03-27 design.tex

--- 2018-01-07 sections/design.tex
@@
 By spending more time together, participants are more likely to overcome barriers of professional terminology.
-On the other hand, events that are too long may lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
-A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and open project work.
+However, events that are too long may lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
+A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and project work.
 
@@
 Hack week outcomes depend strongly on the interest and engagement of participants.
-Some attendees arrive with the goal of writing a scientific article.
-Other attendees plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
+Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
 This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
@@
 Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
-In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also supplementary materials for concrete suggestions.
+In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also supplementary materials for concrete suggestions).
 %can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it

### 2018-03-27 participants.tex

--- 2018-01-07 sections/participants.tex
@@
 To achieve this diversity, we advocate for a selection process that is as quantitative and transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
-%Transparency is necessary for applicants to understand acceptance/rejection decisions, and accountability is of crucial importance for the detection of inherent biases in the selection, which may harm both the event's success as well as the larger community.
 
-Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible~\cite{sunstein2015wiser,bohnet2016}.
+Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible~\cite{sunstein2015wiser}. 
 In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 
@@
 The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
-%For example, organizers may require that half of the participants (or as close as possible to that) be graduate students, while also maximizing the number of different countries from which participants originate.
 It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
@@
 Biased participant sets may still result from selection procedures that fail to include crucial categories.
-%For example, it would be difficult to produce a student-heavy participant set for a summer school if the algorithm has no information about academic seniority, and impossible to correct gender bias in the pool of applicants, if no information is available about the gender of participants.
 
-Blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a very broad population.
+Blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a broad population. 
 For more targeted workshops, organizers should be mindful that blinding to demographics might be counterproductive~\cite{behaghel2015unintended} if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
@@
 
-No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate the different steps in their procedure%.
-For example, merit scores can be interrogated for shared biases, and one can plot the demographics both before and after the initial selection step to diagnose whether certain groups are preferentially selected or excluded
-(see also the Supplementary materials for more details on how the previous hack weeks approached this problem).
-%One way to maximize transparency in the selection process is to minimize human decision making steps that introduce biases, and to . 
-%We work to achieve this by first assessing the merit of each candidate with respect to the overall goals of the hack week.
-%We try to minimize bias in this step by blinding ourselves to a candidate's other attributes, including name and other personal information, and assess their candidacy based soley on questions asked specifically for this purpose.
-%When doing this procedure for a large enough sample, it is unlikely that the resulting pool of acceptable candidates is smaller than the number of available spaces at the workshop.
-
-%The second step in the selection procedure then requires tie-breaking between equally acceptable candidates.
-%It is here where one may impose outside constraints on the selection based on the goals of the workshop.
-%If multiple competing constraints are considered, this task essentially becomes a complex optimization problem, for which algorithms exist that will outperform any human selection procedure.
-
-%One solution to this optimization procedure is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}.
+No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate their procedure (see also the Supplementary materials for more details on how the previous hack weeks approached this problem).

### 2018-03-27 what.tex

--- 2018-01-05 sections/what.tex
@@
 Hackathons originated from the open-source software movement and have historically focused on software and technology development. 
-In recent years hackathons have evolved into a model providing opportunities for intensive, interdisciplinary collaboration \cite{Groen2015-cj,Moller2013-ah} and education \cite{Kienzler2015-zu,Lamers2014-xf} in the sciences.
+In recent years hackathons have evolved into a model providing opportunities for intensive, interdisciplinary collaboration \cite{Groen2015-cj} and education \cite{Kienzler2015-zu,Lamers2014-xf} in the sciences.
 Core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
@@
 However, by embedding these approaches in a more open hackathon framework, we can explore unique ways of enhance the traditional summer school model.
-For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu,Papert1980-fh}.
-Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over the transfer established knowledge. 
-Finally, the social structure of hackathons enable us to priortize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.
+For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu}.
+Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over established knowledge. 
+Finally, the social structure of hackathons enable us to prioritize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.
 
-We have found that these multiple dimensions of the hack week model provide signifcant flexibilty across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
+We have found that the multiple dimensions of the hack week model provide flexibility across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
 For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
@@
 We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
-
-%\begin{bf}
-%Many industry hackathons are focused on delivering solutions to a specific problem or challenge on a very short timescale. 
-%Hack weeks are  more open-ended than these hackathons and allow for a wider variety of outcomes and goals among participants.
-Within academia, the majority of venues for scientific collaboration, learning and communication fall into one of three categories: traditional conferences, summer schools, and, more recently, unconferences.
-All three types share elements with hack weeks, but have goals that are more narrowly defined.
-%Traditional conferences are focused on the exchange of recent results, as well as networking.
-However, collaborative components are usually absent, or restricted to short time intervals (e.g.\ coffee breaks).
-Unconferences are less structured and allow for greater input from participants, but are generally less focused on knowledge transfer and learning, but on achieving measurable outcomes, often in the form of policies, proposals and white papers.
-%\end{bf}
-
-%One criticism of hackathons is that they propel the ``geek'' stereotype and may present a barrier to creating an inclusive working environment, especially for individuals traditionally underrepresented in science and technology \cite{Decker2015}.
-
-%This criticism needs to be actively addressed both through participant selection (see below), and by addressing the possibility that some participants may experience an ``imposter syndrome'' within such circumstances \cite{clance1978imposter} (see supplementary material).
-%Also, while many industry hackathons are competitive, with teams actively competing to solve the same problems for prizes, academic hackathons, and specifically the hack weeks that we have organized are not explicitly competitive.

### 2018-03-29 intro.tex
Restructure introduction; Add figure caption

--- 2018-01-08 sections/intro.tex
@@
 Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-In general, events to date have been designed within either a pedagogical or project-based framework.
-Pedagogically-focused events follow a classic academic model where novices learn new skills from experts, while project-focused workshops emphasize collaborative activities using existing skills.
-A disadvantage of the pedagogical model is that it tends to focus on a one-way flow of information from instructor to student, and can discount the potential contributions by students.
-A disadvantage of the project model is the common perception that the week is designed for technical experts, which may discourage others from attending.
-In 2014, we initiated an alternative model of ``Hack Weeks'' that aim to combine the best of each of these models.
-These week-long events combine pedagogy (often focused on statistical and computational techniques) with hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+In general, events to date have been designed to emphasize either a more traditional pedagogical class and lecture methodology, or emphasize more project-based learning (Figure \ref{fig:hackspectrum}).
+Pedagogically-focused events (e.g., Software Carpentry workshops \cite{b:wilson-swc-lessons-2016} in which no student-led projects are pursued) follow a classic academic model where novices learn new skills from experts.
+This model tends to focus on a one-way flow of information from instructor to student, and are usually targeted towards participants in the training phase of their career.
+On the other end of the spectrum, project-focused workshops emphasize collaborative activities using existing skills, leading to the common perception that they are designed for technical experts, and this may limit their audience.
+To bridge this gap, we describe here a model that we have implemented: ``Hack Weeks'' that aim to capitalize on the advantages of each of these models.
+These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
 
@@
 \begin{center}
-\includegraphics[width=8cm]{HackSpectrum.pdf}
-\caption{Comparison of Extracurricular Workshop Models}
+\includegraphics[width=8cm]{NewHackSpectrum.png}
+\caption{{\bf Comparison of Extracurricular Workshop Models}: Different types of events lie on a spectrum between the emphasis on pedagogy (e.g.
+Software Carpentry workshop) to an emphasis on project-based/hack-based activities (e.g.
+at science-oriented hackathons).
+Hack weeks also vary in the degree of emphasis on projects (Astro hack week, which included unstructured time for group projects from the first day) or pedagogy (Neuro hack week, which included an entire first day of data science tutorials, with less time for projects)}
 \label{fig:hackspectrum}

### 2018-04-10 results.tex
Added Kruskal-Wallis and Mann-Whitney U to analysis, started modifying results section

--- 2018-01-08 sections/results.tex
@@
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (b, c) ).
-More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
+%More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
 We asked participants to subjectively rate their knowledge before the hack week, and find skill levels to be broadly distributed.
@@
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed a regular practice of publishing code before the event (Figure \ref{fig:survey}, (i)).
-The overall behaviour hints at how general conventions and attitudes differ in different fields.
-Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
-Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science (Figure \ref{fig:survey}, (j)): again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
-Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%).
+%at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed a regular practice of publishing code before the event (Figure \ref{fig:survey}, (i)).
+%The overall behaviour hints at how general conventions and attitudes differ in different fields.
+%Astronomy shows the largest degree of openness toward open science, whereas our results indicate that open science is still fairly uncommon in the geosciences, with neuroscience falling in between.
+%Similar attitudes are reflected when asking whether the hack week has made participants more comfortable with open science (Figure \ref{fig:survey}, (j)): again, geoscience shows the large improvement with over 97\% agreeing with this statement to some degree, followed by neuroscience (95\%) and astronomy (72\%).
+%Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible. 
 While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks like open-source software and data sharing.

### 2018-04-12 results.tex
Additional figures for appendix, added table to results section, modified text

@@
 \section*{Results}
+
+\begin{table*}
+%\begin{table*}[hbtp]
+%\renewcommand{\arraystretch}{1.3}
+\footnotesize
+\caption{}
+\begin{threeparttable} 
+\begin{tabularx}{18cm}{p{3.4cm}p{0.3cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}}
+\toprule
+\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ \\ \midrule
+\multicolumn{16}{l}{\textbf{Comparison with career stage}}  \\ \midrule
+New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
+Improvement of research (b) & 61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
+Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\ 
+Teaching (d) & 10 & 61 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
+\multicolumn{16}{l}{\textbf{Gender Identity}} \\ \midrule
+Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
+Teaching (d) & 61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
+Connection building (e) & 61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
+Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
+\multicolumn{16}{l}{\textbf{Racial/ethnic Identity}} \\ \midrule
+Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
+Teaching (d) & 61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
+Connection building (e) & 61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
+Valued contributions (f) & 58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\ 
+\bottomrule
+\end{tabularx}
+   \begin{tablenotes}
+      \item{Respondents to a post-event evaluation survey indicate their level of agreement on a six-point Likert-type scale to a range of statements (corresponding to Figure \ref{fig:survey}).
+Results are for hack weeks held in 2016 and 2017 for AHW and in 2016 for GHW and NHW, respectively.
+The number of respondents for each question is included in the table above.
+Additional figures showing the distributions of scores are available in the Appendix.}
+\end{tablenotes}
+
+\end{threeparttable}
+\label{tab:priors}
+%\end{table*}
+\end{table*}
 
@@
 At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
-We report here about results in three different domains: the development of technical skills (a -- e), collaboration and learning (f,g), and shifts in attitudes towards reproducibility and open science (h -- j)}
+We report here about results in three different domains: the development of technical skills (a -- c), collaboration and teaching (d -- f), and shifts in attitudes towards reproducibility and open science (g, h).}
 \label{fig:survey}
@@
 Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that most participants self-report successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (b, c) ).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist (Figure \ref{fig:survey}, (b, c)).
 %More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
@@
 
-The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, (g)), especially at Neuro Hack Week, where more than 64\% of participants strongly agreed with this statement.
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\%; Figure \ref{fig:survey}, (f)), though responses are not as unequivocal as they are in some of the other categories.
-Similarly to learning outcomes, this is expected given the range of skill levels at the workshops.
-It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
+We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\%; Figure \ref{fig:survey}, (d)), though responses are not as unequivocal as they are in some of the other categories.
+The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, (e)), especially at Neuro Hack Week, where more than 64\% of participants strongly agreed with this statement.
+
+Given the broad diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible.
+At the same time, we put forth the hypothesis earlier in the this paper that peer learning is a major mode of knowledge transfer during hack weeks, and that this type of learning takes place at all career stages.
+However, we find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (a), improvements in day-to-day research (b) or overall improvements in science (c) (all statistical results are presented in Table \ref{tab:survey}).
+Only for GHW we find an indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following \citet{cohen1988}).
+Similarly, we find no strong indication that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (d).
+
+One important question regarding hack week outcomes is whether participants from underrepresented groups thrive at these events, or whether their full participation is impeded.
+Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building would indicate that improvements in workshop facilitation and structure may be required to allow participants of all stripes to participate fully.
+In Section 4 of the Appendix, we lay out various strategies we have employed to facilitate participation, empower members of underrepresented groups and mitigate impostor syndrome.
+Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity (Table \ref{tab:survey}).
+Similarly, participants who self-identified as minorities with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks.
+Both underrepresented groups similarly feel that they built equally valuable connections at all three hack weeks and generally feel that their contributions to their hack teams were valued.
+For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections (p=0.04) with a medium to large effect size ($\phi_c = 0.4$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01; \phi_c = 0.4$).
+These initial results indicate that future work on how demographics interact with hack week attendance may be fruitful.
+We note that since we did not include a question about impostor syndrome itself, we cannot draw causal connections between our results and its prevalence.
+Additionally, our number of respondents is small, and the lack of significant differences may be caused by the small statistical power of our sample.
+%It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
 Participants at all career stages report similar engagement in teaching.
@@
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%).
-%at all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (h)) put code or data created at the hack week into a public repository, while a substantially smaller fraction of participants followed a regular practice of publishing code before the event (Figure \ref{fig:survey}, (i)).
+We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)).
+At all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g) ) put code or data created at the hack week into a public repository.
 %The overall behaviour hints at how general conventions and attitudes differ in different fields.
@@
 In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.
+
+We caution the reader that the survey results should be read only as an indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the exploratory nature of the surveys conducted as part of the hack weeks.
+The most important independent variable--attendance at the hack week--is something that our current survey design cannot account for.
+Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to internal biases of respondents.
+While these surveys were largely of an exploratory nature, future work will include more a more refined survey design as well as attendance versus non-attendance as a control variable.
 

### 2018-04-12 supplementary_material.tex
Additional figures for appendix, added table to results section, modified text

@@
 
+\setcounter{tocdepth}{3}
+
 \begin{document}
@@
 \end{affiliations}
+
+\newpage
+
+\tableofcontents
+
+\newpage
 
@@
 the number of attendees) and event duration are important to consider during the design and planning phase of a hack week.
-All hack weeks held so far have had group sizes between 40-60 participants.
+All hack weeks we have organized so far have had group sizes between 40-60 participants.
 The number of attendees resulted from a number of different design considerations.
@@
 
-All hack weeks organized up to now have been five days.
+All hack weeks organized by us up to now have been five days.
 This duration was chosen to allow enough time for both learning and project work at the same time.
@@
 This is especially important because the group will inevitably include a wide range of personalities, some of which may find it easier to form new bonds and interact with the group than others.
-At all hack weeks, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
+At all hack weeks we have organized, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
 Clearly communicating values and expectations at the beginning for the week can help setting the stage for an environment in which a large range of different personalities can thrive.
@@
 
+\section{Additional Figures}
+
+Here, we present a number of additional figures related to the survey data presented in the main manuscript.
+The stacked bar charts presented here allow a closer look at the data that produced the results quoted in Table 1 of the manuscript.
+
+\subsection{Career stage}
+
+Here we explore a number of survey responses with career stage as an independent variable.
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_3_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_3_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_4_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_2_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\clearpage
+
+\subsection{Gender Identity}
+
+Here we explore a number of survey responses with gender identity as an independent variable.
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_4_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_2_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_5_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q19_5_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\clearpage
+
+\subsection{Racial/Ethnic Identity}
+
+Here we explore a number of survey responses with racial/ethnic identity as an independent variable.
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_4_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_2_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_5_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q19_5_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
+
 \section{References}

### 2018-04-12 results.tex
Changed Appendix to supplementary materials

@@
 The number of respondents for each question is included in the table above.
-Additional figures showing the distributions of scores are available in the Appendix.}
+Additional figures showing the distributions of scores are available in the Supplementary Materials.}
 \end{tablenotes}
@@
 Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building would indicate that improvements in workshop facilitation and structure may be required to allow participants of all stripes to participate fully.
-In Section 4 of the Appendix, we lay out various strategies we have employed to facilitate participation, empower members of underrepresented groups and mitigate impostor syndrome.
+In Section 4 of the Supplementary materials, we lay out various strategies we have employed to facilitate participation, empower members of underrepresented groups and mitigate impostor syndrome.
 Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity (Table \ref{tab:survey}).

### 2018-04-12 participants.tex

@@
 
-In our experience, a participant group that is diverse across categories of diversity, gender, discipline and career track helps to ensure we meet these objectives.
+In our experience, a participant group that is diverse across categories of minority status, geographical origin, gender, discipline and career track helps to ensure we meet these objectives.
 To achieve this diversity, we advocate for a selection process that is as quantitative and transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
+
+Participant selection is an intrinsically difficult process that is fraught with personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
+Traditional selection processes rely heavily on internal heuristics of reviewers, including using unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012} as a cipher for suitability.
+On the other hand, while purely mechanistic procedures may to be able to mitigate some reviewer biases, this may come at a cost of baking in structural inequalities.
+A blind selection purely on merit may be flawed if some groups of applicants have been systematically disadvantaged in certain ways: for example,
 

### 2018-04-12 design.tex
Some wording changes and references to supplementary materials, also changed format of supplementary materials to enable numbering of subsections

@@
 
-Scheduling, group size and venue are important design considerations contributing to the success of a hack week.   
+Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 1 and 2 for practical guidance) .   
 Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges. 
@@
 Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
-Workspaces must also accommodate interactive project work where small teams can gather and work together.    
+Workspaces must also accommodate interactive project work where small teams can gather and work together (for more detailed considerations, see also the Supplementary Materials).    
 %Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
@@
 
-Another important design consideration is group size.
+Another important design consideration is group size (see also SM, Section 2).
 If the group is too large, chances for random participant exchanges are reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
-If the group is too small, it is unlikely to achieve the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines.
-In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.
+If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines.
+In the past, we have found groups with sizes between 50 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group.
+However, we encourage organizers to experiment with group size depending on their workshop goals.
 
@@
 The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week.
+A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 2.3).
 Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
-In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also supplementary materials for concrete suggestions).
-%can lead to an increase in the prevalence of impostor syndrome experienced by many participants (see also supplementary materials), and it
+In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4 for concrete suggestions).

### 2018-04-12 intro.tex
Some wording changes and references to supplementary materials, also changed format of supplementary materials to enable numbering of subsections

--- 2018-01-08 sections/intro.tex
@@
 \begin{center}
-\includegraphics[width=8cm]{HackSpectrum.pdf}
+\includegraphics[width=8cm]{NewHackSpectrum.png}
 \caption{Comparison of Extracurricular Workshop Models}
@@
 As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
-Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines (supplementary materials provide additional details on the practical aspects of organizing these events).
+Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.
+Supplementary materials (SM) provide additional details on the practical aspects of organizing these events.

### 2018-04-12 participants.tex
Some wording changes and references to supplementary materials, also changed format of supplementary materials to enable numbering of subsections

@@
 \section*{Audience and Participant Selection}
+
 Hack weeks differ from traditional conferences or summer schools in that knowledge transfer occurs across many levels of seniority, disciplinary boundaries, and novelty of the topics discussed.
-In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from participants.
+In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from attendees.
 Therefore in order to maximize learning outcomes and collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
@@
 On the other hand, while purely mechanistic procedures may to be able to mitigate some reviewer biases, this may come at a cost of baking in structural inequalities.
-A blind selection purely on merit may be flawed if some groups of applicants have been systematically disadvantaged in certain ways: for example,
+A blind selection purely on from figure of merit will likely be biased if some groups of applicants have been systematically disadvantaged in certain ways: for example, if workshop organizers require a minimum level of programming experience, this could disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
 
@@
 Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
-A merit selection could then perform via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate, ~\cite{highhouse2008} it may be beneficial to subsequently set a fairly tolerant threshold for acceptance, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+A merit selection could then be performed via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
+
+However, blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a broad population, where the initial selection does not directly depend on skill level (as has been true e.g.\ for Astro Hack Week; see also Supplementary Materials).
+For workshops where merit selection is intrinsically tied to skill levels, organizers should be mindful that blinding to demographics can be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
+Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
+In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also Supplementary Materials for details). 
+Because systemic biases likely also enter at the application stage (where underrepresented groups may be less likely to apply) organizers should consider oversampling traditionally disenfranchised groups compared to the population of applicants.
+
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 
@@
 
-Blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a broad population. 
-For more targeted workshops, organizers should be mindful that blinding to demographics might be counterproductive~\cite{behaghel2015unintended} if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
-In this case, it may be beneficial to include demographic characteristics in the first stage. 
-Because systemic biases likely also enter at the application stage (where underrepresented groups may be less likely to apply) organizers should consider oversampling traditionally disenfranchised groups compared to the population of applicants.
-
-No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate their procedure (see also the Supplementary materials for more details on how the previous hack weeks approached this problem).
+No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate their procedure.
+For example, comparing demographic characteristics of the selected versus non-selected groups can unveil unintended biases during the merit-selection base and thus allow adjustments in the procedure to mitigate or fully remove these effects.

### 2018-04-12 results.tex
Some wording changes and references to supplementary materials, also changed format of supplementary materials to enable numbering of subsections

@@
 \textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}$emph{a} & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$\emph{b} & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage}}  \\ \midrule
@@
 Additional figures showing the distributions of scores are available in the Supplementary Materials.}
+      \item[\emph{a}]{The degrees of freedom for Cram\'{e}r's V corrected according to \citet{bergsma2013}}
+      \item[\emph{b}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
+
 \end{tablenotes}
@@
 
-Measuring the success of a hack week objectively is complicated by the variety of objectives that a hack week might have (see above). 
+Measuring the success of a hack week objectively is complicated by the variety of goals that a hack week might have (see above). 
 Additionally, the participant-driven format facilitates knowledge transfer and collaborations in sometimes surprising ways that escape traditional measures of success.
@@
 We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)).
-At all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g) ) put code or data created at the hack week into a public repository.
+At all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
 %The overall behaviour hints at how general conventions and attitudes differ in different fields.
@@
 
-We caution the reader that the survey results should be read only as an indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the exploratory nature of the surveys conducted as part of the hack weeks.
+We caution the reader that the survey results should be read only as an indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the exploratory nature of the surveys conducted as part of past hack weeks we have organized.
 The most important independent variable--attendance at the hack week--is something that our current survey design cannot account for.
 Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to internal biases of respondents.
-While these surveys were largely of an exploratory nature, future work will include more a more refined survey design as well as attendance versus non-attendance as a control variable.
+Future work will include more a more refined survey design as well as attendance versus non-attendance as a control variable.
 

### 2018-04-12 supplementary_material.tex
Some wording changes and references to supplementary materials, also changed format of supplementary materials to enable numbering of subsections

@@
 
-\documentclass{nature}
+\documentclass{aastex62}
 
@@
 
+%\author{Daniela Huppenkothen$^{1,2,3}$, Anthony Arendt$^{4,5}$, David W. Hogg$^{2,1,6,7}$, Karthik Ram$^8$, Jake VanderPlas$^5$, and Ariel Rokem$^5$}
+
+%\makeatletter
+%\let\saved@includegraphics\includegraphics
+%\AtBeginDocument{\let\includegraphics\saved@includegraphics}
+%\renewenvironment*{figure}{\@float{figure}}{\end@float}
+%\makeatother
+
+%\renewcommand*\thesection{\arabic{section}}
+%\renewcommand*\thesubsection{\arabic{subsection}}
+
+\setcounter{tocdepth}{3}
+
+\begin{document}
+
 \title{Supplementary material for "Hack Weeks as a model for Data Science Education and Collaboration"}
@@
 
-\author{Daniela Huppenkothen$^{1,2,3}$, Anthony Arendt$^{4,5}$, David W. Hogg$^{2,1,6,7}$, Karthik Ram$^8$, Jake VanderPlas$^5$, and Ariel Rokem$^5$}
+\author{Daniela Huppenkothen}
+\altaffiliation{dhuppenk@uw.edu}
+\affiliation{DIRAC Institute, Department of Astronomy, University of Washington, 3910 15th Ave NE, Seattle, WA 98195, USA}
+\affiliation{Center for Data Science, New York University, 65 5th Avenue, 7th Floor, New York, NY 10003, USA}
+\affiliation{Center for Cosmology and Particle Physics, New York University, 726 Broadway, 10th Floor, New York, NY 10003, USA}
+\nocollaboration
 
-\makeatletter
-\let\saved@includegraphics\includegraphics
-\AtBeginDocument{\let\includegraphics\saved@includegraphics}
-\renewenvironment*{figure}{\@float{figure}}{\end@float}
-\makeatother
+\author{Anthony Arendt}
+\affiliation{Polar Science Center/Applied Physics Laboratory, University of Washington, 1013 NE 40th Street, Box 355640, Seattle, WA 98105-6698}
+\affiliation{The University of Washington eScience Institute, The WRF Data Science Studio, Physics/Astronomy Tower, 6th Floor, 3910 15th Ave NE, Campus Box 351570, University of Washington, Seattle, WA 98105, USA}
+\nocollaboration
 
-\setcounter{tocdepth}{3}
+\author{David W. Hogg}
+\affiliation{Center for Cosmology and Particle Physics, New York University, 726 Broadway, 10th Floor, New York, NY 10003, USA}
+\affiliation{Center for Data Science, New York University, 65 5th Avenue, 7th Floor, New York, NY 10003, USA}
+\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, USA}
+\nocollaboration
 
-\begin{document}
+\author{Karthik Ram}
+\affiliation{Berkeley Institute for Data Science \& Berkeley Initiative in Global Change Biology, University of California, Berkeley,  Berkeley CA 94720}
+\nocollaboration
 
-\maketitle
+\author{Jake VanderPlas}
+\affiliation{The University of Washington eScience Institute, The WRF Data Science Studio, Physics/Astronomy Tower, 6th Floor, 3910 15th Ave NE, Campus Box 351570, University of Washington, Seattle, WA 98105, USA}
+\nocollaboration
 
-\begin{affiliations}
-\item{Center for Data Science, New York University, 65 5th Avenue, 7th Floor, New York, NY 10003, USA}
-\item{Center for Cosmology and Particle Physics, New York University, 726 Broadway, 10th Floor, New York, NY 10003, USA}
-\item{DIRAC Institute, Department of Astronomy, University of Washington, 3910 15th Ave NE, Seattle, WA 98195, USA}
-\item{Polar Science Center/Applied Physics Laboratory, University of Washington, 1013 NE 40th Street, Box 355640, Seattle, WA 98105-6698}
-\item{The University of Washington eScience Institute, The WRF Data Science Studio, Physics/Astronomy Tower, 6th Floor, 3910 15th Ave NE, Campus Box 351570, University of Washington, Seattle, WA 98105, USA}
-\item{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}
-\item{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, USA}
-\item{Berkeley Institute for Data Science \& Berkeley Initiative in Global Change Biology, University of California, Berkeley,  Berkeley CA 94720}
-\end{affiliations}
+\author{Ariel Rokem}
+\affiliation{The University of Washington eScience Institute, The WRF Data Science Studio, Physics/Astronomy Tower, 6th Floor, 3910 15th Ave NE, Campus Box 351570, University of Washington, Seattle, WA 98105, USA}
+\nocollaboration
+
+%\maketitle
+
+%\begin{affiliations}
+%\item{Center for Data Science, New York University, 65 5th Avenue, 7th Floor, New York, NY 10003, USA}
+%\item{Center for Cosmology and Particle Physics, New York University, 726 Broadway, 10th Floor, New York, NY 10003, USA}
+%\item{DIRAC Institute, Department of Astronomy, University of Washington, 3910 15th Ave NE, Seattle, WA 98195, USA}
+%\item{Polar Science Center/Applied Physics Laboratory, University of Washington, 1013 NE 40th Street, Box 355640, Seattle, WA 98105-6698}
+%\item{The University of Washington eScience Institute, The WRF Data Science Studio, Physics/Astronomy Tower, 6th Floor, 3910 15th Ave NE, Campus Box 351570, University of Washington, Seattle, WA %98105, USA}
+%\item{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}
+%\item{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010, USA}
+%\item{Berkeley Institute for Data Science \& Berkeley Initiative in Global Change Biology, University of California, Berkeley,  Berkeley CA 94720}
+%\end{affiliations}
 
@@
 Similarly, in order to foster active collaboration among many different subfields, we endeavored to admit participants with many different types of overlapping expertise.
-Having a set of participants that is too small would interfere both with these goals of representation and of fostering interdisciplinary collaborations and networking, thus we would advise admitting at least $\sim$30 or so participants if these objectives are an important part of the hack week.
+Having a set of participants that is too small could interfere both with these goals of representation and of fostering interdisciplinary collaborations and networking.
+Smaller groups (of less than $\sim$30 participants or so) will need to attend to participant selection particularly carefully in order to achieve the desired level of diversity across participants.
 
@@
 Meanwhile, neuro hack week opted to ask about gender ethnicity and race, using questions that are standardly used for example, when collecting statistics about study participants.
+
+ \textbf{For any kind of merit-based selection, committees should critically examine their assumptions about reasonable expectations regarding (minimum) skill levels required for participation.
+Candidates from underrepresented groups, smaller institutions or other countries may have had fewer opportunities to learn relevant skills.
+Consequently, in a merit-based selection based on these skills these questions may effectively trace demographic diversity and lead to biased selections even when committees are overall vigilant.
+If a strict merit-based selection is part of the procedure, it should be performed with these caveats in mind.
+It is always a good idea to check the results of any selection procedure for intrinsic biases.
+For example, visual comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can be adjusted and refined.}
 
@@
 
-\section{References}
-\bibliographystyle{naturemag}
+\newpage
+\bibliographystyle{aasjournal}
 \bibliography{paper}

### 2018-04-12 intro.tex
A few more sentences about differences with summer schools

--- 2018-03-29 intro.tex
@@
 Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-In general, events to date have been designed to emphasize either a more traditional pedagogical class and lecture methodology, or emphasize more project-based learning (Figure \ref{fig:hackspectrum}).
+In general, many of these events tend to emphasize either a more traditional pedagogical class and lecture methodology, or emphasize a focus on ad-hoc projects that participants work on during the event (Figure \ref{fig:hackspectrum}).
 Pedagogically-focused events (e.g., Software Carpentry workshops \cite{b:wilson-swc-lessons-2016} in which no student-led projects are pursued) follow a classic academic model where novices learn new skills from experts.
@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+In contrast to many summer schools that also include team project work (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course), one of the explicit goals of the hack weeks that we have organized is to facilitate a multi-directional transfer of knowledge between people in different stages of their career, and with different backgrounds.
+Hence, participants often include not only early-career researchers and trainees, but also research staff members and faculty members at various stages of their careers (see also \ref{sec:participants}).
 
@@
 As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
-Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines (supplementary materials provide additional details on the practical aspects of organizing these events).
+Here we share the philosophy behind the hack week model as we have implemented it, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines (supplementary materials provide additional details on the practical aspects of organizing these events).

### 2018-04-12 results.tex
Fixed some references and typos in table

@@
 \textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}$emph{a} & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$\emph{b} & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}$ & $\phi_c$ \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage}}  \\ \midrule
@@
 Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\ 
-Teaching (d) & 10 & 61 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
+Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
 \multicolumn{16}{l}{\textbf{Gender Identity}} \\ \midrule
@@
 Additional figures showing the distributions of scores are available in the Supplementary Materials.}
-      \item[\emph{a}]{The degrees of freedom for Cram\'{e}r's V corrected according to \citet{bergsma2013}}
+      \item[\emph{a}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
       \item[\emph{b}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
@@
 However, we find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (a), improvements in day-to-day research (b) or overall improvements in science (c) (all statistical results are presented in Table \ref{tab:survey}).
-Only for GHW we find an indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following \citet{cohen1988}).
+Only for GHW we find an indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following~\citep{cohen1988}).
 Similarly, we find no strong indication that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (d).

### 2018-04-12 design.tex
More minor fixes

@@
 Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
-Workspaces must also accommodate interactive project work where small teams can gather and work together (for more detailed considerations, see also the Supplementary Materials).    
+Workspaces must also accommodate interactive project work where small teams can gather and work together (for more detailed considerations, see also the SM, Sections 2 and 3).    
 %Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 

### 2018-04-12 participants.tex
More minor fixes

@@
 
-However, blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a broad population, where the initial selection does not directly depend on skill level (as has been true e.g.\ for Astro Hack Week; see also Supplementary Materials).
+However, blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a broad population, where the initial selection does not directly depend on skill level (as has been true e.g.\ for Astro Hack Week; see also SM, Section 2.1).
 For workshops where merit selection is intrinsically tied to skill levels, organizers should be mindful that blinding to demographics can be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
 Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
-In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also Supplementary Materials for details). 
+In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also SM, Section 2.1). 
 Because systemic biases likely also enter at the application stage (where underrepresented groups may be less likely to apply) organizers should consider oversampling traditionally disenfranchised groups compared to the population of applicants.
 
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritous pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 

### 2018-04-12 results.tex
More minor fixes

@@
 The number of respondents for each question is included in the table above.
-Additional figures showing the distributions of scores are available in the Supplementary Materials.}
+Additional figures showing the distributions of scores are available in the SM, Section 6.}
       \item[\emph{a}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
@@
 \end{threeparttable}
-\label{tab:priors}
+\label{tab:survey}
 %\end{table*}
@@
 Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building would indicate that improvements in workshop facilitation and structure may be required to allow participants of all stripes to participate fully.
-In Section 4 of the Supplementary materials, we lay out various strategies we have employed to facilitate participation, empower members of underrepresented groups and mitigate impostor syndrome.
+In Section 4 of the Supplementary Materials, we lay out various strategies we have employed to facilitate participation, empower members of underrepresented groups and mitigate impostor syndrome.
 Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity (Table \ref{tab:survey}).

### 2018-04-13 participants.tex
Moved Entrofy paragraph into supplementary materials, rephrased audience and participant selection section in main text

@@
 To achieve this diversity, we advocate for a selection process that is as quantitative and transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
+Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible~\cite{sunstein2015wiser}.
 
 Participant selection is an intrinsically difficult process that is fraught with personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
-Traditional selection processes rely heavily on internal heuristics of reviewers, including using unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012} as a cipher for suitability.
+It requires laying out a definition of successful participation, what criteria must be met in order to maximize the likelihood of success, and how those criteria will be assessed given the information about the candidates selected during the application stage.
+Traditional selection processes rely heavily on internal heuristics of reviewers, including using unrelated characteristics as a cipher for suitability.
 On the other hand, while purely mechanistic procedures may to be able to mitigate some reviewer biases, this may come at a cost of baking in structural inequalities.
-A blind selection purely on from figure of merit will likely be biased if some groups of applicants have been systematically disadvantaged in certain ways: for example, if workshop organizers require a minimum level of programming experience, this could disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
 
-Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible~\cite{sunstein2015wiser}. 
-In practice, there are different approaches to counteract intrinsic human biases and provide transparency. 
-Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
+For hack weeks, it is not obvious a priori that prerequisites for attendance in terms of skills should exist, and prerequisites will depend on the objectives of the workshop.
+For example, Astro Hack Week has traditionally accepted participants at all skill levels with respect to data science and did not a merit-based selection, whereas Neuro Hack Week did include skill-based criteria in their selection (see also SM, Section 2.1 for more detail on the individual selection procedures).
+If merit-based selection is part of the process, organizers face the crucial decision whether to assess merit blindly or not.
+Research particularly around hiring procedures has shown that there is no single correct procedure.
+
+Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases related to internal heuristics.
 A merit selection could then be performed via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
+This type of blinded procedure tends to reduce biases during participant selection when committees would otherwise not consider diversity during their selection [REF].
+However, a blind selection purely on a figure of merit will likely still be biased if some groups of applicants have been systematically disadvantaged in certain ways: for example, if workshop organizers require a minimum level of programming experience, this could disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
 
-However, blinding in the initial merit selection step is most effective at counteracting biases when the hack week targets a broad population, where the initial selection does not directly depend on skill level (as has been true e.g.\ for Astro Hack Week; see also SM, Section 2.1).
-For workshops where merit selection is intrinsically tied to skill levels, organizers should be mindful that blinding to demographics can be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
+Blinding to demographics can be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
 Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
@@
 
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritous pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
-
-One solution to the latter problem is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}). 
-The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
-It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
-Any human biases involved in setting these target fractions will be perpetuated in the selection procedure. 
-Secondly, perhaps more obviously, the algorithm can only act on information that has been collected.
-Biased participant sets may still result from selection procedures that fail to include crucial categories.
-
 No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate their procedure.

### 2018-04-13 supplementary_material.tex
Moved Entrofy paragraph into supplementary materials, rephrased audience and participant selection section in main text

@@
 
- \textbf{For any kind of merit-based selection, committees should critically examine their assumptions about reasonable expectations regarding (minimum) skill levels required for participation.
+ For any kind of merit-based selection, committees should critically examine their assumptions about reasonable expectations regarding (minimum) skill levels required for participation.
 Candidates from underrepresented groups, smaller institutions or other countries may have had fewer opportunities to learn relevant skills.
@@
 It is always a good idea to check the results of any selection procedure for intrinsic biases.
-For example, visual comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can be adjusted and refined.}
+For example, visual comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can be adjusted and refined.
+
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritous pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+
+One solution to the latter problem is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}). 
+The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
+It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
+Any human biases involved in setting these target fractions will be perpetuated in the selection procedure. 
+Secondly, perhaps more obviously, the algorithm can only act on information that has been collected.
+Biased participant sets may still result from selection procedures that fail to include crucial categories.
 

### 2018-04-13 supplementary_material.tex
Added 300ppi pdf version of hack spectrum, changed ref in intro, fixed typo in supplementary materials

@@
 
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritous pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 

### 2018-04-13 intro.tex
Added 300ppi pdf version of hack spectrum, changed ref in intro, fixed typo in supplementary materials

@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
-In contrast to many summer schools that also include team project work (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course), one of the explicit goals of the hack weeks that we have organized is to facilitate a multi-directional transfer of knowledge between people in different stages of their career, and with different backgrounds.
+Summer schools have been designed to excel in transfer of knowledge from experts in the field to early-career researchers.
+This does many times come in the form of project work, either as individuals or in groups.
+But in contrast to many summer schools that also include team project work (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course), the hack weeks that we have organized tend to be less structured and more participant-driven.
+This looser structure is designed to facilitate a multi-directional transfer of knowledge between people in different stages of their career, and with different backgrounds.
 Hence, participants often include not only early-career researchers and trainees, but also research staff members and faculty members at various stages of their careers (see also \ref{sec:participants}).
+As suggested by Figure \ref{fig:hackspectrum}, these events exist on a continuum, trading off a focus on pedagogy and a focus on more participant-driven activities (because there are only 24 hours in every day).
 

### 2018-04-16 intro.tex
Rephrased the what section and moved some stuff there from intro

--- 2018-04-12 intro.tex
@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
-In contrast to many summer schools that also include team project work (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course), one of the explicit goals of the hack weeks that we have organized is to facilitate a multi-directional transfer of knowledge between people in different stages of their career, and with different backgrounds.
-Hence, participants often include not only early-career researchers and trainees, but also research staff members and faculty members at various stages of their careers (see also \ref{sec:participants}).
 

### 2018-04-16 what.tex
Rephrased the what section and moved some stuff there from intro

@@
 Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
-Our focus on both pedagogy and open project work aims to fill an educational space we feel is not currently addressed by existing models (Figure \ref{fig:hackspectrum}). 
+In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently addressed by existing models (Figure \ref{fig:hackspectrum}).
+
 The hackathon, a time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events. 
@@
 
+Summer schools have been designed to excel in transfer of knowledge from experts in the field to (early-career) researchers: they often serve as an entry point for scientists who aim to expand their research into a new area or switch fields.
+They are excellent at giving participants a reasonably deep understanding of a topic or field in a short amount of time.
+Within this concept, learning can take many forms, including traditional lecture formats but also hands-on project work, often in teams (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course).
+
 Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning, forming the second axis of our events.
-Teaching within our events draws from a spectrum of approaches such as those traditionally employed in summer schools and software workshops. 
-However, by embedding these approaches in a more open hackathon framework, we can explore unique ways of enhance the traditional summer school model.
-For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu}.
-Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over established knowledge. 
-Finally, the social structure of hackathons enable us to prioritize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.
+They aim to synthesize different goals and strategies from both models: they are much more participant-driven than a summer school, but have a stronger focus on pedagogy than a hackathon.
+Where a summer school is often organized around a framework of lectures and tutorials known in advance, hack weeks leave the majority of time unstructured.
+Tutorials at hack weeks often serve as an entry point into a topic for further exploration and learning, rather than an in-depth study.
+This is closely related to the nature of the material taught at these events.
+Data science is a new, emerging area combining knowledge and skills from many different fields, only a fraction of which can be communicated within a short tutorial.
+As organizers, we make a curated selection of a small number of tutorials each year, and encourage participants to use the unstructured time to self-organize tutorials on any topic as needed, often on topics vastly different to those discussed in the formal lectures.
 
-We have found that the multiple dimensions of the hack week model provide flexibility across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
-For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
-In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.
+Additionally, as we lay out in the subsequent section below, multi-directional knowledge transfer is one of many goals of a hack week, and participants often attend with more than one objective.
+The unstructured schedule and inherent lack of focus on specific topics within a hack week present both an opportunity and a risk.
+They uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number of tutorials and their content.
+This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities.
+They also encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few.
+Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach activities and work aimed at improving the scientific community itself.
+
+There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week can run the danger of not doing any of them well.
+Because tutorials are not the major focus of a hack week, the knowledge gained by participants in these tutorials may not be very deep.
+Additionally, a hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require more preparation from the side of participants than other events (e.g.
+defining project ideas, researching possible connections to other participants, bringing along their own tutorials).
+Because of these risks, organizers face a much larger degree of uncertainty both before and during a hack week, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, keeping track of the goals different attendees may have, and careful facilitation of the large range of different types of activities at any given time.
+
+%Teaching within our events draws from a spectrum of approaches such as those traditionally employed in summer schools and software workshops. 
+%However, by embedding these approaches in a more open hackathon framework, we can explore unique ways of enhance the traditional summer school model.
+%For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu}.
+%Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over established knowledge. 
+%Finally, the social structure of hackathons enable us to prioritize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.
+
+%We have found that the multiple dimensions of the hack week model provide flexibility across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
+%For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
+%In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.
 
 We note that the terminology for these events is constantly evolving, and that the ``hackathon'' concept may have implicit connotations that are disfavored in some communities.
+We also note that all of these events live on a constantly changing continuum, depending on the requirements of the scientific domain within which they live.
+For example, Neuro Hack Week is moving toward a more traditional summer school model, while Astro Hack Week has strengthened its focus on projects and hacks in recent iterations.
+
+%, the hack weeks that we have organized tend to be less structured and more participant-driven.
+This looser structure is designed to facilitate a multi-directional transfer of knowledge between people in different stages of their career, and with different backgrounds.
+Hence, participants often include not only early-career researchers and trainees, but also research staff members and faculty members at various stages of their careers (see also \ref{sec:participants}).
+As suggested by Figure \ref{fig:hackspectrum}, these events exist on a continuum, trading off a focus on pedagogy and a focus on more participant-driven activities (because there are only 24 hours in every day).

### 2018-04-17 design.tex
Shortened design section a little and moved some text around

--- 2018-03-27 design.tex
@@
 If the group is too large, chances for random participant exchanges are reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
-If the group is too small, it is unlikely to achieve the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines.
-In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.
+If the group is too small, reaching the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines may require much more careful managing of the participant selection than is true for larger groups.
+Our events have consisted of groups of around 50-70 people, which we have found anecdotally to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.
 
-As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
-If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
-In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.
+%As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
+%If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
+%In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.
 
-Hack week outcomes depend strongly on the interest and engagement of participants.
-Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
-This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
-This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
+This breadth of possible workshop outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
 The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week.

### 2018-04-17 what.tex
Shortened design section a little and moved some text around

@@
 Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
-In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently addressed by existing models (Figure \ref{fig:hackspectrum}).
+In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).
 
@@
 This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities.
-They also encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few.
-Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach activities and work aimed at improving the scientific community itself.
+They encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few.
+Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach and work aimed at improving the scientific community itself.
 
@@
 defining project ideas, researching possible connections to other participants, bringing along their own tutorials).
-Because of these risks, organizers face a much larger degree of uncertainty both before and during a hack week, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, keeping track of the goals different attendees may have, and careful facilitation of the large range of different types of activities at any given time.
-
-%Teaching within our events draws from a spectrum of approaches such as those traditionally employed in summer schools and software workshops. 
-%However, by embedding these approaches in a more open hackathon framework, we can explore unique ways of enhance the traditional summer school model.
-%For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu}.
-%Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over established knowledge. 
-%Finally, the social structure of hackathons enable us to prioritize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.
-
-%We have found that the multiple dimensions of the hack week model provide flexibility across a spectrum of approaches, the design of which depends on the specific characteristics of each community of practice.
-%For example, the astronomy community is relatively small and has a foundation of shared approaches and software implementations, allowing for a greater focus on project work over formal tutorials.
-%In contrast, both the neuro- and geoscience communities covered a broader range of sub-disciplines and had a less cohesive set of existing practices, calling for greater focus on tutorials and education.
+Because of these risks, organizers face a much larger degree of uncertainty both before and during a hack week, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, keeping track of the goals different attendees may have, and careful facilitation of the large range of different types of activities at any given time (see also the following sections as well as the SM, Section 4).
 
@@
 For example, Neuro Hack Week is moving toward a more traditional summer school model, while Astro Hack Week has strengthened its focus on projects and hacks in recent iterations.
-
-%, the hack weeks that we have organized tend to be less structured and more participant-driven.
-This looser structure is designed to facilitate a multi-directional transfer of knowledge between people in different stages of their career, and with different backgrounds.
-Hence, participants often include not only early-career researchers and trainees, but also research staff members and faculty members at various stages of their careers (see also \ref{sec:participants}).
-As suggested by Figure \ref{fig:hackspectrum}, these events exist on a continuum, trading off a focus on pedagogy and a focus on more participant-driven activities (because there are only 24 hours in every day).

### 2018-04-23 themes.tex

--- 2017-08-17 sections/themes.tex
@@
 
-To date, all organized hack weeks have been subject-specific, i.e.\ aimed to bring together a community with a shared scientific interest, such as neuroscience.
+To date, all hack weeks we have organized have been subject-specific, i.e.\ aimed at bringing together a community with a shared scientific interest, such as neuroscience.
 Advantages to this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.
@@
 For these events, building a shared vocabulary and understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.
-
-%This has several clear advantages.
-%In particular, communities organized by subject generally share the same language.%, which drastically cuts down on the time participants spend simply agreeing on scientific terms.
-%Similarly, the scientific objectives and questions are broadly clear to all participants.%: there is little need to explain in detail why a particular question may be of interest to the field.
-%Finally, there are practical considerations: while even within fields the use of particular data sets and software packages may differ quite widely, there is generally a shared understanding of the type of data usually taken, and the most common ways to analyze this data.
-%This leaves more time for active collaboration on cutting-edge science.
-
-%However, there are also drawbacks to limiting the theme of a hack week to a single field.
-%Homogeneity may be a disadvantage if it leads to \textit{group think} and inhibits new, creative solutions.
-%Within a field, the chance of participants being familiar with each other may be larger, leading to the formation of cliques and insufficient mixing.
-%These disadvantages may be mitigated by organizing a hack week around a technique (e.g. Gaussian Processes) or modality (e.g. imaging) instead.
-%In these case, building a shared vocabulary and understanding the major data analysis problems in each field is a crucial task.%, and should be allocated sufficient time in form of introductory talks or tutorials to foster cross-disciplinary collaboration.
-%On the other hand, cross-disciplinary hack weeks around a specific technique or type of data set allow methods developed in a specific field to diffuse into other subjects and therefore help avoid duplication of method development efforts.
-
-%An example of such events are the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings held at the UC Berkeley Institute for Data Science in the Summer 2016, and at the UW eScience Institute in the Spring of %2017.
-%These three-day events also included a mix of tutorials, talks, and joint work on projects, but in contrast to the hack weeks described in the present paper, these events included participants from a variety of different research fields, with the common thread being that they all have an interest in the analysis of images, as part their research.
-%Thus, audience and speakers included researchers from wide array of domains: astronomy, neuroscience and geosciences, but also material science, chemistry, biology, medicine, engineering, and others.
-%In addition, participants included computer vision researchers, who develop algorithms that can be applied to answer scientific questions.
-%Similarly, a series of events at UC Berkeley (Text XD) focused on the analysis of text data across different domains of research.

### 2018-05-01 participants.tex

@@
 In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from attendees.
-Therefore in order to maximize learning outcomes and collaborative exchanges, it is crucial that the participant selection process be carried out with considerable care.
+In our experience, maximizing learning outcomes and collaborative exchanges at hack weeks requires a participant group that is diverse across categories of minority status, geographical origin, gender, discipline and career track.
 
-In our experience, a participant group that is diverse across categories of minority status, geographical origin, gender, discipline and career track helps to ensure we meet these objectives.
-To achieve this diversity, we advocate for a selection process that is as quantitative and transparent as possible, enabling participants to hold organizers accountable for their selection decisions.
-Research particularly in the hiring literature has shown that cohort selection is most effective and unbiased when selection procedures are as quantitative as possible~\cite{sunstein2015wiser}.
-
-Participant selection is an intrinsically difficult process that is fraught with personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
-It requires laying out a definition of successful participation, what criteria must be met in order to maximize the likelihood of success, and how those criteria will be assessed given the information about the candidates selected during the application stage.
-Traditional selection processes rely heavily on internal heuristics of reviewers, including using unrelated characteristics as a cipher for suitability.
-On the other hand, while purely mechanistic procedures may to be able to mitigate some reviewer biases, this may come at a cost of baking in structural inequalities.
+Traditional selection processes that rely heavily on internal heuristics of reviewers, especially thost that consider characteristics peripheral to the evaluation criteria, are often fraught personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
+In order to maximize diversity and minimize bias, we advocate for a selection process that is as quantitative and transparent as possible~\cite{sunstein2015wiser}, enabling participants to hold organizers accountable for their selection decisions.
+This requires laying out a definition of successful participation, what criteria must be met in order to maximize the likelihood of success, and how those criteria will be assessed given the information about the candidates selected during the application stage.
 
 For hack weeks, it is not obvious a priori that prerequisites for attendance in terms of skills should exist, and prerequisites will depend on the objectives of the workshop.
-For example, Astro Hack Week has traditionally accepted participants at all skill levels with respect to data science and did not a merit-based selection, whereas Neuro Hack Week did include skill-based criteria in their selection (see also SM, Section 2.1 for more detail on the individual selection procedures).
-If merit-based selection is part of the process, organizers face the crucial decision whether to assess merit blindly or not.
-Research particularly around hiring procedures has shown that there is no single correct procedure.
+For example, Astro Hack Week has traditionally accepted participants at all skill levels with respect to data science and did not include a merit-based selection, whereas Neuro Hack Week did include skill-based criteria in their selection (see also SM, Section 2.1 for more detail on the individual selection procedures).
 
+If merit-based selection is part of the evaluation process, organizers face the crucial decision whether to assess merit blindly or not.
 Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases related to internal heuristics.

### 2018-05-01 conclusions.tex

--- 2018-01-08 sections/conclusions.tex
@@
 As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
-Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
+Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives.
+\textbf{However, the most important independent variable--hack week attendance--was not probed in previous surveys.
+Comparisons between equivalent groups of researchers who did or did not attend these workshops are important in order to quantify more precisely the effect a hack week has on individual career progression.}
 Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.

### 2018-05-01 intro.tex

--- 2018-04-13 intro.tex
@@
 \begin{center}
-\includegraphics[width=8cm]{NewHackSpectrum.png}
+\includegraphics[width=8cm]{NewHackSpectrum.pdf}
 \caption{{\bf Comparison of Extracurricular Workshop Models}: Different types of events lie on a spectrum between the emphasis on pedagogy (e.g.
@@
 As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
-Here we share the philosophy behind the hack week model as we have implemented it, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines (supplementary materials provide additional details on the practical aspects of organizing these events).
+Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.
+Supplementary materials (SM) provide additional details on the practical aspects of organizing these events.

### 2018-05-01 results.tex

@@
       \item[\emph{a}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-      \item[\emph{b}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
+      \item[\emph{b}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
 
@@
 One key metric is the number of publications that result from hack week projects, but this is a fairly narrow definition of success, in line with standard academic performance indicators.
-Assuming that participants work largely in the open during a hack week, and that most projects have a strong programming component another indicator of success is the activity of participants in terms of code written and committed to a public code repository.
+Assuming that participants work largely in the open during a hack week, and that most projects have a strong programming component, another indicator of success is the activity of participants in terms of code written and committed to a public code repository.
 Still, these measures ignore learning, community-building as well as networking outcomes, which can be assessed through post-workshop surveys.

### 2018-05-01 why.tex
Shortened design section a little and moved some text around

@@
 \noindent It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
-Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see supplementary materials).
+Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see SM, Section 1.2).

### 2018-05-07 intro.tex

--- 2018-04-16 intro.tex
@@
 In general, many of these events tend to emphasize either a more traditional pedagogical class and lecture methodology, or emphasize a focus on ad-hoc projects that participants work on during the event (Figure \ref{fig:hackspectrum}).
-Pedagogically-focused events (e.g., Software Carpentry workshops \cite{b:wilson-swc-lessons-2016} in which no student-led projects are pursued) follow a classic academic model where novices learn new skills from experts.
+Pedagogically-focused events follow a classic academic model where novices learn new skills from experts.
 This model tends to focus on a one-way flow of information from instructor to student, and are usually targeted towards participants in the training phase of their career.
@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+Our hackweek model shares some similarities to those summer schools (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course) that combine individual or group project work with teaching methods that transfer knowedge from experts to early-career researchers.
+However, a distinguishing factor of our hackweeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
+This looser structure is designed to facilitate a multi-directional transfer of knowledge between a diverse group of people in different stages of their career and with different backgrounds (see also \ref{sec:participants}).
+The flexibility of our approach results in shifts along the continuum of pedagogy and open project work across hackweek disciplines, and from year to year within a specific discipline (Figure \ref{fig:hackspectrum}).
 

### 2018-05-11 intro.tex
Added intro changes from Anthony

@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
-Our hackweek model shares some similarities to those summer schools (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course) that combine individual or group project work with teaching methods that transfer knowedge from experts to early-career researchers.
+Our hackweek model shares some similarities to those summer schools (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course) that combine individual or group project work with teaching methods that transfer knowledge from experts to early-career researchers.
 However, a distinguishing factor of our hackweeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.

### 2018-05-12 intro.tex
Added N to figures, also updated some text

@@
 However, a distinguishing factor of our hackweeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
-This looser structure is designed to facilitate a multi-directional transfer of knowledge between a diverse group of people in different stages of their career and with different backgrounds (see also \ref{sec:participants}).
+This looser structure is designed to facilitate a multi-directional transfer of knowledge between a diverse group of people in different stages of their career and with different backgrounds (see also section Participants below).
 The flexibility of our approach results in shifts along the continuum of pedagogy and open project work across hackweek disciplines, and from year to year within a specific discipline (Figure \ref{fig:hackspectrum}).

### 2018-05-12 participants.tex
Added N to figures, also updated some text

@@
 In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from attendees.
-In our experience, maximizing learning outcomes and collaborative exchanges at hack weeks requires a participant group that is diverse across categories of minority status, geographical origin, gender, discipline and career track.
+In our experience, maximizing learning outcomes and collaborative exchanges at hack weeks requires a participant group that is diverse across categories of minority status, geographical origin, gender, discipline and career track, among others.
 
-Traditional selection processes that rely heavily on internal heuristics of reviewers, especially thost that consider characteristics peripheral to the evaluation criteria, are often fraught personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
+Traditional selection processes that rely heavily on internal heuristics of reviewers, especially those that consider characteristics peripheral to the evaluation criteria, are often fraught personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
 In order to maximize diversity and minimize bias, we advocate for a selection process that is as quantitative and transparent as possible~\cite{sunstein2015wiser}, enabling participants to hold organizers accountable for their selection decisions.
@@
 This type of blinded procedure tends to reduce biases during participant selection when committees would otherwise not consider diversity during their selection [REF].
-However, a blind selection purely on a figure of merit will likely still be biased if some groups of applicants have been systematically disadvantaged in certain ways: for example, if workshop organizers require a minimum level of programming experience, this could disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
 
-Blinding to demographics can be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills.
+However, a blind selection purely on a figure of merit will be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills: for example, if workshop organizers require a minimum level of programming experience, this could disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
 Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.

### 2018-05-12 results.tex
Added intro changes from Anthony
Added N to figures, also updated some text

@@
 \section*{Results}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{f2.eps}
+%\end{subfigure}
+\caption{Post-workshop surveys from three hack weeks: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
+The overall response rates for GHW and NHW were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
+At AHW, the combined overall response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
+Note that individual response rates may be slightly lower (as some questions were optional) and are reported in the panel titles for AHW, GHW and NHW, respectively.
+We report here about results in three different domains: the development of technical skills (a -- c), collaboration and teaching (d -- f), and shifts in attitudes towards reproducibility and open science (g, h).}
+\label{fig:survey}
+%\end{center}
+\end{figure*}
 
@@
  & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
-\multicolumn{16}{l}{\textbf{Comparison with career stage}}  \\ \midrule
+\multicolumn{16}{l}{\textbf{Comparison with career stage (early-career stage; faculty and research scientists; other)}}  \\ \midrule
 New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
@@
 Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
-\multicolumn{16}{l}{\textbf{Gender Identity}} \\ \midrule
+\multicolumn{16}{l}{\textbf{Gender Identity (non-minority; minority$^\emph{c}$)}} \\ \midrule
 Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
@@
 Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
-\multicolumn{16}{l}{\textbf{Racial/ethnic Identity}} \\ \midrule
+\multicolumn{16}{l}{\textbf{Racial/ethnic Identity (white; non-white)}} \\ \midrule
 Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
@@
    \begin{tablenotes}
-      \item{Respondents to a post-event evaluation survey indicate their level of agreement on a six-point Likert-type scale to a range of statements (corresponding to Figure \ref{fig:survey}).
+      \item{Respondents to a post-event evaluation survey indicate their level of agreement on a six-point Likert-type scale to a range of statements (letters correspond to panels in Figure \ref{fig:survey}).
 Results are for hack weeks held in 2016 and 2017 for AHW and in 2016 for GHW and NHW, respectively.
 The number of respondents for each question is included in the table above.
-Additional figures showing the distributions of scores are available in the SM, Section 6.}
-      \item[\emph{a}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-      \item[\emph{b}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
+Additional figures showing the distributions of responses for each question are available in the SM, Section 6.}
+      \item[\emph{a}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
+      \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
+      \item[\emph{c}]{We defined ``minority'' with respect to the field of research and grouped women and non gender-conforming participants into this category.}
 
@@
 \end{table*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{f2.eps}
-%\end{subfigure}
-\caption{Post-workshop surveys from three hack weeks: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-For GHW and NHW, response rates were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
-At AHW, the combined response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
-We report here about results in three different domains: the development of technical skills (a -- c), collaboration and teaching (d -- f), and shifts in attitudes towards reproducibility and open science (g, h).}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
 
@@
 However, we find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (a), improvements in day-to-day research (b) or overall improvements in science (c) (all statistical results are presented in Table \ref{tab:survey}).
-Only for GHW we find an indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following~\citep{cohen1988}).
+Only for GHW we find a weak indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following~\citep{cohen1988}).
 Similarly, we find no strong indication that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (d).

### 2018-05-14 design.tex

--- 2018-04-12 design.tex
@@
 
-As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
-If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
-In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.
+%As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
+%If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
+%In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.
 

### 2018-05-14 intro.tex

--- 2018-05-07 intro.tex
@@
 In general, many of these events tend to emphasize either a more traditional pedagogical class and lecture methodology, or emphasize a focus on ad-hoc projects that participants work on during the event (Figure \ref{fig:hackspectrum}).
-Pedagogically-focused events follow a classic academic model where novices learn new skills from experts.
+Pedagogically-focused events (e.g., Software Carpentry workshops \cite{b:wilson-swc-lessons-2016} in which no student-led projects are pursued) follow a classic academic model where novices learn new skills from experts.
 This model tends to focus on a one-way flow of information from instructor to student, and are usually targeted towards participants in the training phase of their career.
@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
-Our hackweek model shares some similarities to those summer schools (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course) that combine individual or group project work with teaching methods that transfer knowedge from experts to early-career researchers.
-However, a distinguishing factor of our hackweeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
-This looser structure is designed to facilitate a multi-directional transfer of knowledge between a diverse group of people in different stages of their career and with different backgrounds (see also \ref{sec:participants}).
-The flexibility of our approach results in shifts along the continuum of pedagogy and open project work across hackweek disciplines, and from year to year within a specific discipline (Figure \ref{fig:hackspectrum}).
 

### 2018-05-14 conclusions.tex
Minor wording changes

--- 2018-01-08 sections/conclusions.tex
@@
 As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
-Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives, but that the measure of success is field-specific in that it depends to some degree on how much the concepts hack weeks promote were already adopted within the community.
+Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives.
+
 Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.

### 2018-05-14 intro.tex
Minor wording changes

@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+Our hack week model shares some similarities to those summer schools (e.g., Advanced Course on Computational Neuroscience, Okinawa Computational Neuroscience Course, Woods Hole Computational Neuroscience Summer course) that combine individual or group project work with teaching methods that transfer knowledge from experts to early-career researchers.
+However, a distinguishing factor of our hack weeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
+This looser structure is designed to facilitate a multi-directional transfer of knowledge between a diverse group of people in different stages of their career and with different backgrounds (see also section Participants below).
+The flexibility of our approach results in shifts along the continuum of pedagogy and open project work across hack week disciplines, and from year to year within a specific discipline (Figure \ref{fig:hackspectrum}).
 

### 2018-05-14 participants.tex
Minor wording changes

@@
 
-For hack weeks, it is not obvious a priori that prerequisites for attendance in terms of skills should exist, and prerequisites will depend on the objectives of the workshop.
+For hack weeks, prerequisites will depend on the objectives of the workshop and may not exist at all.
 For example, Astro Hack Week has traditionally accepted participants at all skill levels with respect to data science and did not include a merit-based selection, whereas Neuro Hack Week did include skill-based criteria in their selection (see also SM, Section 2.1 for more detail on the individual selection procedures).
@@
 A merit selection could then be performed via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
-This type of blinded procedure tends to reduce biases during participant selection when committees would otherwise not consider diversity during their selection [REF].
+This type of blinded procedure tends to reduce biases during participant selection when committees would otherwise not consider diversity during their selection~\cite{bohnet2016}.
 

### 2018-05-14 results.tex
Minor wording changes

@@
 \caption{Post-workshop surveys from three hack weeks: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-The overall response rates for GHW and NHW were 100\%, with $N_{\mathrm{GHW}}= 83$ and $N_{\mathrm{NHW}} = 86$, respectively.
-At AHW, the combined overall response rate for both years was 76\%, or $N_{\mathrm{AHW}} = 72$ out of $94$ total attendees.
-Note that individual response rates may be slightly lower (as some questions were optional) and are reported in the panel titles for AHW, GHW and NHW, respectively.
+Response rates are reported in the panel titles for AHW, GHW and NHW, respectively.
 We report here about results in three different domains: the development of technical skills (a -- c), collaboration and teaching (d -- f), and shifts in attitudes towards reproducibility and open science (g, h).}
@@
 Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity (Table \ref{tab:survey}).
-Similarly, participants who self-identified as minorities with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks.
+Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks.
 Both underrepresented groups similarly feel that they built equally valuable connections at all three hack weeks and generally feel that their contributions to their hack teams were valued.

### 2018-05-14 supplementary_material.tex
Minor wording changes

@@
 Here we explore a number of survey responses with gender identity as an independent variable.
+We grouped ``gender identity'' into two categories: ``non-minority'' contains all participants who self-identified as male, ``minority'' contains all participants who responded with ``female'' or any other response.
+Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics and geoscience), where they tend to be strongly underrepresented.
 
@@
 Here we explore a number of survey responses with racial/ethnic identity as an independent variable.
+We grouped responses into two final categories, ``white'' (Caucasian participants) ``non-white'' (all others).
 

### 2018-05-14 themes.tex
Minor wording changes

@@
 On the other hand, homogeneity may lead to \textit{group think} and inhibit new, creative solutions. 
-In this case, it may be advantageous to design a hack week around a technique (e.g.\ Gaussian Processes) or modality (e.g.\ imaging), such as the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}} meetings.
+In this case, it may be advantageous to design a hack week around a technique (e.g.\ Gaussian Processes) or modality (e.g.\ imaging), such as the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}}) meetings.
 For these events, building a shared vocabulary and understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.

### 2018-05-14 what.tex
Minor wording changes

--- 2018-03-27 what.tex
@@
 Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks. 
-Our focus on both pedagogy and open project work aims to fill an educational space we feel is not currently addressed by existing models (Figure \ref{fig:hackspectrum}). 
+Our focus on both pedagogy and open project work aims to fill a space we feel is not currently addressed by existing models (Figure \ref{fig:hackspectrum}). 
 The hackathon, a time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events. 
@@
 Teaching within our events draws from a spectrum of approaches such as those traditionally employed in summer schools and software workshops. 
-However, by embedding these approaches in a more open hackathon framework, we can explore unique ways of enhance the traditional summer school model.
+By embedding these approaches in a more open hackathon framework, we can explore unique ways of enhancing the traditional summer school model.
 For example, our hack weeks transfer knowledge across many levels of seniority, with participants routinely leading tutorials, and with numerous opportunities to `learn by doing' within a constructivist educational framework \cite{Bransford2000-lu}.
-Additionally, the content of our tutorials tends to favor cutting-edge topics and experimental approaches over established knowledge. 
+The content of our tutorials tends to favor cutting-edge topics and experimental approaches over established knowledge. 
 Finally, the social structure of hackathons enable us to prioritize participant diversity thereby maximizing the likelihood for the co-creation of learning during the event.

### 2018-05-14 intro.tex

--- 2018-05-14 intro.tex
@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+
+Our hack week model shares some similarities to those summer schools that combine individual or group project work with teaching methods that transfer knowledge from experts to early-career researchers.
+However, a distinguishing factor of our hack weeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
 

### 2018-05-14 what.tex

--- 2018-04-17 what.tex
@@
 
-Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning, forming the second axis of our events.
+Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning.
 They aim to synthesize different goals and strategies from both models: they are much more participant-driven than a summer school, but have a stronger focus on pedagogy than a hackathon.

### 2018-05-14 conclusions.tex
Final changes

@@
 
-As introduced above, hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
+Hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
 Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives.

### 2018-05-14 design.tex
Minor wording changes
Final changes

@@
 
-A hackweek requires a flexible workspace environment with ample opportunity for re-configuration. 
+A hack week requires a flexible workspace environment with ample opportunity for re-configuration. 
 Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
-Workspaces must also accommodate interactive project work where small teams can gather and work together (for more detailed considerations, see also the SM, Sections 2 and 3).    
+Workspaces must also accommodate interactive project work where small teams can gather and work together (for more details, see also the SM, Sections 2 and 3).    
 %Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
@@
 If the group is too large, chances for random participant exchanges are reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
-If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines.
+If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations.
 In the past, we have found groups with sizes between 50 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group.
-However, we encourage organizers to experiment with group size depending on their workshop goals.
+We encourage organizers to experiment with group size depending on their workshop goals.
 

### 2018-05-14 intro.tex
Final changes

@@
 at science-oriented hackathons).
-Hack weeks also vary in the degree of emphasis on projects (Astro hack week, which included unstructured time for group projects from the first day) or pedagogy (Neuro hack week, which included an entire first day of data science tutorials, with less time for projects)}
+Hack weeks also vary in the degree of emphasis on projects (Astro Hack Week, which included unstructured time for group projects from the first day) or pedagogy (Neuro hack week, which included an entire first day of data science tutorials, with less time for projects)}
 \label{fig:hackspectrum}

### 2018-05-14 results.tex
Final changes

@@
 Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
-\multicolumn{16}{l}{\textbf{Gender Identity (non-minority; minority$^\emph{c}$)}} \\ \midrule
+\multicolumn{16}{l}{\textbf{Gender Identity (non-minority; minority)}} \\ \midrule
 Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
@@
       \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-      \item[\emph{c}]{We defined ``minority'' with respect to the field of research and grouped women and non gender-conforming participants into this category.}
-
 \end{tablenotes}
@@
 We also note that there might be systematic effects and biases due to the subjectivity of rating knowledge and learning outcomes.
-
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
 We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\%; Figure \ref{fig:survey}, (d)), though responses are not as unequivocal as they are in some of the other categories.
-The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, (e)), especially at Neuro Hack Week, where more than 64\% of participants strongly agreed with this statement.
+The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, (e)), especially at NHW, where more than 64\% of participants strongly agreed with this statement.
 
 Given the broad diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible.
-At the same time, we put forth the hypothesis earlier in the this paper that peer learning is a major mode of knowledge transfer during hack weeks, and that this type of learning takes place at all career stages.
-However, we find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (a), improvements in day-to-day research (b) or overall improvements in science (c) (all statistical results are presented in Table \ref{tab:survey}).
+At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer during hack weeks, and that this type of learning takes place at all career stages.
+We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (a), improvements in day-to-day research (b) or overall improvements in science (c) (all statistical results are presented in Table \ref{tab:survey}).
 Only for GHW we find a weak indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following~\citep{cohen1988}).

### 2018-05-14 what.tex
Final changes

@@
 Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
+Additionally, they provide opportunities to build new collaborations and network.
 In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).
@@
 Building on these core elements hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}.
-``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
-``Contributive'' hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
-Finally, ``Communal'' hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.
+%``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
+%``Contributive'' hackathons seek to improve to an existing effort through focused work on discrete tasks, for example to make up for deficiencies in an ongoing project.
+%Finally, ``Communal'' hackathons place a strong focus on building a culture of practice and developing resources within an existing community, often defined by a specific domain of knowledge.
 
@@
 Data science is a new, emerging area combining knowledge and skills from many different fields, only a fraction of which can be communicated within a short tutorial.
-As organizers, we make a curated selection of a small number of tutorials each year, and encourage participants to use the unstructured time to self-organize tutorials on any topic as needed, often on topics vastly different to those discussed in the formal lectures.
+As organizers, we make a curated selection of a small number of tutorials each year, and encourage participants to use the unstructured time to self-organize tutorials on any topic as needed.%, often on topics vastly different to those discussed in the formal lectures.
 
-Additionally, as we lay out in the subsequent section below, multi-directional knowledge transfer is one of many goals of a hack week, and participants often attend with more than one objective.
 The unstructured schedule and inherent lack of focus on specific topics within a hack week present both an opportunity and a risk.
@@
 Additionally, a hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require more preparation from the side of participants than other events (e.g.
-defining project ideas, researching possible connections to other participants, bringing along their own tutorials).
+define project ideas, research possible connections to other participants).
 Because of these risks, organizers face a much larger degree of uncertainty both before and during a hack week, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, keeping track of the goals different attendees may have, and careful facilitation of the large range of different types of activities at any given time (see also the following sections as well as the SM, Section 4).

### 2018-05-14 participants.tex

@@
 A merit selection could then be performed via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
-This type of blinded procedure tends to reduce biases during participant selection when committees would otherwise not consider diversity during their selection~\cite{bohnet2016}.
+This type of blinded procedure tends to reduce biases when committees would otherwise not consider diversity during their selection~\cite{bohnet2016}.
 
-However, a blind selection purely on a figure of merit will be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills: for example, if workshop organizers require a minimum level of programming experience, this could disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
+However, a blind selection purely on a figure of merit will be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills: for example, requiring a minimum level of programming experience will likely disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
 Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.

### 2018-05-14 supplementary_material.tex

@@
 
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures, ~\cite{highhouse2008} it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures~\citep{highhouse2008}, it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 
@@
 
-Some of our hack week tutorials follow the structure of the Software Carpentry model \cite{b:wilson-swc-lessons-2016}.
+Some of our hack week tutorials follow the structure of the Software Carpentry model \citep{b:wilson-swc-lessons-2016}.
 The content is divided in to a series of lessons, each having a well-defined set of learning objectives, key questions, and expected outcomes.
@@
 This provides a simple method to deliver help to those students who need it without interrupting the flow of the tutorial.
-We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \cite{kluyver2016jupyter} or following the markdown templates provided by Software Carpentry.
+We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \citep{kluyver2016jupyter} or following the markdown templates provided by Software Carpentry.
 This way the hack week ``curriculum'' can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
@@
 Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometime so complicated that it slows down teaching.
-In some cases, we have experimented with \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \cite{holdgraf2017portable}.
-Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \cite{Holdgraf2017-pd, Titus_Brown_undated-pc}.
+In some cases, we have experimented with \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \citep{holdgraf2017portable}.
+Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \citep{Holdgraf2017-pd, Titus_Brown_undated-pc}.
 
@@
 The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \cite{Clance1978-ef}.
+Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \citep{Clance1978-ef}.
 Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS, namely that participants feel the need to be experts in multiple scientific and technical aspects of the activities pursued during a hack week.
@@
 The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
-This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
+This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
 Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
@@
 \caption{}
-\label{fig:survey}
+\label{fig:survey}utbt
 %\end{center}

### 2018-05-28 conclusions.tex
Reworked version to satisfy editorial requests

@@
 The fast-paced changes of the computational and methodological landscape require traditional fields of science to rapidly adapt to new data analysis challenges.
-Traditional modes of learning, including university curricula, are often too slow to incorporate new developments on short enough time scales to meet their acute need in scientific advancement.
-To address this imbalance, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
+To address these challenges, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
 Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields.
 
-Hack weeks serve multiple purposes, including dissemination of state-of-the-art technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research as well as  promoting open science and reproducibility.
+Hack weeks serve multiple purposes, including dissemination of technological advances through the scientific community, building collaborations between academic subdisciplines and fostering interdisciplinary research.
 Initial results from three events held in 2016 and 2017 in three different fields (astronomy, geosciences and neurosciences) indicate that hack weeks succeed at all of these objectives.
 
-Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes for early-career academics and adoption of new methods.
+Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes and adoption of new methods.
 We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.

### 2018-05-28 design.tex
Reworked version to satisfy editorial requests

@@
 
-Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 1 and 2 for practical guidance) .   
+Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance) .   
 Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges. 
@@
 Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
-Workspaces must also accommodate interactive project work where small teams can gather and work together (for more details, see also the SM, Sections 2 and 3).    
+Workspaces must also accommodate interactive project work where small teams can gather and work together (for more details, see also the SM, Sections 4.1.3 and 4.1.4).    
 %Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
@@
 
-Another important design consideration is group size (see also SM, Section 2).
-If the group is too large, chances for random participant exchanges are reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
+Another important design consideration is group size (see also SM, Section 4.1.4).
+In a large group, chances for random participant exchanges may be reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
 If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations.
-In the past, we have found groups with sizes between 50 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group.
-We encourage organizers to experiment with group size depending on their workshop goals.
+We have found groups with sizes between 50 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group, but we encourage organizers to experiment with group size.
 
@@
 This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
-The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week.
-A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 2.3).
+The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week (see also SM; Section 4.2.4).
+A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 4.1.7).
 Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
-In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4 for concrete suggestions).
+In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).

### 2018-05-28 intro.tex
Reworked version to satisfy editorial requests

@@
 
-Our hack week model shares some similarities to those summer schools that combine individual or group project work with teaching methods that transfer knowledge from experts to early-career researchers.
-However, a distinguishing factor of our hack weeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
+%Our hack week model shares some similarities to those summer schools that combine individual or group project work with teaching methods that transfer knowledge from experts to early-career researchers.
+%However, a distinguishing factor of our hack weeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
 
@@
 at science-oriented hackathons).
-Hack weeks also vary in the degree of emphasis on projects (Astro Hack Week, which included unstructured time for group projects from the first day) or pedagogy (Neuro hack week, which included an entire first day of data science tutorials, with less time for projects)}
+Hack weeks also vary in the degree of emphasis on projects (e.g.
+Astro Hack Week) or pedagogy (e.g.
+Neuro Hack Week)}
 \label{fig:hackspectrum}
@@
 
-As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, two focused on Neuroscience, and two focused on Geoscience.
-Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks in other disciplines.
+As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, and two each focused on Neuroscience and Geoscience.
+Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks.
 Supplementary materials (SM) provide additional details on the practical aspects of organizing these events.

### 2018-05-28 participants.tex
Reworked version to satisfy editorial requests

@@
 
-Hack weeks differ from traditional conferences or summer schools in that knowledge transfer occurs across many levels of seniority, disciplinary boundaries, and novelty of the topics discussed.
+Hack weeks differ from many traditional conferences or summer schools in that knowledge transfer occurs across many levels of seniority and disciplinary boundaries.
 In addition, a substantial amount of hack week content is generated during the event itself, requiring active participation from attendees.
@@
 For hack weeks, prerequisites will depend on the objectives of the workshop and may not exist at all.
-For example, Astro Hack Week has traditionally accepted participants at all skill levels with respect to data science and did not include a merit-based selection, whereas Neuro Hack Week did include skill-based criteria in their selection (see also SM, Section 2.1 for more detail on the individual selection procedures).
+For example, Astro Hack Week has traditionally accepted participants at all skill levels with respect to data science and did not include a merit-based selection, whereas Neuro Hack Week did include skill-based criteria in their selection (see also SM, Section 4.1.5 for more detail on the individual selection procedures).
 
@@
 Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
-In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also SM, Section 2.1). 
+In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also SM, Section 4.1.5). 
 Because systemic biases likely also enter at the application stage (where underrepresented groups may be less likely to apply) organizers should consider oversampling traditionally disenfranchised groups compared to the population of applicants.

### 2018-05-28 results.tex
Reworked version to satisfy editorial requests

@@
 \end{figure*}
-
-\begin{table*}
-%\begin{table*}[hbtp]
-%\renewcommand{\arraystretch}{1.3}
-\footnotesize
-\caption{}
-\begin{threeparttable} 
-\begin{tabularx}{18cm}{p{3.4cm}p{0.3cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}p{0.33cm}}
-\toprule
-\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
-\multicolumn{16}{l}{\textbf{Comparison with career stage (early-career stage; faculty and research scientists; other)}}  \\ \midrule
-New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
-Improvement of research (b) & 61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
-Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\ 
-Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
-\multicolumn{16}{l}{\textbf{Gender Identity (non-minority; minority)}} \\ \midrule
-Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
-Teaching (d) & 61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
-Connection building (e) & 61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
-Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
-\multicolumn{16}{l}{\textbf{Racial/ethnic Identity (white; non-white)}} \\ \midrule
-Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
-Teaching (d) & 61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
-Connection building (e) & 61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
-Valued contributions (f) & 58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\ 
-\bottomrule
-\end{tabularx}
-   \begin{tablenotes}
-      \item{Respondents to a post-event evaluation survey indicate their level of agreement on a six-point Likert-type scale to a range of statements (letters correspond to panels in Figure \ref{fig:survey}).
-Results are for hack weeks held in 2016 and 2017 for AHW and in 2016 for GHW and NHW, respectively.
-The number of respondents for each question is included in the table above.
-Additional figures showing the distributions of responses for each question are available in the SM, Section 6.}
-      \item[\emph{a}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
-      \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-\end{tablenotes}
-
-\end{threeparttable}
-\label{tab:survey}
-%\end{table*}
-\end{table*}
 
@@
 
-Given the broad diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible.
-At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer during hack weeks, and that this type of learning takes place at all career stages.
-We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (a), improvements in day-to-day research (b) or overall improvements in science (c) (all statistical results are presented in Table \ref{tab:survey}).
-Only for GHW we find a weak indication that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.4$ (following~\citep{cohen1988}).
-Similarly, we find no strong indication that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (d).
+Given the diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible.
+At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer for participants of all career stages.
+We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12).
+Only for GHW do we find a very weak suggestion that early-career researchers might agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
+Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)).
 
-One important question regarding hack week outcomes is whether participants from underrepresented groups thrive at these events, or whether their full participation is impeded.
+One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded.
 Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building would indicate that improvements in workshop facilitation and structure may be required to allow participants of all stripes to participate fully.
-In Section 4 of the Supplementary Materials, we lay out various strategies we have employed to facilitate participation, empower members of underrepresented groups and mitigate impostor syndrome.
-Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity (Table \ref{tab:survey}).
-Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks.
-Both underrepresented groups similarly feel that they built equally valuable connections at all three hack weeks and generally feel that their contributions to their hack teams were valued.
-For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections (p=0.04) with a medium to large effect size ($\phi_c = 0.4$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01; \phi_c = 0.4$).
+Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
+Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$).
+In neither case seem the attitudes toward building valuable connections at all three hack weeks or the value of their contributions to their hack teams differ with respect to gender or racial/ethnic identity.
+For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
 These initial results indicate that future work on how demographics interact with hack week attendance may be fruitful.
-We note that since we did not include a question about impostor syndrome itself, we cannot draw causal connections between our results and its prevalence.
-Additionally, our number of respondents is small, and the lack of significant differences may be caused by the small statistical power of our sample.
-%It is notable, however, that we find no correlation between the response to this question and career stage for any of the events surveyed here.
-Participants at all career stages report similar engagement in teaching.
-This suggests there is some evidence for our hypothesis that compared to e.g.\ a traditional summer school, a hack week is less hierarchical and encourages lateral knowledge transfer between participants at different career stages.
-However, we note that the sample sizes per category are relatively small, and future work is needed to confirm this hypothesis.
-We similarly find no correlations with gender identity or race/ethnicity.
 
-We find that the hack weeks have been largely successful at efforts to promote positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)).
-At all three events, more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
+We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
 %The overall behaviour hints at how general conventions and attitudes differ in different fields.
@@
 
-We caution the reader that the survey results should be read only as an indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the exploratory nature of the surveys conducted as part of past hack weeks we have organized.
+We caution the reader that our survey results should be read only as an initial indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the surveys' exploratory nature.
+Our number of respondents is small and the effects likely subtle, thus the lack of significant differences may be caused by the small statistical power of our sample.
 The most important independent variable--attendance at the hack week--is something that our current survey design cannot account for.
 Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to internal biases of respondents.
-Future work will include more a more refined survey design as well as attendance versus non-attendance as a control variable.
+Future work will include a more refined survey design as well as attendance versus non-attendance as a control variable.
 
@@
 \subsubsection*{Example 1: Astro Hack Week}
-In 2015, a small team used AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing well-tested implementations of time series analysis algorithms often used in X-ray astronomy.
-The start of this project was facilitated by the collaborative environment at Astro Hack Week, including expertise in how to start/run open-source projects and role models of successful projects.
-Astro Hack Week enabled participants to seed a new collaboration around a software project needed by the larger community.
+In 2015, a small team used AHW to found a new software project called Stingray\footnote{https://github.com/StingraySoftware/stingray} with the goal of providing implementations of time series analysis algorithms often used in astronomy.
+Astro Hack Week enabled participants to seed a new collaboration around a software project needed by the larger community, facilitated by the collaborative environment at AHW.
 Stingray has since matured into an enduring collaboration within the community with five active maintainers and four Google Summer of Code projects.
@@
 The team used machine learning algorithms in conjunction with the powerful hardware resources provided by Google Earth Engine\footnote{\url{http://georgerichardson.net/2017/04/10/searching-for-energy-in-a-random-forest/}}.
-George Richardson, one of the project leads, now works for a renewable resource company in Seattle.
+%George Richardson, one of the project leads, now works for a renewable resource company in Seattle.
 \subsubsection*{Example 3: Neuro Hack Week}

### 2018-05-28 supplementary_material.tex
Reworked version to satisfy editorial requests

@@
 
-\newpage
+%\newpage
+
+\section{Materials and Methods}
+\label{sec:suppmaterials}
+We performed post-attendance surveys for AHW, GHW and NHW in 2016 and 2017.
+All surveys contained general questions about attitudes towards the workshop as well as open science and reproducibility, shared among all three surveys.
+The NHW and GHW surveys and the AHW 2017 survey were administered on site on the last day of the workshop; AHW participants in 2016 were e-mailed immediately after the workshop, with reminders after several weeks.
+All responses were collected within four weeks of the end of the workshop.
+The experimental procedures were approved by the Institutional Review Boards at UW, NYU and UC Berkeley.
+All participants gave their informed consent.
+
+Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100\% in both years; the response rate for AHW was 71\% (35 out of 49) in 2016 and 82\% (37 out of 45) in 2017.
+The lower response rates for AHW can be explained by the generally lower rates of response for time-delayed survey in 2016.
+In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered.
+For the AHW surveys, we checked whether the demographics of the respondents different significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and find no significant difference for any of the groups except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees).
+Because we group responses in the career stage category into categories ``early-career'' (students and postdoctoral fellows), ``late-career'' (faculty, research scientists) and ``other'' (e.g.
+participants from outside academia), this discrepancy has no bearing on the results we report.
+For gender, we grouped participants into ``minority'' (women and non-binary) and ``non-minority'' (men) participants; similarly, for race and ethnicity, we also chose ``non-white'' (all but Caucasian participants) and ``white'' (Cauciasian) categories.
+Note that demographic data for GHW and NHW is only available for 2016, for AHW for both 2016 and 2017, and responding to any question was optional.
+
+Participants were asked to respond to statements regarding these topics using a six-point Likert-type scale.
+All questions were anonymously recorded.
+No responses were discarded, and no pre-processing was performed on the data.
+We test for correlations between demographic characteristics (independent variable) and question responses (dependent variable) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
+For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
+The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
+
+\clearpage
+
+\section{Additional Statistical Results}
+
+Here we present the full results of the statistical analysis performed for the survey (Table \ref{tab:survey}).
+We searched for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity.
+The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, and (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$).
+The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}
+
+For no combination of question with a demographic variable, we find a $p$-value below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
+In the main text, we put these non-detections into context, quote those tests that near significance, coupled with a moderate to large effect size, as an indication where future research may be fruitful, and present the study's limitations.
+
+\begin{table*}
+%\begin{table*}[hbtp]
+%\renewcommand{\arraystretch}{1.3}
+\footnotesize
+\caption{}
+\begin{threeparttable} 
+\begin{tabularx}{18cm}{p{3.6cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}}
+\toprule
+\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
+\multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
+New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
+Improvement of research (b) & 61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
+Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\ 
+Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
+\multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
+Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
+Teaching (d) & 61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
+Connection building (e) & 61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
+Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
+\multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
+Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
+Teaching (d) & 61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
+Connection building (e) & 61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
+Valued contributions (f) & 58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\ 
+\bottomrule
+\end{tabularx}
+   \begin{tablenotes}
+      \item{Respondents to a post-event evaluation survey indicate their level of agreement on a six-point Likert-type scale to a range of statements (associated letters correspond to panels in Figure 2 in the main text).
+Results are for hack weeks held in 2016 and 2017 for AHW and in 2016 for GHW and NHW, respectively.
+The number of respondents for each question is included in the table above.
+Additional figures showing the distributions of responses for each question are available in Section 3 of the SM below.}
+      \item[\emph{a}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
+      \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
+      \item[\emph{c}]{Responses to the question of career stage were grouped into three responses: early-career stage; faculty and research scientists; other}
+      \item[\emph{d}]{Responses to the question of gender identity were grouped with respect to representation within the field of study into two categories: non-minority; minority}
+      \item[\emph{e}]{Responses to the question of racial and ethnic identity were grouped with respect to representation within the field of study into two categories: white; non-white}
+\end{tablenotes}
+
+\end{threeparttable}
+\label{tab:survey}
+%\end{table*}
+\end{table*}
+
+\clearpage
+
+\section{Additional Figures}
+\label{sec:suppfigures}
+Here, we present a number of additional figures related to the survey data presented in the main manuscript.
+The stacked bar charts presented here allow a closer look at the data that produced the results quoted in Table 1 above in the SM.
+
+\subsection{Career stage}
+
+Here we explore a number of survey responses with career stage as an independent variable.
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_3_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr1}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_3_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr2}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_4_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr3}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_2_Q3_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr4}
+%\end{center}
+\end{figure*}
+
+\clearpage
+
+\subsection{Gender Identity}
+
+Here we explore a number of survey responses with gender identity as an independent variable.
+We grouped ``gender identity'' into two categories: ``non-minority'' contains all participants who self-identified as male, ``minority'' contains all participants who responded with ``female'' or any other response.
+Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics and geoscience), where they tend to be strongly underrepresented.
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_4_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr5}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_2_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr6}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_5_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr7}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q19_5_Q27_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr8}
+%\end{center}
+\end{figure*}
+
+\clearpage
+
+\subsection{Racial/Ethnic Identity}
+
+Here we explore a number of survey responses with racial/ethnic identity as an independent variable.
+We grouped responses into two final categories, ``white'' (Caucasian participants) ``non-white'' (all others).
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_4_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr9}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q23_2_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr10}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q24_5_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr11}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}[h!]
+%\begin{center}
+%\begin{subfigure}[t]{\textwidth}
+\centering
+%\caption{}
+\includegraphics[width=\textwidth]{Q19_5_Q30_stackedbars.eps}
+%\end{subfigure}
+\caption{}
+\label{fig:corr12}
+%\end{center}
+\end{figure*}
+
+\clearpage
+
+\section{Practical Considerations}
+\subsection{What to do before}
 
@@
 
-\section{What to do before}
-
-\subsection{Obtaining permission for tracking and studying participants}
+\subsubsection{Obtaining permission for tracking and studying participants}
 A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes.
@@
 
-\subsection{Determining your objectives}
+\subsubsection{Determining your objectives}
 
@@
 
-\subsection{Hack Week Location}
+\subsubsection{Hack Week Location}
 When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration.
@@
 
-\section{Event Size and Duration}
+\subsubsection{Event Size and Duration}
 
@@
 
-\subsection{Participant selection}
+\subsubsection{Participant selection}
 
@@
 
-\subsection{Funding}
+\subsubsection{Funding}
 
@@
 
-\subsection{Code of Conduct}
+\clearpage 
+\subsubsection{Code of Conduct}
 
@@
 
-\subsection{Communicating with the Outside World}
+\subsubsection{Communicating with the Outside World}
 
@@
 
-\section{During a Hack Week}
+\clearpage 
+\subsection{During a Hack Week}
 
@@
 
-\subsection{Tutorials}
+\subsubsection{Tutorials}
 Tutorials provide a mechanism for rapid knowledge sharing and are often one of the most useful components of a hack week.
@@
 
-\subsection{Break-Out Sessions}
+\subsubsection{Break-Out Sessions}
 
@@
 
-\subsection{Hack Sessions}
+\subsubsection{Hack Sessions}
 
@@
 
-\section{Facilitating Participation}
+\subsubsection{Facilitating Participation}
 
@@
 
-\subsection{Impostor Syndrome}
+\subsubsection{Impostor Syndrome}
 
@@
 
-\section{What to do after}
+\subsection{What to do after}
 
@@
 
-\section{Additional Figures}
-
-Here, we present a number of additional figures related to the survey data presented in the main manuscript.
-The stacked bar charts presented here allow a closer look at the data that produced the results quoted in Table 1 of the manuscript.
-
-\subsection{Career stage}
-
-Here we explore a number of survey responses with career stage as an independent variable.
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q23_3_Q3_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q24_3_Q3_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q24_4_Q3_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q23_2_Q3_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\clearpage
-
-\subsection{Gender Identity}
-
-Here we explore a number of survey responses with gender identity as an independent variable.
-We grouped ``gender identity'' into two categories: ``non-minority'' contains all participants who self-identified as male, ``minority'' contains all participants who responded with ``female'' or any other response.
-Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics and geoscience), where they tend to be strongly underrepresented.
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q24_4_Q27_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q23_2_Q27_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q24_5_Q27_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q19_5_Q27_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}utbt
-%\end{center}
-\end{figure*}
-
-\clearpage
-
-\subsection{Racial/Ethnic Identity}
-
-Here we explore a number of survey responses with racial/ethnic identity as an independent variable.
-We grouped responses into two final categories, ``white'' (Caucasian participants) ``non-white'' (all others).
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q24_4_Q30_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q23_2_Q30_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q24_5_Q30_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\begin{figure*}[h!]
-%\begin{center}
-%\begin{subfigure}[t]{\textwidth}
-\centering
-%\caption{}
-\includegraphics[width=\textwidth]{Q19_5_Q30_stackedbars.eps}
-%\end{subfigure}
-\caption{}
-\label{fig:survey}
-%\end{center}
-\end{figure*}
-
-\newpage
 \bibliographystyle{aasjournal}

### 2018-05-28 what.tex
Reworked version to satisfy editorial requests

@@
 
-Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
+%Our hack weeks combine structured, tutorial-style instruction with informal education and peer learning opportunities occurring within projects and hacks.
 Additionally, they provide opportunities to build new collaborations and network.
+Our hack weeks combine structured, tutorial-style instruction with open-ended project work, providing opportunities for peer learning, networking, and building collaborations.  
 In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).
@@
 Our hack weeks extend the scientifically-focused, communal hackathon model into a space that includes a strong element of pedagogy and peer learning.
-They aim to synthesize different goals and strategies from both models: they are much more participant-driven than a summer school, but have a stronger focus on pedagogy than a hackathon.
-Where a summer school is often organized around a framework of lectures and tutorials known in advance, hack weeks leave the majority of time unstructured.
-Tutorials at hack weeks often serve as an entry point into a topic for further exploration and learning, rather than an in-depth study.
-This is closely related to the nature of the material taught at these events.
+They aim to synthesize different goals and strategies from both models: they are more participant-driven than a summer school, but have a stronger focus on pedagogy than a hackathon.
+Where a summer school is often organized around a framework of lectures and tutorials known in advance, hack weeks leave the majority of time to be designed by participants, under careful facilitation of the organizers.
+Tutorials at hack weeks often serve as an entry point into a topic for further exploration and learning.
+%This is closely related to the nature of the material taught at these events.
 Data science is a new, emerging area combining knowledge and skills from many different fields, only a fraction of which can be communicated within a short tutorial.
@@
 
-The unstructured schedule and inherent lack of focus on specific topics within a hack week present both an opportunity and a risk.
-They uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number of tutorials and their content.
+Hack weeks uniquely allow organizers to tailor the content of the workshop to the needs of the participants in an ad-hoc fashion, including the number and content of tutorials.
 This way, the group as a whole can respond quickly or react to unforeseen challenges and opportunities.
@@
 
-There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week can run the danger of not doing any of them well.
-Because tutorials are not the major focus of a hack week, the knowledge gained by participants in these tutorials may not be very deep.
-Additionally, a hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require more preparation from the side of participants than other events (e.g.
-define project ideas, research possible connections to other participants).
-Because of these risks, organizers face a much larger degree of uncertainty both before and during a hack week, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, keeping track of the goals different attendees may have, and careful facilitation of the large range of different types of activities at any given time (see also the following sections as well as the SM, Section 4).
+There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week might potentially not do any of them well.
+Because tutorials are not necessarily the major focus of a hack week, the knowledge gained by participants in these tutorials may be shallow.
+A hack week carries a much larger risk of failure if objectives and expectations are not set by organizers well in advance, and clearly communicated to participants, because they often require significant preparation from the side of participants.
+Because of these risks, organizers face a much larger degree of uncertainty, and need to be prepared to focus much of their energy on thoughtful selection and management of participants, and facilitation of the large range of different types of activities at any given time (see also the SM, Section 4.2.4).
 

### 2018-05-28 why.tex
Reworked version to satisfy editorial requests

@@
 \item{\textit{Education and Training}:
-While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component that entails discussions about reproducible research.
-Participants gain a strong foundation in open science practices from the diverse group setting and go on the become ambassadors for such practices.
+While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning.
 This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
@@
 
-\noindent It is worth noting that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
-Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see SM, Section 1.2).
+\noindent Note that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
+Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see SM, Section 4.1.2).

### 2018-06-02 supplementary_material.tex
Implemented equivalence test, currently fixing table

@@
 We test for correlations between demographic characteristics (independent variable) and question responses (dependent variable) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
-For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
+For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a medium effect size according to~\citep{cohen1988}.
+
+We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
+For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size.
+We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
+For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $\[0,1\]$, we only require an upper bound} this equivalence bound.
+If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.
+
 The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
@@
 \begin{threeparttable} 
-\begin{tabularx}{18cm}{p{3.6cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}}
+\begin{tabularx}{18cm}{p{3.5cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
 \toprule
-\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+\textbf{Question} & \multicolumn{7}{c}{\textbf{Astro Hack Week}} & \multicolumn{7}{c}{\textbf{GeoHackWeek}} & \multicolumn{7}{c}{\textbf{NeuroHackWeek}} \\ \midrule
  & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule

### 2018-06-04 abstract.tex
Pass through abstract.

--- 2017-08-30 sections/abstract.tex
@@
 \begin{abstract}
-Across almost all scientific disciplines, the instruments that record our experimental data and the methods required for storage and data analysis are rapidly increasing in complexity.
-This gives rise to the need for scientific communities to adapt on shorter time scales than traditional university curricula allow for, and therefore requires new modes of knowledge transfer.
-The universal applicability of data science tools to a broad range of problems has generated new opportunities to foster exchange of ideas and computational workflows across disciplines.
-In recent years, hack weeks have emerged as an effective tool for fostering these exchanges by providing training in modern data analysis workflows.
+Across almost all scientific disciplines, instruments that record experimental data and methods required for storage and data analysis are rapidly increasing in complexity.
+As a result of these rapid developments, scientific communities need to adapt on shorter time scales than traditional university curricula can evolve.
+New modes of knowledge transfer are therefore needed.
+On the other hand, the broad applicability of data science tools to a range of problems has generated new opportunities to foster exchange of ideas and computational workflows across disciplines.
+In recent years, hack weeks have emerged as a novel and effective tool for fostering these exchanges and providing training in modern data analysis workflows.
 While there are variations in hack week implementation, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and project work in a collaborative environment.
-In this paper, we present the concept of a hack week in the larger context of scientific meetings and point out similarities and differences to traditional conferences.
-We motivate the need for such an event and present in detail its strengths and challenges.
-We find that hack weeks are successful at cultivating collaboration and the exchange of knowledge.
+In this paper, we present the concept of a hack week in the larger context of scientific meetings.
+We motivate the need for such an event and present its strengths and challenges.
+We find that hack weeks are successful at cultivating collaboration and facilitating the exchange of knowledge.
 Participants self-report that these events help them both in their day-to-day research as well as their careers.

### 2018-06-04 intro.tex
Edits in intro and what sections.

@@
 In addition, the recent reproducibility crises in several scientific fields has motivated the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,baker2016}.
-Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts such as workshops \cite{demasi2017}.
+Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts in the form of less formal workshops \cite{demasi2017}.
 Well-known examples are the Software and Data Carpentry workshops providing training in research computing skills through a volunteer instructor program \cite{b:wilson-swc-lessons-2016,teal2015data}.
@@
 This model tends to focus on a one-way flow of information from instructor to student, and are usually targeted towards participants in the training phase of their career.
-On the other end of the spectrum, project-focused workshops emphasize collaborative activities using existing skills, leading to the common perception that they are designed for technical experts, and this may limit their audience.
+On the other end of the spectrum, project-focused workshops emphasize collaborative activities using existing skills, leading to the common perception that they are designed for technical experts.
+This may limit their audience.
 To bridge this gap, we describe here a model that we have implemented: ``Hack Weeks'' that aim to capitalize on the advantages of each of these models.

### 2018-06-04 conclusions.tex

@@
 
-The fast-paced changes of the computational and methodological landscape require traditional fields of science to rapidly adapt to new data analysis challenges.
-To address these challenges, new types of workshops, including unconferences, hackathons and bootcamps, have been developed in recent years in various scientific disciplines to exist alongside with and support the existing structure of academic conferences.
+The fast-paced changes in the computational and methodological landscape require that traditional fields of science rapidly adapt to new data analysis challenges.
+To address these challenges, new types of workshops, including unconferences, hackathons and bootcamps have been developed in recent years in various scientific disciplines and now exist alongside with, and support the existing structure of academic conferences, formal classes, and other learning opportunities.
 Here, we introduce one such concept, hack weeks, and detail the underlying philosophical ideas along with experiences from events held in three different fields.
@@
 
-Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years to asses their effect on collaboration networks, career outcomes and adoption of new methods.
-We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts can.
+Hack weeks are still a very young concept, and estimating the long-term impact of these events within the scientific communities they serve will require follow-up over multiple years, to assess their effect on collaboration networks, career outcomes and adoption of new methods.
+We have shown, however, that hack weeks provide an easy-to-implement, fairly low-cost method to introduce new technologies and methods into scientific fields on much shorter time scales than traditional teaching efforts.
 While we focus here on hack weeks in scientific fields, the concept could be extended to other areas, and is more generally useful in any area (1) where useful tools can be learned in short tutorials, (2) where results and outcomes can be produced on the timescale of a few days, and (3) that would benefit from collaborative approaches that cross traditional boundaries.
-Such areas could include for example the social sciences, music and art.
+Such areas could include the social sciences, the humanities, as well as music and art.

### 2018-06-04 participants.tex

@@
 
-Traditional selection processes that rely heavily on internal heuristics of reviewers, especially those that consider characteristics peripheral to the evaluation criteria, are often fraught personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
+Traditional selection processes that rely heavily on internal heuristics of reviewers, especially those that consider characteristics peripheral to the evaluation criteria, are often fraught with personal and structural biases~\cite[e.g.][]{sunstein2015wiser}.
 In order to maximize diversity and minimize bias, we advocate for a selection process that is as quantitative and transparent as possible~\cite{sunstein2015wiser}, enabling participants to hold organizers accountable for their selection decisions.
-This requires laying out a definition of successful participation, what criteria must be met in order to maximize the likelihood of success, and how those criteria will be assessed given the information about the candidates selected during the application stage.
+This requires laying out a definition of successful participation, defining what criteria must be met in order to maximize the likelihood of success, and defining how those criteria will be assessed given the information about the candidates selected during the application stage.
 
@@
 
-If merit-based selection is part of the evaluation process, organizers face the crucial decision whether to assess merit blindly or not.
-Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases related to internal heuristics.
+If merit-based selection is part of the evaluation process, organizers face the crucial decision whether to assess merit blinded to other applicant characteristics.
+Because human decision makers tend to be swayed by unrelated characteristics including name~\cite{bertrand2004} or gender~\cite{mossracusin2012}, an initial merit selection blinded to demographic characteristics can be an effective way to counteract certain biases.
 A merit selection could then be performed via scores given independently by members of the organizing committee based on a set of pre-defined, explicit selection criteria.
@@
 
-However, a blind selection purely on a figure of merit will be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills: for example, requiring a minimum level of programming experience will likely disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
-Additionally, blinding has been found to have negative effects on diversity for committees that have already had a strong commitment to diversity, because they often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
+However, a blind selection based purely on an assessment of merit will be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills: for example, requiring a minimum level of programming experience will likely disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
+Additionally, blinding has been found to have negative effects on diversity for committees that  already have a strong commitment to diversity, because these circumstances allow to correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
 In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also SM, Section 4.1.5).
@@
 No matter the selection procedures used, we encourage organizers to critically examine their cohort selection, experiment with new approaches, and routinely evaluate their procedure.
-For example, comparing demographic characteristics of the selected versus non-selected groups can unveil unintended biases during the merit-selection base and thus allow adjustments in the procedure to mitigate or fully remove these effects.
+For example, comparing demographic characteristics of the selected versus non-selected groups can unveil unintended biases during the merit-selection phase and thus allows adjustments in the procedure to mitigate or fully remove these effects.

### 2018-06-04 results.tex

@@
 One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded.
-Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building would indicate that improvements in workshop facilitation and structure may be required to allow participants of all stripes to participate fully.
+Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow a wide range of participants to participate fully.
 Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
 Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$).
-In neither case seem the attitudes toward building valuable connections at all three hack weeks or the value of their contributions to their hack teams differ with respect to gender or racial/ethnic identity.
+Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ with respect to gender or racial/ethnic identity.
 For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
@@
 %Overall, our results indicate that hack weeks are effective at addressing persisting doubts about making research open and reproducible.
-While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks like open-source software and data sharing.
+While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, such as production of open-source software and data sharing.
 In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.
@@
 We caution the reader that our survey results should be read only as an initial indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the surveys' exploratory nature.
-Our number of respondents is small and the effects likely subtle, thus the lack of significant differences may be caused by the small statistical power of our sample.
+The number of respondents is small and the effects likely subtle.
+This means that the lack of significant differences may be caused by the small statistical power of our sample.
 The most important independent variable--attendance at the hack week--is something that our current survey design cannot account for.
@@
 \subsubsection*{Example 3: Neuro Hack Week}
-Motion of study participants inside of the MRI machine is a major concern in neuroimaging studies.
-During NHW 2016 one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
-To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off were varied.
+Motion of study participants inside the MRI machine is a major concern in neuroimaging studies.
+During NHW 2016, one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
+To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off, were varied.
 %They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four researchers from four institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.
+The team (composed of four researchers from four different institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.

### 2018-06-04 supplementary_material.tex

--- 2018-05-28 supplementary_material.tex
@@
 In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered.
-For the AHW surveys, we checked whether the demographics of the respondents different significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and find no significant difference for any of the groups except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees).
+For the AHW surveys, we checked whether the demographics of the respondents differed significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and found no significant difference for any of the groups except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees).
 Because we group responses in the career stage category into categories ``early-career'' (students and postdoctoral fellows), ``late-career'' (faculty, research scientists) and ``other'' (e.g.
 participants from outside academia), this discrepancy has no bearing on the results we report.
-For gender, we grouped participants into ``minority'' (women and non-binary) and ``non-minority'' (men) participants; similarly, for race and ethnicity, we also chose ``non-white'' (all but Caucasian participants) and ``white'' (Cauciasian) categories.
-Note that demographic data for GHW and NHW is only available for 2016, for AHW for both 2016 and 2017, and responding to any question was optional.
+For gender, we grouped participants into ``minority'' (women and non-binary) and ``non-minority'' (men) participants; similarly, for race and ethnicity, we also chose ``non-white'' (all but Caucasian participants) and ``white'' (Caucasian) categories.
+Note that demographic data for GHW and NHW is only available for 2016, for AHW for both 2016 and 2017, and that responding to any of the questions was optional.
 
@@
 No responses were discarded, and no pre-processing was performed on the data.
-We test for correlations between demographic characteristics (independent variable) and question responses (dependent variable) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
+We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
 For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
@@
 Here we present the full results of the statistical analysis performed for the survey (Table \ref{tab:survey}).
-We searched for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity.
+We tested for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity.
 The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, and (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$).
@@
 
-For no combination of question with a demographic variable, we find a $p$-value below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
-In the main text, we put these non-detections into context, quote those tests that near significance, coupled with a moderate to large effect size, as an indication where future research may be fruitful, and present the study's limitations.
+We found no $p$-values below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials} for any of the combinations of questions and demographic variables.
+In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication where future research may be fruitful.
+The results also discuss the present study's limitations.
 
@@
 \label{sec:suppfigures}
-Here, we present a number of additional figures related to the survey data presented in the main manuscript.
-The stacked bar charts presented here allow a closer look at the data that produced the results quoted in Table 1 above in the SM.
+Here, we present a number of additional figures related to the survey data presented in the main text.
+The stacked bar charts presented here provide anothe view of the data that produced the results quoted in Table 1 above.
 
@@
 We grouped ``gender identity'' into two categories: ``non-minority'' contains all participants who self-identified as male, ``minority'' contains all participants who responded with ``female'' or any other response.
-Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics and geoscience), where they tend to be strongly underrepresented.
+Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics and geoscience), where they tend to be under-represented.
 
@@
 To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities.
-These scientists then reviewed their observations and gave valuable feedback about the hack week outcomes.
+These researchers then reviewed their observations and gave valuable feedback about the hack week outcomes.
 In order to be able to track and study participants, approval from an Internal Review Board is required.
@@
 Is the community to be invited more software-focused or should outcomes skew towards publishable results?
-Does the hack week aim to maximize its output by carrying best practices into universities and institutions that may be particularly underrepresented?
+Does the hack week aim to maximize its output by carrying best practices into universities and institutions that may be particularly under-represented?
 These questions, among others, should have a very clear answer before any further organization takes place.
@@
 \subsubsection{Hack Week Location}
-When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should receive equal consideration.
+When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should be taken into consideration.
 The non-traditional nature of hack weeks means that some students report challenges in receiving support from their supervisor or department to attend.
@@
 Therefore, the search for a location should commence no later than twelve months before the workshop, ideally much earlier.
-The space itself should be large and ideally configurable, with moveable tables and chairs, several screens or projectors, and an ample number of white-/chalkboards.
+The space itself should be large and ideally configurable, with moveable tables and chairs, several screens or projectors, and an ample number of white- or chalk-boards.
 A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.
@@
 On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas.
-On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical for hack week projects.
-It may hence be advisable to allow for diffusion of the group into adjacent rooms, while providing ample venue for the entire group to congregate for tutorials, breaks and reports.
+On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical of hack week projects.
+It may be advisable to allow for diffusion of the group into adjacent rooms, while providing ample room for the entire group to congregate for tutorials, breaks and reports out to the entire group.
 There should be plenty of power outlets for participants to plug in their computers.
-The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access to visitors from other institutions.
+The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access for visitors from other institutions.
 Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available?
@@
 Having a set of participants that is too small could interfere both with these goals of representation and of fostering interdisciplinary collaborations and networking.
-Smaller groups (of less than $\sim$30 participants or so) will need to attend to participant selection particularly carefully in order to achieve the desired level of diversity across participants.
+Organizers of evens for smaller groups (of less than $\sim$30 participants or so) will need to attend to participant selection particularly carefully, in order to achieve the desired level of diversity across participants.
 
-All hack weeks organized by us up to now have been five days.
+All hack weeks organized by us until now have been five days long.
 This duration was chosen to allow enough time for both learning and project work at the same time.
 Participants often arrived with ambitious project ideas, or with the goal of taking newly learned methods and applying them to their data sets right away.
-Thus many projects tended to extend over multiple days.
+This meant that many projects tended to extend over multiple days.
 Attendees who participated in multiple projects often used the hack week to try out several different ideas.
-A five-day hack week allows for ample time for these ambitious projects or experimentation, where a shorter hack week would necessarily have to be more constrained and focused.
+A five-day hack week allows ample time for these ambitious projects or for experimentation, where a shorter hack week would necessarily have to be more constrained and focused.
 At the same time, hack weeks tend to be exhausting.
@@
 Because selection procedures are subject to a whole range of human biases, it is worth thinking critically about established procedures and how to improve them.
-At the same time, hack weeks are comparatively lower-stakes events than other situations where cohort selection is relevant, like e.g.\ graduate school admissions, thus hack weeks may also serve as test beds for experimentation and improvements that can inform selection in other contexts.
+At the same time, hack weeks are comparatively lower-stakes events than other situations where cohort selection is relevant, such as graduate school admissions.
+Thus, hack weeks may also serve as test beds for experimentation and improvements that can inform selection in other contexts.
 
 The hack weeks presented here used different approaches, that nevertheless share several common themes and ideas.
-In particular, they focus on making the selection procedure as quantitative as possible and use the data collected during each stage of the process to critically evaluate the efficacy of the process and its subject to human bias.
-Secondly, much energy is expended early on in the design of the application form, which often use a combination of open-answer and multiple-choice questions to assess candidates' suitability, skill levels and backgrounds.
+In particular, they all focused on making the selection procedure as quantitative as possible and use the data collected during each stage of the process to critically evaluate the efficacy of the process and its subject to human bias.
+Secondly, much energy was devoted early on in the design of the application form, which often use a combination of open-answer and multiple-choice questions to assess candidates' suitability, skill levels and backgrounds.
 Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' or ``expert'' have no absolute scaling and may have different meanings to different applicants.
-They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates (the Dunning Kruger effect).
-It is thus advisable to calibrate skill levels on concrete milestones.
+They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates (the so-called Dunning Kruger effect).
+It is therefore advisable to calibrate skill levels on concrete milestones.
 For example, instead of asking participants whether they are ``beginners'' or ``experts'' at machine learning, we instead asked them whether they had done machine learning for class work, had used techniques from machine learning in their own research, or had developed and implemented their own methods.
@@
 For example, Astro Hack Week only asked whether an applicant considered themselves a minority with respect to gender identity within astronomy, rather than asking them to explicitly self-identify within a specific category.
-Meanwhile, neuro hack week opted to ask about gender ethnicity and race, using questions that are standardly used for example, when collecting statistics about study participants.
+Meanwhile, neuro hack week opted to ask about gender, ethnicity and race using the questions that are standardly used when reporting statistics about study participants.
 
@@
 It is always a good idea to check the results of any selection procedure for intrinsic biases.
-For example, visual comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can be adjusted and refined.
+For example, comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can be adjusted and refined.
 
@@
 The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
-It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest.  
-Any human biases involved in setting these target fractions will be perpetuated in the selection procedure. 
+It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest and any human biases involved in setting these target fractions will be perpetuated in the selection procedure.
 Secondly, perhaps more obviously, the algorithm can only act on information that has been collected.
@@
 Targets within each categories were set according to the different workshop goals.
-For example, we set ambitious goals of admitting 50\% attendees from underrepresented minorities with respect to gender, race and ethnicity with the goal of increasing representation of these groups with astronomy and data science.
+For example, we set ambitious goals of admitting 50\% attendees from under-represented minorities with respect to gender, race and ethnicity with the goal of increasing representation of these groups with astronomy and data science.
 Similarly, we set targets for other categories (e.g.
@@
 
-Neuro hack week took a different approach: participants provided CV as part of their application, in addition to answering questions about their technical and research background, and writing a short statement of the purpose of their participation.
+Neuro hack week took a different approach: participants provided a CV as part of their application, in addition to answering questions about their technical and research background, and writing a short statement of the purpose of their participation.
 Each application was reviewed by several instructors using 1-10 scales on several criteria: technical and computing experience, research experience, the quality of the statement, an assessment of the contribution of participation to the applicant's career, and an assessment of the degree to which participation would increase the representation of groups that are usually under-represented in neuroscience and computing.
@@
 An overarching goal of every hack week is to provide a welcoming, comfortable, and inclusive environment for participants to learn and work.
-However, disparities in backgrounds, knowledge and experience, together with the close collaboration much of the hacking requires, creates the potential for conduct that might offend participants, and preclude them from participating fully.
+However, disparities in backgrounds, knowledge and experience, together with the close collaboration that hacking requires, creates the potential for conduct that might offend participants, and preclude them from participating fully.
 For this reason, a hack week should have a clear and enforceable code of conduct, that can guide the behavior of participants, and help organizers maintain a professional and welcoming environment.
@@
 For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome, defensive and inhibit learning on a larger scale.
-Even if no code of conduct infractions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are thoughtful of creating a welcoming environment, which in turn feeds back to the creation of such a space.
+Even if no code of conduct infractions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are intentional in creating a welcoming environment, which in turn feeds back to the creation of such a space.
 
@@
 Tutorials may be live-streamed via popular services such as youtube, but aside from the technical complexity of setting up a reliable live stream, their interactive format may not easily lend themselves to the one-directional communication of a video feed.
-Even so, live streams have proven to be both useful and popular in the past, but require considerable effort to set up.
+Even so, live streams and video recording posted after the fact have proven to be both useful and popular in the past, even while they require considerable effort to set up.
 At the very least, there should be some method of real-time communication with the in-group.
@@
 We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \citep{kluyver2016jupyter} or following the markdown templates provided by Software Carpentry.
-This way the hack week ``curriculum'' can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
+This allows the hack week ``curriculum'' to be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
 Depending on the content, some instructors convey concepts during the tutorial via ``live-coding'' while others display static code and their results.
-Either way, we find dividing code into short blocks, each with a specific task, helps participants from becoming saturated with information.
+Either way, we find dividing code into short blocks, each with a specific task, helps participants not to become overwhelmed with information.
 In some cases, tutorials require substantial software installation and data.
@@
 We suggest that the primary goals of the tutorials should be to provide an entry point into an exploration of participants' datasets, opening the door to more thorough study outside of the hack week.
-Given the broad span of available tools and topics, the organizing committee should seek to teach content that represents the state-of-the-art and is deemed to be of greatest use to the broadest span of participants.
+Given the broad span of available tools and topics, the organizing committee should seek to teach content that represents the state-of-the-art and is deemed to be of greatest use to the broadest range of participants.
 A common theme across all of our hack weeks is to teach initial tutorials in version control, command line interfaces, data science platforms and practices of reproducible research.
 These tutorials endeavor to provide a common baseline of techniques that participants can build on in other tutorials and project work.
-Subsequent tutorials are then delivered in more domain-specific fields, but can be arranged in a way that builds knowledge constructively through the week.
+Subsequent tutorials are then delivered in more domain-specific fields, but can be arranged in a way that builds knowledge constructively over the course of the week.
 In some cases, it may be advisable to recommend some theoretical texts as reading before attending the hack week, though it is likely that speakers should not count on every participant arriving with the same knowledge baseline.
@@
 
-It is noteworthy that in practice, the mixed audience and interactive setting can often lead to impostor syndrome (see below) among the speakers, who lose their explicit status as expert among the participants.
-It is therefore imperative that speakers are made aware very early about the audience so that they can plan their tutorial accordingly.
-It also helps to make them aware that participants may be taking the role of teaching assistants during exercises and include them in the teaching, rather than see them as adversaries who may seek to correct them.
+It is worth noting that in practice, the mixed audience and interactive setting can often lead to impostor syndrome (see below) among the speakers, who lose their explicit status as expert among the participants.
+It is therefore imperative that speakers are made aware as early as possible about the audience, so that they can plan their tutorial accordingly.
+It also helps to make them aware that participants may be taking the role of teaching assistants during exercises, and include them in the teaching, rather than see them as adversaries who may seek to correct them.
 Conversely, experts may ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and to encourage questions, especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome.
-At Astro Hack Week, we encourage the organizing committee and experienced participants to ask questions, particularly when they know the answer but think the concept may not yet be clear to parts of the audience.
+At Astro Hack Week, we encourage the organizing committee and experienced participants to ask questions, particularly when they know the answer, but think the concept may not yet be clear to parts of the audience.
 This has in the past led to exchanges between experts that have greatly contributed to the tutorials in ways that were both unplanned and unexpected for all involved, beginners and experts alike.
@@
 In the initial stage, participants pitch projects to the group and in some cases request specific expertise (of a certain method, programming language or other skill).
-Ideally, the organizing committee will encourage participants to post potential projects as early as possible online in a central document, which will allow some organization even before the begin of the hack week.
+Ideally, the organizing committee will encourage participants to post potential projects as early as possible online in a central document, which will allow some organization even before the beginning of the hack week.
 If the hack week admits very junior participants (e.g.
-undergraduate students or junior graduate students) it may be advisable for the organizing committee to contact experienced participants in advance and ask them to suggest hacks that are appropriate for those junior participants.
+undergraduate students or junior graduate students) it may be advisable for the organizing committee to contact experienced participants in advance and ask them to suggest hacks that are appropriate for these junior participants.
 At the hack week itself, the time designated for project work will generally be prefaced by a pitch session, where participants with projects give a short description of their project to the group and request help.
 For junior participants and those new to the format, this can be daunting, even if (or especially when) they do not have a project themselves.
-In particular junior participants tend to believe that they do not yet have the skills to contribute meaningfully to any projects.
-Here, it can be useful to require those participants who are not pitching projects to state their expertise and affirm that every participant has expertise in something.
+Junior participants in particular tend to believe that they do not yet have the skills to contribute meaningfully to any projects.
+Here, it can be useful to require the participants who are not pitching projects to state their expertise and affirm that every participant has expertise in something.
 This broad definition of hacking can help the organizers to facilitate inclusion by emphasizing the value in skills such as writing (e.g.
@@
 At the end of the week we provide ample time for participants to present their projects to the community.
-We endeavor to design these presentations in a way that builds confidence, especially for junior researchers, and that welcomes all projects regardless of their level of sophistication.
-The open-ended format of a hack week encourages experimental approaches and new directions, which necessarily will not all be successful.
-Too often, showcases focus on the most successful and impressive projects, which distorts reality and produces both an unreasonable expectation of what the result of a hack week ought to be as well as a pressure to be successful that can in practice inhibit participants.
+We endeavor to design these presentations in a way that builds confidence, especially for junior researchers, and that welcomes all projects regardless of their level of complexity and/or sophistication.
+The open-ended format of a hack week encourages experimental approaches and new directions.
+This means that some projects will not reach a clear or successful conclusions.
+Too often, showcases focus on the most successful and impressive projects.
+This tends to distort the reality of the experience of hack week participants and produces both an unreasonable expectation of what the result of a hack week ought to be, as well as a pressure to be successful that can in practice inhibit participants.
 Therefore, it should be encouraged that participants show failed ideas, not least because the knowledge of why a project failed can be a valuable learning experience beyond the team working on that particular project.
@@
 At all hack weeks we have organized, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
-Clearly communicating values and expectations at the beginning for the week can help setting the stage for an environment in which a large range of different personalities can thrive.
+Clearly communicating values and expectations at the beginning for the week can help set the stage for an environment in which a large range of different personalities can thrive.
 This includes both communication of the code of conduct, putting an emphasis on collaborative rather than competitive behaviour, as well as a discussion of potential issues participants might experience, such as impostor syndrome (see also section below).
@@
 We also designate at least two members of the organizing committee whose role it is to facilitate both tutorials and hack sessions.
-It is their task to for example manage dominant personalities, which can be both a boon and a hindrance to the success of a hack week.
+It is their task to also manage dominant personalities, which can be both a boon and a hindrance to the success of a hack week.
 Often, these attendees can be effective at generating excitement and engagement, but can also intimidate and dominate discussions.
 It is up to the organizers to balance the participation of these attendees with the needs of the rest of the group.
-Similarly, facilitators at Astro Hack Week spend the majority of hacking time circulating the room, connecting participants without a group to a project, help connect groups who run into a problem with experts who can help out, and make sure group work proceeds in such a way that all members feel valued.
+Similarly, facilitators at the hack weeks we have organized spend the majority of hacking time circulating the room, connecting participants without a group to a project, help connect groups who run into a problem with experts who can help out, and make sure group work proceeds in such a way that all members feel valued.
 
-At geohack week we employed ``Liberating Structures'', a facilitation approach in which we introduce small shifts in the way groups engage in order to increase relational coordination and trust.
+At geo hack week we employed ``Liberating Structures'', a facilitation approach in which we introduce small shifts in the way groups engage in order to increase relational coordination and trust.
 We invited a professional facilitator to assist us in leading a series of conversations and activities aimed at fostering group engagement.
 In one activity, participants faced one another and were given a brief statement inviting a two minute conversation, after which the rows rotated to maximize the mixing of ideas.
-Invitations included statements such as ``I belong at geohack week because...'' or ``A challenge I hope we can make progress on is...''.
+Invitations included statements such as ``I belong at geo hack week because...'' or ``A challenge I hope we can make progress on is...''.
 Importantly, event organizers joined in the activity, which we found helped participants feel as if they were not alone in facing some of the challenges inherent in data science studies.
@@
 Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS, namely that participants feel the need to be experts in multiple scientific and technical aspects of the activities pursued during a hack week.
-This particular form of IS hinges to some degree on the design focus on diversity of backgrounds (everyone else seems to know something that you do not!) and might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, and not least produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
+This particular form of IS hinges to some degree on the design focus on diversity of backgrounds in that all the other participants seem to know something that any given participant does not.
+This might be further exacerbated by the expectation that attendants expose their ideas to public scrutiny, find collaborators in a very short amount of time, produce and present a successful hack at the end of the week\footnote{see also this insightful blog post: \url{https://medium.com/astronomy-without-stars/the-horror-of-hack-days-52c6b52cfc3b}}.
 The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
 This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
-Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, to be pro-active about forming new collaborations.
- Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered.
-We are working within our hack weeks to mitigate IS using various techniques.
+Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, or to be proactive about forming new collaborations.
+ Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered by IS.
+ Thus, we are working within our hack weeks to mitigate IS using various techniques.
 With respect to minority participants, ensuring adequate representation can decrease feelings of otherness and may help reduce IS.
@@
 Additionally, computational fields tend to be fast-paced; a tutorial given one year might be outdated a few years later.
-It is thus useful to put some detailed thoughts into the survey.
-Following from discussions the committee has likely had during the initial stage of preparing the announcement and selecting participants, organizers will want to confront their own expectations before the workshop, and design a survey that allows for both measurements of specific outcomes and goals the organizers are interested in, as well as leave enough free space for participants to self-report outcomes that arose spontaneously or were out of scope of the initial design.
-It is helpful to design the survey early on and allow for some time near the end of the workshop for participants to fill it out there; in our experience, this leads to much higher response rates than if the survey is distributed after the end of the workshop.
-While some improvements and changes in the survey design from year to year are inevitable, it is advisable to set a number of core questions that do \textit{not} change from year to year.
+Therefore, it is useful to put some detailed thoughts into the survey.
+Following discussions the organizing committee will have likely had during the initial stage of preparing the announcement and selecting participants, organizers will want to confront their own expectations before the workshop, and design a survey that allows for both measurements of specific outcomes and goals the organizers are interested in, as well as leave enough free space for participants to self-report outcomes that arose spontaneously or were out of the scope of the initial design.
+It is helpful to design the survey early on and allow for some time near the end of the workshop for participants to fill it out while they are still physically present; in our experience, this leads to much higher response rates than if the survey is distributed after the end of the workshop.
+While some improvements and changes in survey design from year to year are inevitable, it is advisable to set a number of core questions that do \textit{not} change from year to year.
 These questions will allow organizers to assess the workshop's function within their community, and assess how changes to the format between one year and another may affect outcomes of interest.

### 2018-06-04 supplementary_material.tex

@@
 For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
+
+We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
+For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size.
+We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
+For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $\[0,1\]$, we only require an upper bound} this equivalence bound.
+If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.
+
 The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.

### 2018-06-04 supplementary_material.tex
Reformatted table in prep for additional data

@@
 \usepackage{tabularx}
+%\usepackage{pdflscape}
 \linespread{1.2}
@@
 We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
-For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $\[0,1\]$, we only require an upper bound} this equivalence bound.
+For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
 If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.
@@
 The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
-
-\clearpage
 
@@
 
-\begin{table*}
+%\begin{landscape}
+\begin{sidewaystable}
 %\begin{table*}[hbtp]
@@
 \begin{threeparttable} 
-\begin{tabularx}{18cm}{p{3.5cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
+\begin{tabularx}{18cm}{p{3.5cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
 \toprule
 \textbf{Question} & \multicolumn{7}{c}{\textbf{Astro Hack Week}} & \multicolumn{7}{c}{\textbf{GeoHackWeek}} & \multicolumn{7}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$  \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
@@
 %\end{table*}
-\end{table*}
+\end{sidewaystable}
+%\end{landscape}
 

### 2018-06-06 supplementary_material.tex
Did some extensive copyediting

--- 2018-06-04 supplementary_material.tex
@@
 \label{sec:suppmaterials}
-We performed post-attendance surveys for AHW, GHW and NHW in 2016 and 2017.
-All surveys contained general questions about attitudes towards the workshop as well as open science and reproducibility, shared among all three surveys.
+We performed post-attendance surveys for AHW, GHW, and NHW in 2016 and 2017.
+All surveys contained general questions about attitudes towards the workshop, as well as open science and reproducibility, shared among all three surveys.
 The NHW and GHW surveys and the AHW 2017 survey were administered on site on the last day of the workshop; AHW participants in 2016 were e-mailed immediately after the workshop, with reminders after several weeks.
-All responses were collected within four weeks of the end of the workshop.
-The experimental procedures were approved by the Institutional Review Boards at UW, NYU and UC Berkeley.
-All participants gave their informed consent.
+All responses were collected within four weeks of the end of their respective workshops.
+The experimental procedures were approved by the Institutional Review Boards at UW, NYU, and UC Berkeley.
+All participants gave informed consent.
 
-Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100\% in both years; the response rate for AHW was 71\% (35 out of 49) in 2016 and 82\% (37 out of 45) in 2017.
-The lower response rates for AHW can be explained by the generally lower rates of response for time-delayed survey in 2016.
+Response rates for NHW (2016: 41 responses; 2017: 45 responses) and GHW (2016: 42 responses; 2017: 41 responses) were 100 percent in both years; the response rate for AHW was 71 percent (35 out of 49) in 2016 and 82 percent (37 out of 45) in 2017.
+The lower response rates for AHW can be explained by the generally lower rates of response for the time-delayed survey in 2016.
 In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered.
-For the AHW surveys, we checked whether the demographics of the respondents differed significantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background and gender identity) and found no significant difference for any of the groups except for career stage in 2017: here, we find that graduate students are underrepresented (27\% of respondents compared to 41\% of attendees), while undergraduate students are overrepresented (25\% of respondents compared to 12\% of attendees).
-Because we group responses in the career stage category into categories ``early-career'' (students and postdoctoral fellows), ``late-career'' (faculty, research scientists) and ``other'' (e.g.
+For the AHW surveys, we checked whether the demographics of the respondents diﬀered signiﬁcantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background, and gender identity) and found no signiﬁcant diﬀerence for any of the groups except for career stage in 2017: Here, we ﬁnd that graduate students are underrepresented (27 percent of respondents compared to 41 percent of attendees), while undergraduate students are overrepresented (25 percent of respondents compared to 12 percent of attendees).
+Because we group responses in the career stage category into the categories “early-career” (students and postdoctoral fellows), “late-career” (faculty, research scientists), and “other” (e.g.
 participants from outside academia), this discrepancy has no bearing on the results we report.
-For gender, we grouped participants into ``minority'' (women and non-binary) and ``non-minority'' (men) participants; similarly, for race and ethnicity, we also chose ``non-white'' (all but Caucasian participants) and ``white'' (Caucasian) categories.
-Note that demographic data for GHW and NHW is only available for 2016, for AHW for both 2016 and 2017, and that responding to any of the questions was optional.
+For gender, we grouped participants into “minority” (women and non-binary) and “non-minority” (men) participants; similarly, for race and ethnicity, we divided participants into the categories “non-white” (all but Caucasian participants) and “white” (Caucasian).
+Note that demographic data for GHW and NHW is only available for 2016, while for AHW, data is available for both 2016 and 2017.
+Responding to any question in the survey was optional.
 
@@
 No responses were discarded, and no pre-processing was performed on the data.
-We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
+We tested for correlations between demographic characteristics (independent variables), questioned responses (dependent variables) using a standard $\chi^2$ test and computed the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
 For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
@@
 
-We found no $p$-values below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials} for any of the combinations of questions and demographic variables.
-In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication where future research may be fruitful.
+For no combination of question questions with a demographic variable, do we ﬁnd a $p$-value below the trial-corrected signiﬁcance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
+In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication of where future research may be fruitful.
 The results also discuss the present study's limitations.
@@
 Here we explore a number of survey responses with gender identity as an independent variable.
-We grouped ``gender identity'' into two categories: ``non-minority'' contains all participants who self-identified as male, ``minority'' contains all participants who responded with ``female'' or any other response.
-Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics and geoscience), where they tend to be under-represented.
+We grouped ``gender identity'' into two categories: ``Non-minority'' contains all participants who self-identified as male; ``minority'' contains all participants who responded with ``female'' or any other response.
+Women are here considered a minority with respect to their chosen field of study (neuroscience, astrophysics, or geoscience), where they tend to be under-represented.
 
@@
 Here we explore a number of survey responses with racial/ethnic identity as an independent variable.
-We grouped responses into two final categories, ``white'' (Caucasian participants) ``non-white'' (all others).
+We grouped responses into two final categories: ``white'' (Caucasian participants) and ``non-white'' (all others).
 
@@
 \subsubsection{Obtaining permission for tracking and studying participants}
-A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and quantify the achievement of learning outcomes.
-To this end, we invited social scientists, interested in mechanisms of scientific collaboration and educational design process, to observe our activities.
+A secondary goal of most of our previous events was to study the development of a hack week, to observe participant behavior, and to quantify the achievement of learning outcomes.
+To this end, we invited social scientists, interested in mechanisms of scientific collaboration and the educational design process to observe our activities.
 These researchers then reviewed their observations and gave valuable feedback about the hack week outcomes.
-In order to be able to track and study participants, approval from an Internal Review Board is required.
+In order to track and study participants, approval from an Internal Review Board is required.
 This should be done as early as possible, and having assistance from researchers in the social sciences is invaluable, in particular for hack weeks in fields that generally do not involve human subject research.
-Ideally, participants should consent to be contacted at the application stage, which opens up the potential of tracking both participants and non-participants and allows for evaluation of the hack week.
+Ideally, participants should consent to be contacted during the application stage, which opens up the potential for tracking both participants and non-participants and allows for evaluation of the hack week.
 
@@
 
-Unlike traditional conferences, where the programme is primarily driven by speaker selection, careful goal planning is essential for a successful hack week.
+Unlike traditional conferences, where the program is primarily driven by speaker selection, careful goal planning is essential for a successful hack week.
 The stated goals will continue to determine almost all of the remaining organizational questions, including space, participant selection and workshop evaluation.
-Is the target audience primarily students, or mature researchers, or a mix of both?
-Should the taught component dominate over project work or should it be only a minor programme point?
+Is the target audience made up primarily of students, or mature researchers, or a mix of both?
+Should the taught component dominate over project work or should it be only a minor program point?
 Is the community to be invited more software-focused or should outcomes skew towards publishable results?
@@
 
-\subsubsection{Hack Week Location}
-When considering space, both the broader location (city and country) as well as the specifics (the room in which the workshop will occur) should be taken into consideration.
-The non-traditional nature of hack weeks means that some students report challenges in receiving support from their supervisor or department to attend.
-Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be a high priority when deciding on a location.
+\subsubsection{Hack week Location}
+When considering space, the broader location (city and country) and the specifics (the room in which the workshop will occur) should be taken into consideration.
+The non-traditional nature of hack weeks means that some students report challenges in receiving support from their supervisors or departments to attend.
+Thus, keeping the costs of attendance low (including travel costs, room and board, and conference fees) should be a high priority when choosing a location.
 Conference centers like the Lorentz Center at the University of Leiden provide funding along with the collaborative space, but have long lead times when applying for their spaces (typically 12-18 months).
 The same can be true for collaborative spaces within universities and active learning environments, which tend to be popular.
-Therefore, the search for a location should commence no later than twelve months before the workshop, ideally much earlier.
-The space itself should be large and ideally configurable, with moveable tables and chairs, several screens or projectors, and an ample number of white- or chalk-boards.
-A useful guiding principle is to book a space that fits a group at least 25\% larger than that envisioned for the workshop, and the less configurable the space, the larger the extra room should be.
+Therefore, the search for a location should commence no later than twelve months before the workshop and ideally much earlier.
+
+The space itself should be large and ideally configurable, with movable tables and chairs, several screens or projectors, and an ample number of white boards or chalkboards.
+Two useful guiding principles are to book a space that fits a group at least 25 percent larger than that envisioned for the workshop, and that the less configurable the space, the larger the extra room should be.
 There is a natural tension between keeping the group together and providing physically separated break-out spaces.
 On the one hand, conducting the workshop in a single large venue improves group cohesion and prevents self-segregation of researchers and ideas.
-On the other hand, the interactive nature of the workshop may lead to an environment where it is difficult for individuals to focus on the highly complex tasks that are typical of hack week projects.
-It may be advisable to allow for diffusion of the group into adjacent rooms, while providing ample room for the entire group to congregate for tutorials, breaks and reports out to the entire group.
+On the other hand, the interactive nature of the workshop may lead to an environment in which it is difficult for individuals to focus on the highly complex tasks that are typical of hack week projects.
+Hence, it may be advisable to allow for diffusion of the group into adjacent rooms, while providing ample space in which the entire group can congregate for tutorials, breaks and reports out to the entire group.
 There should be plenty of power outlets for participants to plug in their computers.
-The space should have ready access to wifi, for activities that require an internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access for visitors from other institutions.
-Accessibility should be taken into consideration: is the space accessible for participants who use wheel-chairs? Are accessible restrooms available?
+The space should have ready access to Wi-Fi, for activities that require an Internet connection (e.g., collaborative software development), and organizers should arrange in advance to provide access for visitors from other institutions.
+Accessibility should be taken into consideration: Is the space accessible for participants who use wheel-chairs? Are accessible restrooms available?
 
-\subsubsection{Event Size and Duration}
+\subsubsection{Event size and duration}
 
@@
 the number of attendees) and event duration are important to consider during the design and planning phase of a hack week.
-All hack weeks we have organized so far have had group sizes between 40-60 participants.
-The number of attendees resulted from a number of different design considerations.
-One practical consideration was the size of the available location: especially the project-focused parts of a hack week generally require a reconfigurable space with more seats than attendees, such that spontaneous organization in sub-groups is easy.
-However, group size is also important in the context of community building, group dynamics and programme decisions.
-We limited group size to a maximum of 70 participants, because previous anecdotal evidence has shown us that this is about the maximum group size where interactions between all participants is still possible.
+All hack weeks we have organized so far have had group sizes ranging from 40-60 participants.
+The number of attendees has depended on a number of different design considerations.
+One practical consideration was the size of the available location: The project-focused parts of a hack week, especially, generally require a reconfigurable space with more seats than attendees, such that spontaneous organization in sub-groups is easy.
+However, group size is also important in the context of community building, group dynamics and program decisions.
+We limited group size to a maximum of 70 participants, because previous anecdotal evidence has shown us that this is about the maximum group size in which interactions between all participants are still possible.
 We wanted to ensure that all participants had an opportunity to at least meet all other participants, in order to facilitate community building and keep the group from fracturing into cliques.
@@
 Similarly, in order to foster active collaboration among many different subfields, we endeavored to admit participants with many different types of overlapping expertise.
-Having a set of participants that is too small could interfere both with these goals of representation and of fostering interdisciplinary collaborations and networking.
+Having too small a set of participants could interfere both with the goal of representation and that of fostering interdisciplinary collaborations and networking.
 Organizers of evens for smaller groups (of less than $\sim$30 participants or so) will need to attend to participant selection particularly carefully, in order to achieve the desired level of diversity across participants.
 
-All hack weeks organized by us until now have been five days long.
-This duration was chosen to allow enough time for both learning and project work at the same time.
+All hack weeks organized by us until now have lasted five days long.
+This duration was chosen to allow enough time for both learning and project work.
 Participants often arrived with ambitious project ideas, or with the goal of taking newly learned methods and applying them to their data sets right away.
@@
 Attendees who participated in multiple projects often used the hack week to try out several different ideas.
-A five-day hack week allows ample time for these ambitious projects or for experimentation, where a shorter hack week would necessarily have to be more constrained and focused.
+A five-day hack week allows for ample time for these ambitious projects or for experimentation, where a shorter hack week would necessarily have to be more constrained and focused.
 At the same time, hack weeks tend to be exhausting.
@@
 
-As we lay out in the main text, participant selection is one of the most important design decisions as part of a hack week.
+As we lay out in the main text, participant selection is one of the most important design decisions made while planning a hack week.
 Conversations about participant selection, and planning of selection strategies should commence early, ideally at the start of the organizing process.
 Because selection procedures are subject to a whole range of human biases, it is worth thinking critically about established procedures and how to improve them.
-At the same time, hack weeks are comparatively lower-stakes events than other situations where cohort selection is relevant, such as graduate school admissions.
+At the same time, hack weeks are comparatively lower-stakes events than other situations in which cohort selection is relevant, such as graduate school admissions.
 Thus, hack weeks may also serve as test beds for experimentation and improvements that can inform selection in other contexts.
@@
 The hack weeks presented here used different approaches, that nevertheless share several common themes and ideas.
-In particular, they all focused on making the selection procedure as quantitative as possible and use the data collected during each stage of the process to critically evaluate the efficacy of the process and its subject to human bias.
-Secondly, much energy was devoted early on in the design of the application form, which often use a combination of open-answer and multiple-choice questions to assess candidates' suitability, skill levels and backgrounds.
-Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' or ``expert'' have no absolute scaling and may have different meanings to different applicants.
+In particular, they all focused on making the selection procedure as quantitative as possible and use the data collected during each stage of the process to critically evaluate the efficacy of the process and  the degree to which it is subject to human bias.
+Secondly, much energy was devoted early on in the design of the application form, which often uses a combination of open-answer and multiple-choice questions to assess candidates' suitability, skill levels and backgrounds.
+Care must be taken when designing questions assessing skill levels, because options such as ``beginner'' and ``expert'' have no absolute scaling and may have different meanings to different applicants.
 They may also be subject to intrinsic biases, if candidates of certain backgrounds systematically underestimate or overestimate their skill levels compared to other candidates (the so-called Dunning Kruger effect).
-It is therefore advisable to calibrate skill levels on concrete milestones.
-For example, instead of asking participants whether they are ``beginners'' or ``experts'' at machine learning, we instead asked them whether they had done machine learning for class work, had used techniques from machine learning in their own research, or had developed and implemented their own methods.
+It is therefore advisable to calibrate skill levels using concrete milestones.
+For example, instead of asking participants whether they are ``beginners'' or ``experts'' at machine learning, we instead asked whether they had done machine learning for class work, had used techniques from machine learning in their own research, or had developed and implemented their own methods.
 Similarly, questions about programming skills may ask whether participants can perform a particular task in one of a list of languages (e.g., ``write a module or a library'').
 Responding to questions about demographic characteristics should generally be optional, and organizers should consider how these questions might interact adversely with workshop goals, e.g.\ via stereotype threat.
-For example, Astro Hack Week only asked whether an applicant considered themselves a minority with respect to gender identity within astronomy, rather than asking them to explicitly self-identify within a specific category.
-Meanwhile, neuro hack week opted to ask about gender, ethnicity and race using the questions that are standardly used when reporting statistics about study participants.
+For example, Astro Hack Week only asked whether applicants considered themselves a minority with respect to gender identity within astronomy, rather than asking them to explicitly self-identify within a specific category.
+Meanwhile, Neuro Hack Week opted to ask about gender, ethnicity and race using the questions that are standard when reporting statistics about study participants.
 
@@
 Candidates from underrepresented groups, smaller institutions or other countries may have had fewer opportunities to learn relevant skills.
-Consequently, in a merit-based selection based on these skills these questions may effectively trace demographic diversity and lead to biased selections even when committees are overall vigilant.
+Consequently, in a merit-based selection based on these skills these questions may effectively trace demographic diversity and lead to biased selections even when committees are vigilant overall.
 If a strict merit-based selection is part of the procedure, it should be performed with these caveats in mind.
 It is always a good idea to check the results of any selection procedure for intrinsic biases.
-For example, comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can be adjusted and refined.
+For example, comparisons of the group selected based on merit versus the group not selected with respect to various demographic variables may easily reveal unintended biases introduced at this stage, and the procedure can then be adjusted and refined.
 
-Since human decision makers also routinely overestimate their ability to forecast future performance of a candidate and underestimate intrinsic variance during selection procedures~\citep{highhouse2008}, it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via e.g.\ via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+Since human decision makers also routinely overestimate their ability to forecast future performance of candidates and underestimate intrinsic variance during selection procedures~\citep{highhouse2008}, it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 
 One solution to the latter problem is implemented in the software \textit{entrofy}\footnote{\url{http://github.com/dhuppenkothen/entrofy}}).
-The algorithm aims to find a group of participants that together match as closely as possible a requested distribution on specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
-It is worth noting that this algorithm is vulnerable toward biases in two ways: firstly, humans will set the target fractions for any category of interest and any human biases involved in setting these target fractions will be perpetuated in the selection procedure.
-Secondly, perhaps more obviously, the algorithm can only act on information that has been collected.
+The algorithm aims to find a group of participants that together match as closely as possible a requested distribution of specified dimensions (e.g., career stage, geographic location, etc.), to meet pre-set fractions set by the organizers.
+It is worth noting that this algorithm is vulnerable toward biases in two ways: First, humans will set the target fractions for any category of interest and any human biases involved in setting these target fractions will be perpetuated in the selection procedure.
+Second, perhaps more obviously, the algorithm can only act on information that has been collected.
 Biased participant sets may still result from selection procedures that fail to include crucial categories.
 
-The overall selection procedure for Astro Hack Week comprised a two-step process.
+The overall selection procedure for Astro Hack Week was made up of a two-step process.
 In the initial step, we performed a blind assessment for suitability using open-ended questions querying candidates' motivations for attending the workshop, and their goals for attending.
-We only excluded a small fraction ($<10\%$) of applicants at this step, because our goals aligned with admitting a wide range of different skill levels and backgrounds.
+We only excluded a small fraction ($<10$ percent) of applicants at this step, because our goals aligned with admitting a wide range of different skill levels and backgrounds.
 Applicants were generally either excluded for being an obvious mismatch with the workshop goals (as in the case of an Android developer applying to work on an unrelated game) or for being duplicate entries.
 We subsequently used a range of categories queried with multiple-choice questions  on the application to maximize diversity across these categories, which included career stage, gender identity, racial/ethnic identity, geographic origin, programming background, machine learning skills, statistics knowledge, and previous attendance at Astro Hack Week.
-Targets within each categories were set according to the different workshop goals.
-For example, we set ambitious goals of admitting 50\% attendees from under-represented minorities with respect to gender, race and ethnicity with the goal of increasing representation of these groups with astronomy and data science.
+Targets within each category were set according to the different workshop goals.
+For example, we set ambitious goals of admitting 50 percent attendees from under-represented minorities with respect to gender, race and ethnicity with the goal of increasing representation of these groups within astronomy and data science.
 Similarly, we set targets for other categories (e.g.
-skill levels) in order to maximize goals around collaboration, networking and fostering interdisciplinary exchange of knowledge.
-Subsequently, we used a greedy optimization algorithm to jointly optimize over all categories simultaneously.
+skill levels) in order to maximize goals around collaboration, networking, and fostering an interdisciplinary exchange of knowledge.
+Subsequently, we used a greedy optimization algorithm to jointly optimize all categories simultaneously.
 We evaluated our results by directly comparing the characteristics of the selected set of participants with both our targets and the distribution of the set of all candidates.
-For demographic characteristics specifically, 42\% of participants in both 2016 and 2017 considered themselves a minority with respect to gender identity in astronomy, and 32\% of participants in 2016 (2017: 33\%) considered themselves a minority with respect to racial or ethnic identity.
-About half of the attendees were students ($\sim$40\% graduate students, $\sim$10\% undergraduate students), another $25\%$ postdoctoral fellow, and the remainder about equally split between research staff, tenured and tenure-track faculty, and participants in non-academic careers.
+For demographic characteristics specifically, 42 percent of participants in both 2016 and 2017 considered themselves a minority with respect to gender identity in astronomy, and 32 percent of participants in 2016 (2017: 33 percent) considered themselves a minority with respect to racial or ethnic identity.
+About half of the attendees were students ($\sim$40 percent graduate students, $\sim$10 percent undergraduate students), another 25 percent were postdoctoral fellows, and the remainder were about equally split between research staff, tenured and tenure-track faculty, and participants in non-academic careers.
 
-Neuro hack week took a different approach: participants provided a CV as part of their application, in addition to answering questions about their technical and research background, and writing a short statement of the purpose of their participation.
-Each application was reviewed by several instructors using 1-10 scales on several criteria: technical and computing experience, research experience, the quality of the statement, an assessment of the contribution of participation to the applicant's career, and an assessment of the degree to which participation would increase the representation of groups that are usually under-represented in neuroscience and computing.
+Neuro Hack Week took a different approach: Participants provided a CV as part of their application, in addition to answering questions about their technical and research background, and writing a short statement of the purpose of their participation.
+Each application was reviewed by several instructors using 1-10 scales to rate several criteria: technical and computing experience, research experience, the quality of the statement, an assessment of the contribution of participation to the applicant's career, and an assessment of the degree to which participation would increase the representation of groups that are usually underrepresented in neuroscience and computing.
 The scores were combined, applicants were ranked according to the combined score, and the 40 highest-scoring applicants were then selected to participate.
 As an example,  this procedure resulted in gender parity in 2016 and near-parity (17 women and 23 men) in 2017.
-In neuro hack week as well about half of the participants were students (though no undergraduate students participated), about $25\%$ were post-docs, with the remainder including research staff and faculty members.
+In Neuro Hack Week, as well about half of the participants were students (though no undergraduate students participated), about 25 percent were post-docs, and the remainder included research staff and faculty members.
 
@@
 We found that minority participants disproportionally decline to attend, and often cite lack of sufficient travel funding as the primary reason.
-In particular if increasing diversity in the more technical aspects of a field is one of the workshop's goals, providing enough travel funding for between five to ten participants can make a significant difference in allowing minority researchers to attend.
+In particular if increasing diversity in the more technical aspects of a field is one of the workshop's goals, providing enough travel funding for five to ten participants can make a significant difference in allowing minority researchers to attend.
 Funding for some activities can be obtained through commercial vendors.
@@
 Federal funding agencies are also sometimes able to fund hack weeks and similar activities under traditional funding mechanisms that support training.
-For example, Neuro hack week 2017 was supported through a grant from the National Institute of Mental Health through their standard R25 training grant mechanism, and this grant continues to support it in extended format in coming years.
+For example, Neuro Hack Week 2017 was supported through a grant from the National Institute of Mental Health through its standard R25 training grant mechanism, and this grant will continue to support it in extended format in the coming years.
 
@@
 
-An overarching goal of every hack week is to provide a welcoming, comfortable, and inclusive environment for participants to learn and work.
-However, disparities in backgrounds, knowledge and experience, together with the close collaboration that hacking requires, creates the potential for conduct that might offend participants, and preclude them from participating fully.
+An overarching goal of every hack week is to provide a welcoming, comfortable, and inclusive environment in which participants may learn and work.
+However, disparities in backgrounds, knowledge, and experience, together with the close collaboration that hacking requires, creates the potential for conduct that might offend participants, and preclude them from participating fully.
 For this reason, a hack week should have a clear and enforceable code of conduct, that can guide the behavior of participants, and help organizers maintain a professional and welcoming environment.
-It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, how it should be enforced, and prepare for different scenarios and their possible resolution in advance.
-A part of this could and should be the recognition that participants (and, indeed, members of the organizing committee) may occasionally violate the code of conduct in minor ways, and allow for tools and mechanisms to resolve these infractions.
+It is the responsibility of the organizing committee to discuss what belongs in the code of conduct, consider how it should be enforced, and prepare for different scenarios and their possible resolution in advance.
+A part of this could and should be the recognition that participants (and, indeed, members of the organizing committee) may occasionally violate the code of conduct in minor ways, and the setting up of tools and mechanisms to resolve these infractions.
 Major infractions, conversely, should be taken seriously and in severe cases lead to exclusion of participants.
 Aside from important provisions banning discrimination on the basis of demographic characteristics, there may be domain- or workshop-specific clauses that may require recognition.
-For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome, defensive and inhibit learning on a larger scale.
-Even if no code of conduct infractions are expected, the very design of such a code sets expectations, and helps signal to participants that organizers are intentional in creating a welcoming environment, which in turn feeds back to the creation of such a space.
+For example, as communities move between programming languages or techniques, shaming participants for their choice of method or language should be strongly discouraged, since it will make participants feel unwelcome and defensive, and will inhibit learning on a larger scale.
+Even if no code of conduct infraction is expected, the very design of such a code sets expectations, and helps signal to participants that organizers have thought about creating a welcoming environment, which in turn feeds back to the creation of such a space.
 
-\subsubsection{Communicating with the Outside World}
+\subsubsection{Communicating with the outside world}
 
 Because hack weeks have enjoyed considerable popularity, they have often been oversubscribed.
-This leaves organizers with an important decision: given the necessity of turning a large fraction of applicants away, what methods does the group wish to employ in order to communicate during the hack week with the wider community not in attendance?
-Tutorials may be live-streamed via popular services such as youtube, but aside from the technical complexity of setting up a reliable live stream, their interactive format may not easily lend themselves to the one-directional communication of a video feed.
+This leaves organizers with an important decision: Given the necessity of turning a large fraction of applicants away, what methods does the group wish to employ in order to communicate during the hack week with the wider community not in attendance?
+Tutorials may be live-streamed via popular services such as YouTube, but aside from the technical complexity of setting up a reliable live stream, their interactive format may not easily lend itself to the one-directional communication of a video feed.
 Even so, live streams and video recording posted after the fact have proven to be both useful and popular in the past, even while they require considerable effort to set up.
 At the very least, there should be some method of real-time communication with the in-group.
-Twitter has proven itself to be an effective medium of general communication with the wider community, including hack pitches, questions during tutorials and collaboration with researchers outside of Astro Hack Week.
+Twitter has proven itself to be an effective medium of general communication with the wider community, including hack pitches, questions during tutorials and collaboration with researchers outside Astro Hack Week.
 However, the closed nature of Twitter to those with an account may not make it the ideal platform for participation.
 Other open-source formats that do not require sign-up, such as Matrix, may be more compatible with the hack week format \footnote{\url{https://dave.cheney.net/2017/04/11/why-slack-is-inappropriate-for-open-source-communications}}.
-Important documents such as programme (and especially programme changes) as well as hack descriptions should be open to the public and communicated clearly to facilitate collaborations with outside researchers.
+Important documents such as program descriptions, program changes, and hack descriptions should be open to the public and communicated clearly to facilitate collaborations with outside researchers.
 
@@
 However, the format rests firmly on a foundation of community-building and creating opportunities for participants to exchange ideas and knowledge.
-While not impossible to achieve in a virtual setting, both would be significantly harder than in a single physical location, to a point where the lack of immediate exchange may severely limit to generate successful new collaborations and ideas.
+While these goals are not impossible to achieve in a virtual setting, both would be significantly harder to accomplish online than in a single physical location, to a point where the lack of immediate exchange may severely limit the generation of successful new collaborations and ideas.
 We thus recommend against running a purely virtual hack week.
@@
 \clearpage
-\subsection{During a Hack Week}
+\subsection{During a hack week}
 
-Unlike traditional conference, where the programme is generally set by pre-selected talks, hack weeks require additional involvement from the organization committee during the hack week itself.
+Unlike a traditional conference, where the program is generally set by pre-selected talks, hack weeks require additional involvement from the organization committee during the hack week itself.
 
@@
 Tutorials provide a mechanism for rapid knowledge sharing and are often one of the most useful components of a hack week.
-Unlike traditional lectures, in which more time is available to explore numerous topics in depth, tutorials aim to deliver only the most essential information while ensuring students remain engaged and interactive.
+Unlike traditional lectures, in which more time is available to explore numerous topics in depth, tutorials aim to deliver only the most essential information while ensuring that students remain engaged and interactive.
 Moreover, the need for an instructor to balance prepared material with spontaneous, interactive learning, while attending to community building and inclusive practices, can be difficult for even the most seasoned educator.
@@
 
-We have experimented with numerous designs in delivering our hack week tutorials in a way that is student-centered and focused on meeting each student at their particular skill level.
-We have observed that, especially during long tutorials and with a particularly diverse group of students, there is a natural separation between participants who are focused on gaining new knowledge as it is being presented, versus those who already possess this knowledge and hence turn their attention to other subjects as the tutorial is ongoing.
+We have experimented with numerous designs in delivering our hack week tutorials in a way that is student-centered and focused on meeting all students at their particular skill levels.
+We have observed that, especially during long tutorials and with a particularly diverse group of students, there is a natural separation between participants who are focused on gaining new knowledge as it is being presented, versus those who already possess this knowledge and hence turn their attention to other subjects as the tutorial is underway.
 In practice, this concern can be mitigated by explicitly involving experienced participants in the tutorials.
 We encourage instructors to call upon experts in the audience to identify themselves and effectively act as teaching assistants.
-Expert participants often find this gratifying because it allows them to test their own understand and improve their skills by learning through teaching.
+Expert participants often find this gratifying because it allows them to test their own understanding and improve their skills by learning through teaching.
 At the same time, beginners benefit by having increased interactions with other hack week participants, thereby increasing team cohesion and building community.
-Another technique to engage expert participants during the tutorial is to invite them to work through more challenging content located in the online lesson, which will not be covered in the session, but for which they can obtain help during the tutorial.
+Another technique for engaging expert participants during the tutorial is to invite them to work through more challenging content located in the online lesson, which will not be covered in the session, but for which they can obtain help during the tutorial.
 
 Some of our hack week tutorials follow the structure of the Software Carpentry model \citep{b:wilson-swc-lessons-2016}.
-The content is divided in to a series of lessons, each having a well-defined set of learning objectives, key questions, and expected outcomes.
-Interspersed throughout each lesson are a series of exercises which participants work through at their own pace.
-We frequently use a method drawn from Software Carpentry whereby students indicate their need for help on an exercise via posting a colored sticky-note on their computer.
-This provides a simple method to deliver help to those students who need it without interrupting the flow of the tutorial.
-We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \citep{kluyver2016jupyter} or following the markdown templates provided by Software Carpentry.
-This allows the hack week ``curriculum'' to be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
+The content is divided into a series of lessons, each having a well-defined set of learning objectives, key questions, and expected outcomes.
+Interspersed throughout each lesson are a series of exercises that participants work through at their own pace.
+We frequently use a method drawn from Software Carpentry, whereby students indicate their need for help with an exercise via posting a colored sticky note on their computer.
+This provides a simple method for identifying those students who need help without interrupting the flow of the tutorial.
+We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \citep{kluyver2016jupyter} or by following the markdown templates provided by Software Carpentry.
+This way, the hack week ``curriculum'' can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
 Depending on the content, some instructors convey concepts during the tutorial via ``live-coding'' while others display static code and their results.
-Either way, we find dividing code into short blocks, each with a specific task, helps participants not to become overwhelmed with information.
+Either way, we find that dividing code into short blocks, each with a specific task, helps keep participants not to become overwhelmed with information.
 In some cases, tutorials require substantial software installation and data.
-Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometime so complicated that it slows down teaching.
-In some cases, we have experimented with \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \citep{holdgraf2017portable}.
+Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometimes so complicated that it slows down teaching.
+In some cases, we have experimented with using \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \citep{holdgraf2017portable}.
 Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \citep{Holdgraf2017-pd, Titus_Brown_undated-pc}.
@@
 In-depth collaboration among instructors in advance of the event is particularly valuable in this regard, to assess which methods or subjects will be most useful for the specific audience.
-We suggest that the primary goals of the tutorials should be to provide an entry point into an exploration of participants' datasets, opening the door to more thorough study outside of the hack week.
+We suggest that the primary goals of the tutorials should be to provide an entry point into an exploration of participants' datasets, opening the door to more thorough study outside the hack week.
 Given the broad span of available tools and topics, the organizing committee should seek to teach content that represents the state-of-the-art and is deemed to be of greatest use to the broadest range of participants.
-A common theme across all of our hack weeks is to teach initial tutorials in version control, command line interfaces, data science platforms and practices of reproducible research.
+A common theme across all our hack weeks is teaching initial tutorials in version control, command line interfaces, data science platforms and practices of reproducible research.
 These tutorials endeavor to provide a common baseline of techniques that participants can build on in other tutorials and project work.
 Subsequent tutorials are then delivered in more domain-specific fields, but can be arranged in a way that builds knowledge constructively over the course of the week.
-In some cases, it may be advisable to recommend some theoretical texts as reading before attending the hack week, though it is likely that speakers should not count on every participant arriving with the same knowledge baseline.
+In some cases, it may be advisable to recommend some theoretical texts as preparatory reading for hack week, though speakers should not count on every participant arriving with the same knowledge baseline.
 Because of their practical nature, however, software requirements should be announced as early and clearly as possible, with the expectation that participants will have installed necessary software in advance.
-It may be helpful to designate some time early on in the meeting to trouble-shoot installations before tutorials begin in order to avoid losing significant time during the actual tutorials.
+It may be helpful to designate a period of time early in the meeting to trouble-shoot installations before tutorials begin in order to avoid losing significant time during the actual tutorials.
 
-It is worth noting that in practice, the mixed audience and interactive setting can often lead to impostor syndrome (see below) among the speakers, who lose their explicit status as expert among the participants.
+It is worth noting that in practice, the mixed audience and interactive setting can often lead to impostor syndrome (see section 4.2.5 below) among the speakers, who lose their explicit status as expert among the participants.
 It is therefore imperative that speakers are made aware as early as possible about the audience, so that they can plan their tutorial accordingly.
-It also helps to make them aware that participants may be taking the role of teaching assistants during exercises, and include them in the teaching, rather than see them as adversaries who may seek to correct them.
+It also helps to make them aware that participants may be taking the role of teaching assistants during exercises, and encourage them to include participants in the teaching, rather than seeing them as adversaries who may seek to correct them.
 Conversely, experts may ask pointed questions about fundamental, important concepts, both in order to help the audience gain a deeper understanding and to encourage questions, especially from junior participants, who may be hesitant to ask questions due to their own impostor syndrome.
-At Astro Hack Week, we encourage the organizing committee and experienced participants to ask questions, particularly when they know the answer, but think the concept may not yet be clear to parts of the audience.
+At Astro Hack Week, we encourage members of the organizing committee and experienced participants to ask questions, particularly when they know the answer, but think the concept may not yet be clear to parts of the audience.
 This has in the past led to exchanges between experts that have greatly contributed to the tutorials in ways that were both unplanned and unexpected for all involved, beginners and experts alike.
@@
 \begin{itemize}
-\item be very clearly tailored to the audience and narrowly scoped,
-\item strictly limit the amount of lecture-style teaching to less than 50\% ,
-\item use experts in the audience to ask key questions and act as teaching assistants during exercises
-\item communicate technical requirements at least a week before the hack week
+\item Be very clearly tailored to the audience and narrowly scoped,
+\item Strictly limit the amount of lecture-style teaching to less than 50 percent,
+\item Use experts in the audience to ask key questions and act as teaching assistants during exercises, and
+\item Communicate technical requirements at least a week before the beginning of the hack week.
 \end{itemize}
@@
 
-Because instructors and organizing committees are unlikely to know in advance what projects and data sets participants will bring, often there will be topics and methods not covered in the pre-planned tutorials that are are of interest to the audience.
-Here, break-out sessions offer an alternative: short (30-45 minute) tutorials that are fairly spontaneously organized (with as little a lead time of a few hours) and often taught by expert participants in the audience.
+Because instructors and organizing committees are unlikely to know in advance what projects and data sets participants will bring, often there will be topics and methods not covered in the pre-planned tutorials that are of interest to the audience.
+Here, break-out sessions offer an alternative: short (30-45 minute) tutorials that are fairly spontaneously organized (with as little lead time as a few hours) and often taught by expert participants in the audience.
 These tutorials can be a more in-depth treatment of one specific method of interest (e.g.
-Gaussian Processes, K-Means Clustering, Deep Learning), or cover a practical skill that may be useful to the audience, but is not formally part of the hack week (this is especially the case for skills related to software-development, such as code testing, documentation, and profiling).
-They are generally more informal than tutorials and can be taught to a subset of the group rather than requiring all participants to attend.
-Break-out sessions should be limited to one or two a day, in order to allow participants to attend break-out sessions without having to sacrifice a significant fraction of their time reserved for project work.
+Gaussian processes, K-means clustering, or deep learning), or cover a practical skill that may be useful to the audience, but is not formally part of the hack week (this is especially the case for skills related to software development, such as code testing, documentation, and profiling).
+Break-out sessions are generally more informal than tutorials and can be taught to a subset of the group rather than requiring all participants to attend.
+They should be limited to one or two a day, in order to allow participants to attend sessions without having to sacrifice a significant fraction of their time reserved for project work.
 It is also possible to intentionally leave a tutorial slot free, to be filled with one or two self-organized break-out sessions instead of a pre-planned tutorial.
 In practical terms, it is advisable to keep a physical board with requests and potential teachers in a prominent location during the week.
-Decisions on which option to choose from the list can be done via an informal voting process, under the condition that a volunteer to hold the tutorial is found.
-Giving a break-out session can be a daunting task: while some experts may have relevant tutorials already prepared from other workshops, often it requires holding a thirty-minute talk with little to no preparation.
-Organizers can and should take steps to provide a positive, encouraging environment for participants to volunteer their knowledge.
-They should particularly encourage junior participants, who may often be the most knowledgable about the topics usually requested, to volunteer for break-out sessions.
-They provide a valuable teaching experience in a friendly environment and offer the opportunity of networking with the larger community.
+Decisions on which option to choose from the list can be done via an informal voting process, under the condition that a volunteer to hold the tutorial has been found.
+Giving a break-out session can be a daunting task: While some experts may have relevant tutorials already prepared from other workshops, often participants are being expected to offer a thirty-minute talk with little to no preparation.
+Organizers can and should take steps to provide a positive, encouraging environment in which participants can volunteer their knowledge.
+They should particularly encourage junior participants, who may often be the most knowledgeable about the topics usually requested, to volunteer for break-out sessions.
+They provide a valuable teaching experience in a friendly environment and offer the opportunity to network with the larger community.
 
@@
 \begin{itemize}
-\item consist of short tutorials about a specific topic or practical skill not covered in the tutorials
-\item are more informal and may be taught to a subset of interested participants
-\item should be limited to one to two per day and no longer than 45 minutes to avoid conflicts with project-work
-\item can be daunting to teach, especially for junior participants, and requires the organizing committee to be pro-active about encouraging volunteer teachers
+\item Consist of short tutorials about a specific topic or practical skill not covered in the tutorials
+\item Are more informal and may be taught to a subset of interested participants
+\item Should be limited to one to two per day and take no longer than 45 minutes to avoid conflicts with project-work
+\item Can be daunting to teach, especially for junior participants, and 
+\item Require the organizing committee to be pro-active about encouraging volunteer teachers.
 \end{itemize}
 
-\subsubsection{Hack Sessions}
+\subsubsection{Hack sessions}
 
@@
 For junior participants and those new to the format, this can be daunting, even if (or especially when) they do not have a project themselves.
-Junior participants in particular tend to believe that they do not yet have the skills to contribute meaningfully to any projects.
+Junior participants, in particular, tend to believe that they do not yet have the skills to contribute meaningfully to any projects.
 Here, it can be useful to require the participants who are not pitching projects to state their expertise and affirm that every participant has expertise in something.
@@
 Tracking ongoing projects can be accomplished online, via real-time collaborative documents, as well as in the physical space, via a diagram of where in the space teams are located.
-Because hack weeks are generally longer than typical hackathons (often limited to just one or two days), this will lead to a variety of shorter projects completed within a day or two, and longer projects going on throughout the week.
-Regular check-in sessions are helpful for pitching new projects, requesting help on projects stuck at a certain point, and re-matching participants after hacks have been completed or abandoned.
+Because hack weeks are generally longer than typical hackathons (which are often limited to just one or two days), this will lead to a variety of shorter projects completed within a day or two, and longer projects going on throughout the week.
+Regular check-in sessions are helpful for pitching new projects, requesting help on projects that are stuck at a certain point, and re-matching participants after hacks have been completed or abandoned.
 These sessions can also be used for completed hacks to be showcased, or intermediate achievements to be presented.
@@
 
-\subsubsection{Facilitating Participation}
+\subsubsection{Facilitating participation}
 
@@
 Community building and careful facilitation of both tutorials and project works are arguably the cornerstones of a successful hack week.
-This is especially important because the group will inevitably include a wide range of personalities, some of which may find it easier to form new bonds and interact with the group than others.
-At all hack weeks we have organized, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
-Clearly communicating values and expectations at the beginning for the week can help set the stage for an environment in which a large range of different personalities can thrive.
-This includes both communication of the code of conduct, putting an emphasis on collaborative rather than competitive behaviour, as well as a discussion of potential issues participants might experience, such as impostor syndrome (see also section below).
+This is especially important because the group will inevitably include a wide range of personalities, some of whom may find it easier to form new bonds and interact with the group than others.
+At all the hack weeks we have organized, we have taken concrete steps to address potential issues and foster a welcoming, inclusive atmosphere.
+Clearly communicating values and expectations at the beginning of the week can help set the stage for an environment in which a large range of different personalities can thrive.
+This includes both communication of the code of conduct, putting an emphasis on collaborative rather than competitive behavior, as well as a discussion of potential issues that participants might experience, such as impostor syndrome (see also  section 4.2.5 below).
 
-From our experience, there are ways in which organizers can influence group dynamics in specific ways to allow timid participants to speak up.
+Our experience indicates that there are ways in which organizers can influence group dynamics in specific ways to allow timid participants to speak up.
 As mentioned above, we ask a few participants known to us to occasionally ask questions during tutorials, in order to encourage other participants to do the same.
-We also designate at least two members of the organizing committee whose role it is to facilitate both tutorials and hack sessions.
+We also designate at least two members of the organizing committee to take on the role of facilitating both tutorials and hack sessions.
 It is their task to also manage dominant personalities, which can be both a boon and a hindrance to the success of a hack week.
-Often, these attendees can be effective at generating excitement and engagement, but can also intimidate and dominate discussions.
+Often, these attendees can be effective at generating excitement and engagement, but can also intimidate others and dominate discussions.
 It is up to the organizers to balance the participation of these attendees with the needs of the rest of the group.
-Similarly, facilitators at the hack weeks we have organized spend the majority of hacking time circulating the room, connecting participants without a group to a project, help connect groups who run into a problem with experts who can help out, and make sure group work proceeds in such a way that all members feel valued.
+Similarly, facilitators at the hack weeks we have organized spend the majority of hacking time circulating the room, connecting participants without a group to a project, helping connect groups who run into a problem with experts who can help out, and making sure that group work proceeds in such a way that all members feel valued.
 
-At geo hack week we employed ``Liberating Structures'', a facilitation approach in which we introduce small shifts in the way groups engage in order to increase relational coordination and trust.
+At Geo Hack Week, we employed ``Liberating Structures'', a facilitation approach in which we introduce small shifts in the way groups engage in order to increase relational coordination and trust.
 We invited a professional facilitator to assist us in leading a series of conversations and activities aimed at fostering group engagement.
 In one activity, participants faced one another and were given a brief statement inviting a two minute conversation, after which the rows rotated to maximize the mixing of ideas.
-Invitations included statements such as ``I belong at geo hack week because...'' or ``A challenge I hope we can make progress on is...''.
-Importantly, event organizers joined in the activity, which we found helped participants feel as if they were not alone in facing some of the challenges inherent in data science studies.
+Invitations included statements such as ``I belong at Geohack Week because...'' or ``A challenge I hope we can make progress on is...''.
+Importantly, event organizers joined in each activity, which we found helped participants feel as if they were not alone in facing some of the challenges inherent in data science studies.
 Another activity was designed to address the challenges of forming project teams.
@@
 The remaining participants were then invited to visit any station that drew their interest, and to rotate between up to two other stations if they desired.
-We found this approach minimized the pressure on participants, giving them time to get a feel for project ideas and group dynamics before joining a team.
+We found that this approach minimized the pressure on participants, giving them time to get a feel for project ideas and group dynamics before joining a team.
 
@@
 This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
-Less severely, another major concern is that IS inhibits risk taking: participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, or to be proactive about forming new collaborations.
- Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, as well as community building are hampered by IS.
+Less severely, another major concern is that IS inhibits risk taking: Participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, and to be pro-active about forming new collaborations.
+ Many of the goals of a hack week, including the successful completion of projects, lateral knowledge transfer, and community building are hampered by IS.
  Thus, we are working within our hack weeks to mitigate IS using various techniques.
@@
 %I think it helped ease the feeling!!``}).
-Additionally, role models are very effective at encouraging positive behaviour: asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
+Additionally, role models are very effective at encouraging positive behavior: Asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.
 We also invited our organizing committees to model open communication about their own handling of IS through the course of their career.
-We found that transparency about one's own uncertainty, espcially from community leaders, helped create a hospitable atmosphere and removed barriers.
+We found that transparency about one's own uncertainty, especially from community leaders, helped create a hospitable atmosphere and removed barriers.
 
-\subsection{What to do after}
+\subsection{What to do after hack week}
 
@@
 Therefore, it is useful to put some detailed thoughts into the survey.
-Following discussions the organizing committee will have likely had during the initial stage of preparing the announcement and selecting participants, organizers will want to confront their own expectations before the workshop, and design a survey that allows for both measurements of specific outcomes and goals the organizers are interested in, as well as leave enough free space for participants to self-report outcomes that arose spontaneously or were out of the scope of the initial design.
-It is helpful to design the survey early on and allow for some time near the end of the workshop for participants to fill it out while they are still physically present; in our experience, this leads to much higher response rates than if the survey is distributed after the end of the workshop.
+Following discussions the organizing committee will have likely had during the initial stage of preparing the announcement and selecting participants, organizers will want to confront their own expectations before the workshop, and design a survey that allows for both measurements of specific outcomes and goals the organizers are interested in, as well as leaving enough free space for participants to self-report outcomes that arose spontaneously or were outside the scope of the initial design.
+It is helpful to design the survey early and allow for some time near the end of the workshop for participants to fill it out while they are still physically present; in our experience, this leads to much higher response rates than if the survey is distributed after the end of the workshop.
 While some improvements and changes in survey design from year to year are inevitable, it is advisable to set a number of core questions that do \textit{not} change from year to year.
@@
 This document might be rewritten as conference proceedings.
-Additionally, organizers should encourage participants to identify written outcomes with the hack week: for publications, acknowledgments provide a convenient venue.
-Code repositories can allow this kind of acknowledgment in a \textit{readme} document; GitHub also allows attaching tags to a repository, thus all code repositories that experienced major contributions during the workshop may include a tag specific to the hack week.
+Additionally, organizers should encourage participants to identify written outcomes of the hack week: For publications, acknowledgments provide a convenient venue.
+Code repositories can allow this kind of acknowledgment in a \textit{readme} document; GitHub also allows attaching tags to a repository, so all code repositories that experienced major contributions during the workshop may include a tag specific to the hack week.
 
-On a larger scale, it might be useful to track measures of success more long-term.
-Those could for example be publication networks for scientists who attended the workshop, their publications as well as their career development.
+On a larger scale, it might be useful to track measures of success that are more long-term.
+Those could for example be publication networks for scientists who attended the workshop, including their publications as well as their career development.
 For this purpose, it is useful to keep a way to contact previous participants, via a mailing list or real-time messaging services like \textit{Matrix} or \textit{Slack}.

### 2018-06-06 supplementary_material.tex
Fixed newly introduced typo

@@
 
-For no combination of question questions with a demographic variable, do we ﬁnd a $p$-value below the trial-corrected signiﬁcance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
+For no combination of questions with a demographic variable, do we ﬁnd a $p$-value below the trial-corrected signiﬁcance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
 In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication of where future research may be fruitful.

### 2018-06-11 design.tex

@@
 
-Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance) .   
+Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance).
 Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges.
 By spending more time together, participants are more likely to overcome barriers of professional terminology.
-However, events that are too long may lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
+However, events that are too long may also lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
 A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and project work.
 
-A hack week requires a flexible workspace environment with ample opportunity for re-configuration. 
+Another important design consideration is the space. A hack week requires a flexible workspace environment with ample opportunity for re-configuration.
 Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops.
@@
 %However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
-Fortunately, many universities are experimenting with other types of spaces: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.
+Fortunately, many universities are experimenting with new types of spaces that allow these kinds of activities: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.
 
@@
 If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations.
-We have found groups with sizes between 50 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group, but we encourage organizers to experiment with group size.
+We have found groups with sizes between 40 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group, but we encourage organizers to experiment with group size.
 
@@
 This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
-The large variety in participant backgrounds and experiences--and the resulting range in personalities of attendees--requires careful, active facilitation of both taught and project components of a hack week (see also SM; Section 4.2.4).
+The large variety in participant backgrounds and experiences--and the resulting range in personalities and objectives of attendees--requires careful, active facilitation of both taught and project components of a hack week (see also SM; Section 4.2.4).
 A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 4.1.7).
-Community building is a core component of a hack week, and designs need to include both very strong personalities and very shy participants.
-In particular the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).
+Community building is a core component of a hack week, and the design of hack weeks needs to include both very strong personalities and very shy participants.
+In particular, the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).

### 2018-06-11 themes.tex

@@
 In this case, it may be advantageous to design a hack week around a technique (e.g.\ Gaussian Processes) or modality (e.g.\ imaging), such as the ImageXD (image processing across domains\footnote{\url{http://http://www.imagexd.org/}}) meetings.
-For these events, building a shared vocabulary and understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.
+For these events, building a shared vocabulary and shared understanding of major data analysis problems is crucial, but they also allow for cross-disciplinary diffusion of techniques into other subjects and therefore decrease the risk of duplication of method development efforts.

### 2018-06-11 what.tex

@@
 Additionally, they provide opportunities to build new collaborations and network.
-Our hack weeks combine structured, tutorial-style instruction with open-ended project work, providing opportunities for peer learning, networking, and building collaborations.  
-In a space spanned by pedagogical focus on one axis and focus on project work on the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).
+Our hack weeks combine structured, tutorial-style instruction with open-ended project work, providing opportunities for peer learning, networking, and for building collaborations.
+In a space spanned by pedagogical focus as one dimension and focus on project work as the other, the hack weeks we have organized are designed to lie somewhere in between traditional summer schools and hackathons, where we believe they fill a space not currently fully addressed by existing models (Figure \ref{fig:hackspectrum}).
 
@@
 Core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
-Building on these core elements hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}.
+Building on these core elements, hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}.
 %``Catalytic'' hackathons seek novel project ideas aimed at solving a tractable, well-defined challenge.
@@
 They encourage participants to self-organize in many different forms: experts working with other experts, mentoring relationships between experts and non-experts, or study groups among non-experts, to name but a few.
-Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: our hack weeks explicitly encourage projects e.g.~ around outreach and work aimed at improving the scientific community itself.
+Hack weeks also allow participants to experiment with projects and ideas beyond their day-to-day research: For example, our hack weeks explicitly encourage projects around outreach and work aimed at improving the scientific community itself.
 
-There is, however, a major risk in the lack of focus: by wanting to be too many things at once, a hack week might potentially not do any of them well.
+There is, however, a major risk in the lack of focus: by wanting to do many things at once, a hack week might potentially not do any of them well.
 Because tutorials are not necessarily the major focus of a hack week, the knowledge gained by participants in these tutorials may be shallow.

### 2018-06-11 why.tex

@@
 While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning.
-This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
+This type of lateral knowledge transfer \cite{b:wilson-swc-lessons-2016} is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
 
@@
 
-\item{\textit{Community Building}: Hack weeks provide a tremendous opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
-These events allow computationally minded researchers to break from the isolation of their academic departments and spark new collaborations.}
+\item{\textit{Community Building}: Hack weeks provide an opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
+These events allow computationally-minded researchers to break from the isolation of their academic departments and spark new collaborations.}
 

### 2018-06-11 results.tex
small corrections to tense

@@
 
-Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that most participants self-report successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
+Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that most participants self-reported successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
 The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist  (Figure \ref{fig:survey}, (b, c)).
@@
 Because peer learning is a major mode of knowledge transfer at hack weeks, we asked participants whether they taught other participants.
-We find that again a majority agrees with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\%; Figure \ref{fig:survey}, (d)), though responses are not as unequivocal as they are in some of the other categories.
+We find that again a majority agreed with this statement to some degree (AHW: 79\%, GHW: 69\%, NHW: 75\%; Figure \ref{fig:survey}, (d)), though responses are not as unequivocal as they are in some of the other categories.
 The majority of participants felt that they built valuable connections to other researchers (Figure \ref{fig:survey}, (e)), especially at NHW, where more than 64\% of participants strongly agreed with this statement.
@@
 Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ with respect to gender or racial/ethnic identity.
-For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
+For GHW, there is a weak indication that participants from racial/ethnic minorities responded more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
 These initial results indicate that future work on how demographics interact with hack week attendance may be fruitful.
 
-We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
+We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reported that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
 %The overall behaviour hints at how general conventions and attitudes differ in different fields.

### 2018-06-21 results.tex
Updated paper with equivalence test results

--- 2018-06-04 results.tex
@@
 Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)).
+However, the confidence intervals on the measured effect sizes are wide, and thus testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size indicates that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).
 
@@
 For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
-These initial results indicate that future work on how demographics interact with hack week attendance may be fruitful.
+Equivalence tests for the hypothesis of a small or non-existent effect reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW).
+These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an initial indication that our careful facilitation strategies may be effective in fostering participation, and that future work on how demographics interact with hack week attendance may be fruitful.
 

### 2018-06-21 supplementary_material.tex
Updated paper with equivalence test results

--- 2018-06-04 supplementary_material.tex
@@
 We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
-For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size.
+For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{}.
 We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
@@
 We found no $p$-values below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials} for any of the combinations of questions and demographic variables.
-In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication where future research may be fruitful.
+In the main text, we place these null results in context, and refer to the tests that were near significance, coupled with a moderate to large effect size, as an indication where future research may be fruitful.
 The results also discuss the present study's limitations.
@@
 \begin{threeparttable} 
-\begin{tabularx}{18cm}{p{3.5cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
+\begin{tabularx}{24cm}{p{3.2cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
 \toprule
-\textbf{Question} & \multicolumn{7}{c}{\textbf{Astro Hack Week}} & \multicolumn{7}{c}{\textbf{GeoHackWeek}} & \multicolumn{7}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $p_{\mathrm{eq}}$  \\ \midrule
+\textbf{Question} & \multicolumn{9}{c}{\textbf{Astro Hack Week}} & \multicolumn{9}{c}{\textbf{GeoHackWeek}} & \multicolumn{9}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ &$\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
-New topics, tools, methods (a) & 	61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 		36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 		41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
-Improvement of research (b) & 		61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 	35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 	40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
-Better science (c) & 				59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 		35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 	40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\
-Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 	40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
+New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0 & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0 & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0 & 0.32 & 0.17 \\
+Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0 & 0.31 & 0.14 \\
+Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0 & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0 & 0.28 & 0.12 \\
+Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0 & 0.33 & 0.19 \\ \midrule
 \multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
-Improvement of research (b) & 		60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 		35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 	40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
-Teaching (d) & 					61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 	36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 	40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
-Connection building (e) & 			61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 		36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 	41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
-Valued contributions (f) & 			57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 	34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 	38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
+Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0 & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0      & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0 & 0.30 & 0.05\\
+Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0 & 0.30 & 0.05 & 		36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.37 & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0 & 0.36 & 0.11\\
+Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0 & 0.28 & 0.03 & 		36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0      & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0 & 0.32 & 0.07 \\
+Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0 & 0.28 & 0.03 &		34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0      & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0 & 0.35 & 0.10\\ \midrule
 \multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
-Improvement of research (b)  & 	59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 		34 &2 & 0.01 & 0.99 & 0.97 & 0 &		39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
-Teaching (d) & 					61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 	34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 		39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
-Connection building (e) & 			61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 		34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 	40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
-Valued contributions (f) & 			58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 	32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 		37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\
+Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0 & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0 & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0 & 0.32 & 0.06 \\
+Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0 & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0 & 0.38 & 0.12 \\
+Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0 & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0 & 0.33 & 0.07 \\
+Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0 & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0 & 0.36 & 0.11\\
 \bottomrule

### 2018-06-21 abstract.tex

@@
 \begin{abstract}
-Across almost all scientific disciplines, instruments that record experimental data and methods required for storage and data analysis are rapidly increasing in complexity.
-As a result of these rapid developments, scientific communities need to adapt on shorter time scales than traditional university curricula can evolve.
-New modes of knowledge transfer are therefore needed.
-On the other hand, the broad applicability of data science tools to a range of problems has generated new opportunities to foster exchange of ideas and computational workflows across disciplines.
-In recent years, hack weeks have emerged as a novel and effective tool for fostering these exchanges and providing training in modern data analysis workflows.
-While there are variations in hack week implementation, all events consist of a common core of three components: tutorials in state-of-the-art methodology, peer-learning and project work in a collaborative environment.
-In this paper, we present the concept of a hack week in the larger context of scientific meetings.
-We motivate the need for such an event and present its strengths and challenges.
+Across almost all scientific disciplines, instruments and methods for recording, storing and analyzing data are rapidly increasing in complexity.
+Numerous data science tools are in development that aim to manage this complexity through development of open, reproducible workflows and the exchange of ideas across disciplines.
+Successful navigation of this new data science landscape requires training in new programming languages and frameworks, as well as immersion in new modes of interaction that foster data sharing and collaborative software development.
+Unfortunately, learning these skills from traditional university curricula can be challenging because most courses are not designed to evolve on time scales that can keep pace with rapidly shifting data science methods.
+Here we present the concept of a hack week as a novel and effective model offering opportunitites for networking and community building, education in state-of-the-art data science methods and immersion in collaborative project work.
 We find that hack weeks are successful at cultivating collaboration and facilitating the exchange of knowledge.

### 2018-06-21 results.tex

@@
 Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow a wide range of participants to participate fully.
+
 Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
@@
 
-We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reports that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
+We find that the hack weeks have been largely successful in promoting positive attitudes towards reproducibility and open science: at all three events, the majority reported that the hack week has made them more comfortable with open science  (GHW: 97\%; NHW: 95\%; AHW: 72\%, Figure \ref{fig:survey}, (h)), and more than 85\% of all participants (AHW: 86\%, GHW: 94\%, NHW: 95\%; Figure \ref{fig:survey}, (g)) put code or data created at the hack week into a public repository.
 %The overall behaviour hints at how general conventions and attitudes differ in different fields.

### 2018-06-21 supplementary_material.tex

--- 2018-06-06 supplementary_material.tex
@@
 No responses were discarded, and no pre-processing was performed on the data.
-We tested for correlations between demographic characteristics (independent variables), questioned responses (dependent variables) using a standard $\chi^2$ test and computed the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
+We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
 For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
+
+We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
+For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{}.
+We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
+For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
+If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.
+
 The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
-
-\clearpage
 
@@
 \begin{threeparttable} 
-\begin{tabularx}{18cm}{p{3.6cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}}
+\begin{tabularx}{24cm}{p{3.2cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
 \toprule
-\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
+\textbf{Question} & \multicolumn{9}{c}{\textbf{Astro Hack Week}} & \multicolumn{9}{c}{\textbf{GeoHackWeek}} & \multicolumn{9}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ &$\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
-New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
-Improvement of research (b) & 61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
-Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\
-Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
+New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0 & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0 & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0 & 0.32 & 0.17 \\
+Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0 & 0.31 & 0.14 \\
+Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0 & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0 & 0.28 & 0.12 \\
+Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0 & 0.33 & 0.19 \\ \midrule
 \multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
-Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
-Teaching (d) & 61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
-Connection building (e) & 61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
-Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
+Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0 & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0      & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0 & 0.30 & 0.05\\
+Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0 & 0.30 & 0.05 & 		36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.37 & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0 & 0.36 & 0.11\\
+Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0 & 0.28 & 0.03 & 		36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0      & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0 & 0.32 & 0.07 \\
+Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0 & 0.28 & 0.03 &		34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0      & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0 & 0.35 & 0.10\\ \midrule
 \multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
-Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
-Teaching (d) & 61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
-Connection building (e) & 61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
-Valued contributions (f) & 58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\
+Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0 & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0 & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0 & 0.32 & 0.06 \\
+Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0 & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0 & 0.38 & 0.12 \\
+Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0 & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0 & 0.33 & 0.07 \\
+Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0 & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0 & 0.36 & 0.11\\
 \bottomrule

### 2018-06-21 results.tex
Minor changes to results

@@
 Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)).
-However, the confidence intervals on the measured effect sizes are wide, and thus testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size indicates that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).
+However, the confidence intervals on the measured effect sizes are wide.
+Testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size suggests that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).
 
 One important question is whether participants from underrepresented groups thrive at hack weeks, or whether their full participation is impeded.
-Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow a wide range of participants to participate fully.
-
-Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
-Similarly, minority participants with respect to gender identity and race or ethnicity do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$).
-Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ with respect to gender or racial/ethnic identity.
+Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow members of these groups to participate fully.
+In our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
+Similarly, minority participants do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$).
+Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ between minority and non-minority groups.
 For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
-Equivalence tests for the hypothesis of a small or non-existent effect reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW).
+Equivalence tests reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW).
 These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an initial indication that our careful facilitation strategies may be effective in fostering participation, and that future work on how demographics interact with hack week attendance may be fruitful.

### 2018-06-21 supplementary_material.tex
More minor corrections

@@
 In 2017, a number of participants did not attend the full week, and thus several attendees were not present on the last day when the survey was administered.
-For the AHW surveys, we checked whether the demographics of the respondents diﬀered signiﬁcantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background, and gender identity) and found no signiﬁcant diﬀerence for any of the groups except for career stage in 2017: Here, we ﬁnd that graduate students are underrepresented (27 percent of respondents compared to 41 percent of attendees), while undergraduate students are overrepresented (25 percent of respondents compared to 12 percent of attendees).
-Because we group responses in the career stage category into the categories “early-career” (students and postdoctoral fellows), “late-career” (faculty, research scientists), and “other” (e.g.
+For the AHW surveys, we checked whether the demographics of the respondents diffÄered signifÅcantly from those of the attendees overall in three relevant categories (career stage, racial/ethnic background, and gender identity) and found no signifiÅcant difference for any of the groups except for career stage in 2017: Here, we find that graduate students are underrepresented (27 percent of respondents compared to 41 percent of attendees), while undergraduate students are overrepresented (25 percent of respondents compared to 12 percent of attendees).
+Because we group responses in the career stage category into the categories ``Äúearly-career''Äù (students and postdoctoral fellows), ``late-career'' (faculty, research scientists), and ``other"Äù (e.g.
 participants from outside academia), this discrepancy has no bearing on the results we report.
-For gender, we grouped participants into “minority” (women and non-binary) and “non-minority” (men) participants; similarly, for race and ethnicity, we divided participants into the categories “non-white” (all but Caucasian participants) and “white” (Caucasian).
+For gender, we grouped participants into ``Äúminority'' (women and non-binary) and ``non-minority''Ä (men) participants; similarly, for race and ethnicity, we divided participants into the categories ``non-white'' (all but Caucasian participants) and ``white'' (Caucasian).
 Note that demographic data for GHW and NHW is only available for 2016, while for AHW, data is available for both 2016 and 2017.
@@
 We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
-For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{}.
-We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
+For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{shiskina2018}.
+We generate 10,000 random data sets from the measured responses by sampling with replacement for each response variable and compute the corresponding effect size in order to generate confidence intervals around the measured Cram\'{e}r's V.
 For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
@@
 We tested for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity.
-The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, and (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$).
+The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$), and (5) the upper 95\% confidence interval for the Cram\'{e}r's V statistic ($\mathrm{CI}_{\mathrm{U}}$) as well as the corresponding $p$-value that the null hypothesis (the presence of an effect with moderate or higher strength) can be rejected, $p_\mathrm{eq}$ .
 The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}
 
-For no combination of questions with a demographic variable, do we ﬁnd a $p$-value below the trial-corrected signiﬁcance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
+For no combination of questions with a demographic variable, do we find a $p$-value below the trial-corrected significance threshold of $p < 0.0007$ quoted in Section \ref{sec:suppmaterials}.
 In the main text, we place these null results in context, and refer to the tests that were near significance, and were coupled with a moderate to large effect size, as an indication of where future research may be fruitful.
@@
 \begin{threeparttable} 
-\begin{tabularx}{24cm}{p{3.2cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
+\begin{tabularx}{24cm}{p{4.1cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{1.0cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{1.0cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{1.0cm}}
 \toprule
-\textbf{Question} & \multicolumn{9}{c}{\textbf{Astro Hack Week}} & \multicolumn{9}{c}{\textbf{GeoHackWeek}} & \multicolumn{9}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ &$\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule
+\textbf{Question} & \multicolumn{8}{c}{\textbf{Astro Hack Week}} & \multicolumn{8}{c}{\textbf{GeoHackWeek}} & \multicolumn{8}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
-New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0 & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0 & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0 & 0.32 & 0.17 \\
-Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0 & 0.31 & 0.14 \\
-Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0 & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0 & 0.28 & 0.12 \\
-Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0 & 0.33 & 0.19 \\ \midrule
+New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0.32 & 0.17 \\
+Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0.31 & 0.14 \\
+Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0.28 & 0.12 \\
+Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0.33 & 0.19 \\ \midrule
 \multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
-Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0 & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0      & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0 & 0.30 & 0.05\\
-Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0 & 0.30 & 0.05 & 		36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.37 & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0 & 0.36 & 0.11\\
-Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0 & 0.28 & 0.03 & 		36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0      & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0 & 0.32 & 0.07 \\
-Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0 & 0.28 & 0.03 &		34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0      & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0 & 0.35 & 0.10\\ \midrule
+Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0.30 & 0.05\\
+Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0.30 & 0.05 & 	36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0.36 & 0.11\\
+Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0.28 & 0.03 & 	36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0.32 & 0.07 \\
+Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0.28 & 0.03 &	34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0.35 & 0.10\\ \midrule
 \multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
-Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0 & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0 & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0 & 0.32 & 0.06 \\
-Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0 & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0 & 0.38 & 0.12 \\
-Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0 & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0 & 0.33 & 0.07 \\
-Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0 & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0 & 0.36 & 0.11\\
+Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0.32 & 0.06 \\
+Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0.38 & 0.12 \\
+Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0.33 & 0.07 \\
+Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0.36 & 0.11\\
 \bottomrule
@@
 More generally, being open about the presence and prevalence of IS can help participants feel more at ease.
-%(e.g.
-one Astro Hack Week participant remarked in a survey, \textit{''I really appreciated the direct acknowledgement of impostor syndrome on the first day.
-%I think it helped ease the feeling!!``}).
 Additionally, role models are very effective at encouraging positive behavior: Asking participants with prior experience at hack weeks at all academic levels to ask questions, even when they might know the answer, can help foster an inclusive environment that rewards risk taking.

### 2018-06-21 abstract.tex
Shorten abstract, neuro example.

@@
 \begin{abstract}
-Across almost all scientific disciplines, instruments and methods for recording, storing and analyzing data are rapidly increasing in complexity.
-Numerous data science tools are in development that aim to manage this complexity through development of open, reproducible workflows and the exchange of ideas across disciplines.
-Successful navigation of this new data science landscape requires training in new programming languages and frameworks, as well as immersion in new modes of interaction that foster data sharing and collaborative software development.
-Unfortunately, learning these skills from traditional university curricula can be challenging because most courses are not designed to evolve on time scales that can keep pace with rapidly shifting data science methods.
+Across many scientific disciplines, methods for recording, storing and analyzing data are rapidly increasing in complexity.
+Facility in data science tools to manage this complexity requires training in new programming languages and frameworks, as well as immersion in new modes of interaction that foster data sharing and collaborative software development and exchange across disciplines.
+Learning these skills from traditional university curricula can be challenging because most courses are not designed to evolve on time scales that can keep pace with rapidly shifting data science methods.
 Here we present the concept of a hack week as a novel and effective model offering opportunitites for networking and community building, education in state-of-the-art data science methods and immersion in collaborative project work.

### 2018-06-21 results.tex
Shorten abstract, neuro example.

@@
 \subsubsection*{Example 3: Neuro Hack Week}
-Motion of study participants inside the MRI machine is a major concern in neuroimaging studies.
-During NHW 2016, one of the teams focused on a large and openly available data-set of MRI data from children\footnote{ABIDE: \url{http://preprocessed-connectomes-project.org/abide}}.
-To test the effect of motion on the results, the team conducted an analysis in which both the number of experimental subjects included, as well as motion cut-off, were varied.
+During NHW 2016, one of the teams analyzed an openly available data-set of MRI data from children\cite{Di_Martino2014-qf}, to test the effects of motion on the results, using varying motion cut-offs.
 %They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (composed of four researchers from four different institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results in the open access journal Research Ideas and Outcomes \cite{leonard2017}.
+The team (four researchers from  different institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results \cite{leonard2017}.

### 2018-06-21 results.tex
Shorten abstract, neuro example.

--- 2018-06-21 results.tex
@@
 Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)).
-However, the confidence intervals on the measured effect sizes are wide, and thus testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size indicates that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).
 
@@
 Significant differences between minorities and non-minorities, even on a self-reported scale, for questions related to learning outcomes, teaching or network building, would indicate that improvements in workshop facilitation and structure may be required to allow a wide range of participants to participate fully.
-
 Based on our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
@@
 Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ with respect to gender or racial/ethnic identity.
-For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
-Equivalence tests for the hypothesis of a small or non-existent effect reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW).
-These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an initial indication that our careful facilitation strategies may be effective in fostering participation, and that future work on how demographics interact with hack week attendance may be fruitful.
+For GHW, there is a weak indication that participants from racial/ethnic minorities responded more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
+These initial results indicate that future work on how demographics interact with hack week attendance may be fruitful.
 

### 2018-06-21 supplementary_material.tex
Fix typo.

--- 2018-06-21 supplementary_material.tex
@@
 \usepackage{tabularx}
-%\usepackage{pdflscape}
 \linespread{1.2}
@@
 No responses were discarded, and no pre-processing was performed on the data.
-We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
+We tested for correlations between demographic characteristics (independent variables), questioned responses (dependent variables) using a standard $\chi^2$ test and computed the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
 For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
+The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
 
-We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
-For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{}.
-We generate 10,000 random samples via bootstrap for each response variable and compute the corresponding effect size in order to generate 90\% confidence intervals around the measured Cram\'{e}r's V.
-For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
-If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.
-
-The full analysis procedure is available online\footnote{See the repository: \url{https://github.com/uwescience/HackWeek-Writeup}}.
+\clearpage
 
@@
 
-%\begin{landscape}
-\begin{sidewaystable}
+\begin{table*}
 %\begin{table*}[hbtp]
@@
 \begin{threeparttable}
-\begin{tabularx}{24cm}{p{3.2cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}p{0.4cm}}
+\begin{tabularx}{18cm}{p{3.6cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}p{0.42cm}}
 \toprule
-\textbf{Question} & \multicolumn{9}{c}{\textbf{Astro Hack Week}} & \multicolumn{9}{c}{\textbf{GeoHackWeek}} & \multicolumn{9}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ &$\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{L}}$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule
+\textbf{Question} & \multicolumn{6}{c}{\textbf{Astro Hack Week}} & \multicolumn{6}{c}{\textbf{GeoHackWeek}} & \multicolumn{6}{c}{\textbf{NeuroHackWeek}} \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ \\ \midrule
 \multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
-New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0 & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0 & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0 & 0.32 & 0.17 \\
-Improvement of research (b) & 		61 &   6 &   9.89 & 0.13 & 1.93 & 0.18 & 0 & 0.24 & 0.08 &	35 & 4   & 12.41 & 0.02 & 1.88 & 0.35 & 0 & 0.28 & 0.13 &		40 &   6 &   7.42 & 0.28 & 1.90 & 0.13 & 0 & 0.31 & 0.14 \\
-Better science (c) & 				59 &   6 &   3.79 & 0.70 & 1.93 & 0      & 0 & 0.25 & 0.09 &	35 & 4   & 10.61 & 0.03 & 1.88 & 0.31 & 0 & 0.27 & 0.12 & 		40 &   4 &   5.06 & 0.29 & 1.90 & 0.11 & 0 & 0.28 & 0.12 \\
-Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0 & 0.33 & 0.19 \\ \midrule
+New topics, tools, methods (a) & 61 & 10 & 5.64 & 0.85 & 1.93 & 0 & 36 & 6 & 2.51 & 0.87 & 1.88 & 0 & 41& 10 & 15.53 & 0.15 & 1.9 & 0.23 \\
+Improvement of research (b) & 61 & 6 &  9.89 & 0.13 & 1.93 & 0.18 & 35 & 4 & 12.41 & 0.02 & 1.88 & 0.35 & 40 & 6 & 7.42 & 0.28 & 1.90 & 0.13 \\
+Better science (c) & 59 & 6 & 3.79 & 0.70 & 1.93 & 0 & 35 & 4 & 10.61 & 0.03 & 1.88 & 0.31 & 40 & 4 &  5.06 & 0.29 & 1.90 & 0.11\\
+Teaching (d) & 61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 40 & 10 & 6.90 & 0.74 & 1.90 & 0 \\ \midrule
 \multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
-Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0 & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0      & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0 & 0.30 & 0.05\\
-Teaching (d) & 					61 &   5 & 7.30 & 0.20 & 0.98 & 0.19 & 0 & 0.30 & 0.05 & 		36 & 5 & 12.71 & 0.26 & 0.97 & 0      & 0.37 & 0.46 & 0.11 & 		40 & 5 & 3.17 & 0.67 & 0.97 & 0      & 0 & 0.36 & 0.11\\
-Connection building (e) & 			61 &   4 & 3.70 & 0.45 & 0.98 & 0      & 0 & 0.28 & 0.03 & 		36 & 3 &   5.89 & 0.12 & 0.97 & 0.28 & 0      & 0.34 & 0.08 &		41 & 3 & 4.42 & 0.22 & 0.98 & 0.18 & 0 & 0.32 & 0.07 \\
-Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0 & 0.28 & 0.03 &		34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0      & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0 & 0.35 & 0.10\\ \midrule
+Improvement of research (b) & 60 & 3 & 2.08 & 0.56 & 0.98 & 0 & 35 & 2 & 5.26 & 0.07 & 0.97 & 0.30 & 40 & 3 & 7.62 & 0.05 & 0.97 & 0.34\\
+Teaching (d) & 61 & 5 & 7.30 & 0.20 & 0.98 & 0.19 & 36 & 5 & 12.71 & 0.26 & 0.97 & 0.46 & 40 &5 & 3.17 & 0.67 & 0.97 & 0 \\
+Connection building (e) & 61 & 4 & 3.70 & 0.45 & 0.98 & 0 & 36 & 3 & 5.89 & 0.12 & 0.97 & 0.28 & 41 & 3 & 4.42 & 0.22 & 0.98 & 0.18\\
+Valued contributions (f) & 57 & 4 & 5.98 & 0.20 & 0.98 & 0.18 & 34 & 5 & 9.46 & 0.09 & 0.97 & 0.36 & 38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 \\ \midrule
 \multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
-Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0 & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0 & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0 & 0.32 & 0.06 \\
-Teaching (d) & 					61 &   5 & 5.63   & 0.34 & 0.98 & 0.09 & 0 & 0.31 & 0.05 &	34 & 4 & 1.29 & 0.86 & 0.97 & 0      & 0 & 0.39 & 0.12 & 			39 & 5 & 3.06 & 0.69 & 0.97 & 0      & 0 & 0.38 & 0.12 \\
-Connection building (e) & 			61 &   4 & 1.64   & 0.80 & 0.98 & 0      & 0 & 0.28 & 0.04 &	34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 0 & 0.36 & 0.09 &			40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 & 0 & 0.33 & 0.07 \\
-Valued contributions (f) & 			58 &   4 & 13.14 & 0.01 & 0.98 & 0.40 & 0 & 0.29 & 0.04 &	32 & 5 & 2.49 & 0.78 & 0.96 & 0      & 0 & 0.40 & 0.14 &			37 & 5 & 4.05 & 0.54 & 0.97 & 0      & 0 & 0.36 & 0.11\\
+Improvement of research (b)  & 59 & 3 & 0.97 & 0.81 & 0.98 & 0 & 34 &2 & 0.01 & 0.99 & 0.97 & 0 & 39 & 3 & 2.79 & 0.43 & 0.97 & 0 \\
+Teaching (d) & 61 & 5 & 5.63 & 0.34 & 0.98 & 0.09 & 34 & 4 & 1.29 & 0.86 & 0.97 & 0 & 39 & 5 & 3.06 & 0.69 & 0.97 & 0 \\
+Connection building (e) & 61 & 4 & 1.64 & 0.80 & 0.98 & 0 & 34 & 3 & 8.42 & 0.04 & 0.97 & 0.40 & 40 & 3 & 3.49 & 0.32 & 0.97 & 0.10 \\
+Valued contributions (f) & 58 & 4 & 13.14 & 0.01 & 0.98 & 0.40 & 32 & 5 & 2.49 & 0.78 & 0.96 & 0 & 37 & 5 & 4.05 & 0.54 & 0.97 & 0 \\
 \bottomrule
@@
 %\end{table*}
-\end{sidewaystable}
-%\end{landscape}
+\end{table*}
 
@@
 Here, we present a number of additional figures related to the survey data presented in the main text.
-The stacked bar charts presented here provide anothe view of the data that produced the results quoted in Table 1 above.
+The stacked bar charts presented here provide another view of the data that produced the results quoted in Table 1 above.
 

### 2018-06-21 abstract.tex
Reword abstract a tiny bit.

@@
 Across many scientific disciplines, methods for recording, storing and analyzing data are rapidly increasing in complexity.
-Facility in data science tools to manage this complexity requires training in new programming languages and frameworks, as well as immersion in new modes of interaction that foster data sharing and collaborative software development and exchange across disciplines.
+Skillfully using data science tools that manage this complexity requires training in new programming languages and frameworks, as well as immersion in new modes of interaction that foster data sharing, collaborative software development and exchange across disciplines.
 Learning these skills from traditional university curricula can be challenging because most courses are not designed to evolve on time scales that can keep pace with rapidly shifting data science methods.

### 2018-06-21 results.tex
Shorten figure caption.

--- 2018-06-21 results.tex
@@
 %\end{subfigure}
-\caption{Post-workshop surveys from three hack weeks: participants in the 2016 astro-, geo- and neuro- hack weeks responded to questions assessing their experiences.
-Response rates are reported in the panel titles for AHW, GHW and NHW, respectively.
-We report here about results in three different domains: the development of technical skills (a -- c), collaboration and teaching (d -- f), and shifts in attitudes towards reproducibility and open science (g, h).}
+\caption{Post-workshop survey responses from the 2016 astro-, geo- and neuro- hack weeks.
+Response rates are in the panel titles.
+Results presented in three different domains: the development of technical skills (a -- c), collaboration and teaching (d -- f), and shifts in attitudes towards reproducibility and open science (g, h).}
 \label{fig:survey}
@@
 \subsubsection*{Example 3: Neuro Hack Week}
-During NHW 2016, one of the teams analyzed an openly available data-set of MRI data from children\cite{Di_Martino2014-qf}, to test the effects of motion on the results, using varying motion cut-offs.
+During NHW 2016, one of the teams analyzed an openly available data-set of MRI data from children\cite{Di_Martino2014-qf}, to test the effects of motion on analysis results, using varying motion cut-offs.
 %They tested both the split-half reliability of an analysis of brain connectivity, as well as an analysis that used machine learning to distinguish between brains of children with and without autism spectrum disorder.
-The team (four researchers from  different institutions) continued to work on this project remotely after the end of NHW, and eventually published a paper describing these results \cite{leonard2017}.
+The team (all from  different institutions) continued to work on this project remotely after the end of NHW, eventually publishing a paper describing these results \cite{leonard2017}.

### 2018-06-21 results.tex
More shorter.

@@
 Given the diversity in skills and backgrounds of participants admitted to our hack weeks, a dependence of learning outcomes and teaching on career stage is plausible.
-At the same time, we suggested earlier in this paper that peer learning is a major mode of knowledge transfer for participants of all career stages.
-We find no strong evidence for a significant difference between early-career researchers and senior participants for any of the three hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12).
-Only for GHW do we find a very weak suggestion that early-career researchers might agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
+At the same time, as suggested earlier, peer learning is a major mode of knowledge transfer for participants of all career stages.
+We find no strong evidence for a significant difference between early-career and senior participants in any of the hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12).
+Only for GHW do we find that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
 Similarly, we find no indication ($p > 0.0007$) that late-career researchers self-report a higher level of teaching at hack weeks compared to early-career researchers (Figure \ref{fig:survey}, (d)).
@@
 Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ between minority and non-minority groups.
-For GHW, there is a weak indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
+For GHW, there is an indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
 Equivalence tests reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW).
-These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an initial indication that our careful facilitation strategies may be effective in fostering participation, and that future work on how demographics interact with hack week attendance may be fruitful.
+These results, while not a decisive exclusion of an effect of race/ethnicity or gender on hack week participation, provide an indication that our facilitation strategies may be effective in fostering participation.
+Future work on how demographics interact with hack week attendance may be fruitful.
 
@@
 While the focus on open science is not necessarily a required component of a hack week, it aligns naturally with many of the goals and values commonly promoted at hack weeks, such as production of open-source software and data sharing.
-In some fields, especially where ethical issues around data sharing and privacy are relevant, this might not be a desired focus of the hack week, or might be replaced or augmented by a discussion of ethical considerations.
+In some fields, especially where ethical issues around data sharing and privacy are relevant, this should be augmented with a discussion of ethical considerations.
 
-We caution the reader that our survey results should be read only as an initial indication of the hypotheses we proposed about the use and outcomes of hack weeks, in line with the surveys' exploratory nature.
-The number of respondents is small and the effects likely subtle.
-This means that the lack of significant differences may be caused by the small statistical power of our sample.
-The most important independent variable--attendance at the hack week--is something that our current survey design cannot account for.
-Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to internal biases of respondents.
-Future work will include a more refined survey design as well as attendance versus non-attendance as a control variable.
+In line with the surveys' exploratory nature, these results should be read only as an initial indication of the hypotheses we proposed about the use and outcomes of hack weeks.
+The number of respondents is small and the effects likely subtle, and lack of significant differences may be due to statistical power in our sample.
+Furthermore, the most important independent variable -- attendance of a hack week -- is not accounted for in our current design.
+Moreover, self-reported learning outcomes are not an objective measure, because they are likely subject to response biases.
+Future work will include more refined survey design and inclusion of control non-attendees.
 
-Because all three events are relatively recent, it is still early to evaluate long-term outcomes, as well as others including publications and collaborations resulting from these events.
+Because all three events are relatively recent, it is still early to evaluate long-term outcomes, including publications and collaborations resulting from these events.
 There are, however, initial indicators that all hack weeks encouraged long-term engagement with new concepts or tools and that they directly resulted in a number of publications \cite{gullysantiago2015,faria2016,keshavan2017,leonard2017,jordan2017,peterson2017,hahn2017,pricewhelan2017}.
-For specific examples, see also below.
+Specific examples follow below.
 

### 2018-06-21 intro.tex
Shorten figure caption.
Got all the references to fit on the last page!

@@
 \dropcap{A}s data become cheaper to gather and store, researchers have become increasingly reliant on computational workflows requiring skills in statistical modeling, machine learning and scalable computation.
-In addition, the recent reproducibility crises in several scientific fields has motivated the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,baker2016}.
-Formal university curricula have been relatively slow to offer courses in these important topics: the slack in this area has often been picked-up by extra-curricular, ad-hoc efforts in the form of less formal workshops \cite{demasi2017}.
-Well-known examples are the Software and Data Carpentry workshops providing training in research computing skills through a volunteer instructor program \cite{b:wilson-swc-lessons-2016,teal2015data}.
-At the same time, there has been a rise in the number of statistical and computational courses designed for specific scientific disciplines.
-Examples include the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, as well as a variety of project-focused (rather than pedagogical) meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, but similar-spirit meetings have been held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
-In general, many of these events tend to emphasize either a more traditional pedagogical class and lecture methodology, or emphasize a focus on ad-hoc projects that participants work on during the event (Figure \ref{fig:hackspectrum}).
+In addition, recent concerns about reproducibility crises motivate the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,baker2016}.
+Formal university curricula have been relatively slow to offer courses in these important topics and this vacuum is often filled by extra-curricular, ad-hoc, less formal workshops \cite{demasi2017}.
+Well-known examples include Software and Data Carpentry workshops, that provide training in research computing through a volunteer instructor program \cite{b:wilson-swc-lessons-2016,teal2015data}.
+Meanwhile, there is an increase in statistical and computational courses designed for specific scientific disciplines, such as the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, as well as a variety of project-focused (rather than pedagogical) meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
+Shorter, meetings are also held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+In general, many of these events tend to emphasize either more traditional pedagogical class and lecture methodologies, or focus on ad-hoc projects developed during the event (Figure \ref{fig:hackspectrum}).
 Pedagogically-focused events follow a classic academic model where novices learn new skills from experts.
@@
 \includegraphics[width=8cm]{NewHackSpectrum.pdf}
-\caption{{\bf Comparison of Extracurricular Workshop Models}: Different types of events lie on a spectrum between the emphasis on pedagogy (e.g.
+\caption{Different types of events lie on a spectrum between the emphasis on pedagogy (e.g.
 Software Carpentry workshop) to an emphasis on project-based/hack-based activities (e.g.
@@
 Astro Hack Week) or pedagogy (e.g.
-Neuro Hack Week)}
+Neuro Hack Week).}
 \label{fig:hackspectrum}

### 2018-06-21 why.tex
So much white-space wasted in that itemized list.

--- 2018-05-28 why.tex
@@
 
-\begin{itemize}
-\item{\textit{Education and Training}:
-While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning.
-This type of lateral knowledge transfer is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.}
+\noindent\textbf{Education and Training}: While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning.
+This type of lateral knowledge transfer \cite{b:wilson-swc-lessons-2016} is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.
 
-\item{\textit{Tool Development}: Hack weeks present an opportunity for scientific software developers to meaningfully engage with users and critically evaluate applications to particular scientific issues.}
+\noindent\textbf{Tool Development}: Hack weeks present an opportunity for scientific software developers to meaningfully engage with users and critically evaluate applications to particular scientific issues.
 
-\item{\textit{Community Building}: Hack weeks provide a tremendous opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
-These events allow computationally minded researchers to break from the isolation of their academic departments and spark new collaborations.}
+\noindent\textbf{Community Building}: Hack weeks provide an opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
+These events allow computationally-minded researchers to break from the isolation of their academic departments and spark new collaborations.
 
-\item{\textit{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
-Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \cite{Sung2003-go}}.
+\noindent\textbf{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
+Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \cite{Sung2003-go}.
 
-\item{\textit{Recruitment and Networking}: Hack weeks are often a melting pot of participants from academia, government, and industry and provide numerous opportunities for networking.
-Close collaboration in diverse groups exposes skills that might be suitable for careers outside of one's narrow domain.}
+\noindent\textbf{Recruitment and Networking}: Hack weeks are often a melting pot of participants from academia, government, and industry and provide numerous opportunities for networking.
+Close collaboration in diverse groups exposes skills that might be suitable for careers outside of one's narrow domain.
 
-\item{\textit{It's fun}: Hack weeks provide a respite from day-to-day research activities and provide a low-stress venue to learn new skills and attempt high-risk projects.}
-
-\end{itemize}
+\noindent\textbf{It's fun}: Hack weeks provide a respite from day-to-day research activities and provide a low-stress venue to learn new skills and attempt high-risk projects.
 

### 2018-06-21 why.tex
It all fits now!

@@
 
-There are several reasons to run a hack week of the sort described here.
-
-\noindent\textbf{Education and Training}: While some hack weeks are focused more on education than others (see Figure \ref{fig:hackspectrum}), there is often a skill-development component in the form of tutorials and informal and peer learning.
-This type of lateral knowledge transfer \cite{b:wilson-swc-lessons-2016} is a core attribute of a hack week, and provides an opportunity to learn skills that are not described in papers and software implementations.
+\noindent\textbf{Education and Training}: While some hack weeks are more focused more on education than others (see Figure \ref{fig:hackspectrum}), skill-development cin the form of tutorials and informal and peer learning is often a component.
+Furthermore, lateral knowledge transfer \cite{b:wilson-swc-lessons-2016} through collaboration provides an opportunity to learn skills that are not described in papers and software implementations.
 
@@
 
-\noindent\textbf{Community Building}: Hack weeks provide an opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
-These events allow computationally-minded researchers to break from the isolation of their academic departments and spark new collaborations.
+\noindent\textbf{Community Building}: Hack weeks are an opportunity to catalyze community development through shared interest in solving computational challenges with open source software.
+They allow computationally-minded researchers to break from the isolation of their institutions and spark new collaborations.
 
-\noindent\textbf{Interdisciplinary research}: Intensive, time-bounded collaborative events are an excellent opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
-Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in risk averse traditional academia \cite{Sung2003-go}.
+\noindent\textbf{Interdisciplinary research}: Intensive, time-bounded collaborative events are an opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
+Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in academic environments \cite{Sung2003-go}.
 
-\noindent\textbf{Recruitment and Networking}: Hack weeks are often a melting pot of participants from academia, government, and industry and provide numerous opportunities for networking.
-Close collaboration in diverse groups exposes skills that might be suitable for careers outside of one's narrow domain.
+\noindent\textbf{Recruitment and Networking}: Hack weeks are a melting pot of participants from academia, government, and industry and provide numerous opportunities for networking.
+Close collaboration in diverse groups exposes skills that might be suitable for careers outside of a narrow domain.
 
-\noindent\textbf{It's fun}: Hack weeks provide a respite from day-to-day research activities and provide a low-stress venue to learn new skills and attempt high-risk projects.
+\noindent\textbf{It's fun}: Hack weeks provide a respite from routine and provide a low-stress venue to learn new skills and attempt high-risk projects.
 

### 2018-06-21 design.tex

@@
 
-Scheduling, group size and venue are important design considerations contributing to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance).
-Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges.
+Several design elements contribute to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance).
+For example, scheduling: longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges.
 By spending more time together, participants are more likely to overcome barriers of professional terminology.
-However, events that are too long may also lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
-A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and project work.
+But events that are too long may also lead to fatigue, resulting in a drop in positive outcomes later in the workshop.
+A well-designed hack week will have a clear schedule limiting the number of parallel sessions, and balancing the duration of taught components and project work.
 
-Another important design consideration is the space. A hack week requires a flexible workspace environment with ample opportunity for re-configuration.
-Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops.
-Workspaces must also accommodate interactive project work where small teams can gather and work together (for more details, see also the SM, Sections 4.1.3 and 4.1.4).
+The space used is also an important consideration.
+A hack week requires a flexible workspace, that allows re-configuration, accommodating both lectures combined with interactive exchanges and individual work on laptops, as well as project work in small teams (see SM 4.1.3, 4.1.4).
 %Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure.
 %However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
-Fortunately, many universities are experimenting with new types of spaces that allow these kinds of activities: for example, the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.
+Fortunately, many universities are experimenting with new types of spaces that allow these kinds of activities, and the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.
 
-Another important design consideration is group size (see also SM, Section 4.1.4).
+Another design consideration is group size (see SM, 4.1.4).
 In a large group, chances for random participant exchanges may be reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
@@
 Hack week outcomes depend strongly on the interest and engagement of participants.
-Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic, such as machine learning, or to analyze a specific data using the tools covered in the tutorials.
+Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic (e.g., machine learning), or to analyze a specific data using the tools covered in the tutorials.
 This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
-This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.
-The large variety in participant backgrounds and experiences--and the resulting range in personalities and objectives of attendees--requires careful, active facilitation of both taught and project components of a hack week (see also SM; Section 4.2.4).
-A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (more information in the SM, Section 4.1.7).
+This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible design.
+The large variety in participant backgrounds and experiences--and the resulting range in personalities and objectives of attendees--requires careful, active facilitation of both taught and project components of a hack week (see SM 4.2.4).
+A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (see SM 4.1.7).
 Community building is a core component of a hack week, and the design of hack weeks needs to include both very strong personalities and very shy participants.

### 2018-06-23 design.tex
Some small corrections

@@
 The space used is also an important consideration.
-A hack week requires a flexible workspace, that allows re-configuration, accommodating both lectures combined with interactive exchanges and individual work on laptops, as well as project work in small teams (see SM 4.1.3, 4.1.4).
+A hack week requires a flexible workspace that allows re-configuration, accommodating both lectures combined with interactive exchanges and individual work on laptops, as well as project work in small teams (see SM 4.1.3, 4.1.4).
 %Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure.
@@
 A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (see SM 4.1.7).
-Community building is a core component of a hack week, and the design of hack weeks needs to include both very strong personalities and very shy participants.
+Community building is a core component of a hack week, and facilitation efforts need to consider both very strong personalities and very shy participants.
 In particular, the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).

### 2018-06-23 intro.tex
Some small corrections

@@
 Meanwhile, there is an increase in statistical and computational courses designed for specific scientific disciplines, such as the \textit{Summer School in Statistics for Astronomers}\footnote{\url{http://astrostatistics.psu.edu/su16/}}, the Google Earth Engine User Summits\footnote{\url{https://events.withgoogle.com/google-earth-engine-user-summit-2017/}}, as well as a variety of project-focused (rather than pedagogical) meetings, such as the dotAstronomy meetings\footnote{\url{http://dotastronomy.com}}.
-Shorter, meetings are also held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
+Shorter meetings are also held in conjunction with conferences, such as the Hack Days at the annual American Astronomical Society meetings, the Brainhack hackathons associated with the meetings of the Organization for Human Brain Mapping and the Society for Neuroscience\cite{Cameron_Craddock2016-wc}, and a hackathon at the American Geophysical Union meeting\footnote{\url{http://onlinelibrary.wiley.com/doi/10.1002/2014EO480004/pdf}}.
 In general, many of these events tend to emphasize either more traditional pedagogical class and lecture methodologies, or focus on ad-hoc projects developed during the event (Figure \ref{fig:hackspectrum}).
@@
 To bridge this gap, we describe here a model that we have implemented: ``Hack Weeks'' that aim to capitalize on the advantages of each of these models.
-These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) with less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) and less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
 
@@
 \includegraphics[width=8cm]{NewHackSpectrum.pdf}
-\caption{Different types of events lie on a spectrum between the emphasis on pedagogy (e.g.
-Software Carpentry workshop) to an emphasis on project-based/hack-based activities (e.g.
+\caption{Different types of events lie on a spectrum between an emphasis on pedagogy (e.g.
+Software Carpentry workshop) and an emphasis on project-based/hack-based activities (e.g.
 at science-oriented hackathons).
@@
 
-As of the publication of this paper, we have run eight such hack week events: four focused on Astronomy, and two each focused on Neuroscience and Geoscience.
+We have run eight such hack week events: four focused on Astronomy, and two each focused on Neuroscience and Geoscience.
 Here we share the philosophy behind the hack week model, results from surveys of participants, practical lessons we have learned in organizing these events, and recommendations for future hack weeks.

### 2018-06-23 supplementary_material.tex
Fix typo.
Some small corrections

--- 2018-06-21 supplementary_material.tex
@@
 We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
-For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a small effect size according to~\citep{cohen1988}.
+For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a medium effect size according to~\citep{cohen1988}.
 
@@
 For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{shiskina2018}.
-We generate 10,000 random data sets from the measured responses by sampling with replacement for each response variable and compute the corresponding effect size in order to generate confidence intervals around the measured Cram\'{e}r's V.
+Following a bootstrap procedure, we generate 10,000 random data sets from the measured responses by sampling with replacement for each response variable and compute the corresponding effect size in order to generate confidence intervals around the measured Cram\'{e}r's V.
 For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
@@
 Here we present the full results of the statistical analysis performed for the survey (Table \ref{tab:survey}).
-We tested for correlations between responses to a number of questions and demographic factors, most notably gender identity and racial/ethnic identity.
+We tested for correlations between responses to a number of questions and demographic factors, most notably career stage, gender identity and racial/ethnic identity.
 The table also shows (1) the number of responses $N$ to each question for each hack week, (2) value of the $\chi^2$ and Cram\'{e}r's V statistics, $\chi^2$ and $\phi_c$, respectively, (3) the $p$-value associated with the $\chi^2$ statistic, (4) the degrees of freedom for both $\chi^2$-test ($\mathrm{dof}_{\chi^2}$) and Cram\'{e}r's V ($\mathrm{dof}_{\phi_c}$), and (5) the upper 95\% confidence interval for the Cram\'{e}r's V statistic ($\mathrm{CI}_{\mathrm{U}}$) as well as the corresponding $p$-value that the null hypothesis (the presence of an effect with moderate or higher strength) can be rejected, $p_\mathrm{eq}$ .
-The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}
+The data sets used in this data analysis are presented in Section \ref{sec:suppfigures}, Figures \ref{fig:corr1} through \ref{fig:corr12}.
 
@@
       \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-      \item[\emph{c}]{Responses to the question of career stage were grouped into three responses: early-career stage; faculty and research scientists; other}
+      \item[\emph{c}]{Responses to the question of career stage were grouped into three categories: early-career stage; faculty and research scientists; other}
       \item[\emph{d}]{Responses to the question of gender identity were grouped with respect to representation within the field of study into two categories: non-minority; minority}

### 2018-06-23 what.tex
Some small corrections

@@
 
-The hackathon, a time-bounded, collaborative events that bring together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events.
+The hackathon, a time-bounded, collaborative event that brings together participants around a shared challenge or learning objective \cite{Decker2015}, forms one primary axis of our events.
 Hackathons originated from the open-source software movement and have historically focused on software and technology development.
 In recent years hackathons have evolved into a model providing opportunities for intensive, interdisciplinary collaboration \cite{Groen2015-cj} and education \cite{Kienzler2015-zu,Lamers2014-xf} in the sciences.
-Core element of all hackathons include opportunities for networking, strengthening of social ties, and the building of community connections, both within and across disciplines.
+Core elements of hackathons include opportunities for networking, strengthening social ties, and building community connections, both within and across disciplines.
 Building on these core elements, hackathons have been implemented in different ways depending on the overall purpose, mode of participation, style of work environment and participant motivation \cite{Drouhard2017}.

### 2018-06-30 participants.tex

@@
 However, a blind selection based purely on an assessment of merit will be counterproductive if it excludes participants who might have had less exposure to certain technologies or fewer opportunities to learn certain skills: for example, requiring a minimum level of programming experience will likely disadvantage candidates who have had fewer opportunities to learn programming due to structural inequalities.
-Additionally, blinding has been found to have negative effects on diversity for committees that  already have a strong commitment to diversity, because these circumstances allow to correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
+Additionally, blinding has been found to have negative effects on diversity for committees that  already have a strong commitment to diversity, because these committees often correct for structural inequalities by considering demographic variables during merit selection~\cite{behaghel2015unintended}.
 In this case, it may be beneficial to construct selection criteria that explicitly consider diversity and inclusivity (as Neuro Hack Week has done; see also SM, Section 4.1.5).

### 2018-06-30 results.tex

@@
 Focusing on the outcomes of astro-, geo- and neuro- hack weeks (AHW, GHW, NHW, respectively) from 2016 and 2017, we find that most participants self-reported successful learning outcomes (AHW: 76\%, GHW: 89\%, NHW: 79\% for responses ``somewhat agree'', ``agree'' and ``strongly agree''; Figure \ref{fig:survey}, (a)).
-The overwhelming majority of respondents at the hack weeks ($>95\%$ for all three events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist  (Figure \ref{fig:survey}, (b, c)).
+The overwhelming majority of respondents at the hack weeks ($>95\%$ for all events) believed that they learned things that improve their day-to-day research, and that attendance has made them a better scientist  (Figure \ref{fig:survey}, (b, c)).
 %More specifically, we compared learning outcomes in data visualization (the only topic explicitly shared between all three hack weeks; Figure \ref{fig:survey}, (d, e).
@@
 At the same time, as suggested earlier, peer learning is a major mode of knowledge transfer for participants of all career stages.
-We find no strong evidence for a significant difference between early-career and senior participants in any of the hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) for all with zero or small effect size; data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12).
+We find no strong evidence for a significant difference between early-career and senior participants in any of the hack weeks for questions regarding learning outcomes of new tools and topics (Figure \ref{fig:survey}, panel (a)), improvements in day-to-day research (Figure \ref{fig:survey}, (b)) or overall improvements in science (Figure \ref{fig:survey}, (c)) ($p > 0.0007$ (trial-corrected significance threshold) with zero or small effect size for all).
+Data and all (including non-significant) correlations are presented in SM, Section 2, Table 1 and SM Section 3, Figures 1-12.
 Only for GHW do we find that early-career researchers agree more strongly that the hack week improved their day-to-day research (p=0.02) with a large effect size of $\phi_c = 0.35$ for $1.88$ degrees of freedom (following~\citep{cohen1988}).
@@
 However, the confidence intervals on the measured effect sizes are wide.
-Testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size suggests that the data are not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).
+Testing for the absence of an effect using an equivalence test on the effect size with an equivalence bound corresponding to a moderate effect size suggests that the data are currently not conclusive enough to reject a moderate to large effect ($ p_\mathrm{eq} > 0.0007$ for all four questions above).
 
@@
 In our surveys, we find no significant dependence of self-reported learning outcomes on gender identity or race/ethnicity ($p > 0.0007$).
-Similarly, minority participants do not significantly differ in their answers with respect to teaching outcomes for any of the hack weeks ($p > 0.0007$).
-Attitudes towards building valuable connections at all three hack weeks or the value of their contributions to their hack teams also do not differ between minority and non-minority groups.
-For GHW, there is an indication that participants from racial/ethnic minorities respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
+Similarly, for none of the hack weeks do minority participants differ significantly in their answers with respect to teaching outcomes, building valuable connections or the value of their contributions to their hack teams  ($p > 0.0007$).
+For GHW, there is an indication that participants from racial/ethnic minorities may respond more positively when asked about building connections ($p=0.04$) with a medium to large effect size ($\phi_c = 0.4$; $\mathrm{dof}_{\phi_c} = 0.97$), while for AHW, responses regarding the value of contributions to hack teams may be spread more widely for participants from racial/ethnic minorities than for Caucasian participants ($p = 0.01$; $\phi_c = 0.4$, $\mathrm{dof}_{\phi_c} = 0.98$).
 Equivalence tests reveal systematically small, though non-significant, $p$-values in the range of $p=0.02-0.05$ for AHW for all four questions above in conjunction with gender or ethnic/racial identity, while for GHW and NHW $p > 0.05$ for the same questions (note that the sample size for GHW and NHW was smaller by a factor of 2 compared to AHW).

### 2018-06-30 supplementary_material.tex

@@
 \textbf{Question} & \multicolumn{8}{c}{\textbf{Astro Hack Week}} & \multicolumn{8}{c}{\textbf{GeoHackWeek}} & \multicolumn{8}{c}{\textbf{NeuroHackWeek}} \\ \midrule
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ & $p_{\mathrm{eq}}$ 
- & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}$ &$p_{\mathrm{eq}}$  \\ \midrule
-\multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{c}$}}  \\ \midrule
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}^{\emph{c}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}^{\emph{c}}$ & $p_{\mathrm{eq}}$ 
+ & $N$ & $\mathrm{dof}_{\chi^2}^\emph{a}$ & $\chi^2$ & $p$ & $\mathrm{dof}_{\phi_c}^\emph{b}$ & $\phi_c$ & $\mathrm{CI}_{\mathrm{U}}^{\emph{c}}$ &$p_{\mathrm{eq}}$  \\ \midrule
+\multicolumn{16}{l}{\textbf{Comparison with career stage$^\emph{d}$}}  \\ \midrule
 New topics, tools, methods (a) & 	61 & 10 &   5.64 & 0.85 & 1.93 & 0      & 0.26 & 0.12 &	36 & 6   &   2.51 & 0.87 & 1.88 & 0      & 0.31 & 0.16 &		41 & 10 & 15.53 & 0.15 & 1.90 & 0.23 & 0.32 & 0.17 \\
@@
 Teaching (d) & 					61 & 10 & 13.65 & 0.19 & 1.93 & 0.17 & 0.28 & 0.13 &	36 & 10 & 13.68 & 0.19 & 1.88 & 0.22 & 0.34 & 0.20 & 		40 & 10 &   6.90 & 0.74 & 1.90 &      0 & 0.33 & 0.19 \\ \midrule
-\multicolumn{16}{l}{\textbf{Gender Identity$^\emph{d}$}} \\ \midrule
+\multicolumn{16}{l}{\textbf{Gender Identity$^\emph{e}$}} \\ \midrule
 Improvement of research (b) & 		60 &   3 & 2.08 & 0.56 & 0.98 & 0      & 0.26 & 0.02&		35 & 2 &   5.26 & 0.07 & 0.97 & 0.30 & 0.33 & 0.07 &		40 & 3 & 7.62 & 0.05 & 0.97 & 0.34 & 0.30 & 0.05\\
@@
 Valued contributions (f) & 			57 &   4 & 5.98 & 0.20 & 0.98 & 0.18 & 0.28 & 0.03 &	34 & 5 &   9.46 & 0.09 & 0.97 & 0.36 & 0.34 & 0.08 &		38 & 5 & 6.75 & 0.34 & 0.97 & 0.21 & 0.35 & 0.10\\ \midrule
-\multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{e}$}} \\ \midrule
+\multicolumn{16}{l}{\textbf{Racial/ethnic Identity$\emph{f}$}} \\ \midrule
 Improvement of research (b)  & 	59 &   3 & 0.97   & 0.81 & 0.98 & 0      & 0.27 & 0.03 &	34 & 2 & 0.01 & 0.99 & 0.97 & 0      & 0.34 & 0.08 &			39 & 3 & 2.79 & 0.43 & 0.97 & 0      & 0.32 & 0.06 \\
@@
       \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-      \item[\emph{c}]{Responses to the question of career stage were grouped into three categories: early-career stage; faculty and research scientists; other}
-      \item[\emph{d}]{Responses to the question of gender identity were grouped with respect to representation within the field of study into two categories: non-minority; minority}
-      \item[\emph{e}]{Responses to the question of racial and ethnic identity were grouped with respect to representation within the field of study into two categories: white; non-white}
+      \item[\emph{c}]{The bootstrapped upper $95\%$ confidence intervals for the Cram\'{e}r's V statistic derived according to~\citep{shiskina2018}.}
+      \item[\emph{d}]{Responses to the question of career stage were grouped into three categories: early-career stage; faculty and research scientists; other}
+      \item[\emph{e}]{Responses to the question of gender identity were grouped with respect to representation within the field of study into two categories: non-minority; minority}
+      \item[\emph{f}]{Responses to the question of racial and ethnic identity were grouped with respect to representation within the field of study into two categories: white; non-white}
 \end{tablenotes}

### 2018-06-30 themes.tex

@@
 To date, all hack weeks we have organized have been subject-specific, i.e.\ aimed at bringing together a community with a shared scientific interest, such as neuroscience.
-Advantages to this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.
+Advantages of this approach include shared language and scientific objectives within communities organized by subject, leaving more time for active collaboration on cutting-edge science.
 On the other hand, homogeneity may lead to \textit{group think} and inhibit new, creative solutions.

### 2018-06-30 why.tex

@@
 
-\noindent\textbf{Education and Training}: While some hack weeks are more focused more on education than others (see Figure \ref{fig:hackspectrum}), skill-development cin the form of tutorials and informal and peer learning is often a component.
+\noindent\textbf{Education and Training}: While some hack weeks are more focused more on education than others (see Figure \ref{fig:hackspectrum}), skill-development in the form of tutorials as well as informal and peer learning is often a component.
 Furthermore, lateral knowledge transfer \cite{b:wilson-swc-lessons-2016} through collaboration provides an opportunity to learn skills that are not described in papers and software implementations.
@@
 
-\noindent\textbf{Community Building}: Hack weeks are an opportunity to catalyze community development through shared interest in solving computational challenges with open source software.
+\noindent\textbf{Community Building}: Hack weeks are an opportunity to catalyze community development through a shared interest in solving computational challenges with open source software.
 They allow computationally-minded researchers to break from the isolation of their institutions and spark new collaborations.
@@
 \noindent\textbf{Interdisciplinary research}: Intensive, time-bounded collaborative events are an opportunity to experiment with concepts, questions, and methods that span boundaries within and across disciplines.
-Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in academic environments \cite{Sung2003-go}.
+Despite the fact that interdisciplinary experiments are impactful \cite{Hall2012-hi}, they are often discouraged in traditional academic environments \cite{Sung2003-go}.
 
@@
 
-\noindent\textbf{It's fun}: Hack weeks provide a respite from routine and provide a low-stress venue to learn new skills and attempt high-risk projects.
+\noindent\textbf{It's fun}: Hack weeks provide a respite from routine and a low-stress venue to learn new skills and attempt high-risk projects.
 
 \noindent Note that the reasons for participants to attend a hack week are as diverse as the reasons for running such an event.
-Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focusing on an existing project already in progress (for more details on setting objectives, see SM, Section 4.1.2).
+Beginner participants may attend primarily to learn a new technique, while others may attend to gain experience in mentoring, or to focus on an existing project already in progress (for more details on setting objectives, see SM, Section 4.1.2).

### 2018-07-07 intro.tex
Version as submitted to PNAS
Minor changes to make editorial office happy

@@
 In addition, recent concerns about reproducibility crises motivate the acquisition of skills in open science and the design of reproducible workflows \cite[e.g.][]{pashler2012,baker2016}.
-Formal university curricula have been relatively slow to offer courses in these important topics and this vacuum is often filled by extra-curricular, ad-hoc, less formal workshops \cite{demasi2017}.
+Formal university curricula have been relatively slow to offer courses in these important topics and this vacuum is often filled by extra-curricular, ad-hoc, less formal workshops.
 Well-known examples include Software and Data Carpentry workshops, that provide training in research computing through a volunteer instructor program \cite{b:wilson-swc-lessons-2016,teal2015data}.
@@
 These week-long events combine structured periods focused on pedagogy (often with an emphasis on statistical and computational techniques) and less structured periods devoted to hacks and creative projects, with the goal of encouraging collaboration and learning among people at various stages of their career.
+\renewcommand{\floatpagefraction}{.3}%
+\renewcommand{\dblfloatpagefraction}{.3}%
 
-%Our hack week model shares some similarities to those summer schools that combine individual or group project work with teaching methods that transfer knowledge from experts to early-career researchers.
-%However, a distinguishing factor of our hack weeks in that they tend to be less structured and more participant-driven, so that the learning content is co-created between organizers and participants based on the unique experience of people in the room.
-
-\begin{figure}
+\begin{figure}[h!]
 \begin{center}

### 2018-07-07 supplementary_material.tex
Minor changes to make editorial office happy

@@
 \usepackage{amssymb}	% Extra maths symbols
-%\usepackage{natbib}
+\usepackage{natbib}
 
@@
 No responses were discarded, and no pre-processing was performed on the data.
-We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \citep{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
-For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a medium effect size according to~\citep{cohen1988}.
+We test for correlations between demographic characteristics (independent variables) and question responses (dependent variables) using a standard $\chi^2$ test and compute the effect sizes via a bias-corrected version of Cram\'{e}r's V \cite{cramer1946,bergsma2013}, denoted $\phi_c$ in the manuscript.
+For a fairly permissive significant threshold of $\alpha < 0.05$, we require a $p$-value of $p < 0.0007$, corrected for $N_\mathrm{trials} = 36$ trials, to claim a significant effect as well as at least a medium effect size according to~\cite{cohen1988}.
 
 We note that while a non-significant $p$-value can reject a hypothesis, this is not automatically evidence for the corresponding null hypothesis, since a lack of statistical power could also originate in a too small sample size.
-For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\citep{shiskina2018}.
+For cases where the null hypothesis corresponds to the scientific hypothesis of interest, we perform an equivalence test based on the Cram\'{e}r's V effect size following~\cite{shiskina2018}.
 Following a bootstrap procedure, we generate 10,000 random data sets from the measured responses by sampling with replacement for each response variable and compute the corresponding effect size in order to generate confidence intervals around the measured Cram\'{e}r's V.
-For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \citet{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
+For each hypothesis, we determine an equivalence bound for Cram\'{e}r's V for a medium-strength effect size based on \cite{cohen1988} given the number of degrees of freedom and calculate the $p$-value that the measured effect size could extend above\footnote{Note that because Cram\'{e}r's V is defined on the interval $[0,1]$, we only require an upper bound} this equivalence bound.
 If the latter hypothesis can be rejected, this indicates that the observed effect falls within the equivalence bounds and is statistically smaller than any effect deemed worthwhile.
@@
       \item[\emph{a}]{While the range of Likert items was the same for every question, the degrees of freedom for $\chi^2$ and Cram\'{e}r's V may differ depending on the number of items with non-zero responses, and the number of non-zero items in the independent variable.}
-      \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\citep{bergsma2013}}
-      \item[\emph{c}]{The bootstrapped upper $95\%$ confidence intervals for the Cram\'{e}r's V statistic derived according to~\citep{shiskina2018}.}
+      \item[\emph{b}]{The degrees of freedom for Cram\'{e}r's V corrected according to~\cite{bergsma2013}}
+      \item[\emph{c}]{The bootstrapped upper $95\%$ confidence intervals for the Cram\'{e}r's V statistic derived according to~\cite{shiskina2018}.}
       \item[\emph{d}]{Responses to the question of career stage were grouped into three categories: early-career stage; faculty and research scientists; other}
@@
 
-Since human decision makers also routinely overestimate their ability to forecast future performance of candidates and underestimate intrinsic variance during selection procedures~\citep{highhouse2008}, it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
+Since human decision makers also routinely overestimate their ability to forecast future performance of candidates and underestimate intrinsic variance during selection procedures~\cite{highhouse2008}, it may be beneficial to set a fairly tolerant threshold for acceptance based on merit criteria, and select the cohort from this meritorious pool via an algorithm, imposing outside constraints on the selection based on the goals of the workshop.
 
@@
 
-Some of our hack week tutorials follow the structure of the Software Carpentry model \citep{b:wilson-swc-lessons-2016}.
+Some of our hack week tutorials follow the structure of the Software Carpentry model \cite{b:wilson-swc-lessons-2016}.
 The content is divided into a series of lessons, each having a well-defined set of learning objectives, key questions, and expected outcomes.
@@
 This provides a simple method for identifying those students who need help without interrupting the flow of the tutorial.
-We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \citep{kluyver2016jupyter} or by following the markdown templates provided by Software Carpentry.
+We encourage instructors to develop online material to accompany their tutorials, either in the form of Jupyter Notebooks \cite{kluyver2016jupyter} or by following the markdown templates provided by Software Carpentry.
 This way, the hack week ``curriculum'' can be built up over time and students can refer to online content after the tutorial if they need to revisit certain topics.
@@
 Though we encourage participants to install these software dependencies on their own computers, so that they can continue using them after the tutorial, this is sometimes so complicated that it slows down teaching.
-In some cases, we have experimented with using \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \citep{holdgraf2017portable}.
-Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \citep{Holdgraf2017-pd, Titus_Brown_undated-pc}.
+In some cases, we have experimented with using \emph{Jupyterhub}\footnote{https://jupyterhub.readthedocs.io/en/latest/} to provide students with immediate access to a computational environment that provides teaching materials (e.g., Jupyter notebooks), software dependencies, and data \cite{holdgraf2017portable}.
+Providing access to an ephemeral consistent environment for the purpose of teaching has become even simpler with the recent advent of the \emph{Binder} system \cite{Holdgraf2017-pd, Titus_Brown_undated-pc}.
 
@@
 The \textit{impostor phenomenon}, or \textit{impostor syndrome} (IS) is a dissonant feeling experienced by certain high-achieving individuals, that despite objective evidence to the contrary, they are in fact not as intelligent or capable as they appear.
-Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \citep{Clance1978-ef}.
+Individuals with IS thus experience a fear of being ``found out'', shamed and expelled from their environment \cite{clance1978imposter}.
 Initial observations of the first Astro Hack Week conducted by data science ethnographer Brittany Fiore-Gartland\footnote{\url{http://astrohack week.org/blog/ethnographic-notes.html}} suggested that hack weeks are an environment prone to a particular kind of IS, namely that participants feel the need to be experts in multiple scientific and technical aspects of the activities pursued during a hack week.
@@
 The prevalence of IS at a hack week may be endemic to the format, and should thus be a major concern for any organizing committee.
-This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \citep{Parkman2016-ro}.
+This is because of the chilling effect it tends to have on participants and the community as a whole, and particularly on women (in particular in fields in which women are under-represented) and members of ethnic and racial minorities, correlating with anxiety and other forms of mental distress \cite{Parkman2016-ro}.
 Less severely, another major concern is that IS inhibits risk taking: Participants experiencing it will be less likely to ask a question, to put forward an idea for a hack, and to be pro-active about forming new collaborations.
@@
 
-\bibliographystyle{aasjournal}
+\bibliographystyle{unsrt}
 \bibliography{paper}
